### 2022-03-13 rmd/suppl_info.Rmd
edit: rmd

---
output:
  pdf_document:
    latex_engine: xelatex
    includes:
      before_body: title.tex
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
header-includes:
  \pagenumbering{gobble}
  \usepackage{amsmath}
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```

```{=tex}
\newpage
\pagenumbering{arabic}
```

```{r table_sensitivity}

source(here::here("code/table_stvy.R"))

j <- 0

## conditional text
c_text <- c("ecosystem size",
            "ecosystem complexity")

table_stvy_chunk <- knit_chunk <- NULL

for(i in 1:length(table_stvy)) {

  table_legend <- paste0("Sensitivity analysis of the ", c_text[i]," effect on the FCL. Parameter estimates of linear regression models are shown. Response variables are Spearman's rank correlation between food chain length and ", c_text[i],". Response and explanatory variables (i.e., simulation parameters) were standardized (mean = 0, SD = 1) before the analysis.")

  knit_chunk <- paste0("## Table S", i + j, " Sensitivity analysis for ", c_text[i]," \n",
                       table_legend,
                       "\n```{r table_s", i + j,", results='asis'}
                       \n\n
                       table_stvy[[", i,"]]
                       \n\n
                       ```\n",
                       "\n\\pagebreak\n")

  table_stvy_chunk <- c(table_stvy_chunk, knit_chunk)

}

```

`r paste(knit(text = table_stvy_chunk), collapse = '\n')`

### 2022-03-13 rmd/title.tex
edit: rmd

%\begin{titlepage}
\begin{center}
\LARGE{\textbf{Supplementary Materials for:}}\\[7.5mm]
\LARGE{Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks}\\[7.5mm]
\end{center}
\begin{flushleft}
\textbf{This PDF file includes:}
\begin{itemize}
\item XXX
\item References
\end{itemize}
\end{flushleft}
% \end{titlepage}

\newpage

### 2022-03-21 code/figure_si_fcl.R
edit: figure si

@@ -13,12 +13,12 @@ load(file = here::here("output/result_main.RData"))
 df_sim <- sim_main_result %>% 
   group_by(param_set) %>% 
   filter(!(theta == 1 &
-           sd_disturb_source == 5 &
+           sd_disturb_source == 3 &
            sd_disturb_lon == 0.1)) %>% 
-  mutate(igp = case_when(e_bp == 0 ~ "Chain",
+  mutate(omn = case_when(e_bp == 0 ~ "Chain",
                          e_bp == 2 ~ "Weak",
                          e_bp == 4 ~ "Strong"),
-         igp = factor(igp, levels = c("Chain",
+         omn = factor(omn, levels = c("Chain",
                                       "Weak",
                                       "Strong")),
          productivity = recode(r_b,
@@ -38,10 +38,10 @@ df_sim <- sim_main_result %>%
   )
 
 df_param <- expand.grid(theta = 0.1,
-                        sd_disturb_source = c(0.1, 5),
-                        sd_disturb_lon = c(0.1, 5)) %>% 
+                        sd_disturb_source = c(0.1, 3),
+                        sd_disturb_lon = c(0.1, 3)) %>% 
   as_tibble() %>% 
-  filter(sd_disturb_source != sd_disturb_lon)
+  filter(sd_disturb_source > sd_disturb_lon)
 
 
 # figure ------------------------------------------------------------------
@@ -56,7 +56,7 @@ list_g_np <- foreach(i = seq_len(nrow(df_param))) %do% {
            sd_disturb_source == df_param$sd_disturb_source[i]) %>% 
     ggplot(aes(x = n_patch,
                y = fcl,
-               linetype = igp)) +
+               linetype = omn)) +
     geom_smooth(method = "loess",
                 size = 0.5,
                 color = "salmon",
@@ -79,7 +79,7 @@ list_g_pb <- foreach(i = seq_len(nrow(df_param))) %do% {
            sd_disturb_source == df_param$sd_disturb_source[i]) %>% 
     ggplot(aes(x = p_branch,
                y = fcl,
-               linetype = igp)) +
+               linetype = omn)) +
     geom_smooth(method = "loess",
                 size = 0.5,
                 color = "steelblue",
@@ -93,3 +93,8 @@ list_g_pb <- foreach(i = seq_len(nrow(df_param))) %do% {
   
   return(g_pb)
 }
+
+
[…]

### 2022-03-21 code/library.R
edit: misc

@@ -1,7 +1,6 @@
 
 # common packages
 pacman::p_load(mcbrnet,
-               tidyverse,
                tidygraph,
                foreach,
                doParallel,
@@ -12,4 +11,5 @@ pacman::p_load(mcbrnet,
                ggraph,
                NetIndices,
                patchwork,
-               igraph)
+               igraph,
+               tidyverse)

### 2022-03-21 code/table_stvy.R

@@ -49,10 +49,10 @@ table_stvy <- foreach(i = 1:2) %do% {
     tidy() %>%
     mutate(Term = param_label,
            Interpretation = term_label) %>% 
-    select(Term,
-           Interpretation,
-           Estimate =estimate,
-           SE = std.error) %>% 
+    dplyr::select(Term,
+                  Interpretation,
+                  Estimate =estimate,
+                  SE = std.error) %>% 
     kable(format = "markdown",
           digits = 3)
 }

### 2022-03-21 code/run_sim_stvy.R
edit: stvy
change analysis setup

@@ -13,8 +13,8 @@ registerDoSNOW(cl)
 n_para <- 500
 df_param <- tibble(# parameters for brnet
                    mean_disturb_source = 0.8,
-                   sd_disturb_source = runif(n_para, 0.01, 1),
-                   sd_disturb_lon = runif(n_para, 0.01, 1),
+                   sd_disturb_source = runif(n_para, 0.05, 5),
+                   sd_disturb_lon = runif(n_para, 0.05, 5),
                    
                    # carring capacity
                    base_k = 100,

### 2022-03-21 code/table_param.R
add: table param

@@ -1,0 +1,83 @@
+
+# setup -------------------------------------------------------------------
+
+rm(list = ls())
+pacman::p_load(tidyverse)
+
+
+# table -------------------------------------------------------------------
+
+load(file = here::here("output/result_main.RData"))
+
+df_param <- sim_main_result %>% 
+  dplyr::select(-n_rep,
+                -mc_capacity,
+                -n_patch,
+                -p_branch,
+                -param_set,
+                -base_k,
+                -z,
+                -n_timestep,
+                -n_warmup,
+                -n_burnin,
+                -fcl,
+                -p_basal,
+                -p_igprey,
+                -p_igpred) %>% 
+  summarize(across(.fns = function(x) list(unique(x)))) %>% 
+  pivot_longer(cols = everything(),
+               names_to = "param",
+               values_to = "Main") %>%
+  mutate(Parameter = case_when(param == "r_b" ~ "$r_b$",
+                               param == "e_bc" ~"$e_{bc}$",
+                               param == "e_bp" ~"$e_{bp}$",
+                               param == "e_cp" ~"$e_{cp}$",
+                               param == "a_bc" ~"$a_{bc}$",
+                               param == "a_bp" ~"$a_{bp}$",
+                               param == "a_cp" ~"$a_{cp}$",
+                               param == "h_bc" ~"$h_{bc}$",
+                               param == "h_bp" ~"$h_{bp}$",
+                               param == "h_cp" ~"$h_{cp}$",
+                               param == "theta" ~"$\\theta$",
+                               param == "p_disturb" ~"$p_m$",
+                               param == "mean_disturb_source" ~ "$m$",
+                               param == "sd_disturb_source" ~ "$\\sigma_h$",
+                               param == "sd_disturb_lon" ~ "$\\sigma_l$",
+                               param == "s0" ~ "$s_0$",
+                               param == "p_dispersal" ~ "$p_d$"),
+         Group = case_when(param %in% c("r_b",
+                                        "s0",
+                                        "e_bc",
+                                        "e_bp",
+                                        "e_cp",
+                                        "a_bc",
+                                        "a_bp",
+                                        "a_cp",
+                                        "h_bc",
+                                        "h_bp",
+                                        "h_cp") ~ "Food web",
+                           param %in% c("theta",
[…]

### 2022-03-21 rmd/suppl_info.Rmd
edit
Update suppl_info.Rmd

@@
 
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE,
+                      message = FALSE)
 
@@
 Parameter estimates of linear regression models are shown.
-Response variables are Spearman's rank correlation between food chain length and ", c_text[i],".
+The response variable is the Spearman's rank correlation between food chain length and ", c_text[i],".
 Response and explanatory variables (i.e., simulation parameters) were standardized (mean = 0, SD = 1) before the analysis.")
@@
 `r paste(knit(text = table_stvy_chunk), collapse = '\n')`
+
+\pagebreak
+
+```{r si_figure_fcl}
+
+source(here::here("code/figure_si_fcl.R"))
+
+j <- 0
+
+## conditional text
+c_text <- c("ecosystem size",
+            "ecosystem complexity")
+
+fig_fcl_chunk <- knit_chunk <- NULL
+
+for(i in 1:length(list_g_fcl)) {
+
+  figure_legend <- paste0(str_to_sentence(c_text[i])," effect on food chain length.
+We used the following values of parameters: $\\theta$ = `r df_sim$theta[i]`, $\\sigma_l$ = `r df_sim$sd_disturb_lon[i]`, and $\\sigma_h$ = `r df_sim$sd_disturb_source[i]`")
+
+  knit_chunk <- paste0("## Figure S", i + j, " ", str_to_sentence(c_text[i]), " effect on food chain length", " \n",
+                       "\n```{r figure_s", i + j,", fig.height = 6.5, fig.width = 8}
+                       \n\n
+                       list_g_fcl[[", i,"]]
+                       \n\n
+                       ```\n",
+                       figure_legend,
+                       "\n\\pagebreak\n")
+
+  fig_fcl_chunk <- c(fig_fcl_chunk, knit_chunk)
+
+}
+
+```
+
+`r paste(knit(text = fig_fcl_chunk), collapse = '\n')`

### 2022-03-21 code/table_param.R
edit: table param

@@ -80,4 +80,5 @@ df_param <- sim_main_result %>%
                                  param == "s0" ~ "Unif(0.5, 1)",
                                  param == "p_dispersal" ~ "Unif(0, 0.1)")) %>% 
   arrange(Group, Parameter) %>% 
+  mutate(Group = ifelse(duplicated(Group), NA, Group)) %>% 
   dplyr::select(Group, Parameter, Main, Sensitivity)

### 2022-03-23 code/table_param.R
edit: table

@@ -2,14 +2,14 @@
 # setup -------------------------------------------------------------------
 
 rm(list = ls())
-pacman::p_load(tidyverse)
+source(here::here("code/library.R"))
 
 
 # table -------------------------------------------------------------------
 
 load(file = here::here("output/result_main.RData"))
 
-df_param <- sim_main_result %>% 
+df_value <- sim_main_result %>% 
   dplyr::select(-n_rep,
                 -mc_capacity,
                 -n_patch,
@@ -24,7 +24,8 @@ df_param <- sim_main_result %>%
                 -p_basal,
                 -p_igprey,
                 -p_igpred) %>% 
-  summarize(across(.fns = function(x) list(unique(x)))) %>% 
+  filter(sd_disturb_source > sd_disturb_lon) %>% 
+  summarize(across(.fns = function(x) list(sort(unique(x))))) %>% 
   pivot_longer(cols = everything(),
                names_to = "param",
                values_to = "Main") %>%
@@ -78,7 +79,24 @@ df_param <- sim_main_result %>%
                                  param == "sd_disturb_source" ~ "Unif(0.05, 5)",
                                  param == "sd_disturb_lon" ~ "Unif(0.05, 5)",
                                  param == "s0" ~ "Unif(0.5, 1)",
-                                 param == "p_dispersal" ~ "Unif(0, 0.1)")) %>% 
+                                 param == "p_dispersal" ~ "Unif(0, 0.1)"),
+         Interpretation = case_when(param == "r_b" ~ "Reproductive rate of basal species",
+                                    param == "e_bc" ~ "Conversion efficiency (B to C)",
+                                    param == "e_bp" ~ "Conversion efficiency (B to P)",
+                                    param == "e_cp" ~ "Conversion efficiency (C to P)",
+                                    param == "a_bc" ~ "Attack rate (C on B)",
+                                    param == "a_bp" ~ "Attack rate (P on B)",
+                                    param == "a_cp" ~ "Attack rate (P on C)",
+                                    param == "h_bc" ~ "Handling time (C on B)",
+                                    param == "h_bp" ~ "Handling time (P on B)",
+                                    param == "h_cp" ~ "Handling time (P on C)",
+                                    param == "theta" ~ "Inverse of mean dispersal distance",
+                                    param == "p_disturb" ~ "Disturbance prob.",
+                                    param == "mean_disturb_source" ~ "Disturbance intensity",
[…]

### 2022-03-23 code/table_stvy.R

@@ -8,48 +8,65 @@ source(here::here("code/analysis_stvy.R"))
 
 # table -------------------------------------------------------------------
 
-term_label <- c("Intercept",
-                "Disturbance prob.",
-                "Disturbance variation at headwaters",
-                "Local disturbance variation",
-                "Reproductive rate of basal species",
-                "Conversion efficiency (B to C)",
-                "Conversion efficiency (B to P)",
-                "Conversion efficiency (C to P)",
-                "Attack rate (B to C)",
-                "Attack rate (B to P)",
-                "Attack rate (C to P)",
-                "Handling time (B to C)",
-                "Handling time (B to P)",
-                "Handling time (C to P)",
-                "Survival prob.",
-                "Dispersal prob.",
-                "Inverse of mean dispersal distance")
-
-param_label <- c("",
-                 "$p_{m}$",
-                 "$\\sigma_{s}$",
-                 "$\\sigma_{l}$",
-                 "$r_b$",
-                 "$e_{bc}$",
-                 "$e_{bp}$",
-                 "$e_{cp}$",
-                 "$a_{bc}$",
-                 "$a_{bp}$",
-                 "$a_{cp}$",
-                 "$h_{bc}$",
-                 "$h_{bp}$",
-                 "$h_{cp}$",
-                 "$s_{0}$",
-                 "$p_{d}$",
-                 "$\\theta$")
-
 table_stvy <- foreach(i = 1:2) %do% {
   fit_sense$fit[[i]] %>% 
     tidy() %>%
-    mutate(Term = param_label,
-           Interpretation = term_label) %>% 
-    dplyr::select(Term,
+    mutate(Group = case_when(term %in% c("r_b",
+                                         "s0",
+                                         "e_bc",
+                                         "e_bp",
+                                         "e_cp",
+                                         "a_bc",
+                                         "a_bp",
+                                         "a_cp",
+                                         "h_bc",
+                                         "h_bp",
+                                         "h_cp") ~ "Food web",
+                             term %in% c("theta",
+                                         "p_dispersal") ~ "Dispersal",
+                             term %in% c("p_disturb",
+                                         "mean_disturb_source",
+                                         "sd_disturb_source",
+                                         "sd_disturb_lon") ~ "Disturbance"),
+           Term = case_when(term == "r_b" ~ "$r_b$",
+                            term == "e_bc" ~"$e_{bc}$",
[…]

### 2022-03-23 code/figure_si_fcl.R
edit: si figure

@@ -12,9 +12,6 @@ load(file = here::here("output/result_main.RData"))
 
 df_sim <- sim_main_result %>% 
   group_by(param_set) %>% 
-  filter(!(theta == 1 &
-           sd_disturb_source == 3 &
-           sd_disturb_lon == 0.1)) %>% 
   mutate(omn = case_when(e_bp == 0 ~ "Chain",
                          e_bp == 2 ~ "Weak",
                          e_bp == 4 ~ "Strong"),
@@ -34,10 +31,21 @@ df_sim <- sim_main_result %>%
                           `0.1` = sprintf('"High disturbance"~(p[m]=="%.2f")',
                                           p_disturb)),
          disturb = factor(disturb,
-                          levels = rev(levels(factor(disturb))))
-  )
+                          levels = rev(levels(factor(disturb))))) %>% 
+  pivot_longer(cols = c(fcl, p_basal, p_igprey, p_igpred),
+               names_to = "y",
+               values_to = "value") %>% 
+  mutate(y = case_when(y == "fcl" ~ "Food~chain~length",
+                       y == "p_basal" ~ "Basal~species~occupancy",
+                       y == "p_igprey" ~ "IG-prey~occupancy",
+                       y == "p_igpred" ~ "IG-predator~occupancy"),
+         y = factor(y,
+                    levels = c("Food~chain~length",
+                               "Basal~species~occupancy",
+                               "IG-prey~occupancy",
+                               "IG-predator~occupancy")))
 
-df_param <- expand.grid(theta = 0.1,
+df_param <- expand.grid(theta = c(0.1, 1),
                         sd_disturb_source = c(0.1, 3),
                         sd_disturb_lon = c(0.1, 3)) %>% 
   as_tibble() %>% 
@@ -48,6 +56,7 @@ df_param <- expand.grid(theta = 0.1,
 
 theme_set(plt_theme)
 
+## ecosystem size
 list_g_np <- foreach(i = seq_len(nrow(df_param))) %do% {
   
   g_np <- df_sim %>% 
@@ -55,22 +64,24 @@ list_g_np <- foreach(i = seq_len(nrow(df_param))) %do% {
            sd_disturb_lon == df_param$sd_disturb_lon[i],
            sd_disturb_source == df_param$sd_disturb_source[i]) %>% 
     ggplot(aes(x = n_patch,
-               y = fcl,
+               y = value,
                linetype = omn)) +
     geom_smooth(method = "loess",
                 size = 0.5,
[…]

### 2022-03-23 code/figure_si_fcl.R
Update figure_si_fcl.R

@@ -65,18 +65,20 @@ list_g_np <- foreach(i = seq_len(nrow(df_param))) %do% {
            sd_disturb_source == df_param$sd_disturb_source[i]) %>% 
     ggplot(aes(x = n_patch,
                y = value,
-               linetype = omn)) +
+               linetype = omn,
+               color = y,
+               fill = y)) +
     geom_smooth(method = "loess",
-                size = 0.5,
-                color = "salmon",
-                fill = "salmon") +
+                size = 0.5) +
     facet_grid(rows = vars(y),
                cols = vars(disturb, productivity),
                labeller = label_parsed,
                scales = "free_y") +
     labs(x = "Ecosystem size (number of patches)",
          y = "Value",
-         linetype = "Omnivory")
+         linetype = "Omnivory") +
+    guides(color = "none",
+           fill = "none")
   
   return(g_np)
 }
@@ -90,18 +92,20 @@ list_g_pb <- foreach(i = seq_len(nrow(df_param))) %do% {
            sd_disturb_source == df_param$sd_disturb_source[i]) %>% 
     ggplot(aes(x = p_branch,
                y = value,
-               linetype = omn)) +
+               linetype = omn,
+               color = y,
+               fill = y)) +
     geom_smooth(method = "loess",
-                size = 0.5,
-                color = "steelblue",
-                fill = "steelblue") +
+                size = 0.5) +
     facet_grid(rows = vars(y),
                cols = vars(disturb, productivity),
                labeller = label_parsed,
                scales = "free_y") +
     labs(x = "Ecosystem complexity (branching prob.)",
          y = "Value",
-         linetype = "Omnivory")
+         linetype = "Omnivory") +
+    guides(color = "none",
+           fill = "none")
   
   return(g_pb)
 }

### 2022-03-23 rmd/suppl_info.Rmd
Update suppl_info.Rmd

@@
 
+options(knitr.kable.NA = '')
+
 ```
@@
 ```
+## Table S1 Parameter values
+
+Parameter values used for the main and sensitivity analysis.
+
+```{r table_param}
+
+source(here::here("code/table_param.R"))
+knitr::kable(df_value, format = "markdown")
+
+```
+
+\pagebreak
 
@@
 
-j <- 0
+j <- 1
 
@@
 
-  table_legend <- paste0("Sensitivity analysis of the ", c_text[i]," effect on the FCL.
+  table_legend <- paste0("Sensitivity analysis of the ", c_text[i]," effect on food chain length.
 Parameter estimates of linear regression models are shown.
@@
 
-  figure_legend <- paste0(str_to_sentence(c_text[i])," effect on food chain length.
-We used the following values of parameters: $\\theta$ = `r df_sim$theta[i]`, $\\sigma_l$ = `r df_sim$sd_disturb_lon[i]`, and $\\sigma_h$ = `r df_sim$sd_disturb_source[i]`")
+  K <- length(list_g_fcl[[i]])
 
-  knit_chunk <- paste0("## Figure S", i + j, " ", str_to_sentence(c_text[i]), " effect on food chain length", " \n",
-                       "\n```{r figure_s", i + j,", fig.height = 6.5, fig.width = 8}
+  for(k in 1:K) {
+
+      figure_legend <- paste0(str_to_sentence(c_text[i])," effect on food chain length and occupancies of constituent species.
+Lines and shades represent loess curves fitted to the simulated data and 95% confidence intervals.
+In this simulation, the following values of parameters were used: $\\theta$ = ", df_param$theta[k], ", $\\sigma_l$ = ", df_param$sd_disturb_lon[k], ", and $\\sigma_h$ = ", df_param$sd_disturb_source[k], ".")
+
+  knit_chunk <- paste0("## Figure S", K * (i - 1) + k + j, " ", str_to_sentence(c_text[i]), " effect on food chain length", " \n",
+                       "\n```{r figure_s", K * (i - 1) + k + j,", fig.height = 10, fig.width = 11.5}
                        \n\n
-                       list_g_fcl[[", i,"]]
+                       list_g_fcl[[", i, "]][[", k, "]]
                        \n\n
@@
 
+}
+
 ```

### 2022-03-24 code/figure_main_sim.R
edit: figures

@@ -45,17 +45,20 @@ theme_set(plt_theme)
 g_np <- df_sim %>% 
   ggplot(aes(x = n_patch,
              y = fcl,
-             linetype = omn)) +
+             color = omn,
+             fill = omn)) +
   geom_smooth(method = "loess",
-              size = 0.5,
-              color = "salmon",
-              fill = "salmon") +
+              size = 0.5) +
   facet_grid(rows = vars(disturb),
              cols = vars(productivity),
              labeller = label_parsed) +
   labs(x = "Ecosystem size (number of patches)",
        y = "Food chain length",
-       linetype = "Omnivory")
+       color = "Omnivory",
+       fill = "Omnivory") +
+  MetBrewer::scale_color_met_d("Hiroshige") +
+  MetBrewer::scale_fill_met_d("Hiroshige")
+
 
 ggsave(g_np,
        filename = here::here("output/figure_np_main.pdf"),
@@ -65,17 +68,20 @@ ggsave(g_np,
 g_pb <- df_sim %>% 
   ggplot(aes(x = p_branch,
              y = fcl,
-             linetype = omn)) +
+             color = omn,
+             fill = omn)) +
   geom_smooth(method = "loess",
-              size = 0.5,
-              color = "steelblue",
-              fill = "steelblue") +
+              size = 0.5) +
   facet_grid(rows = vars(disturb),
              cols = vars(productivity),
              labeller = label_parsed) +
   labs(x = "Ecosystem complexity (branching prob.)",
        y = "Food chain length",
-       linetype = "Omnivory")
+       color = "Omnivory",
+       fill = "Omnivory") +
+  MetBrewer::scale_color_met_d("Hiroshige") +
+  MetBrewer::scale_fill_met_d("Hiroshige")
+
 
 ggsave(g_pb,
        filename = here::here("output/figure_pb_main.pdf"),

### 2022-03-24 code/figure_network.R

@@ -46,8 +46,9 @@ ng <- foreach(i = seq_len(length(net))) %do% {
                     color = grey(0.5),
                     stroke = 0.1,
                     size = 2) +
-    scale_fill_continuous(high = "steelblue",
-                          low = grey(0.95)) +
+    MetBrewer::scale_fill_met_c("Hiroshige") +
+    #scale_fill_continuous(high = "steelblue",
+    #                      low = grey(0.95)) +
     guides(fill = ifelse(i == 3, guide_legend(), "none")) +
     labs(subtitle = paste("Branching prob. =",
                           para$p_branch[i]),
@@ -78,5 +79,5 @@ g_disturb <- bind_rows(mutate(net[[1]]$df_patch, p_branch = 0.2),
              linetype = "dashed") +
   labs(x = "Disturbance intensity",
        y = "Density",
-       fill = "Branching prob.") +
-  guides(color = "none")
+       fill = "Branching prob.",
+       color = "Branching prob.")

### 2022-03-24 code/figure_si_fcl.R

@@ -65,9 +65,8 @@ list_g_np <- foreach(i = seq_len(nrow(df_param))) %do% {
            sd_disturb_source == df_param$sd_disturb_source[i]) %>% 
     ggplot(aes(x = n_patch,
                y = value,
-               linetype = omn,
-               color = y,
-               fill = y)) +
+               color = omn,
+               fill = omn)) +
     geom_smooth(method = "loess",
                 size = 0.5) +
     facet_grid(rows = vars(y),
@@ -77,8 +76,8 @@ list_g_np <- foreach(i = seq_len(nrow(df_param))) %do% {
     labs(x = "Ecosystem size (number of patches)",
          y = "Value",
          linetype = "Omnivory") +
-    guides(color = "none",
-           fill = "none")
+    MetBrewer::scale_color_met_d("Hiroshige") +
+    MetBrewer::scale_fill_met_d("Hiroshige")
   
   return(g_np)
 }
@@ -92,9 +91,8 @@ list_g_pb <- foreach(i = seq_len(nrow(df_param))) %do% {
            sd_disturb_source == df_param$sd_disturb_source[i]) %>% 
     ggplot(aes(x = p_branch,
                y = value,
-               linetype = omn,
-               color = y,
-               fill = y)) +
+               color = omn,
+               fill = omn)) +
     geom_smooth(method = "loess",
                 size = 0.5) +
     facet_grid(rows = vars(y),
@@ -104,8 +102,8 @@ list_g_pb <- foreach(i = seq_len(nrow(df_param))) %do% {
     labs(x = "Ecosystem complexity (branching prob.)",
          y = "Value",
          linetype = "Omnivory") +
-    guides(color = "none",
-           fill = "none")
+    MetBrewer::scale_color_met_d("Hiroshige") +
+    MetBrewer::scale_fill_met_d("Hiroshige")
   
   return(g_pb)
 }

### 2022-03-24 rmd/suppl_info.Rmd
Update suppl_info.Rmd

@@
   pdf_document:
+    toc: true
     latex_engine: xelatex
@@
 
+# Sensitivity data
+load(file = here::here("output/result_stvy.RData"))
+sim_sv <- sim_stvy_result
+n_sim_stvy <- unique(sim_sv$n_warmup) + unique(sim_sv$n_burnin) + unique(sim_sv$n_timestep)
+
 ```
@@
 ```
+# Supplementary text
+
+## Sensitivity analysis
+
+We performed a sensitivity analysis of the trophic community simulation to identify key simulation parameters that strongly affect the relationships between food chain length and ecosystem properties (the number of habitat patches $N_{p}$ and branching probability $P_{b}$.
+We generated 500 sets of parameter combinations by randomly drawing values of 16 simulation parameters from uniform distributions (**Table S1**).
+For each parameter combination, we generated `r max(sim_stvy_result$n_rep)` branching networks with the gradients of ecosystem size ($N_p \sim Unif(10, 150)$) and complexity ($P_b \sim Unif(0.01, 0.99)$).
+This results in a total of `r nrow(sim_stvy_result)` simulation replicates.
+In each simulation replicate, we ran `r n_sim_stvy` time steps of trophic community dynamics.
+We obtained temporal means of food chain length for the last `r unique(sim_sv$n_timestep)` time steps.
+The first `r n_sim_stvy - unique(sim_sv$n_timestep)` time steps were discarded as initialization and burn-in periods.
+
+For each parameter combination, we estimated the Spearman's rank correlation between food chain length and ecosystem properties ($\psi_1$ for $N_{p}$ and $\psi_2$ for $P_{b}$).
+To examine influences of simulation parameters (**Table S1**), we developed the following regression model taking $\psi_1$ or $\psi_2$ as a response variable $u_n$ (parameter combination $n$):
+
+```{=tex}
+\begin{align}
+  u_{n} &\sim Normal(\mu_n, \sigma^2)\\
+  \mu_n &= \zeta_0 + \sum_{k}{\zeta_k x_{k, n}}
+\end{align}
+```
+where $\zeta_k~(k = 0...14)$ are the intercept and regression coefficients, and $x_{k,n}$ are the simulation parameters drawn from uniform distributions (**Table S1**).
+Explanatory variables were standardized to a mean of zero and a standard deviation of one, so that regression coefficients are comparable.
+
+The sensitivity analysis revealed key simulation parameters.
+For ecosystem size, dispersal distance $\theta$, local disturbance variation $\sigma_{l}$, and the reproductive rate of basal species $r_b$ were influential.
+However, parameters that influence food web structure were less influential (attack rate, conversion efficiency, and handling time).
+For ecosystem complexity, dispersal distance $\theta$ and probability $p_d$, local disturbance variation $\sigma_{l}$, and the reproductive rate of basal species $r_b$ were influential.
+Food web parameters were also important (attack rate, conversion efficiency, and handling time).
+
+Based on the results, we considered ecological scenarios relevant for putative controls of food chain: low and high productivity ($r_b = 4, 8$), low and high disturbance ($p_b = 0.01, 0.1$), and three food web architecture ("Chain" [$\{e_{bc}, e_{bp}, e_{cp}\} = \{4, 0, 4\}, \{a_{bc}, a_{bp}, a_{cp}\} = \{0.5, 0.1, 0.5\}$], "Weak omnivory" [$\{e_{bc}, e_{bp}, e_{cp}\} = \{4, 2, 4\}, \{a_{bc}, a_{bp}, a_{cp}\} = \{0.5, 0.1, 0.5\}$], and "Strong omnivory" [$\{e_{bc}, e_{bp}, e_{cp}\} = \{4, 4, 2\}, \{a_{bc}, a_{bp}, a_{cp}\} = \{0.5, 0.5, 0.1\}$]).
+Handling times were fixed at $0.5$ to reduce the number of parameter combinations.
+
+\pagebreak
+
+# Tables
+
 ## Table S1 Parameter values
@@
 
+# Figures
+
 ```{r si_figure_fcl}
@@
       figure_legend <- paste0(str_to_sentence(c_text[i])," effect on food chain length and occupancies of constituent species.
-Lines and shades represent loess curves fitted to the simulated data and 95% confidence intervals.
+Rows are different response variables, and columns are different simulation scenarios.
+Lines and shades represent loess curves fitted to the simulated data and associated 95% confidence intervals.
 In this simulation, the following values of parameters were used: $\\theta$ = ", df_param$theta[k], ", $\\sigma_l$ = ", df_param$sd_disturb_lon[k], ", and $\\sigma_h$ = ", df_param$sd_disturb_source[k], ".")

### 2022-03-24 code/library.R
edit: misc

@@ -12,4 +12,5 @@ pacman::p_load(mcbrnet,
                NetIndices,
                patchwork,
                igraph,
-               tidyverse)
+               tidyverse,
+               MetBrewer)

### 2022-03-24 rmd/manuscript.Rmd
Create manuscript.Rmd

---
Title: "Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks"
output:
  pdf_document:
    latex_engine: xelatex
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
---

**Author:** Justin P.F. Pomeranz^1,2,\*^, Jacques C. Finlay^3^, Akira Terui^1,\*^

^1^ Department of Biology, University of North Carolina at Greensboro; ^2^ Department of Environmental Sciences, Colorado Mesa University; ^3^ Department of Ecology, Evolution and Behavior, University of Minnesota

**\*Corresponding Author:** JPF Pomeranz, jfpomeranz\@gmail.com; A Terui, hanabi0111\@gmail.com

# Abstract

Understanding the drivers of food chain length in natural communities has intrigued ecologists since the publication of 'food cycles' by Elton in the early 20th century. Proposed drivers of food chain length have included extrinsic controls such as productivity, disturbance regime, and ecosystem size, as well as intrinsic factors including trophic omnivory. However, current theories have largely assumed simple, two-dimensional habitat architectures, and may not be adequate to predict food chain length in ecosystems which have a complex, branching structure. Here, we develop a spatially explicit theoretical model which provides an integrated framework for understanding variation in food chain length in branching networks. We show independent, positive influences of both ecosystem size and complexity on food chain length. However, these effects were contingent upon other proposed extrinsic and intrinsic drivers such as disturbance regime, productivity, and trophic omnivory. Our results show that ecosystem complexity is an important driver of food chain length that may explain enigmatic variation can reconcile inconsistent results from empirical studies of food chain length in riverbranching ecosystems.

**Keywords:** Food chain length, omnivory, branching networks, fractals, intraguild predation, complex network architecture

# Introduction

The origin of variation in food chain length (FCL), the number of feeding links from the basal species to top predators, has intrigued ecologists for decades due to its importance in regulating top-down effects 1, energy flow2, and contaminant concentrations in top predators that humans often consume3. Hypotheses regarding controls of food chain length can be organized into three broad categories. First, the resource availability hypothesis states that basal productivity controls food chain length because imperfect energy conversion through predator-prey interactions restricts energy available to higher trophic levels4. Second, the disturbance hypothesis (or the dynamical stability hypothesis) predicts longer food chains are vulnerable to environmental perturbations (i.e., the dynamical constraint). Thus, habitats with frequent and/or severe disturbance should support fewer trophic levels5. Lastly, the ecosystem size hypothesis posits that larger ecosystems should harbor longer food chains through a suite of mechanisms, including enhanced colonization rates6 increased habitat heterogeneity7 and surpassing spatial constraints inhibiting metapopulation dynamics.

Among the three hypotheses, the ecosystem size hypothesis has been consistently supported in both theoretical and empirical explorations, likely because it encompasses multiple mechanisms which underlie the positive association8,9. For example, large ecosystems have increased species richness (species-area relationship8), buffer against disturbances9, and have greater numbers of immigrant sources in the ecosystem, harboring stable metapopulations of constituent species that make up long food chains 6,10,11. Meanwhile, the theoretical basis for the other two hypotheses rest on implicit assumptions of linear food chains, which are often violated in natural systems. As such, the effects of disturbance and resource availability on food chain length have been found to be highly variable among ecosystem types and geographic regions 12--14.

More recently, food web ecologists have begun to recognize that FCL rarely responds to a single driver, but instead is determined by interactions of multiple factors. For example, Ward and McCann15 showed that FCL was more responsive to increasing productivity in large ecosystems compared with small ecosystems, whereas FCL was more sensitive to size in low productivity systems . Such context dependent  responses to extrinsic factors seem to be tightly coupled with the arrangement of trophic links within a food web, often referred to as food web motifs16,17. For example, omnivory (feeding at multiple trophic levels) is a strong intrinsic driver of food chain length 6,15,18,19. When omnivory is absent in three-species communities, the trophic structure assumes a simple linear food chain with maximum length equal to 3 (**Figure 1**). However, as omnivory increases, FCL decreases proportionally as the top predator eats lower in the food web 15,20. Theoretical work predicts that food webs are destabilized when omnivory is prevalent, leading to the loss of species and shorter food chains even in productive, large ecosystems. Conversely, weak omnivory stabilize the food web and increase persistence of species within a community, particularly at intermediate levels of productivity and high disturbance levels 20,21. Omnivory is a dominant module in natural communities 22 but is often not accounted for in theoretical studies20, but see 15. Due to the context-dependency of FCL to multiple drivers, and the potential for interactive effects, a unified framework of multiple food chain drivers is needed to better understand when and where FCL will respond to these extrinsic and intrinsic factors.

While recent research  efforts of food web ecologists have substantially advanced our understanding of food chain drivers, most of the previous work has been conducted in simple, two-dimensional systems, such as lakes7,23,24 and oceanic islands 25,26. However, many natural habitats have a high degree of spatial complexity that cannot be represented simply by ecosystem size. Branching ecosystems, for example, are ubiquitous yet overlooked landscape structures 27, in which geomorphological or biological processes form naturally fractal branching patterns, where the geometric arrangement of patches is similar across spatial scales28 (e.g., rivers, trees, caves, mountain ranges). While we know that food web interactions are spatially influenced29,30, the theory to account for it has lagged, particularly in branching habitat architectures.  Here, we theoretically explore drivers of food chain length in rivers, an excellent example of branching ecosystems.

Ample evidence suggests that branching structure controls environmental heterogeneity and dispersal pathways of organisms within river ecosystems. For example, water and materials propagate downstream as small streams hierarchically join to form larger streams and rivers. Consequently, environmental signals are highly variable among tributaries while showing a strong within-tributary correlation e.g., carrying capacity, flood disturbance31,32. Confluences serve as merging points where branch-specific dynamics aggregate to form distinct environmental conditions in downstream habitat patches 33. Habitat heterogeneity therefore increases with branching complexity (**Figure 1**), which we define here as the probability of branching per unit river distance. Hence, highly-branched river networks provide diverse habitats that buffer the impact of large-scale environmental fluctuations 34, broadening opportunities for species to recolonize habitat patches post-disturbance. Thus, ecosystem complexity in river networks may be a widespread yet overlooked extrinsic environmental driver which needs to be incorporated to understand food chain length.

We develop a model in which trophic community dynamics within branching habitats are simulated. We investigate extrinsic controls such as ecosystem size (number of patches), complexity (branching probability), disturbance regime and productivity level (carrying capacity). In addition, our simulation model incorporates intrinsic drivers of food chain length including food web motifs (degree of omnivory) and dispersal ability, allowing us to investigate the context dependency on drivers of food chain length across extrinsic factors. We find that ecosystem size and complexity are both positively related to FCL, but the form of this relationship is dependent on other extrinsic and intrinsic factors. Trophic structure (degree of omnivory) and disturbance regime are both strong drivers of FCL on their own and have interactive effects. We also find that productivity and dispersal ability were both positively related to FCL. Our model is an important step in unifying drivers of FCL across extrinsic and intrinsic drivers and parsing their context-dependency in spatially complex ecosystems.

# Methods

## **Network Generation**

We generatedsimulated branching networks of connected habitat patches as a series of nodes (Figure 1). The geometric arrangement of networks was controlled with two parameters: ecosystem size, the number of patches (NpatchNP); and ecosystem complexity, branching probability (PbB). Nodes (habitat patches) were assigned to be a confluence (or upstream terminal node) with probability PbB, or an in-branch node with probability 1-Pb 1 - PB. Branches (tributaries) are a series of connected nodes between confluences (or upstream terminal nodes). The number of patches within a branch,np, q q, is a random variable drawn from a geometric distribution np,q Ge(Pb)np,q∼Ge(Pb)  (a geometric distribution is the discrete form of an exponential distribution) but conditional on q=1Nbnp,q=Np (Nb is the number of branches). The number of branches NbB Under this assumption, the number of branches is identical to the number of branching nodes or the number of "success" out of Npatch trials. This translates into the assumption that the variable Nb follows a Binomial distribution asis a random variable drawn from a Binomial distribution: NbBinomial(Npatch, Pb).NbB \~ Binomial(NP, PbB). Generated river networks preserve important geometric scaling relationships (i.e., Horton's law). This framework preserves the fractal nature of branching patterns. For each scenario described below, we generated 1000 networks across a gradient of ecosystem size and complexity by drawing NP and Pba random variable from uniform distributions as: NpatchUnif(5, 150) NP \~ Unif(5, 150) and PbUnif(0.01, 0.99)PbB \~ Unif(0.01, 0.99) (Np was rounded to the nearest integer before running simulations).

For each terminal branch (i.e., headwater branch), h, within the network, a disturbance value, mh, was sampled .  The distribution of mh values was sampled from a normal distribution in a logit scale. This was accomplished by transforming the proportional mean disturbance value, µm, to the logit scale and sampling a value for each upstream terminal node (headwater) as logit mh Normal(logit m, m2).mh \~ Normal( logit( µm), σ2m). Here, we varied the mean source disturbance value, m = (0.1, 0.9) µm = (0.1, 0.9), based on disturbance regime (see **Simulations**) and set the variation of source disturbance as mσ2m = 2. We added lLocal environmental noise (i.e., noise occuring at each node) towas included in the disturbance value of each node as a spatial autoregressive process: logit mdown = logit mup+. mdown and mup are the disturbance values at down- and upstream nodes, and the local environmental noise at the downstream node that is drawn from a normal distribution as Normal(0, 2). This autoregressive process resembles natural disturbance regime, where flood or drought disturbance cascades downstrem with some local variation along the channel.

s moving downstream by sampling a random variable from Normal(0, σ2l). For the main simulation results presented here, σ2l = 0.01. At confluence nodes, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, s (the number of upstream contributing nodespatches): logit mdown= logit m1,up+(1-) logit m2,upmdown = ωm1,up + (1 - ω)m2,up, where mdownmdown is the disturbance magnitude of the focal patch (the confluence), m1,upm1, up and m2,upm2, up are the disturbance magnitude of upstream adjacentcontributing nodestributaries, and ω the relative size of one tributary over another defined as =s1/(s1+s2). = s1/(s1 + s2). Disturbance value distributions had medians close to µm with long tails (**Fig S1**). with long right or left tails (**Fig S1**), respectively, i.e. For example, when m=0.1µm = 0.1 (e.g., a 10% reduction in species densities), most sites experienced low-magnitude disturbances and a few sites experienced high-magnitude disturbances. Likewise, whenm=0.9,  µm = 0.9, most sites experienced high-magnitude disturbances with a few patches receiving low-magnitude disturbances. The low-magnitude patches can be viewed as refugia and a source to supply individuals to colonize impacted sites in subsequent timesteps. Example disturbance value distributions for simulation scenarios reported here are available in the SI.

Carrying capacity in patch x , Kx ,was assumed to increase as a power function of ecosystem area (i.e., watershed area)58:

Kx= Kbase+C\*Npatch, up

Where Kbase=100, C is a constant, Npatch, up number of patches upstream, and a random variable sampled from Unif(1.1, 1.2). At upstream terminal nodes (headwaters) K=Kbase. The R package mcbrnet was used to simulate the networks (Citation).

## **Tri-trophic dynamic simulation model**

We simulated trophic community dynamics in branching river networks using a discretemixed-time model which models predation continuously and has discrete, synchronous reproduction. This is a life-history common to many lotic species, including macroinvertebrates and fish. Accounting for this stage structure is an important aspect which is often missed in continuous-time models57. We describe the model in detail below.

Our simulation model consists of three species communities; a basal species B which has positive logistic growth up to a carrying capacity, K; a primary consumer species, C which grows based on the consumption of B; and a predator P, which can consume both B and C. In our model, the realized species abundance Nix(t+1) for species i (either B, C, or P) at patch x and time t is expressed as a composite of within-patch trophic interactions, immigration, and emigration:

Nix(t+1)∼ Poisson(max{nix(t)+Iix(t)-Eix(t), 0})

Carrying capacity, KK, in patch xx , Kx ,was assumed to increase basen onexponentially as a power function of ecosystem area (i.e., watershed area)58:

Kx= Kbase+C\*N

Where KBase = 100, C is a constant, N = number of patches upstream, and ψ was a random variable sampled from Unif(1.1, 1.2). At upstream terminal nodes (headwaters) K = KBase.

where nix(t) is the expected species abudance after within-patch trophic interactions, Iix(t) the expected number of immigrants into patch x, and Eix(t) the expected number of emigrants from patch x. The realized discrete number of individuals is drawn from a Poisson distribution to account for stochasticity in demographic and dispersal processes.Let Nix(t) Nix(t) represent the population density of species i in patch x at time t. The model starts with some initial nonzero population densities for all three species in each patch. The initial population abundances for each species i in each patch x were independently sampled from a random Poisson distribution Poisson(ix)Poisson(𝝺ix), where ix𝝺ix for B, C and P was 0.8Kx \* K, 0.5Kx \*K and 0.25Kx \* K, respectively.  First, dispersal from all patches is accounted for. The number of emigrants, Eix(t) for species i in from patch x at time t, is estimated as the product of dispersal probability, Pdispersal and species abundance Nix(t). The dispersal probability was assumed to be constant for all three species in each simulation run. Then, the number of emigrants from patch x is divided into immigrants, Iix(t) Ixy(t) to each patch y. The immigration probability at patch x and time t, ix(t) xy ξxy from patch x to y is calculated considering network structure and species dispersal capability as followsbased on the habitat structure among patches according to:

xy(t) =y,y≠xExy(t)e-θdxyxy,y≠xExy(t)e-θdxy

ix(t) = y,yxEiy(t)e-dxyxy,yxEiy(t)e-dxy

Where dxy is the distance between patch x and y, and θ controls the rate of decay. Then, the expected number of immigrants into patch x is calculated as Iix(t)=ix(t)xEix.

The realized number of individuals Nix(t+1) (species i at patch x and time t+1) is given as:

Nix(t+1)∼ Poisson(nix(t)+Iix(t)-Eix(t))

where Nix(t) is the number of individuals at time t,(i.e., before dispersal) Iix(t) the expected number of immigrants to patch x, and Eix(t) the expected number of emigrants from patch x. The realized discrete number of individuals is drawn from a Poisson distribution to account for stochasticity in demographic and dispersal processes. Note that a value of 0 is assigned to Nix(t + 1) for any non-positive value of 𝝺 59.

After dispersal, trophic interactions occur.continuous predation occurs. We model the tri-trophic interactions with intraguild predation as follows:

nBx(t+1)=(rmax1+υNBx(t)')(1-mx)				Basal species

nCx(t+1)=WBC, xNCx(t)'(1-mx)			Intraguild prey

nPx(t+1)=(eBPWBP, x+eCPWCP, x)NPx(t)'(1-mx)	Intraguild predator

The above equations indicate that the basal species exhibit a  where rmax is the maximum population growth of basal species, υ the term controlling density-dependence (rmax-1Kx), and mx the proportional mortality when disturbance occurs (disturbance occurs when =1, otherwise 0).

\

The number of prey i eaten by consumer j, Wij, in a season is calculated as:

\
Wij= ijNiNjijNj+Ni

where Ni and Nj are resource and consumer densities, respectively and αij and βij are coefficients describing the functional response as the ratio of resources, i, to consumers, j, increases. βij describes how fast the predation curve reaches the asymptote of the Type II functional response, and αij/βij determines the location of the asymptote57.

The predator species P, searches for and consumes both species in a patch. Its total effort is controlled by parameters α.p and β.p. The preference parameter δ, controls the proportional search effort of species P for resource B and C:

WBP=δ(Wij)

WCP=(1 - δ)(Wij)

Predator preference for resource B is determined by the equation:

δ= eBPNBeBPNB+ eCPNC

Where eBP and eCP are the conversion efficiencies of predator P for resource B and C, and NB and NC are resource densities, respectively. This formulation allows for flexible diet shifts of the predator based on realized energy gain (conversion efficiency times prey availability) from the prey species60.

At each time-step within a patch, FCL was defined as 0 if no species were present, 1 if only B was present and 2 if only B + C or B + P were present. When all three species were present, FCL was calculated based on the proportion of each resource in P's diet (𝜆i) according to:

FCL = (1 \* B+ 2 \*C) +1

When P only consumes B or C, FCL = 2 or 3, respectively. When P consumes both resources equally, FCL = 2.5. In each simulation, FCL was averaged across all patches and time-steps.

The number of individuals in each patch is reduced by the number consumed Wij (predation effect), as well as by a base survival probability, s0:

NB'=s0(NB- WBC -  WBP)

NC'=s0(NC'-WCP)

P'=s0P

Finally, those individuals remaining reproduce according to the following:

NB, t+1=(rmax1 + υ NB' NB' )(1-φmx)

NC, t+1=(WBCNC')(1-φmx)

NP, t+1=(eBPWBP+ eCPWCP)NP'(1-φmx)

where rmax is the maximum per-capita reproduction rate of the basal species and υ = rmax-1K determines the asymptotic relationship of population growth, where K is carrying capacity. In the absence of predation, population growth of the basal species follows a logistic growth relationship. For the consumer species (C and P) eij is the efficiency of converting resource i into new consumer j. The last term in the equation accounts for the proportional mortality in the event of a disturbance, where mx = the disturbance value (see **Network Generation**) and 𝜑 is an indicator variable of disturbance:

φ =0, no disturbance

φ =1, disturbance occurs

If species B is extirpated from a patch, both the other species were also assumed to be extirpated from that patch during that time step. Species B can colonize any open patch, whereas species C and P can only successfully colonize patches where B was already present (i.e., prey resource is necessary for successful establishment).

## **Simulations**

We used 24 combinations of parameter sets (**SI Table**). The parameter sets include two environmental factors, and twelve ecological scenarios. For environmental scenarios, we used two parameter sets that controlled the frequency and magnitude of disturbances (low and high disturbance regime, respectively). In the low disturbance regime, the probability of a disturbance occurring at each time step was 0.0001 (i.e., a 10,000 year recurrence), and the mean disturbance magnitude was μm = 0.1 (10% reduction in population densities). In contrast, the high disturbance regime had a probability of 0.01 (i.e., 100 year recurrence), and the mean disturbance magnitude was μm = 0.9 (90% reduction in population densities61)

For the ecological scenarios, we used two productivity levels, two dispersal abilities, and three levels of omnivory. Productivity was varied by setting the carrying capacity of the basal species, B (K = 50, 150). Dispersal ability varies the effective distance between patches and can control the probability that a patch is re-colonized post-disturbance. Two levels of dispersal ability were simulated. The high dispersal ability was set with parameters Pdispersal = 0.1 and θ = 0.1 (i.e., high probability of individuals emigrating, and far distance traveled). The low dispersal scenario was set with parameters Pdispersal = 0.01 and θ = 1  (i.e., low probability of dispersal, short dispersal distance). These values were chosen to represent extreme scenarios. Preliminary analyses using intermediate parameter sets had responses between these two scenarios.

Finally, we investigated three food web motifs (**Figure 1B**). Trophic interaction strengths can be controlled by varying the attack rate or conversion efficiency by the consumer P (Ward and McCann 2017). To simplify model simulations, we fixed the attack rate for both C and P (parameters 𝛼 = 8 and 𝛽 = 20), and the conversion efficiency for C as ebc = 6. We controlled the presence and strength of trophic interactions for the predator, P, by varying the conversion efficiency for both prey resources (i.e., varied prey resource quality). The scenario names and parameter levels were as follows: (1) "Food chain": ebp = 0, ecp = 4; (2) "Strong IGP": ebp = 2, ecp = 4; (3) "Weak IGP": ebp = 4, ecp = 2.

Under each scenario, we simulate 1200 timesteps of trophic dynamics (including 200 steps of burn-in) in 1000 branching networks across a gradient of ecosystem size (number of habitat patches, from 5 to 150) and ecosystem complexity (branching probability, from 0.01 to 0.99). Simulations ran for 200 time steps as a burn-in period to minimize influences of initial conditions. The last 1000 timesteps were recorded to estimate the temporal average of food chain length across all patches within a network.

Values of simulation parameters are summarized in (SI). We performed simulations using the mcbrnet and IGPtoy packages (blinded citation, in preparation) in R version 4.0.3. R Code to reproduce simulations, figures, and tables are available at <https://github.com/Jpomz/priv-igp-sens>.

\
The model starts with some initial nonzero population densities for all three species in each patch. The initial population abundances for each species i in each patch x were independently sampled from a random Poisson distribution Poisson(ix)Poisson(𝝺ix), where ix𝝺ix for B, C and P was 0.8Kx \* K, 0.5Kx \*K and 0.25Kx \* K, respectively.

# Results

First, we depicted branching ecosystems as connected habitat patches, which local communities inhabit. Each local community is composed of basal (B), primary consumer (C), and top predator (P), and these constituent species disperse along the network corridors following exponential dispersal kernels. The arrangement and strength of trophic interactions can be varied based on the simulation parameters (food web motifs, **Figure 1**, see **Methods**). Periodically, local communities are impacted by disturbance, whose strength is highly correlated within a tributary. Therefore, our simulation framework resembles environmental dynamics of natural river systems.

We explored how food web motifs affected FCL across extrinsic factors. FCL is related to the degree of omnivory of the predator P (**Figure 1**). When P only consumes C, its trophic level is 3. However, as the proportion of B in P's diet increases, FCL decreases, reaching a minimum of 2 when B constitutes all of P's diet. We assume that P searches for both resources proportionally based on the conversion efficiency for each prey resource (i.e., prey quality) and the relative densities of each resource (see **Methods**). To control the level of omnivory, we varied the conversion efficiencies of both prey resources.

In addition to food web motifs, we also investigated the effects of species dispersal ability and disturbance regime. To simplify our simulations, we looked at two scenarios of dispersal ability. 1) Low dispersal ability, in which a small proportion of the population of all three species emigrate, and the likelihood of successful immigration to neighboring patches declines exponentially with increasing distance (i.e., short-distance dispersal). 2) High dispersal ability, in which a larger proportion of the population emigrates, and the likelihood of successful immigration to neighboring patches declines less-rapidly with increasing distance (i.e., long-distance dispersal; see **Methods**).

The effect of disturbance was also investigated by looking at low and high disturbance regimes. The low disturbance regime was defined as having a low probability of a relatively weak magnitude disturbance occurring throughout the habitat network in each time step. This can be likened to spring-fed streams which are relatively stable through time but are subjected to occasional disturbances such as floods from high precipitation events. In contrast, we investigated a high disturbance regime in which the probability and magnitude of disturbances occurring throughout the habitat network was much higher. This scenario could represent "flashy" streams which are predominantly driven by surface runoff and precipitation events.

We examined interactive effects of extrinsic and intrinsic factors controlling food chain length. We simulated habitat networks over a gradient of both ecosystem size and complexity (branching probability). For each habitat network, we simulated trophic dynamics under two different regimes of disturbance (rare or frequent), productivity (low or high carrying capacity of the basal species), and dispersal ability (low and high), and three local food web structures (the degree of omnivory, **Figure 1**), for a total of 24 scenario combinations. We simulated meta-food web dynamics over 1200 time steps, including 200 steps for burn in, and yielded estimates of food chain length averaged in space and time (see **Methods**). Varying the disturbance regime resulted in markedly different results across ecosystem size and complexity. Thus, we present our results for each disturbance regime separately below.

## **Low disturbance regime**

In ecosystem networks with a low disturbance regime (i.e., dynamical stability), FCL was largely determined by food web motifs (colored lines in **Figure 2**). Generally, FCL was negatively related to the degree of omnivory, with simple linear food chains (P only consumes C) having the longest FCL, and strong omnivory (P consumes very little C) having the shortest FCL. When carrying capacity and dispersal ability was low, there was a slight increase in FCL with the omnivorous motifs across increasing ecosystem size, and when dispersal ability was high, there was a moderate increase for the strong omnivory module (**Figure 2A**). In addition, there was a slight increase in FCL with high dispersal ability across both carrying capacities, but the effect was greater at low carrying capacity. However, when carrying capacity was increased, FCL was largely invariant across gradients of ecosystem size.

There was very little effect of ecosystem complexity on FCL in the low disturbance regime. In all scenarios, FCL did not vary in response to increasing ecosystem complexity, although there was a very slight decline for all three food web modules when carrying capacity and dispersal ability were low. FCL was predominantly determined by food web motifs and carrying capacity, with FCL being negatively related to omnivory, and positively related to carrying capacity. Like the results across ecosystem size, there was a positive effect of high dispersal ability on FCL, and this effect was greater in ecosystems with low carrying capacity (interactive effect).

## **High disturbance regime**

Ecosystem networks with a high disturbance regime had much different responses of FCL across both ecosystem size and complexity. FCL increased across both gradients under all combinations of carrying capacity and dispersal ability, and in most food web motifs (**Figure 3**). However, the strength of the relationship was variable for some motifs, and when dispersal ability and carrying capacity were low, and omnivory was absent, FCL was largely invariant across both gradients (top right panel in **Figure 3A and 3B**. Once again, FCL was largely determined by the degree of omnivory, with simple linear food chains generally having longer FCL, and weak IGP having the shortest FCL.

When dispersal ability and carrying capacity were low, however, a simple linear food chain structure resulted in the shortest FCL across both ecosystem size and complexity. It is likely that this simple motif with a few strong links was unstable and could not persist when exposed to frequent, high magnitude disturbances (e.g., supports dynamical stability hypothesis). In these environments, having prey choice (omnivory) appears to be a necessary requirement to maintain the persistence of species.

Increasing dispersal ability and carrying capacity had positive effects on FCL across both ecosystem size and complexity. Similar to the results in the low disturbance regime, increasing dispersal ability in ecosystems with low carrying capacity resulted in a large increase in FCL, suggesting an interactive effect.

# Discussion

The ecosystem size hypothesis of food chain length has received support from both theoretical and empirical work. Because ecosystem size encompasses several distinct mechanisms that could influence food chain structure, work so far has not elaborated the underlying drivers. Ecosystem complexity is an extrinsic factor which is scale-invariant and related to ecosystem size yet has not been incorporated in theoretical explorations of food chain length to date. Here, we show that food chain length responds positively and independently to increasing ecosystem size and complexity. This supports our hypothesis that scale-dependent theories of FCL are inadequate to correctly predict FCL in ecosystems with a complex structural geometry.  However, the relationship between FCL and ecosystem complexity varied based on additional extrinsic and intrinsic factors, supporting the emerging paradigm that there is no one single driver of FCL in ecosystems 15, but moreover that there is context dependency for which factor is the primary driver of the relationship. Although we focus on river networks here, we expect our results to be transferable to other branching structures, including mountain ranges and cave systems.

While increased disturbance intensity and frequency consistently reduced food chain length as predicted by dynamical stability hypothesis 5, more important findings in this study were evident from their synergistic effects with ecosystem structure. Our low disturbance regime scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independent of ecosystem size and complexity and FCL was predominantly determined by food web motifs. However, under high disturbance regimes, only large and complex networks could provide diverse and heterogeneous habitats with potential refugia, allowing species to thrive and maintain long food chains. This result has important implications for ecosystem management because anthropogenic climate change is predicted to increase the frequency and magnitude of disturbances 35,36. River networks are especially vulnerable due to their propensity to go dry during droughts37 as well as susceptibility to more intense flood disturbance with increasing storm intensity38. Increased dispersal ability alleviated the reduction in FCL from increased disturbance regime in our simulation, and reduced the negative impacts of disturbance and climate change in empirical studies 39. However, human activities have altered the habitat heterogeneity, connectivity, and dispersal ability of species within riverine networks 40,41. In-stream habitat in rivers and streams have been homogenized and lateral connections to the floodplain have been severed by the removal of large woody debris, channelization and straightening of river corridors, and through flow regulation 42,43, and this has affected community assembly44 and trophic structure 45.

Although dispersal has received little attention in food chain research, our simulation highlighted its importance in regulating food chain length. In particular, high dispersal was required for ecosystem size and complexity to have consistent positive effects on FCL under high disturbance regimes. This is probably because sufficient propagules are needed to successfully recolonize after disturbance and maintain long food chains in large and/or complex networks. Human activities that affect habitat connectivity and dispersal ability of species within riverine networks may compromise the buffering effects of ecosystem size and complexity against disturbance by limiting dispersing propagules. For example, dams and diversion have altered the connectivity between suitable habitat patches, or have completely blocked dispersal ability 46,47. Light pollution has also negatively affected the ability of adult aquatic insects to disperse between habitat patches 48.

The context dependency of disturbance, productivity and dispersal has important implications for the conservation and management of freshwater ecosystems. This adds an important distinction in the discussion of the suitability of assigning protection or conservation status to regions. Current conservation schemes generally emphasize size, aiming to protect the largest areas possible. While this is certainly a useful attribute generally, designating regions with a high degree of variability, or many small patches, may be more beneficial than a large, homogenous area 49, and accounting for ecosystem complexity is particularly suited to ecosystem architectures in which branches have variable environmental conditions. Conservation and management activities should focus on activities that increase the habitat heterogeneity and connectivity of suitable habitat patches to increase resilience to predicted increases in disturbance frequency and magnitude.

The role of food web motifs on FCL largely followed our predictions (**Figure 1**). Specifically, FCL declined as the value of the resource B, and thus the proportion in the diet, to the predator P increased. Essentially, the relative qualitative values of the resources to the predator controlled the level of omnivory and FCL. Although FCL was greatest when the community was arranged as a simple linear food chain, this arrangement was not the most stable. Specifically, when disturbance regime was high, and productivity and dispersal ability were low, the observed FCL across networks was \~1 (although increasing dispersal ability and/or carrying capacity resulted in food chains of \~2.5-3). Therefore, in the absence of prey resource choice, strong, top-down trophic interactions were destabilizing, leading to the extirpation of consumer species at higher trophic levels across the networks.

Changes in community composition are occurring worldwide and in all habitats. Species extirpations and extinctions, coupled with species introductions and invasions are creating novel communities worldwide. In Lake Tahoe (USA), for example, replacement of the top fish predator and the introduction of a freshwater shrimp substantially altered the food web structure 50. Understanding the interactions within these novel communities, and how this could affect food chain lengths is an important goal of contemporary ecologists. The strength of omnivory in these novel communities can be increased or decreased, depending on the introduced species and the extant species pool or community. For example, non-indigenous fishes in the Eastern Mediterranean Sea outcompete native species for high quality prey items, causing native species to increase consumption of non-preferred prey items51. Introductions of non-native fishes in the Colorado River basin have led to longer FCL and narrower niches (e.g., lower degree of omnivory) of native species52. Alternatively, the introduction of hyper-successful basal or primary consumer species can weaken or remove the omnivory link by becoming the dominant prey resource. For example, nonnative New Zealand mud snail dominated the macroinvertebrate biomass in streams in Yellowstone National Park and likely affected food web structure53.  Having a single prey resource dominate the community is akin to reducing the degree of omnivory (i.e., removing prey choice), which destabilizes the community, particularly under high disturbance regimes.

It is abundantly clear that ecosystem complexity in natural systems is the rule, not the exception. Early theoretical and empirical work has assumed simple habitat structure as a necessity. However, habitat heterogeneity, such as the presence of prey refugia, and dispersal between connected habitat patches allow communities to stabilize and persist through space and time 54--56. Complex ecosystem structure is an extrinsic factor that can affect food chain length in communities. Incorporating a realistic understanding of ecosystem structural complexity could alter our perception of communities in naturally complex systems, potentially leading to paradigm shifts in macroecology.

\

### 2022-03-24 code/figure_concept.R
edit: figure

@@ -15,8 +15,8 @@ layout <- "
 AAAB
 CCCD
 "
-g_concept <- wrap_plots(g_web) + g_delta + ggtitle("B") +
-  wrap_plots(g_net) + g_disturb + ggtitle("D") +
+g_concept <-  wrap_plots(g_net) + g_disturb + ggtitle("B") +
+  wrap_plots(g_web) + g_delta + ggtitle("D") +
   plot_layout(design = layout)
 
 ggsave(g_concept,

### 2022-03-24 code/figure_food_web.R

@@ -72,4 +72,4 @@ g_fw <- foreach(i = 1:3) %do% {
   return(g)
 }
 
-g_web <- (g_fw[[1]] + ggtitle("A")) | g_fw[[2]] | g_fw[[3]]
+g_web <- (g_fw[[1]] + ggtitle("C")) | g_fw[[2]] | g_fw[[3]]

### 2022-03-24 code/figure_network.R

@@ -60,7 +60,7 @@ ng <- foreach(i = seq_len(length(net))) %do% {
   return(g)
 }
 
-g_net <- (ng[[1]] + labs(title = "C")) + ng[[2]] + ng[[3]] + plot_layout(guide = "collect")
+g_net <- (ng[[1]] + labs(title = "A")) + ng[[2]] + ng[[3]] + plot_layout(guide = "collect")
 
 
 # disturbance distribution ------------------------------------------------

### 2022-03-24 rmd/manuscript.Rmd
edit: manuscript

@@
       })
+bibliography: references.bib
 ---
@@
 
-We generatedsimulated branching networks of connected habitat patches as a series of nodes (Figure 1).
-The geometric arrangement of networks was controlled with two parameters: ecosystem size, the number of patches (NpatchNP); and ecosystem complexity, branching probability (PbB).
-Nodes (habitat patches) were assigned to be a confluence (or upstream terminal node) with probability PbB, or an in-branch node with probability 1-Pb 1 - PB.
-Branches (tributaries) are a series of connected nodes between confluences (or upstream terminal nodes).
-The number of patches within a branch,np, q q, is a random variable drawn from a geometric distribution np,q Ge(Pb)np,q∼Ge(Pb)  (a geometric distribution is the discrete form of an exponential distribution) but conditional on q=1Nbnp,q=Np (Nb is the number of branches).
-The number of branches NbB Under this assumption, the number of branches is identical to the number of branching nodes or the number of "success" out of Npatch trials.
-This translates into the assumption that the variable Nb follows a Binomial distribution asis a random variable drawn from a Binomial distribution: NbBinomial(Npatch, Pb).NbB \~ Binomial(NP, PbB).
-Generated river networks preserve important geometric scaling relationships (i.e., Horton's law).
-This framework preserves the fractal nature of branching patterns.
-For each scenario described below, we generated 1000 networks across a gradient of ecosystem size and complexity by drawing NP and Pba random variable from uniform distributions as: NpatchUnif(5, 150) NP \~ Unif(5, 150) and PbUnif(0.01, 0.99)PbB \~ Unif(0.01, 0.99) (Np was rounded to the nearest integer before running simulations).
+We generated branching networks as connected habitat patches (**Figure 1A**).
+The geometric arrangement of networks was controlled by two parameters: ecosystem size, the number of patches ($N_{p}$); and ecosystem complexity, branching probability ($P_{b}$).
+Habitat patches were assigned to be a confluence (or upstream terminal patch) with probability $P_{b}$, or an in-branch patch with probability $1-P_{b}$.
+Branches (tributaries) are a series of connected patches between confluences (or upstream terminal patches).
+The number of patches within branch $q$, $n_{p,q}$, is a random variable drawn from a geometric distribution $n_{p,q} \sim Ge(P_{b})$ but conditional on $N_{p} = \sum_{q}^{N_{b}} n_{p,q}$ ($N_{b}$ is the number of branches in a network).
+Full details of the network generation process is described in @teruiEmergentDualScaling2021.
+For each ecological scenario described below, we generated 1000 networks across a gradient of ecosystem size and complexity by drawing $N_{p}$ and $P_{b}$ from uniform distributions as: $N_{p} \sim Unif(10, 150)$ and $P_{b} \sim Unif(0.01, 0.99)$ ($N_{p}$ was rounded to the nearest integer before running simulations).
 
-For each terminal branch (i.e., headwater branch), h, within the network, a disturbance value, mh, was sampled .  The distribution of mh values was sampled from a normal distribution in a logit scale.
-This was accomplished by transforming the proportional mean disturbance value, µm, to the logit scale and sampling a value for each upstream terminal node (headwater) as logit mh Normal(logit m, m2).mh \~ Normal( logit( µm), σ2m).
-Here, we varied the mean source disturbance value, m = (0.1, 0.9) µm = (0.1, 0.9), based on disturbance regime (see **Simulations**) and set the variation of source disturbance as mσ2m = 2.
-We added lLocal environmental noise (i.e., noise occuring at each node) towas included in the disturbance value of each node as a spatial autoregressive process: logit mdown = logit mup+.
-mdown and mup are the disturbance values at down- and upstream nodes, and the local environmental noise at the downstream node that is drawn from a normal distribution as Normal(0, 2).
-This autoregressive process resembles natural disturbance regime, where flood or drought disturbance cascades downstrem with some local variation along the channel.
+We simulated the spatial distribution of disturbance intensity (i.e., proportional mortality in the event of stochastic regional disturbance) as follows.
+A disturbance value, $m_{h}$, was assigned randomly to upstream terminal patch $h$ as $logit~m_{h} \sim Normal(logit~m, \sigma_{h}^2)$, where $m$ is the average of disturbance intensity at headwaters in an ordinary scale.
+The disturbance values propagate downstream through a spatial autoregressive process as $logit~m_{down} = logit~m_{up}+ \epsilon$, where $m_{down}$ and $m_{up}$ are the disturbance values at down- and upstream patches, and the local environmental noise at the downstream patch $\epsilon$ was drawn from $Normal(0, \sigma_{l}^2)$.
+This autoregressive process resembles natural disturbance regime, where flood or drought disturbance cascades downstream with some local variation along the channel.
+At confluence patches, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, $Q$ (the number of upstream contributing patches): $logit~m_{down} = \omega~logit~m_{1,up} +(1-\omega)~logit~m_{2,up}$, where $m_{1,up}$ and $m_{2,up}$ are the disturbance values of upstream contributing tributaries, and $\omega$ the relative size of one tributary over another defined as $\omega = \frac{Q_{1}}{Q_{1} + Q_{2}}$.
+Thus, the larger tributary has a greater influence on the downstream environment, as observed in natural river networks [@bendaNetworkDynamicsHypothesis2004].
+Disturbance value distributions had medians close to $m$ with long tails (**Figure 1B**).
+The low-magnitude patches can be viewed as refugia, which supply individuals to impacted sites in subsequent time steps.
 
-s moving downstream by sampling a random variable from Normal(0, σ2l).
-For the main simulation results presented here, σ2l = 0.01.
-At confluence nodes, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, s (the number of upstream contributing nodespatches): logit mdown= logit m1,up+(1-) logit m2,upmdown = ωm1,up + (1 - ω)m2,up, where mdownmdown is the disturbance magnitude of the focal patch (the confluence), m1,upm1, up and m2,upm2, up are the disturbance magnitude of upstream adjacentcontributing nodestributaries, and ω the relative size of one tributary over another defined as =s1/(s1+s2).
-= s1/(s1 + s2).
-Disturbance value distributions had medians close to µm with long tails (**Fig S1**).
-with long right or left tails (**Fig S1**), respectively, i.e.
-For example, when m=0.1µm = 0.1 (e.g., a 10% reduction in species densities), most sites experienced low-magnitude disturbances and a few sites experienced high-magnitude disturbances.
-Likewise, whenm=0.9,  µm = 0.9, most sites experienced high-magnitude disturbances with a few patches receiving low-magnitude disturbances.
-The low-magnitude patches can be viewed as refugia and a source to supply individuals to colonize impacted sites in subsequent timesteps.
-Example disturbance value distributions for simulation scenarios reported here are available in the SI.
-
-Carrying capacity in patch x , Kx ,was assumed to increase as a power function of ecosystem area (i.e., watershed area)58:
-
-Kx= Kbase+C\*Npatch, up
-
-Where Kbase=100, C is a constant, Npatch, up number of patches upstream, and a random variable sampled from Unif(1.1, 1.2).
-At upstream terminal nodes (headwaters) K=Kbase.
-The R package mcbrnet was used to simulate the networks (Citation).
+We also mimicked the downstream increase of local carrying capacity [@teruiThreeEcologicalFactors2016; @carraraComplexInteractionDendritic2014; @rahelFishAssemblagesHabitat1991].
+Carrying capacity in patch $x$, $K_x$, was assumed to increase as a power function of ecosystem area (i.e., watershed area) [@finlayStreamSizeHuman2011; @koenigEmergentProductivityRegimes2019]: $$
+K_{x} = K_{base}N_{p,up}^z
+$$where $K_{base}$ is the carrying capacity at the upstream terminal patch, $N_{p,up}$ the number of patches upstream, and $z$ is the scaling exponent.
+We used $K_{base} = 100$ and $z = 0.54$ based on empirical relationships of gross primary production and watershed area [@finlayStreamSizeHuman2011].
 
@@
 
-We simulated trophic community dynamics in branching river networks using a discretemixed-time model which models predation continuously and has discrete, synchronous reproduction.
-This is a life-history common to many lotic species, including macroinvertebrates and fish.
-Accounting for this stage structure is an important aspect which is often missed in continuous-time models57.
-We describe the model in detail below.
+We employed a discrete-time model to describe tri-trophic dynamics in each habitat patch.
+A discrete model is suitable for organisms with predictable reproduction cycles, a common life-history for many lotic species including macroinvertebrates and fish.
+Our simulation model consists of three species communities; a basal species $B$ with positive logistic growth up to a carrying capacity, $K$; a primary consumer species or intraguild prey, $C$, which grows based on the consumption of $B$; and an intraguild predator $P$, which can consume both $B$ and $C$ (**Figure 1C**).
+In our model, the realized species abundance $A_{ix}(t+1)$ for species i (either $B,C,~or~P$) at patch $x$ and time $t+1$ is expressed as a composite of within-patch trophic interactions, immigration, and emigration:
 
-Our simulation model consists of three species communities; a basal species B which has positive logistic growth up to a carrying capacity, K; a primary consumer species, C which grows based on the consumption of B; and a predator P, which can consume both B and C.
-In our model, the realized species abundance Nix(t+1) for species i (either B, C, or P) at patch x and time t is expressed as a composite of within-patch trophic interactions, immigration, and emigration:
+$$
+A_{ix}(t+1) \sim Poisson(\bar{A}_{ix}(t) + I_{ix}(t) - E_{ix}(t))
+$$
 
-Nix(t+1)∼ Poisson(max{nix(t)+Iix(t)-Eix(t), 0})
+where $\bar{A}_{ix}$ is the expected species abudance after within-patch trophic interactions, \$I\_{ix} the expected number of immigrants into patch $x$, and $E_{ix}(t)$ the expected number of emigrants from patch $x$.
+The realized discrete number of individuals is drawn from a Poisson distribution to account for stochasticity in demographic and dispersal processes.
 
-Carrying capacity, KK, in patch xx , Kx ,was assumed to increase basen onexponentially as a power function of ecosystem area (i.e., watershed area)58:
-
-Kx= Kbase+C\*N
-
-Where KBase = 100, C is a constant, N = number of patches upstream, and ψ was a random variable sampled from Unif(1.1, 1.2). At upstream terminal nodes (headwaters) K = KBase.
-
-where nix(t) is the expected species abudance after within-patch trophic interactions, Iix(t) the expected number of immigrants into patch x, and Eix(t) the expected number of emigrants from patch x.
-The realized discrete number of individuals is drawn from a Poisson distribution to account for stochasticity in demographic and dispersal processes.Let Nix(t) Nix(t) represent the population density of species i in patch x at time t.
+Let Nix(t) Nix(t) represent the population density of species i in patch x at time t.
 The model starts with some initial nonzero population densities for all three species in each patch.

### 2022-03-24 rmd/manuscript.Rmd
edit: manuscript

@@
 bibliography: references.bib
+header-includes:
+  \usepackage{textgreek}
 ---
@@
 
-The origin of variation in food chain length (FCL), the number of feeding links from the basal species to top predators, has intrigued ecologists for decades due to its importance in regulating top-down effects 1, energy flow2, and contaminant concentrations in top predators that humans often consume3.
-Hypotheses regarding controls of food chain length can be organized into three broad categories.
-First, the resource availability hypothesis states that basal productivity controls food chain length because imperfect energy conversion through predator-prey interactions restricts energy available to higher trophic levels4.
-Second, the disturbance hypothesis (or the dynamical stability hypothesis) predicts longer food chains are vulnerable to environmental perturbations (i.e., the dynamical constraint).
-Thus, habitats with frequent and/or severe disturbance should support fewer trophic levels5.
-Lastly, the ecosystem size hypothesis posits that larger ecosystems should harbor longer food chains through a suite of mechanisms, including enhanced colonization rates6 increased habitat heterogeneity7 and surpassing spatial constraints inhibiting metapopulation dynamics.
-
-Among the three hypotheses, the ecosystem size hypothesis has been consistently supported in both theoretical and empirical explorations, likely because it encompasses multiple mechanisms which underlie the positive association8,9.
-For example, large ecosystems have increased species richness (species-area relationship8), buffer against disturbances9, and have greater numbers of immigrant sources in the ecosystem, harboring stable metapopulations of constituent species that make up long food chains 6,10,11.
-Meanwhile, the theoretical basis for the other two hypotheses rest on implicit assumptions of linear food chains, which are often violated in natural systems.
-As such, the effects of disturbance and resource availability on food chain length have been found to be highly variable among ecosystem types and geographic regions 12--14.
-
-More recently, food web ecologists have begun to recognize that FCL rarely responds to a single driver, but instead is determined by interactions of multiple factors.
-For example, Ward and McCann15 showed that FCL was more responsive to increasing productivity in large ecosystems compared with small ecosystems, whereas FCL was more sensitive to size in low productivity systems .
-Such context dependent  responses to extrinsic factors seem to be tightly coupled with the arrangement of trophic links within a food web, often referred to as food web motifs16,17.
-For example, omnivory (feeding at multiple trophic levels) is a strong intrinsic driver of food chain length 6,15,18,19.
-When omnivory is absent in three-species communities, the trophic structure assumes a simple linear food chain with maximum length equal to 3 (**Figure 1**).
-However, as omnivory increases, FCL decreases proportionally as the top predator eats lower in the food web 15,20.
-Theoretical work predicts that food webs are destabilized when omnivory is prevalent, leading to the loss of species and shorter food chains even in productive, large ecosystems.
-Conversely, weak omnivory stabilize the food web and increase persistence of species within a community, particularly at intermediate levels of productivity and high disturbance levels 20,21.
-Omnivory is a dominant module in natural communities 22 but is often not accounted for in theoretical studies20, but see 15.
-Due to the context-dependency of FCL to multiple drivers, and the potential for interactive effects, a unified framework of multiple food chain drivers is needed to better understand when and where FCL will respond to these extrinsic and intrinsic factors.
-
-While recent research  efforts of food web ecologists have substantially advanced our understanding of food chain drivers, most of the previous work has been conducted in simple, two-dimensional systems, such as lakes7,23,24 and oceanic islands 25,26.
-However, many natural habitats have a high degree of spatial complexity that cannot be represented simply by ecosystem size.
-Branching ecosystems, for example, are ubiquitous yet overlooked landscape structures 27, in which geomorphological or biological processes form naturally fractal branching patterns, where the geometric arrangement of patches is similar across spatial scales28 (e.g., rivers, trees, caves, mountain ranges).
-While we know that food web interactions are spatially influenced29,30, the theory to account for it has lagged, particularly in branching habitat architectures.  Here, we theoretically explore drivers of food chain length in rivers, an excellent example of branching ecosystems.
-
-Ample evidence suggests that branching structure controls environmental heterogeneity and dispersal pathways of organisms within river ecosystems.
-For example, water and materials propagate downstream as small streams hierarchically join to form larger streams and rivers.
-Consequently, environmental signals are highly variable among tributaries while showing a strong within-tributary correlation e.g., carrying capacity, flood disturbance31,32.
-Confluences serve as merging points where branch-specific dynamics aggregate to form distinct environmental conditions in downstream habitat patches 33.
-Habitat heterogeneity therefore increases with branching complexity (**Figure 1**), which we define here as the probability of branching per unit river distance.
-Hence, highly-branched river networks provide diverse habitats that buffer the impact of large-scale environmental fluctuations 34, broadening opportunities for species to recolonize habitat patches post-disturbance.
-Thus, ecosystem complexity in river networks may be a widespread yet overlooked extrinsic environmental driver which needs to be incorporated to understand food chain length.
-
-We develop a model in which trophic community dynamics within branching habitats are simulated.
-We investigate extrinsic controls such as ecosystem size (number of patches), complexity (branching probability), disturbance regime and productivity level (carrying capacity).
-In addition, our simulation model incorporates intrinsic drivers of food chain length including food web motifs (degree of omnivory) and dispersal ability, allowing us to investigate the context dependency on drivers of food chain length across extrinsic factors.
-We find that ecosystem size and complexity are both positively related to FCL, but the form of this relationship is dependent on other extrinsic and intrinsic factors.
-Trophic structure (degree of omnivory) and disturbance regime are both strong drivers of FCL on their own and have interactive effects.
-We also find that productivity and dispersal ability were both positively related to FCL.
-Our model is an important step in unifying drivers of FCL across extrinsic and intrinsic drivers and parsing their context-dependency in spatially complex ecosystems.
-
 # Methods
@@
 We simulated the spatial distribution of disturbance intensity (i.e., proportional mortality in the event of stochastic regional disturbance) as follows.
-A disturbance value, $m_{h}$, was assigned randomly to upstream terminal patch $h$ as $logit~m_{h} \sim Normal(logit~m, \sigma_{h}^2)$, where $m$ is the average of disturbance intensity at headwaters in an ordinary scale.
+A disturbance value, $m_{h}$, was assigned randomly to upstream terminal patch $h$ as $logit~m_{h} \sim Normal(logit~\mu_m, \sigma_{h}^2)$, where $\mu_m$ is the average of disturbance intensity at headwaters in an ordinary scale.
 The disturbance values propagate downstream through a spatial autoregressive process as $logit~m_{down} = logit~m_{up}+ \epsilon$, where $m_{down}$ and $m_{up}$ are the disturbance values at down- and upstream patches, and the local environmental noise at the downstream patch $\epsilon$ was drawn from $Normal(0, \sigma_{l}^2)$.
 This autoregressive process resembles natural disturbance regime, where flood or drought disturbance cascades downstream with some local variation along the channel.
-At confluence patches, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, $Q$ (the number of upstream contributing patches): $logit~m_{down} = \omega~logit~m_{1,up} +(1-\omega)~logit~m_{2,up}$, where $m_{1,up}$ and $m_{2,up}$ are the disturbance values of upstream contributing tributaries, and $\omega$ the relative size of one tributary over another defined as $\omega = \frac{Q_{1}}{Q_{1} + Q_{2}}$.
+At confluence patches, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, $Q$ (the number of upstream contributing patches): $logit~m_{down} = \omega~\{logit~m_{1,up} +\epsilon_1\}+(1-\omega)~\{logit~m_{2,up} +\epsilon_2\}$, where $m_{1,up}$ and $m_{2,up}$ are the disturbance values of upstream contributing tributaries, $\epsilon_1$ and $\epsilon_2$ are independent local noises, and $\omega$ the relative size of one tributary over another defined as $\omega = \frac{Q_{1}}{Q_{1} + Q_{2}}$.
 Thus, the larger tributary has a greater influence on the downstream environment, as observed in natural river networks [@bendaNetworkDynamicsHypothesis2004].
@@
 We also mimicked the downstream increase of local carrying capacity [@teruiThreeEcologicalFactors2016; @carraraComplexInteractionDendritic2014; @rahelFishAssemblagesHabitat1991].
-Carrying capacity in patch $x$, $K_x$, was assumed to increase as a power function of ecosystem area (i.e., watershed area) [@finlayStreamSizeHuman2011; @koenigEmergentProductivityRegimes2019]: $$
+Carrying capacity in patch $x$, $K_x$, was assumed to increase as a power function of ecosystem area (i.e., watershed area) [@finlayStreamSizeHuman2011; @koenigEmergentProductivityRegimes2019]:
+
+```{=tex}
+\begin{align}
 K_{x} = K_{base}N_{p,up}^z
-$$where $K_{base}$ is the carrying capacity at the upstream terminal patch, $N_{p,up}$ the number of patches upstream, and $z$ is the scaling exponent.
+\end{align}
+```
+where $K_{base}$ is the carrying capacity at the upstream terminal patch, $N_{p,up}$ the number of patches upstream, and $z$ is the scaling exponent.
 We used $K_{base} = 100$ and $z = 0.54$ based on empirical relationships of gross primary production and watershed area [@finlayStreamSizeHuman2011].
@@
 Our simulation model consists of three species communities; a basal species $B$ with positive logistic growth up to a carrying capacity, $K$; a primary consumer species or intraguild prey, $C$, which grows based on the consumption of $B$; and an intraguild predator $P$, which can consume both $B$ and $C$ (**Figure 1C**).
-In our model, the realized species abundance $A_{ix}(t+1)$ for species i (either $B,C,~or~P$) at patch $x$ and time $t+1$ is expressed as a composite of within-patch trophic interactions, immigration, and emigration:
+In our model, the realized species abundance $A_{i,x}(t+1)$ for species i (either $B,C,~or~P$) at patch $x$ and time $t+1$ is expressed as a composite of within-patch trophic interactions, immigration, and emigration:
+
+```{=tex}
+\begin{align}
+A_{i,x}(t+1) \sim Poisson(\bar{A}_{i,x}(t) + I_{i,x}(t) - E_{i,x}(t))
+\end{align}
+```
+where $\bar{A}_{i,x}$ is the expected species abundance after within-patch trophic interactions, $I_{i,x}(t)$ the expected number of immigrants into patch $x$, and $E_{i,x}(t)$ the expected number of emigrants from patch $x$.
+The realized discrete number of individuals is drawn from a Poisson distribution to account for stochasticity in demographic and dispersal processes.
+
+In each time step, internal dynamics occur first. The basal species is exposed to predation as well as other sources of mortality:
+
+```{=tex}
+\begin{align}
+A_{B,x}'(t) = s_{0,B}\{1-z_{B,x}(t)\}A_{B,x}(t)
+\end{align}
+```
+where $s_{0,i}$ is the background survival probability (density independent) and $z_{B,x}(t)$ the proportional mortality caused by predation. Subsequently, a density-dependent recruitment occurs:
+
+```{=tex}
+\begin{align}
+\bar{A}_{B,x}(t) = \frac{r_{b}A_{B,x}'(t)}{1 + \frac{r_b - 1}{K_{x}}A_{B,x}'(t)}(1-\psi(t)~m_{x})
+\end{align}
+```
+$r_b$ is the reproductive rate for survived individuals, $\psi(t)$ the random binary variable indicating the occurrence of regional disturbance (disturbance occurs if $\psi(t)=1$ otherwise $0$), and $m_x$ the proportional mortality caused by disturbance.
+$\psi(t)$ was drawn from a Bernoulli distribution with disturbance probability $p_m$.
+The predator-induced mortality $z_{B,x}(t)$ is determined as follows:
+
+```{=tex}
+\begin{equation}
+z_{B,x}(t) = z_{BC,x}(t) + z_{BP,x}(t) =
+\left\{
+\begin{aligned}
+&\frac{W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)}{A_{B,x}(t)} &&\text{if~$W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t) < A_{B,x}(t)$}\\
+&1 &&\text{if~$W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t) > A_{B,x}(t)$}\\
+\end{aligned}
+\right.
+\end{equation}
+```
+where $W_{ij,x}(t)$ is the total consumption of consumer $j$ on prey $i$ (number of individuals). We assumed the Holling Type-II functional response for predator-prey interactions:
+
+```{=tex}
+\begin{align}
+W_{ij,x} = \frac{a_{ij}A_{i,x}A_{j,x}}{A_{j,x} + a_{ij}h_{ij}A_{i,x}}
+\end{align}
+```
+$a_{ij}$ is the attack rate and $h_{ij}$ the handling time.
+The parameter $\delta_{x}(t)$ is the relative preference of intraguild predator $P$ to basal species $B$ over intraguild prey $C$.
+To be realistic, we assume that the preference parameter depends on the relative potential amount of energy obtained from each prey:
+
+```{=tex}
+\begin{align}
+\delta_{x} = \frac{e_{BP} f_{BP}(A_{B,x})}{e_{BP} f_{BP}(A_{B,x}) + e_{CP} f_{CP}(A_{C,x})}
+\end{align}
+```
+where $e_{ij}$ is the conversion efficiency from prey $i$ to consumer $j$, and $f_{ij}(A_{i,x}) = \frac{a_{ij}A_{i,x}}{1 + a_{ij}h_{ij}A_{i,x}}$.
+An analogous approach has been applied to a model with multiple prey choices [@shotanishijimaRolesAlternativePrey2014].
+
+The intraguild prey $C$ is controlled by both prey availability and predation.
 
 $$
-A_{ix}(t+1) \sim Poisson(\bar{A}_{ix}(t) + I_{ix}(t) - E_{ix}(t))
+\bar{A}_{C,x}(t) = e_{BC}s_{0,C}(1-z_{CP,x})z_{BC,x}A_{B,x}
 $$
-
-where $\bar{A}_{ix}$ is the expected species abudance after within-patch trophic interactions, \$I\_{ix} the expected number of immigrants into patch $x$, and $E_{ix}(t)$ the expected number of emigrants from patch $x$.
-The realized discrete number of individuals is drawn from a Poisson distribution to account for stochasticity in demographic and dispersal processes.
-
-Let Nix(t) Nix(t) represent the population density of species i in patch x at time t.
-The model starts with some initial nonzero population densities for all three species in each patch.
-The initial population abundances for each species i in each patch x were independently sampled from a random Poisson distribution Poisson(ix)Poisson(𝝺ix), where ix𝝺ix for B, C and P was 0.8Kx \* K, 0.5Kx \*K and 0.25Kx \* K, respectively.  First, dispersal from all patches is accounted for.
-The number of emigrants, Eix(t) for species i in from patch x at time t, is estimated as the product of dispersal probability, Pdispersal and species abundance Nix(t).
-The dispersal probability was assumed to be constant for all three species in each simulation run.
-Then, the number of emigrants from patch x is divided into immigrants, Iix(t) Ixy(t) to each patch y.
-The immigration probability at patch x and time t, ix(t) xy ξxy from patch x to y is calculated considering network structure and species dispersal capability as followsbased on the habitat structure among patches according to:
-
-xy(t) =y,y≠xExy(t)e-θdxyxy,y≠xExy(t)e-θdxy
-
-ix(t) = y,yxEiy(t)e-dxyxy,yxEiy(t)e-dxy
-
-Where dxy is the distance between patch x and y, and θ controls the rate of decay. Then, the expected number of immigrants into patch x is calculated as Iix(t)=ix(t)xEix.
-
-The realized number of individuals Nix(t+1) (species i at patch x and time t+1) is given as:
-
-Nix(t+1)∼ Poisson(nix(t)+Iix(t)-Eix(t))
-
-where Nix(t) is the number of individuals at time t,(i.e., before dispersal) Iix(t) the expected number of immigrants to patch x, and Eix(t) the expected number of emigrants from patch x.
-The realized discrete number of individuals is drawn from a Poisson distribution to account for stochasticity in demographic and dispersal processes.
-Note that a value of 0 is assigned to Nix(t + 1) for any non-positive value of 𝝺 59.
-
-After dispersal, trophic interactions occur.continuous predation occurs. We model the tri-trophic interactions with intraguild predation as follows:
-
-nBx(t+1)=(rmax1+υNBx(t)')(1-mx)				Basal species
-
-nCx(t+1)=WBC, xNCx(t)'(1-mx)			Intraguild prey
-
-nPx(t+1)=(eBPWBP, x+eCPWCP, x)NPx(t)'(1-mx)	Intraguild predator
-
-The above equations indicate that the basal species exhibit a  where rmax is the maximum population growth of basal species, υ the term controlling density-dependence (rmax-1Kx), and mx the proportional mortality when disturbance occurs (disturbance occurs when =1, otherwise 0).
-
-\
-
-The number of prey i eaten by consumer j, Wij, in a season is calculated as:
-
-\
-Wij= ijNiNjijNj+Ni
-
-where Ni and Nj are resource and consumer densities, respectively and αij and βij are coefficients describing the functional response as the ratio of resources, i, to consumers, j, increases.
-βij describes how fast the predation curve reaches the asymptote of the Type II functional response, and αij/βij determines the location of the asymptote57.
-
-The predator species P, searches for and consumes both species in a patch.
-Its total effort is controlled by parameters α.p and β.p.
-The preference parameter δ, controls the proportional search effort of species P for resource B and C:
-
-WBP=δ(Wij)
-
-WCP=(1 - δ)(Wij)
-
-Predator preference for resource B is determined by the equation:
-
-δ= eBPNBeBPNB+ eCPNC
-
-Where eBP and eCP are the conversion efficiencies of predator P for resource B and C, and NB and NC are resource densities, respectively.
-This formulation allows for flexible diet shifts of the predator based on realized energy gain (conversion efficiency times prey availability) from the prey species60.
-
-At each time-step within a patch, FCL was defined as 0 if no species were present, 1 if only B was present and 2 if only B + C or B + P were present.
-When all three species were present, FCL was calculated based on the proportion of each resource in P's diet (𝜆i) according to:
-
-FCL = (1 \* B+ 2 \*C) +1
-
-When P only consumes B or C, FCL = 2 or 3, respectively. When P consumes both resources equally, FCL = 2.5. In each simulation, FCL was averaged across all patches and time-steps.
-
-The number of individuals in each patch is reduced by the number consumed Wij (predation effect), as well as by a base survival probability, s0:
-
-NB'=s0(NB- WBC -  WBP)
-
-NC'=s0(NC'-WCP)
-
-P'=s0P
-
-Finally, those individuals remaining reproduce according to the following:
-
-NB, t+1=(rmax1 + υ NB' NB' )(1-φmx)
-
-NC, t+1=(WBCNC')(1-φmx)
-
-NP, t+1=(eBPWBP+ eCPWCP)NP'(1-φmx)
-
-where rmax is the maximum per-capita reproduction rate of the basal species and υ = rmax-1K determines the asymptotic relationship of population growth, where K is carrying capacity.
-In the absence of predation, population growth of the basal species follows a logistic growth relationship.
-For the consumer species (C and P) eij is the efficiency of converting resource i into new consumer j.
-The last term in the equation accounts for the proportional mortality in the event of a disturbance, where mx = the disturbance value (see **Network Generation**) and 𝜑 is an indicator variable of disturbance:
-
-φ =0, no disturbance
-
-φ =1, disturbance occurs
-
-If species B is extirpated from a patch, both the other species were also assumed to be extirpated from that patch during that time step.
-Species B can colonize any open patch, whereas species C and P can only successfully colonize patches where B was already present (i.e., prey resource is necessary for successful establishment).
-
-## **Simulations**
-
-We used 24 combinations of parameter sets (**SI Table**).
-The parameter sets include two environmental factors, and twelve ecological scenarios.
-For environmental scenarios, we used two parameter sets that controlled the frequency and magnitude of disturbances (low and high disturbance regime, respectively).
-In the low disturbance regime, the probability of a disturbance occurring at each time step was 0.0001 (i.e., a 10,000 year recurrence), and the mean disturbance magnitude was μm = 0.1 (10% reduction in population densities).
-In contrast, the high disturbance regime had a probability of 0.01 (i.e., 100 year recurrence), and the mean disturbance magnitude was μm = 0.9 (90% reduction in population densities61)
-
-For the ecological scenarios, we used two productivity levels, two dispersal abilities, and three levels of omnivory.
-Productivity was varied by setting the carrying capacity of the basal species, B (K = 50, 150).
-Dispersal ability varies the effective distance between patches and can control the probability that a patch is re-colonized post-disturbance.
-Two levels of dispersal ability were simulated.
-The high dispersal ability was set with parameters Pdispersal = 0.1 and θ = 0.1 (i.e., high probability of individuals emigrating, and far distance traveled).
-The low dispersal scenario was set with parameters Pdispersal = 0.01 and θ = 1  (i.e., low probability of dispersal, short dispersal distance).
-These values were chosen to represent extreme scenarios.
-Preliminary analyses using intermediate parameter sets had responses between these two scenarios.
-
-Finally, we investigated three food web motifs (**Figure 1B**).
-Trophic interaction strengths can be controlled by varying the attack rate or conversion efficiency by the consumer P (Ward and McCann 2017).
-To simplify model simulations, we fixed the attack rate for both C and P (parameters 𝛼 = 8 and 𝛽 = 20), and the conversion efficiency for C as ebc = 6.
-We controlled the presence and strength of trophic interactions for the predator, P, by varying the conversion efficiency for both prey resources (i.e., varied prey resource quality).
-The scenario names and parameter levels were as follows: (1) "Food chain": ebp = 0, ecp = 4; (2) "Strong IGP": ebp = 2, ecp = 4; (3) "Weak IGP": ebp = 4, ecp = 2.
-
-Under each scenario, we simulate 1200 timesteps of trophic dynamics (including 200 steps of burn-in) in 1000 branching networks across a gradient of ecosystem size (number of habitat patches, from 5 to 150) and ecosystem complexity (branching probability, from 0.01 to 0.99).
-Simulations ran for 200 time steps as a burn-in period to minimize influences of initial conditions.
-The last 1000 timesteps were recorded to estimate the temporal average of food chain length across all patches within a network.
-
-Values of simulation parameters are summarized in (SI).
-We performed simulations using the mcbrnet and IGPtoy packages (blinded citation, in preparation) in R version 4.0.3.
-R Code to reproduce simulations, figures, and tables are available at <https://github.com/Jpomz/priv-igp-sens>.
-
-\
-The model starts with some initial nonzero population densities for all three species in each patch.
-The initial population abundances for each species i in each patch x were independently sampled from a random Poisson distribution Poisson(ix)Poisson(𝝺ix), where ix𝝺ix for B, C and P was 0.8Kx \* K, 0.5Kx \*K and 0.25Kx \* K, respectively.
 
@@
 
-First, we depicted branching ecosystems as connected habitat patches, which local communities inhabit.
-Each local community is composed of basal (B), primary consumer (C), and top predator (P), and these constituent species disperse along the network corridors following exponential dispersal kernels.
-The arrangement and strength of trophic interactions can be varied based on the simulation parameters (food web motifs, **Figure 1**, see **Methods**).
-Periodically, local communities are impacted by disturbance, whose strength is highly correlated within a tributary.
-Therefore, our simulation framework resembles environmental dynamics of natural river systems.
-
-We explored how food web motifs affected FCL across extrinsic factors.
-FCL is related to the degree of omnivory of the predator P (**Figure 1**).
-When P only consumes C, its trophic level is 3.
-However, as the proportion of B in P's diet increases, FCL decreases, reaching a minimum of 2 when B constitutes all of P's diet.
-We assume that P searches for both resources proportionally based on the conversion efficiency for each prey resource (i.e., prey quality) and the relative densities of each resource (see **Methods**).
-To control the level of omnivory, we varied the conversion efficiencies of both prey resources.
-
-In addition to food web motifs, we also investigated the effects of species dispersal ability and disturbance regime.
-To simplify our simulations, we looked at two scenarios of dispersal ability.
-1) Low dispersal ability, in which a small proportion of the population of all three species emigrate, and the likelihood of successful immigration to neighboring patches declines exponentially with increasing distance (i.e., short-distance dispersal).
-2) High dispersal ability, in which a larger proportion of the population emigrates, and the likelihood of successful immigration to neighboring patches declines less-rapidly with increasing distance (i.e., long-distance dispersal; see **Methods**).
-
-The effect of disturbance was also investigated by looking at low and high disturbance regimes.
-The low disturbance regime was defined as having a low probability of a relatively weak magnitude disturbance occurring throughout the habitat network in each time step.
-This can be likened to spring-fed streams which are relatively stable through time but are subjected to occasional disturbances such as floods from high precipitation events.
-In contrast, we investigated a high disturbance regime in which the probability and magnitude of disturbances occurring throughout the habitat network was much higher.
-This scenario could represent "flashy" streams which are predominantly driven by surface runoff and precipitation events.
-
-We examined interactive effects of extrinsic and intrinsic factors controlling food chain length.
-We simulated habitat networks over a gradient of both ecosystem size and complexity (branching probability).
-For each habitat network, we simulated trophic dynamics under two different regimes of disturbance (rare or frequent), productivity (low or high carrying capacity of the basal species), and dispersal ability (low and high), and three local food web structures (the degree of omnivory, **Figure 1**), for a total of 24 scenario combinations.
-We simulated meta-food web dynamics over 1200 time steps, including 200 steps for burn in, and yielded estimates of food chain length averaged in space and time (see **Methods**).
-Varying the disturbance regime resulted in markedly different results across ecosystem size and complexity.
-Thus, we present our results for each disturbance regime separately below.
-
-## **Low disturbance regime**
-
-In ecosystem networks with a low disturbance regime (i.e., dynamical stability), FCL was largely determined by food web motifs (colored lines in **Figure 2**).
-Generally, FCL was negatively related to the degree of omnivory, with simple linear food chains (P only consumes C) having the longest FCL, and strong omnivory (P consumes very little C) having the shortest FCL.
-When carrying capacity and dispersal ability was low, there was a slight increase in FCL with the omnivorous motifs across increasing ecosystem size, and when dispersal ability was high, there was a moderate increase for the strong omnivory module (**Figure 2A**).
-In addition, there was a slight increase in FCL with high dispersal ability across both carrying capacities, but the effect was greater at low carrying capacity.
-However, when carrying capacity was increased, FCL was largely invariant across gradients of ecosystem size.
-
-There was very little effect of ecosystem complexity on FCL in the low disturbance regime.
-In all scenarios, FCL did not vary in response to increasing ecosystem complexity, although there was a very slight decline for all three food web modules when carrying capacity and dispersal ability were low.
-FCL was predominantly determined by food web motifs and carrying capacity, with FCL being negatively related to omnivory, and positively related to carrying capacity.
-Like the results across ecosystem size, there was a positive effect of high dispersal ability on FCL, and this effect was greater in ecosystems with low carrying capacity (interactive effect).
-
-## **High disturbance regime**
-
-Ecosystem networks with a high disturbance regime had much different responses of FCL across both ecosystem size and complexity.
-FCL increased across both gradients under all combinations of carrying capacity and dispersal ability, and in most food web motifs (**Figure 3**).
-However, the strength of the relationship was variable for some motifs, and when dispersal ability and carrying capacity were low, and omnivory was absent, FCL was largely invariant across both gradients (top right panel in **Figure 3A and 3B**.
-Once again, FCL was largely determined by the degree of omnivory, with simple linear food chains generally having longer FCL, and weak IGP having the shortest FCL.
-
-When dispersal ability and carrying capacity were low, however, a simple linear food chain structure resulted in the shortest FCL across both ecosystem size and complexity.
-It is likely that this simple motif with a few strong links was unstable and could not persist when exposed to frequent, high magnitude disturbances (e.g., supports dynamical stability hypothesis).
-In these environments, having prey choice (omnivory) appears to be a necessary requirement to maintain the persistence of species.
-
-Increasing dispersal ability and carrying capacity had positive effects on FCL across both ecosystem size and complexity.
-Similar to the results in the low disturbance regime, increasing dispersal ability in ecosystems with low carrying capacity resulted in a large increase in FCL, suggesting an interactive effect.
-
 # Discussion
-
-The ecosystem size hypothesis of food chain length has received support from both theoretical and empirical work.
-Because ecosystem size encompasses several distinct mechanisms that could influence food chain structure, work so far has not elaborated the underlying drivers.
-Ecosystem complexity is an extrinsic factor which is scale-invariant and related to ecosystem size yet has not been incorporated in theoretical explorations of food chain length to date.
-Here, we show that food chain length responds positively and independently to increasing ecosystem size and complexity.
-This supports our hypothesis that scale-dependent theories of FCL are inadequate to correctly predict FCL in ecosystems with a complex structural geometry.  However, the relationship between FCL and ecosystem complexity varied based on additional extrinsic and intrinsic factors, supporting the emerging paradigm that there is no one single driver of FCL in ecosystems 15, but moreover that there is context dependency for which factor is the primary driver of the relationship.
-Although we focus on river networks here, we expect our results to be transferable to other branching structures, including mountain ranges and cave systems.
-
-While increased disturbance intensity and frequency consistently reduced food chain length as predicted by dynamical stability hypothesis 5, more important findings in this study were evident from their synergistic effects with ecosystem structure.
-Our low disturbance regime scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independent of ecosystem size and complexity and FCL was predominantly determined by food web motifs.
-However, under high disturbance regimes, only large and complex networks could provide diverse and heterogeneous habitats with potential refugia, allowing species to thrive and maintain long food chains.
-This result has important implications for ecosystem management because anthropogenic climate change is predicted to increase the frequency and magnitude of disturbances 35,36.
-River networks are especially vulnerable due to their propensity to go dry during droughts37 as well as susceptibility to more intense flood disturbance with increasing storm intensity38.
-Increased dispersal ability alleviated the reduction in FCL from increased disturbance regime in our simulation, and reduced the negative impacts of disturbance and climate change in empirical studies 39.
-However, human activities have altered the habitat heterogeneity, connectivity, and dispersal ability of species within riverine networks 40,41.
-In-stream habitat in rivers and streams have been homogenized and lateral connections to the floodplain have been severed by the removal of large woody debris, channelization and straightening of river corridors, and through flow regulation 42,43, and this has affected community assembly44 and trophic structure 45.
-
-Although dispersal has received little attention in food chain research, our simulation highlighted its importance in regulating food chain length.
-In particular, high dispersal was required for ecosystem size and complexity to have consistent positive effects on FCL under high disturbance regimes.
-This is probably because sufficient propagules are needed to successfully recolonize after disturbance and maintain long food chains in large and/or complex networks.
-Human activities that affect habitat connectivity and dispersal ability of species within riverine networks may compromise the buffering effects of ecosystem size and complexity against disturbance by limiting dispersing propagules.
-For example, dams and diversion have altered the connectivity between suitable habitat patches, or have completely blocked dispersal ability 46,47.
-Light pollution has also negatively affected the ability of adult aquatic insects to disperse between habitat patches 48.
-
-The context dependency of disturbance, productivity and dispersal has important implications for the conservation and management of freshwater ecosystems.
-This adds an important distinction in the discussion of the suitability of assigning protection or conservation status to regions.
-Current conservation schemes generally emphasize size, aiming to protect the largest areas possible.
-While this is certainly a useful attribute generally, designating regions with a high degree of variability, or many small patches, may be more beneficial than a large, homogenous area 49, and accounting for ecosystem complexity is particularly suited to ecosystem architectures in which branches have variable environmental conditions.
-Conservation and management activities should focus on activities that increase the habitat heterogeneity and connectivity of suitable habitat patches to increase resilience to predicted increases in disturbance frequency and magnitude.
-
-The role of food web motifs on FCL largely followed our predictions (**Figure 1**).
-Specifically, FCL declined as the value of the resource B, and thus the proportion in the diet, to the predator P increased.
-Essentially, the relative qualitative values of the resources to the predator controlled the level of omnivory and FCL.
-Although FCL was greatest when the community was arranged as a simple linear food chain, this arrangement was not the most stable.
-Specifically, when disturbance regime was high, and productivity and dispersal ability were low, the observed FCL across networks was \~1 (although increasing dispersal ability and/or carrying capacity resulted in food chains of \~2.5-3).
-Therefore, in the absence of prey resource choice, strong, top-down trophic interactions were destabilizing, leading to the extirpation of consumer species at higher trophic levels across the networks.
-
-Changes in community composition are occurring worldwide and in all habitats.
-Species extirpations and extinctions, coupled with species introductions and invasions are creating novel communities worldwide.
-In Lake Tahoe (USA), for example, replacement of the top fish predator and the introduction of a freshwater shrimp substantially altered the food web structure 50.
-Understanding the interactions within these novel communities, and how this could affect food chain lengths is an important goal of contemporary ecologists.
-The strength of omnivory in these novel communities can be increased or decreased, depending on the introduced species and the extant species pool or community.
-For example, non-indigenous fishes in the Eastern Mediterranean Sea outcompete native species for high quality prey items, causing native species to increase consumption of non-preferred prey items51.
-Introductions of non-native fishes in the Colorado River basin have led to longer FCL and narrower niches (e.g., lower degree of omnivory) of native species52.
-Alternatively, the introduction of hyper-successful basal or primary consumer species can weaken or remove the omnivory link by becoming the dominant prey resource.
-For example, nonnative New Zealand mud snail dominated the macroinvertebrate biomass in streams in Yellowstone National Park and likely affected food web structure53.  Having a single prey resource dominate the community is akin to reducing the degree of omnivory (i.e., removing prey choice), which destabilizes the community, particularly under high disturbance regimes.
-
-It is abundantly clear that ecosystem complexity in natural systems is the rule, not the exception.
-Early theoretical and empirical work has assumed simple habitat structure as a necessity.
-However, habitat heterogeneity, such as the presence of prey refugia, and dispersal between connected habitat patches allow communities to stabilize and persist through space and time 54--56.
-Complex ecosystem structure is an extrinsic factor that can affect food chain length in communities.
-Incorporating a realistic understanding of ecosystem structural complexity could alter our perception of communities in naturally complex systems, potentially leading to paradigm shifts in macroecology.
-
-\

### 2022-03-26 rmd/manuscript.Rmd
edit: manuscript

@@
 ---
-Title: "Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks"
+title: "Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks"
 output:
@@
 A disturbance value, $m_{h}$, was assigned randomly to upstream terminal patch $h$ as $logit~m_{h} \sim Normal(logit~\mu_m, \sigma_{h}^2)$, where $\mu_m$ is the average of disturbance intensity at headwaters in an ordinary scale.
-The disturbance values propagate downstream through a spatial autoregressive process as $logit~m_{down} = logit~m_{up}+ \epsilon$, where $m_{down}$ and $m_{up}$ are the disturbance values at down- and upstream patches, and the local environmental noise at the downstream patch $\epsilon$ was drawn from $Normal(0, \sigma_{l}^2)$.
+The disturbance values propagate downstream through a spatial autoregressive process as:
+
+```{=tex}
+\begin{align}
+logit~m_{down} = logit~m_{up}+ \epsilon
+\end{align}
+```
+where $m_{down}$ and $m_{up}$ are the disturbance values at down- and upstream patches, and the local environmental noise at the downstream patch $\epsilon$ was drawn from $Normal(0, \sigma_{l}^2)$.
 This autoregressive process resembles natural disturbance regime, where flood or drought disturbance cascades downstream with some local variation along the channel.
-At confluence patches, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, $Q$ (the number of upstream contributing patches): $logit~m_{down} = \omega~\{logit~m_{1,up} +\epsilon_1\}+(1-\omega)~\{logit~m_{2,up} +\epsilon_2\}$, where $m_{1,up}$ and $m_{2,up}$ are the disturbance values of upstream contributing tributaries, $\epsilon_1$ and $\epsilon_2$ are independent local noises, and $\omega$ the relative size of one tributary over another defined as $\omega = \frac{Q_{1}}{Q_{1} + Q_{2}}$.
+At confluence patches, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, $Q$ (the number of upstream contributing patches):
+
+```{=tex}
+\begin{align}
+logit~m_{down} = \omega~(logit~m_{1,up} +\epsilon_1)+(1-\omega)~(logit~m_{2,up} +\epsilon_2)
+\end{align}
+```
+$m_{1,up}$ and $m_{2,up}$ are the disturbance values of upstream contributing tributaries, $\epsilon_1$ and $\epsilon_2$ are independent local noises, and $\omega$ the relative size of one tributary over another defined as $\omega = \frac{Q_{1}}{Q_{1} + Q_{2}}$.
 Thus, the larger tributary has a greater influence on the downstream environment, as observed in natural river networks [@bendaNetworkDynamicsHypothesis2004].
@@
 \begin{align}
-K_{x} = K_{base}N_{p,up}^z
+K_{x} = K_{base}(N_{p,up} + 1)^z
 \end{align}
@@
 where $K_{base}$ is the carrying capacity at the upstream terminal patch, $N_{p,up}$ the number of patches upstream, and $z$ is the scaling exponent.
-We used $K_{base} = 100$ and $z = 0.54$ based on empirical relationships of gross primary production and watershed area [@finlayStreamSizeHuman2011].
+We used $K_{base} = 100$ and $z = 0.54$ based on an empirical relationship of gross primary production and watershed area [@finlayStreamSizeHuman2011].
 
@@
 ```
-where $\bar{A}_{i,x}$ is the expected species abundance after within-patch trophic interactions, $I_{i,x}(t)$ the expected number of immigrants into patch $x$, and $E_{i,x}(t)$ the expected number of emigrants from patch $x$.
+where $\bar{A}_{i,x}(t+1)$ is the expected species abundance after within-patch trophic interactions, $I_{i,x}(t)$ the expected number of immigrants into patch $x$, and $E_{i,x}(t)$ the expected number of emigrants from patch $x$.
 The realized discrete number of individuals is drawn from a Poisson distribution to account for stochasticity in demographic and dispersal processes.
 
-In each time step, internal dynamics occur first. The basal species is exposed to predation as well as other sources of mortality:
+***Internal dynamics.*** Each time step starts with internal dynamics. The basal species is exposed to predation as well as other sources of mortality:
 
@@
 \begin{align}
-A_{B,x}'(t) = s_{0,B}\{1-z_{B,x}(t)\}A_{B,x}(t)
+A_{B,x}'(t) = s_{0,B}[1-z_{B,x}(t)]A_{B,x}(t)
 \end{align}
 ```
-where $s_{0,i}$ is the background survival probability (density independent) and $z_{B,x}(t)$ the proportional mortality caused by predation. Subsequently, a density-dependent recruitment occurs:
-
-```{=tex}
-\begin{align}
-\bar{A}_{B,x}(t) = \frac{r_{b}A_{B,x}'(t)}{1 + \frac{r_b - 1}{K_{x}}A_{B,x}'(t)}(1-\psi(t)~m_{x})
-\end{align}
-```
-$r_b$ is the reproductive rate for survived individuals, $\psi(t)$ the random binary variable indicating the occurrence of regional disturbance (disturbance occurs if $\psi(t)=1$ otherwise $0$), and $m_x$ the proportional mortality caused by disturbance.
-$\psi(t)$ was drawn from a Bernoulli distribution with disturbance probability $p_m$.
-The predator-induced mortality $z_{B,x}(t)$ is determined as follows:
+where $s_{0,i}$ is the background survival probability (density independent) and $z_{B,x}(t)$ the proportional mortality caused by predation.
+Let $W_{B,x}(t)$ be the total number of basal species consumed by $C$ and $P$.
+The predator-induced mortality $z_{B,x}(t)$, the sum of proportional predation by $C$ ($z_{BC,x}(t)$) and $P$ ($z_{BP,x}(t)$), is then:
 
@@
 \begin{aligned}
-&\frac{W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)}{A_{B,x}(t)} &&\text{if~$W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t) < A_{B,x}(t)$}\\
-&1 &&\text{if~$W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t) > A_{B,x}(t)$}\\
+&\frac{W_{B,x}(t)}{A_{B,x}(t)} &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\
+&1 &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\
 \end{aligned}
@@
 ```
-where $W_{ij,x}(t)$ is the total consumption of consumer $j$ on prey $i$ (number of individuals). We assumed the Holling Type-II functional response for predator-prey interactions:
+and
 
@@
 \begin{align}
-W_{ij,x} = \frac{a_{ij}A_{i,x}A_{j,x}}{A_{j,x} + a_{ij}h_{ij}A_{i,x}}
+W_{B,x}(t) = W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)
 \end{align}
 ```
-$a_{ij}$ is the attack rate and $h_{ij}$ the handling time.
-The parameter $\delta_{x}(t)$ is the relative preference of intraguild predator $P$ to basal species $B$ over intraguild prey $C$.
-To be realistic, we assume that the preference parameter depends on the relative potential amount of energy obtained from each prey:
+where $W_{ij,x}(t)$ is the population-wide consumption of consumer $j$ on prey $i$ (number of individuals), and the parameter $\delta_{x}(t)$ is the relative preference of intraguild predator to basal species over intraguild prey.
+We assume the Holling Type-II functional response for predator-prey interactions:
 
@@
 \begin{align}
-\delta_{x} = \frac{e_{BP} f_{BP}(A_{B,x})}{e_{BP} f_{BP}(A_{B,x}) + e_{CP} f_{CP}(A_{C,x})}
+W_{ij,x}(t) = \frac{a_{ij}A_{i,x}(t)A_{j,x}(t)}{A_{j,x}(t) + a_{ij}h_{ij}A_{i,x}(t)}
 \end{align}
 ```
-where $e_{ij}$ is the conversion efficiency from prey $i$ to consumer $j$, and $f_{ij}(A_{i,x}) = \frac{a_{ij}A_{i,x}}{1 + a_{ij}h_{ij}A_{i,x}}$.
+$a_{ij}$ is the attack rate and $h_{ij}$ the handling time. To be realistic, the preference parameter is assumed to depend on the relative potential amount of energy obtained from each prey:
+
+```{=tex}
+\begin{align}
+\delta_{x}(t) = \frac{e_{BP} f_{BP}(A_{B,x}(t))}{e_{BP} f_{BP}(A_{B,x}(t)) + e_{CP} f_{CP}(A_{C,x}(t))}
+\end{align}
+```
+where $e_{ij}$ is the conversion efficiency from prey $i$ to consumer $j$, and $f_{ij}(A_{i,x}(t)) = \frac{a_{ij}A_{i,x}(t)}{1 + a_{ij}h_{ij}A_{i,x}(t)}$.
 An analogous approach has been applied to a model with multiple prey choices [@shotanishijimaRolesAlternativePrey2014].
+Subsequently, a density-dependent recruitment occurs:
 
-The intraguild prey $C$ is controlled by both prey availability and predation.
+```{=tex}
+\begin{align}
+\hat{A}_{B,x}(t) = \frac{r_{b} A_{B,x}'(t)}{1 + \frac{r_b - 1}{K_{x}}A_{B,x}'(t)}
+\end{align}
+```
+$r_b$ is the reproductive rate for survived individuals.
 
-$$
-\bar{A}_{C,x}(t) = e_{BC}s_{0,C}(1-z_{CP,x})z_{BC,x}A_{B,x}
-$$
+The population growth of the intraguild prey is controlled by conversion efficiency, number of prey captured, background mortality, and predation:
+
+```{=tex}
+\begin{equation}
+\hat{A}_{C,x}(t) =
+\left\{
+\begin{aligned}
+&s_{0,C}(1-z_{CP,x})e_{BC}W_{BC,x}(t) &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\
+&s_{0,C}(1-z_{CP,x})\frac{e_{BC}W_{BC,x}(t)}{W_{B,x}(t)}A_{B,x} &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\
+\end{aligned}
+\right.
+\end{equation}
+```
+Equation 11 assumes competition between $C$ and $P$ when their total consumption $W_{B,x}(t)$ exceeds the abundance of basal species $A_{B,x}(t)$, and the relative consumption $\frac{W_{BC,x}(t)}{W_{B,x}(t)}$ rests on $a_{ij}$, $h_{ij}$, and $A_{j,x}(t)$.
+
+The population of the intraguild predator is fueled by two prey species:
+
+```{=tex}
+\begin{equation}
+\hat{A}_{P,x}(t) =
+\left\{
+\begin{aligned}
+&s_{0,P}\left[e_{BP}\delta_{x}(t)W_{BP,x}(t) + e_{CP}(1 - \delta_{x})W_{CP,x}(t)\right] &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\
+&s_{0,P}\left[\frac{e_{BP}\delta_{x}(t)W_{BP,x}(t)}{W_{B,x}(t)}A_{B,x} + e_{CP}(1 - \delta_{x})W_{CP,x}(t)\right] &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\
+\end{aligned}
+\right.
+\end{equation}
+```
+After recruitment, the constituent species experience stochastic regional disturbance, which causes proportional loss of $m$ to populations:
+
+```{=tex}
+\begin{align}
+\bar{A}_{i,x}(t) = [1 - \psi(t)m_{x}]\hat{A}_{i,x}(t)
+\end{align}
+```
+The binary random variable $\psi(t)$ expresses the occurrence of regional disturbance (disturbance occurs if $\psi(t)=1$ otherwise $0$), and $m_x$ is the disturbance intensity at patch $x$ as determined by the network generation process.
+$\psi(t)$ was drawn from a Bernoulli distribution with disturbance probability $p_m$.
+
+***Dispersal.*** We simulated distance-dependent dispersal along a network.
+The number of emigrants from patch x at time t, $E_{i,x}(t)$, is estimated as the product of dispersal probability $p_{d}$ and species abundance $A_{i,x}$.
+The dispersal probability was assumed to be constant for all three species in each simulation run.
+Then, the number of emigrants from patch x is divided into immigrants, Iix(t) Ixy(t) to each patch y.
+The immigration probability at patch x and time t, ix(t) xy ξxy from patch x to y is calculated considering network structure and species dispersal capability as followsbased on the habitat structure among patches according to:
+
+xy(t) =y,y≠xExy(t)e-θdxyxy,y≠xExy(t)e-θdxy
+
+ix(t) = y,yxEiy(t)e-dxyxy,yxEiy(t)e-dxy
+
+Where dxy is the distance between patch x and y, and θ controls the rate of decay. Then, the expected number of immigrants into patch x is calculated as Iix(t)=ix(t)xEix
 

### 2022-03-27 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 \begin{aligned}
-&s_{0,C}(1-z_{CP,x})e_{BC}W_{BC,x}(t) &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\
-&s_{0,C}(1-z_{CP,x})\frac{e_{BC}W_{BC,x}(t)}{W_{B,x}(t)}A_{B,x} &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\
+&s_{0,C} (1 - z_{CP,x}) e_{BC}W_{BC,x}(t) &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\
+&s_{0,C} (1 - z_{CP,x}) e_{BC} \left( \frac{W_{BC,x}(t)}{W_{B,x}(t)} A_{B,x} \right) &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\
 \end{aligned}
@@
 \begin{aligned}
-&s_{0,P}\left[e_{BP}\delta_{x}(t)W_{BP,x}(t) + e_{CP}(1 - \delta_{x})W_{CP,x}(t)\right] &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\
-&s_{0,P}\left[\frac{e_{BP}\delta_{x}(t)W_{BP,x}(t)}{W_{B,x}(t)}A_{B,x} + e_{CP}(1 - \delta_{x})W_{CP,x}(t)\right] &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\
+&s_{0,P} \left[e_{BP} \delta_{x}(t) W_{BP,x}(t) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) \right] &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\
+&s_{0,P} \left[e_{BP} \left(\frac{\delta_{x}(t)W_{BP,x}(t)}{W_{B,x}(t)}A_{B,x}\right) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) \right] &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\
 \end{aligned}
@@
 \begin{align}
-\bar{A}_{i,x}(t) = [1 - \psi(t)m_{x}]\hat{A}_{i,x}(t)
+\bar{A}_{i,x}(t) = [1 - \psi(t) m_{x}] \hat{A}_{i,x}(t)
 \end{align}
@@
 The binary random variable $\psi(t)$ expresses the occurrence of regional disturbance (disturbance occurs if $\psi(t)=1$ otherwise $0$), and $m_x$ is the disturbance intensity at patch $x$ as determined by the network generation process.
-$\psi(t)$ was drawn from a Bernoulli distribution with disturbance probability $p_m$.
+$\psi(t)$ is drawn from a Bernoulli distribution with disturbance probability $p_m$.
 
 ***Dispersal.*** We simulated distance-dependent dispersal along a network.
-The number of emigrants from patch x at time t, $E_{i,x}(t)$, is estimated as the product of dispersal probability $p_{d}$ and species abundance $A_{i,x}$.
-The dispersal probability was assumed to be constant for all three species in each simulation run.
-Then, the number of emigrants from patch x is divided into immigrants, Iix(t) Ixy(t) to each patch y.
-The immigration probability at patch x and time t, ix(t) xy ξxy from patch x to y is calculated considering network structure and species dispersal capability as followsbased on the habitat structure among patches according to:
+The number of emigrants from patch $x$ at time $t$, $E_{i,x}(t)$, is estimated as the product of dispersal probability $p_{d}$ and species abundance after internal dynamics $\bar{A}_{i,x}(t)$.
+The dispersal probability was assumed to be constant for all three species.
+The relative immigration potential $\xi_{i,x}(t)$ is calculated considering network structure and species dispersal capability as follows:
 
-xy(t) =y,y≠xExy(t)e-θdxyxy,y≠xExy(t)e-θdxy
+```{=tex}
+\begin{align}
+\xi_{i,x}(t)=\frac{\sum_{y, y \ne x}E_{i,y}(t)exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}E_{i,y}(t)exp(-\theta d_{xy})}
+\end{align}
+```
+where $d_{xy}$ is the network distance between patch $x$ and $y$ (shortest path), and $\theta$ indicates the inverse of mean dispersal distance (i.e., larger numbers mean shorter dispersal distance).
+The immigration potential at patch $x$, $\sum_{y, y \ne x}E_{i,y}(t)exp(-\theta d_{xy})$, is scaled by the sum of immigration potential across patches so that $\sum_{x} \xi_{i,x}=1$.
+The expected number of immigrants into patch $x$ is calculated as $I_{i,x}(t)=\xi_{i,x}(t) \sum_{x}E_{i,x}(t)$.
 
-ix(t) = y,yxEiy(t)e-dxyxy,yxEiy(t)e-dxy
+## Simulation
 
-Where dxy is the distance between patch x and y, and θ controls the rate of decay. Then, the expected number of immigrants into patch x is calculated as Iix(t)=ix(t)xEix
+We used 24 combinations of parameter sets (**SI Table**).
+The parameter sets include two environmental factors, and twelve ecological scenarios.
+For environmental scenarios, we used two parameter sets that controlled the frequency and magnitude of disturbances (low and high disturbance regime, respectively).
+In the low disturbance regime, the probability of a disturbance occurring at each time step was 0.0001 (i.e., a 10,000 year recurrence), and the mean disturbance magnitude was μm = 0.1 (10% reduction in population densities).
+In contrast, the high disturbance regime had a probability of 0.01 (i.e., 100 year recurrence), and the mean disturbance magnitude was μm = 0.9 (90% reduction in population densities61)
+
+For the ecological scenarios, we used two productivity levels, two dispersal abilities, and three levels of omnivory.
+Productivity was varied by setting the carrying capacity of the basal species, B (K = 50, 150).
+Dispersal ability varies the effective distance between patches and can control the probability that a patch is re-colonized post-disturbance.
+Two levels of dispersal ability were simulated.
+The high dispersal ability was set with parameters Pdispersal = 0.1 and θ = 0.1 (i.e., high probability of individuals emigrating, and far distance traveled).
+The low dispersal scenario was set with parameters Pdispersal = 0.01 and θ = 1  (i.e., low probability of dispersal, short dispersal distance).
+These values were chosen to represent extreme scenarios.
+Preliminary analyses using intermediate parameter sets had responses between these two scenarios.
+
+Finally, we investigated three food web motifs (**Figure 1B**).
+Trophic interaction strengths can be controlled by varying the attack rate or conversion efficiency by the consumer P (Ward and McCann 2017).
+To simplify model simulations, we fixed the attack rate for both C and P (parameters 𝛼 = 8 and 𝛽 = 20), and the conversion efficiency for C as ebc = 6.
+We controlled the presence and strength of trophic interactions for the predator, P, by varying the conversion efficiency for both prey resources (i.e., varied prey resource quality).
+The scenario names and parameter levels were as follows: (1) "Food chain": ebp = 0, ecp = 4; (2) "Strong IGP": ebp = 2, ecp = 4; (3) "Weak IGP": ebp = 4, ecp = 2.
+
+Under each scenario, we simulate 1200 timesteps of trophic dynamics (including 200 steps of burn-in) in 1000 branching networks across a gradient of ecosystem size (number of habitat patches, from 5 to 150) and ecosystem complexity (branching probability, from 0.01 to 0.99).
+Simulations ran for 200 time steps as a burn-in period to minimize influences of initial conditions.
+The last 1000 timesteps were recorded to estimate the temporal average of food chain length across all patches within a network.
+
+\
 

### 2022-03-28 rmd/manuscript.Rmd
edit: manuscript

@@
 bibliography: references.bib
+csl: the-american-naturalist.csl
 header-includes:
@@
 ---
+
+```{r setup, include=FALSE}
+
+library(tidyverse)
+
+knitr::opts_chunk$set(echo = FALSE,
+                      message = FALSE)
+
+# Sensitivity data
+load(file = here::here("output/result_main.RData"))
+
+n_rep <- max(sim_main_result$n_rep)
+n_scenario <- sim_main_result %>% 
+  filter(sd_disturb_source > sd_disturb_lon) %>% 
+  summarize(n = n_distinct(param_set)) %>% 
+  pull(n)
+n_sim <- sim_main_result %>% 
+  mutate(n_sim = n_timestep + n_burnin + n_warmup) %>% 
+  pull() %>% 
+  unique()
+```
 
@@
 
-## **Network Generation**
+## **Network generation**
 
@@
 Full details of the network generation process is described in @teruiEmergentDualScaling2021.
-For each ecological scenario described below, we generated 1000 networks across a gradient of ecosystem size and complexity by drawing $N_{p}$ and $P_{b}$ from uniform distributions as: $N_{p} \sim Unif(10, 150)$ and $P_{b} \sim Unif(0.01, 0.99)$ ($N_{p}$ was rounded to the nearest integer before running simulations).
 
@@
 where $e_{ij}$ is the conversion efficiency from prey $i$ to consumer $j$, and $f_{ij}(A_{i,x}(t)) = \frac{a_{ij}A_{i,x}(t)}{1 + a_{ij}h_{ij}A_{i,x}(t)}$.
-An analogous approach has been applied to a model with multiple prey choices [@shotanishijimaRolesAlternativePrey2014].
+An analogous approach has been applied to a model with multiple prey choices [@nishijimaRolesAlternativePrey2014].
 Subsequently, a density-dependent recruitment occurs:
@@
 
+***Food chain length.*** At each patch and time step, food chain length was determined as the highest trophic position ($TP_{x}(t)$) of the existing species.
+When the intraguild predator exists, the trophic position was determined as follows:
+
+```{=tex}
+\begin{align}
+TP_{P,x}(t) = \delta_{x}(t) + 2(1 - \delta_{x}(t)) + 1
+\end{align}
+```
+This definition accounts for the influence of omnivory.
+
 ## Simulation
 
-We used 24 combinations of parameter sets (**SI Table**).
-The parameter sets include two environmental factors, and twelve ecological scenarios.
-For environmental scenarios, we used two parameter sets that controlled the frequency and magnitude of disturbances (low and high disturbance regime, respectively).
-In the low disturbance regime, the probability of a disturbance occurring at each time step was 0.0001 (i.e., a 10,000 year recurrence), and the mean disturbance magnitude was μm = 0.1 (10% reduction in population densities).
-In contrast, the high disturbance regime had a probability of 0.01 (i.e., 100 year recurrence), and the mean disturbance magnitude was μm = 0.9 (90% reduction in population densities61)
+Prior to the main simulation, we performed an extensive sensitivity analysis to identify influential parameters (see **Supplementary Text** and **Table S2-S3**).
+Our sensitivity analysis revealed that ecosystem size and/or complexity effects were sensitive to disturbance, productivity, dispersal, and omnivory.
+Therefore, we ran the main simulation with `r n_scenario` parameter sets differing in these ecological contexts (two levels of disturbance, two levels of productivity, two levels of dispersal distance, and three levels of omnivory).
 
-For the ecological scenarios, we used two productivity levels, two dispersal abilities, and three levels of omnivory.
-Productivity was varied by setting the carrying capacity of the basal species, B (K = 50, 150).
-Dispersal ability varies the effective distance between patches and can control the probability that a patch is re-colonized post-disturbance.
-Two levels of dispersal ability were simulated.
-The high dispersal ability was set with parameters Pdispersal = 0.1 and θ = 0.1 (i.e., high probability of individuals emigrating, and far distance traveled).
-The low dispersal scenario was set with parameters Pdispersal = 0.01 and θ = 1  (i.e., low probability of dispersal, short dispersal distance).
-These values were chosen to represent extreme scenarios.
-Preliminary analyses using intermediate parameter sets had responses between these two scenarios.
+We controlled disturbance level with disturbance probability $p_{m}$ (low disturbance, `r min(sim_main_result$p_disturb)`; high disturbance, `r max(sim_main_result$p_disturb)`).
+We used the identical value of mean disturbance intensity at headwaters ($\mu_{m}=$ `r max(sim_main_result$mean_disturb_source)`), spatial variation in disturbance intensity at headwaters ($\sigma_{h}=$ `r min(sim_main_result$sd_disturb_source)`), and local disturbance variation ($\sigma_{l}=$ `r unique(sim_main_result$sd_disturb_lon)`) regardless of disturbance scenario.
+Productivity was varied by setting different values of basal species' growth rate ($r_{b}=$ `r unique(sim_main_result$r_b)`).
+For dispersal, we examined short- and long-distance dispersal scenarios (short-distance dispersal $\theta=$ `r max(sim_main_result$theta)`; long-distance dispersal $\theta=$ `r min(sim_main_result$theta)`), while using constant dispersal probability ($p_{d}=$ `r unique(sim_main_result$p_dispersal)`).
 
-Finally, we investigated three food web motifs (**Figure 1B**).
-Trophic interaction strengths can be controlled by varying the attack rate or conversion efficiency by the consumer P (Ward and McCann 2017).
-To simplify model simulations, we fixed the attack rate for both C and P (parameters 𝛼 = 8 and 𝛽 = 20), and the conversion efficiency for C as ebc = 6.
-We controlled the presence and strength of trophic interactions for the predator, P, by varying the conversion efficiency for both prey resources (i.e., varied prey resource quality).
-The scenario names and parameter levels were as follows: (1) "Food chain": ebp = 0, ecp = 4; (2) "Strong IGP": ebp = 2, ecp = 4; (3) "Weak IGP": ebp = 4, ecp = 2.
+Finally, we investigated three levels of omnivory (**Figure 1D**).
+Trophic interaction strengths can be controlled by varying the attack rate and/or conversion efficiency [@wardMechanisticTheoryAquatic2017].
+We changed attack rate ($a_{BP}$ and $a_{CP}$) and conversion efficiency ($e_{BP}$ and $e_{CP}$) of two prey items for intraguild predator $P$.
+Specifically, we used the following combinations to represent three levels of omnivory: "Chain", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[1], max(sim_main_result$e_cp))` $\}$; "Weak omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[2], max(sim_main_result$e_cp))` $\}$; and "Strong omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(max(sim_main_result$a_bp), min(sim_main_result$a_cp), unique(sim_main_result$e_bp)[3], min(sim_main_result$e_cp))` $\}$.
+To simplify simulation setup, we fixed the following parameters: $e_{BC}=$ `r unique(sim_main_result$e_bc)` , $a_{BC}=$ `r unique(sim_main_result$a_bc)`, and $h_{ij}=$ `r unique(sim_main_result$h_bc)`.
 
-Under each scenario, we simulate 1200 timesteps of trophic dynamics (including 200 steps of burn-in) in 1000 branching networks across a gradient of ecosystem size (number of habitat patches, from 5 to 150) and ecosystem complexity (branching probability, from 0.01 to 0.99).
-Simulations ran for 200 time steps as a burn-in period to minimize influences of initial conditions.
-The last 1000 timesteps were recorded to estimate the temporal average of food chain length across all patches within a network.
-
-\
+Under each scenario, we ran `r max(sim_main_result$n_rep)` simulation replicates with `r n_sim` timesteps of trophic dynamics (including `r n_sim - unique(sim_main_result$n_timestep)` steps of warmup and burn-in).
+During the warmup period (`r unique(sim_main_result$n_warmup)` time steps), we introduced propagules into habitat patches every 10 time steps to allow species to establish populations.
+Propagules are random draws from $Poisson(0.5)$.
+Subsequently, we ran a burn-in period with no propagules for `r unique(sim_main_result$n_burnin)` time steps to remove the effects of initial conditions.
+We obtained the mean food chain length across all habitat patches for the last `r unique(sim_main_result$n_timestep)` time steps.
+Note that simulation replicates are differ in ecosystem size ($N_{p}\sim Unif(10,150)$) and ecosystem complexity ($P_{b} \sim Unif(0.01, 0.99)$) so we can estimate their effects on food chain length.
 
@@
 
+Our simulation highlighted the interactive nature of putative controls of food chain length.
+In particular, disturbance regimes had the strongest influence on the qualitative relationships between FCL and ecosystem properties.
+Below, we report our results for low and high disturbance regimens separately.
+
+## **Low disturbance regime**
+
+In ecosystems with a low disturbance regime, FCL was largely determined by omnivory with limited effects of ecosystem size and complexity (colored lines in **Figure 2**).
+Unsurprisingly, FCL was negatively related to the degree of omnivory, with no-omnivory food chains ($P$ only consumes $C$) having the longest FCL, and strong omnivory ($P$ consumes very little $C$) having the shortest FCL.
+These trends were consistent across dispersal scenarios (**Figure 2**).
+
+Although the effects of ecosystem properties on FCL were minimal, there were noticeable effects on the occupancies of constituent species in some scenarios.
+For example, the occupancies of $C$ and $P$ increased with ecosystem size when omnivory was strong (see the most left column in **Figures S1 and S2**).
+However, these changes were not translated into changes in FCL due to the structural constraint imposed by omnivory ($C$ and $P$ occupy a similar trophic position; **Figure 1D**).
+Similarly, the occupancy of $P$ decreased with ecosystem complexity when omnivory was strong, but FCL remained almost the same as the increase of $C$ could compensate for the loss of $P$ for the same reason (**Figure S3 and S4**).
+
+## **High disturbance regime**
+
+A high disturbance regime produced much different responses of FCL.
+FCL increased across both gradients of ecosystem size and complexity under all combinations of productivity, the degree of omnivory, and dispersal distance (**Figure 3**).
+However, the effects of ecosystem size and complexity were mediated by ecological factors differently.
+Ecosystem size effects were stronger with longer dispersal distance ($\theta = 0.1$), while the opposite was true for ecosystem complexity (**Figure 3**).
+In addition, the effect of ecosystem complexity was pronounced in food webs with weaker omnivory (**Figure 3B**).
+
+Interestingly, the positive effects of ecosystem size and complexity were driven by different mechanisms.
+For ecosystem size, the underlying mechanism was consistent when productivity was high; increased occupancies of all the three species resulted in longer food chains.
+However, when productivity was low, the mechanisms were more complex, particularly with strong omnivory (the third column in **Figure S1 and S2**).
+In this scenario, the occupancies of $B$ and $C$ decreased at larger ecosystem sizes, while $P$ increased its occupancy (**Figure S1 and S2**).
+Therefore, $P$ compensated for the loss of $B$ and $C$ to maintain the positive relationship between FCL and ecosystem size.
+For ecosystem complexity, all the three species increased with ecosystem complexity when productivity was high, translating into longer food chains in more complex ecosystems (**Figure S3 and S4**).
+However, when food webs contained strong omnivory with low productivity (the third column in **Figure S3 and S4**), the increased occupacy of $C$ maintained the positive relationship between FCL and ecosystem complexity by offsetting the loss of $P$.
+Hence, the contrasting mechanisms underlie the positive effects of ecosystem size and complexity on FCL.
+
 # Discussion
+
+Discussion to be added
+
+# References

### 2022-03-28 rmd/suppl_info.Rmd
edit: manuscript

@@
 
-Based on the results, we considered ecological scenarios relevant for putative controls of food chain: low and high productivity ($r_b = 4, 8$), low and high disturbance ($p_b = 0.01, 0.1$), and three food web architecture ("Chain" [$\{e_{bc}, e_{bp}, e_{cp}\} = \{4, 0, 4\}, \{a_{bc}, a_{bp}, a_{cp}\} = \{0.5, 0.1, 0.5\}$], "Weak omnivory" [$\{e_{bc}, e_{bp}, e_{cp}\} = \{4, 2, 4\}, \{a_{bc}, a_{bp}, a_{cp}\} = \{0.5, 0.1, 0.5\}$], and "Strong omnivory" [$\{e_{bc}, e_{bp}, e_{cp}\} = \{4, 4, 2\}, \{a_{bc}, a_{bp}, a_{cp}\} = \{0.5, 0.5, 0.1\}$]).
+Based on the results, we considered ecological scenarios relevant for putative controls of food chain: low and high productivity ($r_b = 4, 8$), low and high disturbance ($p_m = 0.01, 0.1$), and three food web architecture ("Chain" [$\{e_{BC}, e_{BP}, e_{CP}\} = \{4, 0, 4\}, \{a_{BC}, a_{BP}, a_{CP}\} = \{0.5, 0.1, 0.5\}$], "Weak omnivory" [$\{e_{BC}, e_{BP}, e_{CP}\} = \{4, 2, 4\}, \{a_{BC}, a_{BP}, a_{CP}\} = \{0.5, 0.1, 0.5\}$], and "Strong omnivory" [$\{e_{BC}, e_{BP}, e_{CP}\} = \{4, 4, 2\}, \{a_{BC}, a_{BP}, a_{CP}\} = \{0.5, 0.5, 0.1\}$]).
 Handling times were fixed at $0.5$ to reduce the number of parameter combinations.

### 2022-03-29 rmd/figures.Rmd
edit: figures

---
title: "Figures"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE)
```

```{r figure1, fig.width=12, fig.height=6}
source(here::here("code/figure_concept.R"))
print(g_concept)
```

**Figure 1** (A) Example of simulated branching networks. Node color is reflective of the disturbance value calculated in network generation. (B) Distribution of disturbance values in branching networks. (C) Trophic interactions in a three-species community. $B$ refers to a basal species whose population level is controlled by a density-dependent function. The consumer species (or intraguild prey), $C$, consumes $B$ as a resource. The predator (or intraguild predator), $P$, is able to consume both $B$ and $C$. The structure and strength of these trophic interactions determined by functional responses. The strength of trophic interactions involving $P$ is represented by the darkness of the arrows (strength of the $B$-$C$ interaction is fixed), and food chain length varies in response to the proportion that each resource makes to the diet of $P$. On the far left, a simple linear food chain exists when the predator does not consume the basal species. Moving to the right, basal species' biomass is converted to predator biomass at higher efficiencies with greater attack rates. (D) Preference of predator $P$ to basal species $B$ over consumer $C$ under different omnivory scenarios. The preference changes with the relative density of $B$ to $C$ ($C=50$ in this example), and the rate of change is determined by attack rate and handling time.

\pagebreak

```{r figure2, fig.width=12, fig.height=5.5}
source(here::here("code/figure_main_sim.R"))
print(g_ld)
```

**Figure 2** Ecosystem size (A) and complexity (B) have little influence on food chain length under low disturbance regime. Lines are loess curves fitted to simulated data and are colored based on omnivory. Shaded ribbons are 95% confidence intervals. Parameters for low disturbance regime: disturbance probability $p_d = 0.01$ and mean proportional mortality at headwaters $\mu_{m}=0.8$.

\pagebreak

```{r figure3, fig.width=12, fig.height=5.5}
source(here::here("code/figure_main_sim.R"))
print(g_hd)
```

**Figure 3** Ecosystem size (A) and complexity (B) influence food chain length under high disturbance regime. Lines are loess curves fitted to simulated data and are colored based on omnivory. Shaded ribbons are 95% confidence intervals. Parameters for high disturbance regime: disturbance probability $p_{d} = 0.01$ and mean proportional mortality at headwaters $\mu_{m}=0.8$.

### 2022-03-29 code/figure_delta.R

@@ -78,5 +78,5 @@ g_delta <- df_m %>%
              linetype = omn)) +
   geom_line(color = grey(0.4)) +
   labs(x = "Basal species density",
-       y = expression("Preference to basal over IG-prey"~~(delta)),
+       y = expression("Preference to B over C"~~(delta)),
        linetype = "Omnivory")

### 2022-03-29 code/figure_main_sim.R

@@ -12,8 +12,7 @@ load(file = here::here("output/result_main.RData"))
 
 df_sim <- sim_main_result %>% 
   group_by(param_set) %>% 
-  filter(theta == 1,
-         sd_disturb_source == 3,
+  filter(sd_disturb_source == 3,
          sd_disturb_lon == 0.1) %>% 
   mutate(omn = case_when(e_bp == 0 ~ "Chain",
                          e_bp == 2 ~ "Weak",
@@ -34,56 +33,74 @@ df_sim <- sim_main_result %>%
                           `0.1` = sprintf('"High disturbance"~(p[m]=="%.2f")',
                                           p_disturb)),
          disturb = factor(disturb,
-                          levels = rev(levels(factor(disturb))))
-         )
+                          levels = rev(levels(factor(disturb)))),
+         dispersal = recode(theta,
+                            `0.1` = sprintf('"Long dispersal"~(theta=="%.2f")',
+                                            theta),
+                            `1` = sprintf('"Short dispersal"~(theta=="%.2f")',
+                                          theta)),
+         
+  )
 
 
 # figure ------------------------------------------------------------------
 
 theme_set(plt_theme)
+param <- c(0.01, 0.1)
 
-g_np <- df_sim %>% 
-  ggplot(aes(x = n_patch,
-             y = fcl,
-             color = omn,
-             fill = omn)) +
-  geom_smooth(method = "loess",
-              size = 0.5) +
-  facet_grid(rows = vars(disturb),
-             cols = vars(productivity),
-             labeller = label_parsed) +
-  labs(x = "Ecosystem size (number of patches)",
-       y = "Food chain length",
-       color = "Omnivory",
-       fill = "Omnivory") +
-  MetBrewer::scale_color_met_d("Hiroshige") +
-  MetBrewer::scale_fill_met_d("Hiroshige")
+## Ecosystem size
+g_np_list <- foreach(i = seq_len(length(param))) %do% {
+  df_sim %>% 
+    filter(p_disturb == param[i]) %>% 
+    ggplot(aes(x = n_patch,
+               y = fcl,
+               color = omn,
+               fill = omn)) +
+    geom_smooth(method = "loess",
+                size = 0.5) +
+    facet_grid(rows = vars(dispersal),
+               cols = vars(productivity),
[…]

### 2022-03-29 code/figure_network.R

@@ -40,13 +40,13 @@ ng <- foreach(i = seq_len(length(net))) %do% {
                                    flip.y = FALSE,
                                    root = 1)) +
     geom_edge_link(color = "steelblue",
-                   edge_width = 0.01) +
+                   edge_width = 0.1) +
     geom_node_point(aes(fill = disturb),
                     shape = 21,
                     color = grey(0.5),
                     stroke = 0.1,
                     size = 2) +
-    MetBrewer::scale_fill_met_c("Hiroshige") +
+    MetBrewer::scale_fill_met_c("Hiroshige", direction = -1) +
     #scale_fill_continuous(high = "steelblue",
     #                      low = grey(0.95)) +
     guides(fill = ifelse(i == 3, guide_legend(), "none")) +

### 2022-03-29 rmd/manuscript.Rmd
edit: manuscript

@@
 n_rep <- max(sim_main_result$n_rep)
+
 n_scenario <- sim_main_result %>% 
@@
   pull(n)
+
 n_sim <- sim_main_result %>% 
@@
 # Introduction
+
+The origin of variation in food chain length (FCL), the number of feeding links from the basal species to top predators, has intrigued ecologists for decades due to its importance in regulating top-down effects [@paceTrophicCascadesRevealed1999], energy flow [@wangBiodiversityEcosystemFunctioning2018], and contaminant concentrations in top predators that humans often consume [@kiddEffectsTrophicPosition1998].
+Hypotheses regarding controls of food chain length can be organized into three broad categories.
+First, the resource availability hypothesis states that basal productivity controls food chain length because imperfect energy conversion through predator-prey interactions restricts energy available to higher trophic levels [@oksanenExploitationEcosystemsGradients1981].
+Second, the disturbance hypothesis (or the dynamical stability hypothesis) predicts longer food chains are vulnerable to environmental perturbations (i.e., the dynamical constraint).
+Thus, habitats with frequent and/or severe disturbance should support fewer trophic levels [@pimmNumberTrophicLevels1977].
+Lastly, the ecosystem size hypothesis posits that larger ecosystems should harbor longer food chains [@takimotoEffectsProductivityDisturbance2012; @teruiSpatialDisturbanceSynchrony2019; @postEcosystemSizeDetermines2000].
+
+Among the three hypotheses, the ecosystem size hypothesis has been consistently supported in empirical explorations, likely because multiple mechanisms may underlie the positive association [@takimotoEnvironmentalDeterminantsFoodchain2013].
+For example, large ecosystems have increased species richness (species-area relationship [@lomolinoEcologyMostGeneral2000]), buffer against disturbances [@saboRoleDischargeVariation2010], and have greater numbers of immigrant sources in the ecosystem with stable metapopulations [@takimotoEnvironmentalDeterminantsFoodchain2013].
+Meanwhile, the theoretical basis for the other two hypotheses rest on implicit assumptions of linear food chains, which are often violated in natural systems.
+As such, the effects of disturbance and resource availability on food chain length have been found to be highly variable among ecosystem types and geographic regions [@takimotoEffectsProductivityDisturbance2012; @diehlIntraguildPreySuffer2001; @takimotoEnvironmentalDeterminantsFoodchain2013; @postLongShortFoodchain2002].
+
+More recently, food web ecologists have begun to recognize that FCL rarely responds to a single driver, but instead is determined by interactions of multiple factors.
+For example, @wardMechanisticTheoryAquatic2017 showed that FCL was more responsive to increasing productivity in large ecosystems compared with small ecosystems, whereas FCL was more sensitive to size in low productivity systems.
+Such context dependent responses to extrinsic factors seem to be tightly coupled with the arrangement of trophic links within a food web.
+For example, omnivory (feeding at multiple trophic levels) is a strong intrinsic driver of food chain length [@takimotoEffectsProductivityDisturbance2012; @postProximateStructuralMechanisms2007; @woottonOmnivoryStabilityFreshwater2017].
+When omnivory is absent in three-species communities, the trophic structure assumes a simple linear food chain with maximum length equal to 3 (**Figure 1D**).
+However, as omnivory increases, FCL decreases proportionally as the top predator eats lower in the food web [@wardMechanisticTheoryAquatic2017; @woottonOmnivoryStabilityFreshwater2017].
+Theoretical work predicts that food webs are destabilized when omnivory is prevalent, leading to the loss of species and shorter food chains even in productive, large ecosystems.
+Conversely, weak omnivory stabilize the food web and increase persistence of species within a community, particularly at intermediate levels of productivity and high disturbance levels [@woottonOmnivoryStabilityFreshwater2017; @woottonManyWeakInteractions2016].
+Due to the context-dependency of FCL to multiple drivers, and the potential for interactive effects, a unified framework of multiple food chain drivers is needed to better understand when and where FCL will respond to these extrinsic and intrinsic factors.
+
+While recent research efforts of food web ecologists have substantially advanced our understanding of food chain drivers, most of the previous work has been conducted in simple, two-dimensional systems, such as lakes [@postEcosystemSizeDetermines2000; @doiResourceAvailabilityEcosystem2009] and oceanic islands [@takimotoEcosystemSizeNot2008; @pringlePredatorinducedCollapseNiche2019].
+However, many natural habitats have a high degree of spatial complexity that cannot be represented simply by ecosystem size.
+Branching ecosystems, for example, are ubiquitous yet overlooked landscape structures (e.g., rivers, trees, caves, mountain ranges) [@grantLivingBranchesPopulation2007; @altermattDiversityRiverineMetacommunities2013].
+While we know that food web interactions are spatially influenced [@calcagnoConstraintsFoodChain2011; @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012; @teruiSpatialDisturbanceSynchrony2019], the theory to account for it has lagged, particularly in branching habitat architectures.
+Here, we theoretically explore drivers of food chain length in rivers, an excellent example of branching ecosystems.
+
+Ample evidence suggests that branching structure controls environmental heterogeneity and dispersal pathways of organisms within river ecosystems [@altermattDiversityRiverineMetacommunities2013; @tonkinRoleDispersalRiver2018].
+Water and materials propagate downstream as small streams hierarchically join to form larger streams and rivers.
+Consequently, environmental signals are highly variable among tributaries while showing a strong within-tributary correlation e.g., flood disturbance [@nakamuraDisturbanceRegimesStream2000; @isaakSlowClimateVelocities2016; @mcguireNetworkAnalysisReveals2014; @abbottUnexpectedSpatialStability2018].
+Confluences serve as merging points where tributary dynamics aggregate to form distinct environmental conditions in downstream habitat patches.
+Habitat heterogeneity therefore increases with branching complexity (**Figure 1A**), which we define here as the probability of branching per unit river distance.
+Hence, highly-branched river networks provide diverse habitats that buffer the impact of large-scale environmental fluctuations, broadening opportunities for species to recolonize habitat patches post-disturbance [@teruiMetapopulationStabilityBranching2018].
+Thus, ecosystem complexity in river networks may be a widespread yet overlooked extrinsic environmental driver which needs to be incorporated to understand food chain length.
+
+We develop a model in which trophic community dynamics within branching habitats are simulated.
+We investigate extrinsic controls such as ecosystem size (number of patches), complexity (branching probability), and disturbance regime.
+In addition, our simulation model incorporates intrinsic drivers of food chain length including food web structure, productivity level, and dispersal ability, allowing us to investigate the context dependency on drivers of food chain length across extrinsic factors.
+We find that ecosystem size and complexity are both positively related to FCL, but the form of this relationship is dependent on other extrinsic and intrinsic factors.
+Our model is an important step in unifying drivers of FCL across extrinsic and intrinsic drivers and parsing their context-dependency in spatially complex ecosystems.
 

### 2022-03-30 rmd/manuscript.Rmd
edit: manuscript

@@
 Unsurprisingly, FCL was negatively related to the degree of omnivory, with no-omnivory food chains ($P$ only consumes $C$) having the longest FCL, and strong omnivory ($P$ consumes very little $C$) having the shortest FCL.
+In addition, productivity played a secondary role in increasing FCL, particularly when omnivory was strong (compare left and right panels in **Figure 2**).
+Dispersal had no apparent effects on FCL (compare top and bottom panels in **Figure 2**).
 These trends were consistent across dispersal scenarios (**Figure 2**).
@@
 
-A high disturbance regime produced much different responses of FCL.
+On average, increased disturbance frequency caused the dominance of short food chains (**Figures 2 and 3**).
+More importantly, however, a high disturbance regime produced much different responses of FCL to ecosystem properties.
 FCL increased across both gradients of ecosystem size and complexity under all combinations of productivity, the degree of omnivory, and dispersal distance (**Figure 3**).
-However, the effects of ecosystem size and complexity were mediated by ecological factors differently.
-Ecosystem size effects were stronger with longer dispersal distance ($\theta = 0.1$), while the opposite was true for ecosystem complexity (**Figure 3**).
-In addition, the effect of ecosystem complexity was pronounced in food webs with weaker omnivory (**Figure 3B**).
+The effect of ecosystem complexity was generally clearer than that of ecosystem size.
+It is important to note that the positive relationships emerged in a high disturbance regime because smaller and simpler ecosystems were more vulnerable to increased disturbance frequency (compare far left in **Figures 2 and 3**).
+
+Dispersal buffered the reduction of FCL due to increased disturbance, but it influenced FCL differently across the gradients of ecosystem structure.
+The greater buffering effect of dispersal emerged in larger ecosystems (compare top and bottom panels in **Figure 3**) because greater dispersal capability is required to reach distant patches.
+In the meantime, simpler ecosystems benefitted more from increased dispersal because habitat patches are more isolated from one another in linearly arranged networks (**Figure 3**).
+Productivity increased FCL as observed in a low disturbance regime, but it did not modulate the qualitative relationships between FCL and ecosystem properties (compare left and right panels in **Figure 3**).
 
 Interestingly, the positive effects of ecosystem size and complexity were driven by different mechanisms.
-For ecosystem size, the underlying mechanism was consistent when productivity was high; increased occupancies of all the three species resulted in longer food chains.
+For ecosystem size, the underlying mechanism was consistent when productivity was high; increased occupancies of all the three species resulted in longer food chains (i.e., "additive mechanism").
 However, when productivity was low, the mechanisms were more complex, particularly with strong omnivory (the third column in **Figure S1 and S2**).
-In this scenario, the occupancies of $B$ and $C$ decreased at larger ecosystem sizes, while $P$ increased its occupancy (**Figure S1 and S2**).
+In this scenario, the occupancies of $B$ and $C$ decreased at larger ecosystem sizes, while $P$ increased its occupancy ("insertion mechanism"; **Figure S1 and S2**).
 Therefore, $P$ compensated for the loss of $B$ and $C$ to maintain the positive relationship between FCL and ecosystem size.
@@
 However, when food webs contained strong omnivory with low productivity (the third column in **Figure S3 and S4**), the increased occupacy of $C$ maintained the positive relationship between FCL and ecosystem complexity by offsetting the loss of $P$.
-Hence, the contrasting mechanisms underlie the positive effects of ecosystem size and complexity on FCL.
+Hence, the contrasting structural mechanisms underlie the positive association of ecosystem size and complexity on FCL.
 
@@
 
-Discussion to be added
+The ecosystem size hypothesis of food chain length has received support from both theoretical and empirical research [@vanderzandenPatternsFoodChain1999; @postEcosystemSizeDetermines2000; @takimotoEffectsProductivityDisturbance2012], although empirical work so far has rarely elaborated the underlying drivers [but see @saboRoleDischargeVariation2010; @wardMechanisticTheoryAquatic2017].
+In contrast, while the importance of ecosystem complexity is increasingly recognized in ecology [@grantLivingBranchesPopulation2007; @altermattDiversityRiverineMetacommunities2013; @teruiMetapopulationStabilityBranching2018; @teruiEmergentDualScaling2021], it has not been incorporated in theoretical explorations of food chain length to date.
+Here, we show that food chain length responds positively to both increasing ecosystem size and complexity.
+This supports our hypothesis that ecosystem complexity may represent an overlooked yet important extrinsic control of food chain length. However, the relationship between FCL and ecosystem complexity varied based on additional extrinsic and intrinsic factors, supporting the emerging paradigm that there is no one single driver of FCL in ecosystems [@mchughDualInfluencesEcosystem2010; @wardMechanisticTheoryAquatic2017].
+Although we focus on river networks here, we expect our results to be transferable to other branching structures, such as mountain ranges and cave systems.
+
+While increased disturbance frequency consistently reduced food chain length as predicted by dynamical stability hypothesis [@pimmNumberTrophicLevels1977], more important findings in this study were evident from their synergistic effects with ecosystem structure.
+Our low disturbance regime scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independent of ecosystem size and complexity.
+Consequently, FCL was predominantly determined by food web structure.
+However, under high disturbance regimes, only large and complex networks could provide diverse and heterogeneous habitats with potential refugia, allowing species to thrive and maintain long food chains.
+While empirical studies rarely looked at the influence of ecosystem structure on FCL, several studies provide some support for this theoretical prediction.
+For example, @sullivanEcosystemStructureEmerges2015 reported FCL is controlled parimarily by the number of tributary confluences in Idaho streams.
+Similarly, it has been shown that the merging of major tributaries lengthened FCL in the Colorado River, especially during a post-monsoon turbid water phase [@saboPulsedFlowsTributary2018].
+Clearly, more empirical research is needed to confirm our theoretical predictions.
+
+Although dispersal has received little attention in food chain research, our simulation highlighted its importance in regulating food chain length.
+In particular, high dispersal was required for large, linearly-arranged ecosystems to maintain long food chains under high disturbance regimes.
+This is probably because sufficient propagules are needed to successfully recolonize after disturbance.
+Human activities that affect habitat connectivity and dispersal ability of species within riverine networks may compromise the buffering effects of ecosystem size and complexity against disturbance by limiting dispersing propagules.
+For example, dams and diversion have altered the connectivity between suitable habitat patches, or have completely blocked dispersal ability 46,47.
+Light pollution has also negatively affected the ability of adult aquatic insects to disperse between habitat patches 48.
+
+The interactive effect of disturbance and ecosystem structure has important implications for ecosystem management because anthropogenic climate change is predicted to increase the frequency and magnitude of disturbances [@brunnerFutureStreamflowRegime2020; @mittalImpactHumanIntervention2016].
+River networks are especially vulnerable due to their propensity to go dry during droughts [@zipperPervasiveChangesStream2021] as well as susceptibility to more intense flood disturbance with increasing storm intensity [@schiermeierIncreasedFloodRisk2011].
+However, human activities have compromised the habitat heterogeneity and structural connectivity within riverine networks [@ishiyamaPredictingEcologicalImpacts2018; @moritaEffectsHabitatFragmentation2002; @barbarossaImpactsCurrentFuture2020].
+Our theory implies that these human-induced alterations that simplfy river networks may make riverine food webs more vulnerable to the anticipated environmental changes.
+
+The context dependency of disturbance, productivity and dispersal has important implications for the conservation and management of freshwater ecosystems.
+This adds an important distinction in the discussion of the suitability of assigning protection or conservation status to regions.
+Current conservation schemes generally emphasize size, aiming to protect the largest areas possible.
+While this is certainly a useful attribute generally, designating regions with a high degree of variability, or many small patches, may be more beneficial than a large, homogenous area 49, and accounting for ecosystem complexity is particularly suited to ecosystem architectures in which branches have variable environmental conditions.
+Conservation and management activities should focus on activities that increase the habitat heterogeneity and connectivity of suitable habitat patches to increase resilience to predicted increases in disturbance frequency and magnitude.
+
+The role of food web motifs on FCL largely followed our predictions (**Figure 1**).
+Specifically, FCL declined as the value of the resource B, and thus the proportion in the diet, to the predator P increased.
+Essentially, the relative qualitative values of the resources to the predator controlled the level of omnivory and FCL.
+Although FCL was greatest when the community was arranged as a simple linear food chain, this arrangement was not the most stable.
+Specifically, when disturbance regime was high, and productivity and dispersal ability were low, the observed FCL across networks was \~1 (although increasing dispersal ability and/or carrying capacity resulted in food chains of \~2.5-3).
+Therefore, in the absence of prey resource choice, strong, top-down trophic interactions were destabilizing, leading to the extirpation of consumer species at higher trophic levels across the networks.
+
+Changes in community composition are occurring worldwide and in all habitats.
+Species extirpations and extinctions, coupled with species introductions and invasions are creating novel communities worldwide.
+In Lake Tahoe (USA), for example, replacement of the top fish predator and the introduction of a freshwater shrimp substantially altered the food web structure 50.
+Understanding the interactions within these novel communities, and how this could affect food chain lengths is an important goal of contemporary ecologists.
+The strength of omnivory in these novel communities can be increased or decreased, depending on the introduced species and the extant species pool or community.
+For example, non-indigenous fishes in the Eastern Mediterranean Sea outcompete native species for high quality prey items, causing native species to increase consumption of non-preferred prey items51.
+Introductions of non-native fishes in the Colorado River basin have led to longer FCL and narrower niches (e.g., lower degree of omnivory) of native species52.
+Alternatively, the introduction of hyper-successful basal or primary consumer species can weaken or remove the omnivory link by becoming the dominant prey resource.
+For example, nonnative New Zealand mud snail dominated the macroinvertebrate biomass in streams in Yellowstone National Park and likely affected food web structure53.  Having a single prey resource dominate the community is akin to reducing the degree of omnivory (i.e., removing prey choice), which destabilizes the community, particularly under high disturbance regimes.
+
+It is abundantly clear that ecosystem complexity in natural systems is the rule, not the exception.
+Early theoretical and empirical work has assumed simple habitat structure as a necessity.
+However, habitat heterogeneity, such as the presence of prey refugia, and dispersal between connected habitat patches allow communities to stabilize and persist through space and time 54--56.
+Complex ecosystem structure is an extrinsic factor that can affect food chain length in communities.
+Incorporating a realistic understanding of ecosystem structural complexity could alter our perception of communities in naturally complex systems, potentially leading to paradigm shifts in macroecology.
 

### 2022-03-30 code/figure_stvy.R
edit: figure

@@ -66,7 +66,7 @@ g_pb <- df_r %>%
   #           col = "blue") +
   geom_smooth(method = "lm") +
   facet_wrap(facets = ~param,
-             scales = "free",
+             scales = "free_x",
              ncol = 3) +
   labs(y = "Spearman corr. between FCL and P_b") +
   theme_bw()

### 2022-03-31 code/table_param.R
edit: table

@@ -41,7 +41,7 @@ df_value <- sim_main_result %>%
                                param == "h_cp" ~"$h_{cp}$",
                                param == "theta" ~"$\\theta$",
                                param == "p_disturb" ~"$p_m$",
-                               param == "mean_disturb_source" ~ "$m$",
+                               param == "mean_disturb_source" ~ "$\\mu_m$",
                                param == "sd_disturb_source" ~ "$\\sigma_h$",
                                param == "sd_disturb_lon" ~ "$\\sigma_l$",
                                param == "s0" ~ "$s_0$",

### 2022-03-31 rmd/manuscript.Rmd
edit: manuscript

--- 2022-03-29 rmd/manuscript.Rmd
@@
 The origin of variation in food chain length (FCL), the number of feeding links from the basal species to top predators, has intrigued ecologists for decades due to its importance in regulating top-down effects [@paceTrophicCascadesRevealed1999], energy flow [@wangBiodiversityEcosystemFunctioning2018], and contaminant concentrations in top predators that humans often consume [@kiddEffectsTrophicPosition1998].
-Hypotheses regarding controls of food chain length can be organized into three broad categories.
-First, the resource availability hypothesis states that basal productivity controls food chain length because imperfect energy conversion through predator-prey interactions restricts energy available to higher trophic levels [@oksanenExploitationEcosystemsGradients1981].
+Hypotheses regarding controls of FCL can be organized into three broad categories.
+First, the resource availability hypothesis states that basal productivity controls FCL because imperfect energy conversion through predator-prey interactions restricts energy available to higher trophic levels [@oksanenExploitationEcosystemsGradients1981].
 Second, the disturbance hypothesis (or the dynamical stability hypothesis) predicts longer food chains are vulnerable to environmental perturbations (i.e., the dynamical constraint).
 Thus, habitats with frequent and/or severe disturbance should support fewer trophic levels [@pimmNumberTrophicLevels1977].
-Lastly, the ecosystem size hypothesis posits that larger ecosystems should harbor longer food chains [@takimotoEffectsProductivityDisturbance2012; @teruiSpatialDisturbanceSynchrony2019; @postEcosystemSizeDetermines2000].
+Lastly, the ecosystem size hypothesis posits that larger ecosystems should harbor longer food chains [@takimotoEffectsProductivityDisturbance2012a; @teruiSpatialDisturbanceSynchrony2019; @postEcosystemSizeDetermines2000].
 
@@
 Meanwhile, the theoretical basis for the other two hypotheses rest on implicit assumptions of linear food chains, which are often violated in natural systems.
-As such, the effects of disturbance and resource availability on food chain length have been found to be highly variable among ecosystem types and geographic regions [@takimotoEffectsProductivityDisturbance2012; @diehlIntraguildPreySuffer2001; @takimotoEnvironmentalDeterminantsFoodchain2013; @postLongShortFoodchain2002].
+As such, the effects of disturbance and resource availability on FCL have been found to be highly variable among ecosystem types and geographic regions [@takimotoEffectsProductivityDisturbance2012a; @diehlIntraguildPreySuffer2001; @takimotoEnvironmentalDeterminantsFoodchain2013; @postLongShortFoodchain2002].
 
@@
 Such context dependent responses to extrinsic factors seem to be tightly coupled with the arrangement of trophic links within a food web.
-For example, omnivory (feeding at multiple trophic levels) is a strong intrinsic driver of food chain length [@takimotoEffectsProductivityDisturbance2012; @postProximateStructuralMechanisms2007; @woottonOmnivoryStabilityFreshwater2017].
+For example, omnivory (feeding at multiple trophic levels) is a strong intrinsic driver of FCL [@takimotoEffectsProductivityDisturbance2012a; @postProximateStructuralMechanisms2007; @woottonOmnivoryStabilityFreshwater2017].
 When omnivory is absent in three-species communities, the trophic structure assumes a simple linear food chain with maximum length equal to 3 (**Figure 1D**).
@@
 Branching ecosystems, for example, are ubiquitous yet overlooked landscape structures (e.g., rivers, trees, caves, mountain ranges) [@grantLivingBranchesPopulation2007; @altermattDiversityRiverineMetacommunities2013].
-While we know that food web interactions are spatially influenced [@calcagnoConstraintsFoodChain2011; @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012; @teruiSpatialDisturbanceSynchrony2019], the theory to account for it has lagged, particularly in branching habitat architectures.
-Here, we theoretically explore drivers of food chain length in rivers, an excellent example of branching ecosystems.
+While we know that food web interactions are spatially influenced [@calcagnoConstraintsFoodChain2011; @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012a; @teruiSpatialDisturbanceSynchrony2019], the theory to account for it has lagged, particularly in branching habitat architectures.
+Here, we theoretically explore drivers of FCL in rivers, an excellent example of branching ecosystems.
 
@@
 Hence, highly-branched river networks provide diverse habitats that buffer the impact of large-scale environmental fluctuations, broadening opportunities for species to recolonize habitat patches post-disturbance [@teruiMetapopulationStabilityBranching2018].
-Thus, ecosystem complexity in river networks may be a widespread yet overlooked extrinsic environmental driver which needs to be incorporated to understand food chain length.
+Thus, ecosystem complexity in river networks may be a widespread yet overlooked extrinsic environmental driver which needs to be incorporated to understand FCL.
 
@@
 We investigate extrinsic controls such as ecosystem size (number of patches), complexity (branching probability), and disturbance regime.
-In addition, our simulation model incorporates intrinsic drivers of food chain length including food web structure, productivity level, and dispersal ability, allowing us to investigate the context dependency on drivers of food chain length across extrinsic factors.
+In addition, our simulation model incorporates intrinsic drivers of FCL including food web structure, productivity level, and dispersal ability, allowing us to investigate the context dependency on drivers of FCL across extrinsic factors.
 We find that ecosystem size and complexity are both positively related to FCL, but the form of this relationship is dependent on other extrinsic and intrinsic factors.
@@
 
-***Food chain length.*** At each patch and time step, food chain length was determined as the highest trophic position ($TP_{x}(t)$) of the existing species.
+***Food chain length.*** At each patch and time step, FCL was determined as the highest trophic position ($TP_{x}(t)$) of the existing species.
 When the intraguild predator exists, the trophic position was determined as follows:
@@
 Subsequently, we ran a burn-in period with no propagules for `r unique(sim_main_result$n_burnin)` time steps to remove the effects of initial conditions.
-We obtained the mean food chain length across all habitat patches for the last `r unique(sim_main_result$n_timestep)` time steps.
-Note that simulation replicates are differ in ecosystem size ($N_{p}\sim Unif(10,150)$) and ecosystem complexity ($P_{b} \sim Unif(0.01, 0.99)$) so we can estimate their effects on food chain length.
+We obtained the mean FCL across all habitat patches for the last `r unique(sim_main_result$n_timestep)` time steps.
+Note that simulation replicates are differ in ecosystem size ($N_{p}\sim Unif(10,150)$) and ecosystem complexity ($P_{b} \sim Unif(0.01, 0.99)$) so we can estimate their effects on FCL.
 
@@
 
-Our simulation highlighted the interactive nature of putative controls of food chain length.
+Our simulation highlighted the interactive nature of putative controls of FCL.
 In particular, disturbance regimes had the strongest influence on the qualitative relationships between FCL and ecosystem properties.
@@
 In ecosystems with a low disturbance regime, FCL was largely determined by omnivory with limited effects of ecosystem size and complexity (colored lines in **Figure 2**).
-Unsurprisingly, FCL was negatively related to the degree of omnivory, with no-omnivory food chains ($P$ only consumes $C$) having the longest FCL, and strong omnivory ($P$ consumes very little $C$) having the shortest FCL.
+FCL was negatively related to the degree of omnivory, with no-omnivory food chains ($P$ only consumes $C$) having the longest FCL, and strong omnivory ($P$ consumes very little $C$) having the shortest FCL.
+The primary mechanism for the decreased FCL was different in low and high productivity scenarios.
+With low productivity, FCL became shorter with stronger omnivory because it intensifies competition between consumer and predator species.
+This effect can be seen in the decreased occupancies of consumer and predator species in a low productivity scenario (most left column in **Figure S1-S4**).
+However, when productivity was high, the occupancies of both species were \~1.0, indicating that the structural constraint of strong omnivory made FCL shorter (second left column in **Figure S1-S4**).
+Productivity played a secondary role in increasing FCL, particularly when omnivory was strong (compare left and right panels in **Figure 2**).
+Dispersal had no apparent effects on FCL (compare top and bottom panels in **Figure 2**).
 These trends were consistent across dispersal scenarios (**Figure 2**).
@@
 For example, the occupancies of $C$ and $P$ increased with ecosystem size when omnivory was strong (see the most left column in **Figures S1 and S2**).
-However, these changes were not translated into changes in FCL due to the structural constraint imposed by omnivory ($C$ and $P$ occupy a similar trophic position; **Figure 1D**).
+However, these changes were not translated into FCL because $C$ and $P$ occupy a similar trophic position (**Figure 1D**).
 Similarly, the occupancy of $P$ decreased with ecosystem complexity when omnivory was strong, but FCL remained almost the same as the increase of $C$ could compensate for the loss of $P$ for the same reason (**Figure S3 and S4**).
@@
 
-A high disturbance regime produced much different responses of FCL.
-FCL increased across both gradients of ecosystem size and complexity under all combinations of productivity, the degree of omnivory, and dispersal distance (**Figure 3**).
-However, the effects of ecosystem size and complexity were mediated by ecological factors differently.
-Ecosystem size effects were stronger with longer dispersal distance ($\theta = 0.1$), while the opposite was true for ecosystem complexity (**Figure 3**).
-In addition, the effect of ecosystem complexity was pronounced in food webs with weaker omnivory (**Figure 3B**).
+On average, increased disturbance frequency decreased FCL, and omnivory had similar effects on FCL (**Figures 3**).
+More importantly, a high disturbance regime produced much different responses of FCL to ecosystem properties.
+FCL increased across both gradients of ecosystem size and complexity under all combinations of productivity, omnivory, and dispersal distance (**Figure 3**).
+The effect of ecosystem complexity was generally clearer than that of ecosystem size.
+It is important to note that the positive relationships emerged in a high disturbance regime because smaller and simpler ecosystems were disproportionately vulnerable to increased disturbance frequency (compare far left in **Figures 2 and 3**).
 
-Interestingly, the positive effects of ecosystem size and complexity were driven by different mechanisms.
-For ecosystem size, the underlying mechanism was consistent when productivity was high; increased occupancies of all the three species resulted in longer food chains.
-However, when productivity was low, the mechanisms were more complex, particularly with strong omnivory (the third column in **Figure S1 and S2**).
+Dispersal buffered the disturbance effects on FCL, and the magnitude of buffering differed by ecosystem structure.
+The greater buffering effect of dispersal emerged in larger ecosystems (compare top and bottom panels in **Figure 3**) because greater dispersal capability is required to reach distant patches.
+In the meantime, simpler, less branched ecosystems benefited more from increased dispersal because habitat patches are more isolated from one another in linearly arranged networks (**Figure 3**).
+Productivity increased FCL as observed in a low disturbance regime, but it did not modulate the qualitative relationships between FCL and ecosystem properties (compare left and right panels in **Figure 3**).
+
+The positive effects of ecosystem size and complexity were driven by different processes.
+For ecosystem size, the underlying process was consistent when productivity was high; increased occupancies of all the three species resulted in longer food chains.
+However, when productivity was low, the processes were more complex, particularly with strong omnivory (the third column in **Figure S1 and S2**).
 In this scenario, the occupancies of $B$ and $C$ decreased at larger ecosystem sizes, while $P$ increased its occupancy (**Figure S1 and S2**).
@@
 For ecosystem complexity, all the three species increased with ecosystem complexity when productivity was high, translating into longer food chains in more complex ecosystems (**Figure S3 and S4**).
-However, when food webs contained strong omnivory with low productivity (the third column in **Figure S3 and S4**), the increased occupacy of $C$ maintained the positive relationship between FCL and ecosystem complexity by offsetting the loss of $P$.
+However, when food webs contained strong omnivory with low productivity (the third column in **Figure S3 and S4**), the increased occupancy of $C$ maintained the positive relationship between FCL and ecosystem complexity by offsetting the loss of $P$.
 Hence, the contrasting mechanisms underlie the positive effects of ecosystem size and complexity on FCL.
@@
 
-Discussion to be added
+The ecosystem size hypothesis of FCL has received support from both theoretical and empirical research [@vanderzandenPatternsFoodChain1999; @postEcosystemSizeDetermines2000; @takimotoEffectsProductivityDisturbance2012a], although empirical work so far has rarely elaborated the underlying drivers [but see @saboRoleDischargeVariation2010; @wardMechanisticTheoryAquatic2017].
+In contrast, ecosystem complexity has not been incorporated in theoretical explorations of FCL to date.
+Here, we show that FCL responds positively to both increasing ecosystem size and complexity.
+This supports our hypothesis that ecosystem complexity may represent an overlooked yet important extrinsic control of FCL. However, the relationship between FCL and ecosystem complexity varied based on additional extrinsic and intrinsic factors, supporting the emerging paradigm that there is no one single driver of FCL in ecosystems [@mchughDualInfluencesEcosystem2010; @wardMechanisticTheoryAquatic2017].
+Although we focus on river networks here, we expect our results to be transferable to other branching structures, such as mountain ranges and cave systems.
+
+While increased disturbance frequency consistently reduced FCL as predicted by dynamical stability hypothesis [@pimmNumberTrophicLevels1977], more important findings in this study were evident from their synergistic effects with ecosystem structure.
+Our low disturbance regime scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independent of ecosystem size and complexity.
+Consequently, FCL was predominantly determined by food web structure.
+However, under high disturbance regimes, only large and complex networks could provide diverse and heterogeneous habitats with potential refugia, allowing species to thrive and maintain long food chains.
+While empirical studies rarely looked at the influence of ecosystem structure on FCL, several studies provide some support for this theoretical prediction.
+For example, @sullivanEcosystemStructureEmerges2015 reported FCL is controlled primarily by the number of tributary confluences in Idaho streams.
+Similarly, it has been shown that the merging of major tributaries lengthened FCL in the Colorado River, especially during a post-monsoon turbid water phase [@saboPulsedFlowsTributary2018].
+Clearly, more empirical research is needed to confirm our theoretical predictions.
+
+The interactive effect of disturbance and ecosystem structure has important implications for ecosystem management because anthropogenic climate change is predicted to increase the frequency and magnitude of disturbances [@brunnerFutureStreamflowRegime2020; @mittalImpactHumanIntervention2016].
+River networks are especially vulnerable due to their propensity to go dry during droughts [@zipperPervasiveChangesStream2021] as well as susceptibility to more intense flood disturbance with increasing storm intensity [@schiermeierIncreasedFloodRisk2011].
+Therefore, the preservation of habitat diversity conferred by ecosystem size and complexity is critical to biodiversity conservation under rapid environmental changes.
+In addition, maintaining habitat connectivity is particularly important for large, linearly-arranged networks.
+Although dispersal has received less attention in food chain research [but see @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012a; @wangMetapopulationCapacityDetermines2021], our simulation highlighted food webs in these networks are more vulnerable to habitat fragmentation because habitat patches are inherently isolated from one another.
+Dams and road crossings have altered the connectivity between suitable habitat patches, or have completely blocked dispersal ability [@ishiyamaPredictingEcologicalImpacts2018; @moritaEffectsHabitatFragmentation2002; @perkinFragmentationAltersStream2012a].
+Light pollution has also negatively affected the ability of adult aquatic insects to disperse between habitat patches [@perkinEffectsArtificialLighting2014].
+These human activities may have disproportionate impacts on food webs in large watersheds with simple network structure.
+
+It is not surprising that FCL declined with increasing omnivory because it effectively compresses a food web (**Figure 1C**) and/or intensifies competition between consumer and predator species.
+However, how omnivory influences the associations with ecosystem structure has rarely been explored [@takimotoEffectsProductivityDisturbance2012a; @wardMechanisticTheoryAquatic2017].
+An important result in our simulation is that omnivory does not change the qualitative relationships between FCL and ecosystem properties, as opposed to productivity effects.
+This provides a possible explanation for why the ecosystem size hypothesis has been consistently supported in various ecosystems [@takimotoEnvironmentalDeterminantsFoodchain2013] and raises the possibility of ecosystem complexity as an important control of FCL in branching networks.
+In addition, our results imply that the size and complexity effects are robust to food web reorganization, which is often associated with species extirpations or introductions.
+The strength of omnivory in these novel communities can be increased or decreased, depending on the introduced species and the extant species pool or community.
+For example, non-indigenous fishes in the Eastern Mediterranean Sea outcompete native species for high quality prey items, causing native species to increase consumption of non-preferred prey items [@fanelliDepictingNovelEastern2015].
+Such substantial changes in food web structure, however, may have limited influences on how ecosystem structure dictates FCL.
+
+Our findings add an important distinction in the discussion of the suitability of assigning protection or conservation status to regions.
+Current conservation schemes generally emphasize size, aiming to protect the largest areas possible.
+While this is certainly a useful attribute generally, designating networks with a high degree of variability, or many small patches, can be more beneficial than a large, homogenous area [@fahrigWhySeveralSmall2020].
+Accounting for ecosystem complexity is particularly suited to ecosystem architectures in which branches have variable environmental conditions [@koningNetworkGrassrootsReserves2020; @teruiEmergentDualScaling2021].
+Conservation and management activities should focus on activities that increase the habitat heterogeneity and connectivity of suitable habitat patches to increase resilience to predicted increases in disturbance frequency and magnitude.
+
+It is abundantly clear that ecosystem complexity in natural systems is the rule, not the exception.
+Early theoretical and empirical work has assumed simple habitat structure as a necessity.
+However, habitat heterogeneity, such as the presence of prey refugia, and dispersal between connected habitat patches allow communities to stabilize and persist through space and time [@pedersenNonhierarchicalDispersalPromotes2016; @mougiFoodwebComplexityMetacommunity2016].
+Complex ecosystem structure is an extrinsic factor that can affect FCL in communities.
+Incorporating a realistic understanding of ecosystem structural complexity may improve our perception of communities in naturally complex systems.
 

### 2022-03-31 rmd/manuscript.Rmd
edit: manuscript

@@
 
-Understanding the drivers of food chain length in natural communities has intrigued ecologists since the publication of 'food cycles' by Elton in the early 20th century.
-Proposed drivers of food chain length have included extrinsic controls such as productivity, disturbance regime, and ecosystem size, as well as intrinsic factors including trophic omnivory.
-However, current theories have largely assumed simple, two-dimensional habitat architectures, and may not be adequate to predict food chain length in ecosystems which have a complex, branching structure.
+Understanding the drivers of food chain length in natural communities has intrigued ecologists since Elton publicized 'food cycles' in the early 20th century.
+Proposed drivers of food chain length have included productivity, disturbance regime, ecosystem size, and trophic omnivory.
+However, current theories have largely assumed simple, two-dimensional habitat architectures and may not be adequate to predict food chain length in ecosystems with a complex, branching structure.
 Here, we develop a spatially explicit theoretical model which provides an integrated framework for understanding variation in food chain length in branching networks.
-We show independent, positive influences of both ecosystem size and complexity on food chain length.
-However, these effects were contingent upon other proposed extrinsic and intrinsic drivers such as disturbance regime, productivity, and trophic omnivory.
-Our results show that ecosystem complexity is an important driver of food chain length that may explain enigmatic variation can reconcile inconsistent results from empirical studies of food chain length in riverbranching ecosystems.
+We show independent, positive influences of ecosystem size and complexity on food chain length.
+However, these effects were contingent upon other proposed drivers, such as the disturbance regime.
+Our results suggest that ecosystem complexity is an important yet overlooked driver of food chain length in branching ecosystems.
 
@@
 
-The origin of variation in food chain length (FCL), the number of feeding links from the basal species to top predators, has intrigued ecologists for decades due to its importance in regulating top-down effects [@paceTrophicCascadesRevealed1999], energy flow [@wangBiodiversityEcosystemFunctioning2018], and contaminant concentrations in top predators that humans often consume [@kiddEffectsTrophicPosition1998].
+The origin of variation in food chain length (FCL), the number of feeding links from the basal species to top predators, has intrigued ecologists for decades due to its importance in regulating top-down effects [@paceTrophicCascadesRevealed1999], energy flow [@wangBiodiversityEcosystemFunctioning2018], and contaminant concentrations in top predators [@kiddEffectsTrophicPosition1998].
 Hypotheses regarding controls of FCL can be organized into three broad categories.
@@
 For example, large ecosystems have increased species richness (species-area relationship [@lomolinoEcologyMostGeneral2000]), buffer against disturbances [@saboRoleDischargeVariation2010], and have greater numbers of immigrant sources in the ecosystem with stable metapopulations [@takimotoEnvironmentalDeterminantsFoodchain2013].
-Meanwhile, the theoretical basis for the other two hypotheses rest on implicit assumptions of linear food chains, which are often violated in natural systems.
+Meanwhile, the theoretical basis for the other two hypotheses rests on implicit assumptions of linear food chains, which are often violated in natural systems.
 As such, the effects of disturbance and resource availability on FCL have been found to be highly variable among ecosystem types and geographic regions [@takimotoEffectsProductivityDisturbance2012a; @diehlIntraguildPreySuffer2001; @takimotoEnvironmentalDeterminantsFoodchain2013; @postLongShortFoodchain2002].
@@
 More recently, food web ecologists have begun to recognize that FCL rarely responds to a single driver, but instead is determined by interactions of multiple factors.
-For example, @wardMechanisticTheoryAquatic2017 showed that FCL was more responsive to increasing productivity in large ecosystems compared with small ecosystems, whereas FCL was more sensitive to size in low productivity systems.
-Such context dependent responses to extrinsic factors seem to be tightly coupled with the arrangement of trophic links within a food web.
+For example, @wardMechanisticTheoryAquatic2017 showed that FCL was more responsive to increasing productivity in large ecosystems than small ones, whereas FCL was more sensitive to size in low-productivity systems.
+Such context-dependent responses to extrinsic factors seem to be tightly coupled with the arrangement of trophic links within a food web.
 For example, omnivory (feeding at multiple trophic levels) is a strong intrinsic driver of FCL [@takimotoEffectsProductivityDisturbance2012a; @postProximateStructuralMechanisms2007; @woottonOmnivoryStabilityFreshwater2017].
-When omnivory is absent in three-species communities, the trophic structure assumes a simple linear food chain with maximum length equal to 3 (**Figure 1D**).
-However, as omnivory increases, FCL decreases proportionally as the top predator eats lower in the food web [@wardMechanisticTheoryAquatic2017; @woottonOmnivoryStabilityFreshwater2017].
+When omnivory is absent in three-species communities, the trophic structure assumes a simple linear food chain with a maximum length of three (**Figure 1D**).
+However, FCL decreases proportionally as the top predator eats lower in the food web as omnivory increases [@wardMechanisticTheoryAquatic2017; @woottonOmnivoryStabilityFreshwater2017].
 Theoretical work predicts that food webs are destabilized when omnivory is prevalent, leading to the loss of species and shorter food chains even in productive, large ecosystems.
-Conversely, weak omnivory stabilize the food web and increase persistence of species within a community, particularly at intermediate levels of productivity and high disturbance levels [@woottonOmnivoryStabilityFreshwater2017; @woottonManyWeakInteractions2016].
-Due to the context-dependency of FCL to multiple drivers, and the potential for interactive effects, a unified framework of multiple food chain drivers is needed to better understand when and where FCL will respond to these extrinsic and intrinsic factors.
+Conversely, weak omnivory stabilizes the food web and increases the persistence of species within a community, particularly at intermediate levels of productivity and high disturbance levels [@woottonOmnivoryStabilityFreshwater2017; @woottonManyWeakInteractions2016].
+Due to the context-dependency of FCL to multiple drivers and the potential for interactive effects, a unified framework of multiple food chain drivers is needed to better understand when and where FCL will respond to these extrinsic and intrinsic factors.
 
@@
 
-Ample evidence suggests that branching structure controls environmental heterogeneity and dispersal pathways of organisms within river ecosystems [@altermattDiversityRiverineMetacommunities2013; @tonkinRoleDispersalRiver2018].
+Ample evidence suggests that branching structure controls environmental heterogeneity and dispersal pathways for organisms within river ecosystems [@altermattDiversityRiverineMetacommunities2013; @tonkinRoleDispersalRiver2018].
 Water and materials propagate downstream as small streams hierarchically join to form larger streams and rivers.
-Consequently, environmental signals are highly variable among tributaries while showing a strong within-tributary correlation e.g., flood disturbance [@nakamuraDisturbanceRegimesStream2000; @isaakSlowClimateVelocities2016; @mcguireNetworkAnalysisReveals2014; @abbottUnexpectedSpatialStability2018].
-Confluences serve as merging points where tributary dynamics aggregate to form distinct environmental conditions in downstream habitat patches.
-Habitat heterogeneity therefore increases with branching complexity (**Figure 1A**), which we define here as the probability of branching per unit river distance.
-Hence, highly-branched river networks provide diverse habitats that buffer the impact of large-scale environmental fluctuations, broadening opportunities for species to recolonize habitat patches post-disturbance [@teruiMetapopulationStabilityBranching2018].
-Thus, ecosystem complexity in river networks may be a widespread yet overlooked extrinsic environmental driver which needs to be incorporated to understand FCL.
+Consequently, environmental signals (e.g., flood disturbance) are highly variable among tributaries while showing a strong within-tributary correlation  [@nakamuraDisturbanceRegimesStream2000; @isaakSlowClimateVelocities2016; @mcguireNetworkAnalysisReveals2014; @abbottUnexpectedSpatialStability2018].
+Confluences are merging points where tributary dynamics aggregate to form distinct environmental conditions in downstream habitat patches.
+Hence, habitat heterogeneity increases with branching complexity (**Figure 1A**), which we define here as the probability of branching per unit river distance.
+Highly-branched river networks provide diverse habitats that buffer the impact of large-scale environmental fluctuations, broadening opportunities for species to recolonize habitat patches post-disturbance [@teruiMetapopulationStabilityBranching2018].
+Thus, ecosystem complexity in river networks may be a widespread yet overlooked extrinsic environmental driver of FCL.
 
-We develop a model in which trophic community dynamics within branching habitats are simulated.
-We investigate extrinsic controls such as ecosystem size (number of patches), complexity (branching probability), and disturbance regime.
-In addition, our simulation model incorporates intrinsic drivers of FCL including food web structure, productivity level, and dispersal ability, allowing us to investigate the context dependency on drivers of FCL across extrinsic factors.
+We simulate trophic community dynamics within branching habitats to investigate the effects of extrinsic controls of FCL, such as ecosystem size (number of patches), complexity (branching probability), and disturbance regime.
+In addition, our simulation model incorporates intrinsic drivers of FCL, including the degree of omnivory, productivity level, and species' dispersal ability, allowing us to assess the context dependency on drivers of FCL across extrinsic factors.
 We find that ecosystem size and complexity are both positively related to FCL, but the form of this relationship is dependent on other extrinsic and intrinsic factors.
@@
 where $m_{down}$ and $m_{up}$ are the disturbance values at down- and upstream patches, and the local environmental noise at the downstream patch $\epsilon$ was drawn from $Normal(0, \sigma_{l}^2)$.
-This autoregressive process resembles natural disturbance regime, where flood or drought disturbance cascades downstream with some local variation along the channel.
+This autoregressive process resembles a natural disturbance regime, where flood or drought disturbance cascades downstream with some local variation along the channel.
 At confluence patches, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, $Q$ (the number of upstream contributing patches):
@@
 We employed a discrete-time model to describe tri-trophic dynamics in each habitat patch.
-A discrete model is suitable for organisms with predictable reproduction cycles, a common life-history for many lotic species including macroinvertebrates and fish.
+A discrete model is suitable for organisms with predictable reproduction cycles, a common life history for many lotic species, including macroinvertebrates and fish.
 Our simulation model consists of three species communities; a basal species $B$ with positive logistic growth up to a carrying capacity, $K$; a primary consumer species or intraguild prey, $C$, which grows based on the consumption of $B$; and an intraguild predator $P$, which can consume both $B$ and $C$ (**Figure 1C**).
@@
 
-***Dispersal.*** We simulated distance-dependent dispersal along a network.
+***Dispersal.*** We simulated distance-dependent dispersal along with a network.
 The number of emigrants from patch $x$ at time $t$, $E_{i,x}(t)$, is estimated as the product of dispersal probability $p_{d}$ and species abundance after internal dynamics $\bar{A}_{i,x}(t)$.
 The dispersal probability was assumed to be constant for all three species.
-The relative immigration potential $\xi_{i,x}(t)$ is calculated considering network structure and species dispersal capability as follows:
+We calculated the relative immigration potential $\xi_{i,x}(t)$ considering network structure and species dispersal capability as follows:
 
@@
 
-***Food chain length.*** At each patch and time step, FCL was determined as the highest trophic position ($TP_{x}(t)$) of the existing species.
-When the intraguild predator exists, the trophic position was determined as follows:
+***Food chain length.*** At each patch and time step, we determined FCL  as the highest trophic position ($TP_{x}(t)$) of the existing species.
+When the intraguild predator exists, the trophic position is determined as follows:
 
@@
 
-Prior to the main simulation, we performed an extensive sensitivity analysis to identify influential parameters (see **Supplementary Text** and **Table S2-S3**).
+Before the main simulation, we performed an extensive sensitivity analysis to identify influential parameters (see **Supplementary Text** and **Table S2-S3**).
 Our sensitivity analysis revealed that ecosystem size and/or complexity effects were sensitive to disturbance, productivity, dispersal, and omnivory.
@@
 Our simulation highlighted the interactive nature of putative controls of FCL.
-In particular, disturbance regimes had the strongest influence on the qualitative relationships between FCL and ecosystem properties.
-Below, we report our results for low and high disturbance regimens separately.
+In particular, disturbance regimes strongly influenced the qualitative relationships between FCL and ecosystem properties.
+Below, we report our results for low and high disturbance regimes separately.
 
@@
 
-In ecosystems with a low disturbance regime, FCL was largely determined by omnivory with limited effects of ecosystem size and complexity (colored lines in **Figure 2**).
-FCL was negatively related to the degree of omnivory, with no-omnivory food chains ($P$ only consumes $C$) having the longest FCL, and strong omnivory ($P$ consumes very little $C$) having the shortest FCL.
+In ecosystems with a low disturbance regime, omnivory largely determined FCL (colored lines in **Figure 2**), with no-omnivory food chains ($P$ only consumes $C$) having the longest FCL, and strong omnivory ($P$ consumes very little $C$) having the shortest FCL.
 The primary mechanism for the decreased FCL was different in low and high productivity scenarios.
-With low productivity, FCL became shorter with stronger omnivory because it intensifies competition between consumer and predator species.
-This effect can be seen in the decreased occupancies of consumer and predator species in a low productivity scenario (most left column in **Figure S1-S4**).
+With low productivity, strong omnivory caused shorter FCL because it intensifies competition between consumer and predator species.
+We can see this effect in the decreased occupancies of consumer and predator species in a low productivity scenario (most left column in **Figure S1-S4**).
 However, when productivity was high, the occupancies of both species were \~1.0, indicating that the structural constraint of strong omnivory made FCL shorter (second left column in **Figure S1-S4**).
@@
 For example, the occupancies of $C$ and $P$ increased with ecosystem size when omnivory was strong (see the most left column in **Figures S1 and S2**).
-However, these changes were not translated into FCL because $C$ and $P$ occupy a similar trophic position (**Figure 1D**).
-Similarly, the occupancy of $P$ decreased with ecosystem complexity when omnivory was strong, but FCL remained almost the same as the increase of $C$ could compensate for the loss of $P$ for the same reason (**Figure S3 and S4**).
+However, these changes did not translate into FCL because $C$ and $P$ occupy a similar trophic position (**Figure 1D**).
+Meanwhile, the occupancy of $P$ decreased with ecosystem complexity when omnivory was strong, but FCL remained almost the same as the increase of $C$ could compensate for the loss of $P$ (**Figure S3 and S4**).
 
@@
 On average, increased disturbance frequency decreased FCL, and omnivory had similar effects on FCL (**Figures 3**).
-More importantly, a high disturbance regime produced much different responses of FCL to ecosystem properties.
-FCL increased across both gradients of ecosystem size and complexity under all combinations of productivity, omnivory, and dispersal distance (**Figure 3**).
+More importantly, a high disturbance regime produced different responses of FCL to ecosystem properties.
+FCL increased across gradients of ecosystem size and complexity under all productivity, omnivory, and dispersal scenarios (**Figure 3**).
 The effect of ecosystem complexity was generally clearer than that of ecosystem size.
@@
 
-Dispersal buffered the disturbance effects on FCL, and the magnitude of buffering differed by ecosystem structure.
+Dispersal buffered the disturbance effects on FCL and the magnitude of buffering differed by ecosystem structure.
 The greater buffering effect of dispersal emerged in larger ecosystems (compare top and bottom panels in **Figure 3**) because greater dispersal capability is required to reach distant patches.
-In the meantime, simpler, less branched ecosystems benefited more from increased dispersal because habitat patches are more isolated from one another in linearly arranged networks (**Figure 3**).
+In the meantime, simpler, less branched ecosystems benefited more from increased dispersal because habitat patches are more isolated in linearly arranged networks (**Figure 3**).
 Productivity increased FCL as observed in a low disturbance regime, but it did not modulate the qualitative relationships between FCL and ecosystem properties (compare left and right panels in **Figure 3**).
 
-The positive effects of ecosystem size and complexity were driven by different processes.
-For ecosystem size, the underlying process was consistent when productivity was high; increased occupancies of all the three species resulted in longer food chains.
-However, when productivity was low, the processes were more complex, particularly with strong omnivory (the third column in **Figure S1 and S2**).
-In this scenario, the occupancies of $B$ and $C$ decreased at larger ecosystem sizes, while $P$ increased its occupancy (**Figure S1 and S2**).
+Different processes drove the positive effects of ecosystem size and complexity on FCL when strong omnivory exists in low-productivity systems.
+In this scenario, the occupancies of $B$ and $C$ decreased at larger ecosystem sizes, while $P$ increased its occupancy (third column in **Figure S1 and S2**).
 Therefore, $P$ compensated for the loss of $B$ and $C$ to maintain the positive relationship between FCL and ecosystem size.
-For ecosystem complexity, all the three species increased with ecosystem complexity when productivity was high, translating into longer food chains in more complex ecosystems (**Figure S3 and S4**).
-However, when food webs contained strong omnivory with low productivity (the third column in **Figure S3 and S4**), the increased occupancy of $C$ maintained the positive relationship between FCL and ecosystem complexity by offsetting the loss of $P$.
-Hence, the contrasting mechanisms underlie the positive effects of ecosystem size and complexity on FCL.
+In contrast, the increased occupancy of $C$ maintained the positive relationship between FCL and ecosystem complexity by offsetting the loss of $P$.
 
@@
 
-The ecosystem size hypothesis of FCL has received support from both theoretical and empirical research [@vanderzandenPatternsFoodChain1999; @postEcosystemSizeDetermines2000; @takimotoEffectsProductivityDisturbance2012a], although empirical work so far has rarely elaborated the underlying drivers [but see @saboRoleDischargeVariation2010; @wardMechanisticTheoryAquatic2017].
+The ecosystem size hypothesis of FCL has received support from both theoretical and empirical research [@vanderzandenPatternsFoodChain1999; @postEcosystemSizeDetermines2000; @takimotoEffectsProductivityDisturbance2012a], although empirical work so far has rarely elaborated on the underlying drivers [but see @saboRoleDischargeVariation2010; @wardMechanisticTheoryAquatic2017].
 In contrast, ecosystem complexity has not been incorporated in theoretical explorations of FCL to date.
-Here, we show that FCL responds positively to both increasing ecosystem size and complexity.
-This supports our hypothesis that ecosystem complexity may represent an overlooked yet important extrinsic control of FCL. However, the relationship between FCL and ecosystem complexity varied based on additional extrinsic and intrinsic factors, supporting the emerging paradigm that there is no one single driver of FCL in ecosystems [@mchughDualInfluencesEcosystem2010; @wardMechanisticTheoryAquatic2017].
+Here, we show that FCL responds positively to increasing ecosystem size and complexity.
+This supports our hypothesis that ecosystem complexity may represent an overlooked yet important extrinsic control of FCL. However, the relationship between FCL and ecosystem complexity varied based on additional extrinsic and intrinsic factors, supporting the emerging paradigm that there is no single driver of FCL in ecosystems [@mchughDualInfluencesEcosystem2010; @wardMechanisticTheoryAquatic2017].
 Although we focus on river networks here, we expect our results to be transferable to other branching structures, such as mountain ranges and cave systems.
 
-While increased disturbance frequency consistently reduced FCL as predicted by dynamical stability hypothesis [@pimmNumberTrophicLevels1977], more important findings in this study were evident from their synergistic effects with ecosystem structure.
-Our low disturbance regime scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independent of ecosystem size and complexity.
-Consequently, FCL was predominantly determined by food web structure.
+While increased disturbance frequency consistently reduced FCL as predicted by the dynamical stability hypothesis [@pimmNumberTrophicLevels1977], more important findings in this study were evident from their synergistic effects on ecosystem structure.
+Our low disturbance regime scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independently of ecosystem size and complexity.
+Consequently, the food web structure largely controlled FCL.
 However, under high disturbance regimes, only large and complex networks could provide diverse and heterogeneous habitats with potential refugia, allowing species to thrive and maintain long food chains.
@@
 Similarly, it has been shown that the merging of major tributaries lengthened FCL in the Colorado River, especially during a post-monsoon turbid water phase [@saboPulsedFlowsTributary2018].
-Clearly, more empirical research is needed to confirm our theoretical predictions.
+More empirical research is needed to confirm our theoretical predictions.
 
@@
 Although dispersal has received less attention in food chain research [but see @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012a; @wangMetapopulationCapacityDetermines2021], our simulation highlighted food webs in these networks are more vulnerable to habitat fragmentation because habitat patches are inherently isolated from one another.
-Dams and road crossings have altered the connectivity between suitable habitat patches, or have completely blocked dispersal ability [@ishiyamaPredictingEcologicalImpacts2018; @moritaEffectsHabitatFragmentation2002; @perkinFragmentationAltersStream2012a].
+Dams and road crossings have altered the connectivity between suitable habitat patches or have completely blocked dispersal [@ishiyamaPredictingEcologicalImpacts2018; @moritaEffectsHabitatFragmentation2002; @perkinFragmentationAltersStream2012a].
 Light pollution has also negatively affected the ability of adult aquatic insects to disperse between habitat patches [@perkinEffectsArtificialLighting2014].
-These human activities may have disproportionate impacts on food webs in large watersheds with simple network structure.
+These human activities may disproportionately impact food webs in large watersheds with simple network structures.
 
@@
 An important result in our simulation is that omnivory does not change the qualitative relationships between FCL and ecosystem properties, as opposed to productivity effects.
-This provides a possible explanation for why the ecosystem size hypothesis has been consistently supported in various ecosystems [@takimotoEnvironmentalDeterminantsFoodchain2013] and raises the possibility of ecosystem complexity as an important control of FCL in branching networks.
+This result provides a possible explanation for why the ecosystem size hypothesis has been consistently supported in various ecosystems [@takimotoEnvironmentalDeterminantsFoodchain2013] and raises the possibility of ecosystem complexity as an important control of FCL in branching networks.
 In addition, our results imply that the size and complexity effects are robust to food web reorganization, which is often associated with species extirpations or introductions.
 The strength of omnivory in these novel communities can be increased or decreased, depending on the introduced species and the extant species pool or community.
-For example, non-indigenous fishes in the Eastern Mediterranean Sea outcompete native species for high quality prey items, causing native species to increase consumption of non-preferred prey items [@fanelliDepictingNovelEastern2015].
-Such substantial changes in food web structure, however, may have limited influences on how ecosystem structure dictates FCL.
+For example, non-indigenous fishes in the Eastern Mediterranean Sea outcompete native species for high-quality prey items, causing native species to increase their consumption of non-preferred prey items [@fanelliDepictingNovelEastern2015].
+However, such substantial changes in food web structure may have limited influences on how ecosystem structure dictates FCL.
 
@@
 Early theoretical and empirical work has assumed simple habitat structure as a necessity.
-However, habitat heterogeneity, such as the presence of prey refugia, and dispersal between connected habitat patches allow communities to stabilize and persist through space and time [@pedersenNonhierarchicalDispersalPromotes2016; @mougiFoodwebComplexityMetacommunity2016].
+However, habitat heterogeneity, such as prey refugia, and dispersal between connected habitat patches allow communities to stabilize and persist through space and time [@pedersenNonhierarchicalDispersalPromotes2016; @mougiFoodwebComplexityMetacommunity2016].
 Complex ecosystem structure is an extrinsic factor that can affect FCL in communities.

### 2022-03-31 rmd/manuscript.Rmd
edit: manuscript

@@
 ```
-where $W_{ij,x}(t)$ is the population-wide consumption of consumer $j$ on prey $i$ (number of individuals), and the parameter $\delta_{x}(t)$ is the relative preference of intraguild predator to basal species over intraguild prey.
+where $W_{ij,x}(t)$ is the population-wide consumption of consumer $j$ on prey $i$ (number of individuals), and the parameter $\delta_{x}(t)$ is the relative preference of predator to basal species over consumer.
 We assume the Holling Type-II functional response for predator-prey interactions:
@@
 
-The population growth of the intraguild prey is controlled by conversion efficiency, number of prey captured, background mortality, and predation:
+The population growth of the consumer is controlled by conversion efficiency, number of prey captured, background mortality, and predation:
 
@@
 
-The population of the intraguild predator is fueled by two prey species:
+The population of the predator is fueled by two prey species:
 
@@
 ***Food chain length.*** At each patch and time step, we determined FCL as the highest trophic position ($TP_{x}(t)$) of the existing species.
-When the intraguild predator exists, the trophic position is determined as follows:
+When the predator exists, the trophic position is determined as follows:
 
@@
 Trophic interaction strengths can be controlled by varying the attack rate and/or conversion efficiency [@wardMechanisticTheoryAquatic2017].
-We changed attack rate ($a_{BP}$ and $a_{CP}$) and conversion efficiency ($e_{BP}$ and $e_{CP}$) of two prey items for intraguild predator $P$.
+We changed attack rate ($a_{BP}$ and $a_{CP}$) and conversion efficiency ($e_{BP}$ and $e_{CP}$) of two prey items for predator $P$.
 Specifically, we used the following combinations to represent three levels of omnivory: "Chain", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[1], max(sim_main_result$e_cp))` $\}$; "Weak omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[2], max(sim_main_result$e_cp))` $\}$; and "Strong omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(max(sim_main_result$a_bp), min(sim_main_result$a_cp), unique(sim_main_result$e_bp)[3], min(sim_main_result$e_cp))` $\}$.
@@
 
+# **Acknowledgements**
+
+We thank C Perovich for thoughtful discussions in the early stages of this project.
+This material is based upon work supported by the National Science Foundation through the Division of Environmental Biology (DEB 2015634).
+
+# **Author contributions**
+
+JP developed the theoretical model of trophic dynamics, ran the simulations, and made the figures with input from AT.
+JP wrote the first draft, while AT and JF contributed significantly to the final version.
+
+# **Data availability**
+
+No data were used or produced in this research.
+R functions for simulations are provided as the R package *mcbrnet* (available at <https://github.com/aterui/mcbrnet>).
+Codes for simulations, statistical analysis, figures, and tables are available at <https://github.com/aterui/public-proj_fcl-river-theory>.
+
 # References

### 2022-03-31 rmd/manuscript.Rmd
edit: manuscript

@@
 output:
-  pdf_document:
-    latex_engine: xelatex
+  #pdf_document:
+  #  latex_engine: xelatex
+  word_document:
+    reference_docx: word_format.docx
 knit: (function(inputFile, encoding) {
@@
 
-```{=tex}
-\begin{align}
+$$
 logit~m_{down} = logit~m_{up}+ \epsilon
-\end{align}
-```
+$$
+
 where $m_{down}$ and $m_{up}$ are the disturbance values at down- and upstream patches, and the local environmental noise at the downstream patch $\epsilon$ was drawn from $Normal(0, \sigma_{l}^2)$.
@@
 
-```{=tex}
-\begin{align}
+$$
 logit~m_{down} = \omega~(logit~m_{1,up} +\epsilon_1)+(1-\omega)~(logit~m_{2,up} +\epsilon_2)
-\end{align}
-```
+$$
+
 $m_{1,up}$ and $m_{2,up}$ are the disturbance values of upstream contributing tributaries, $\epsilon_1$ and $\epsilon_2$ are independent local noises, and $\omega$ the relative size of one tributary over another defined as $\omega = \frac{Q_{1}}{Q_{1} + Q_{2}}$.
@@
 
-```{=tex}
-\begin{align}
+$$
 K_{x} = K_{base}(N_{p,up} + 1)^z
-\end{align}
-```
+$$
+
 where $K_{base}$ is the carrying capacity at the upstream terminal patch, $N_{p,up}$ the number of patches upstream, and $z$ is the scaling exponent.
@@
 
-```{=tex}
-\begin{align}
+$$
 A_{i,x}(t+1) \sim Poisson(\bar{A}_{i,x}(t) + I_{i,x}(t) - E_{i,x}(t))
-\end{align}
-```
+$$
+
 where $\bar{A}_{i,x}(t+1)$ is the expected species abundance after within-patch trophic interactions, $I_{i,x}(t)$ the expected number of immigrants into patch $x$, and $E_{i,x}(t)$ the expected number of emigrants from patch $x$.
@@
 
-```{=tex}
-\begin{align}
+$$
 A_{B,x}'(t) = s_{0,B}[1-z_{B,x}(t)]A_{B,x}(t)
-\end{align}
-```
+$$
+
 where $s_{0,i}$ is the background survival probability (density independent) and $z_{B,x}(t)$ the proportional mortality caused by predation.
@@
 
-```{=tex}
-\begin{equation}
-z_{B,x}(t) = z_{BC,x}(t) + z_{BP,x}(t) =
-\left\{
-\begin{aligned}
-&\frac{W_{B,x}(t)}{A_{B,x}(t)} &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\
-&1 &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\
-\end{aligned}
-\right.
-\end{equation}
-```
+$$
+z_{B,x}(t) = z_{BC,x}(t) + z_{BP,x}(t) =\left\{\begin{aligned}&\frac{W_{B,x}(t)}{A_{B,x}(t)} &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\&1 &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
+$$
+
 and
 
-```{=tex}
-\begin{align}
+$$
 W_{B,x}(t) = W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)
-\end{align}
-```
+$$
+
 where $W_{ij,x}(t)$ is the population-wide consumption of consumer $j$ on prey $i$ (number of individuals), and the parameter $\delta_{x}(t)$ is the relative preference of predator to basal species over consumer.
@@
 
-```{=tex}
-\begin{align}
+$$
 W_{ij,x}(t) = \frac{a_{ij}A_{i,x}(t)A_{j,x}(t)}{A_{j,x}(t) + a_{ij}h_{ij}A_{i,x}(t)}
-\end{align}
-```
+$$
+
 $a_{ij}$ is the attack rate and $h_{ij}$ the handling time. To be realistic, the preference parameter is assumed to depend on the relative potential amount of energy obtained from each prey:
 
-```{=tex}
-\begin{align}
+$$
 \delta_{x}(t) = \frac{e_{BP} f_{BP}(A_{B,x}(t))}{e_{BP} f_{BP}(A_{B,x}(t)) + e_{CP} f_{CP}(A_{C,x}(t))}
-\end{align}
-```
+$$
+
 where $e_{ij}$ is the conversion efficiency from prey $i$ to consumer $j$, and $f_{ij}(A_{i,x}(t)) = \frac{a_{ij}A_{i,x}(t)}{1 + a_{ij}h_{ij}A_{i,x}(t)}$.
@@
 
-```{=tex}
-\begin{align}
+$$
 \hat{A}_{B,x}(t) = \frac{r_{b} A_{B,x}'(t)}{1 + \frac{r_b - 1}{K_{x}}A_{B,x}'(t)}
-\end{align}
-```
+$$
+
 $r_b$ is the reproductive rate for survived individuals.
@@
 
-```{=tex}
-\begin{equation}
-\hat{A}_{C,x}(t) =
-\left\{
-\begin{aligned}
-&s_{0,C} (1 - z_{CP,x}) e_{BC}W_{BC,x}(t) &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\
-&s_{0,C} (1 - z_{CP,x}) e_{BC} \left( \frac{W_{BC,x}(t)}{W_{B,x}(t)} A_{B,x} \right) &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\
-\end{aligned}
-\right.
-\end{equation}
-```
-Equation 11 assumes competition between $C$ and $P$ when their total consumption $W_{B,x}(t)$ exceeds the abundance of basal species $A_{B,x}(t)$, and the relative consumption $\frac{W_{BC,x}(t)}{W_{B,x}(t)}$ rests on $a_{ij}$, $h_{ij}$, and $A_{j,x}(t)$.
+$$
+\hat{A}_{C,x}(t) =\left\{\begin{aligned}&s_{0,C} (1 - z_{CP,x}) e_{BC}W_{BC,x}(t) &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\&s_{0,C} (1 - z_{CP,x}) e_{BC} \left( \frac{W_{BC,x}(t)}{W_{B,x}(t)} A_{B,x} \right) &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
+$$
+
+The above equation assumes competition between $C$ and $P$ when their total consumption $W_{B,x}(t)$ exceeds the abundance of basal species $A_{B,x}(t)$, and the relative consumption $\frac{W_{BC,x}(t)}{W_{B,x}(t)}$ rests on $a_{ij}$, $h_{ij}$, and $A_{j,x}(t)$.
 
@@
 
-```{=tex}
-\begin{equation}
-\hat{A}_{P,x}(t) =
-\left\{
-\begin{aligned}
-&s_{0,P} \left[e_{BP} \delta_{x}(t) W_{BP,x}(t) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) \right] &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\
-&s_{0,P} \left[e_{BP} \left(\frac{\delta_{x}(t)W_{BP,x}(t)}{W_{B,x}(t)}A_{B,x}\right) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) \right] &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\
-\end{aligned}
-\right.
-\end{equation}
-```
+$$
+\hat{A}_{P,x}(t) =\left\{\begin{aligned}&s_{0,P} \left[e_{BP} \delta_{x}(t) W_{BP,x}(t) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) \right] &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\&s_{0,P} \left[e_{BP} \left(\frac{\delta_{x}(t)W_{BP,x}(t)}{W_{B,x}(t)}A_{B,x}\right) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) \right] &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
+$$
+
 After recruitment, the constituent species experience stochastic regional disturbance, which causes proportional loss of $m$ to populations:
 
-```{=tex}
-\begin{align}
+$$
 \bar{A}_{i,x}(t) = [1 - \psi(t) m_{x}] \hat{A}_{i,x}(t)
-\end{align}
-```
+$$
+
 The binary random variable $\psi(t)$ expresses the occurrence of regional disturbance (disturbance occurs if $\psi(t)=1$ otherwise $0$), and $m_x$ is the disturbance intensity at patch $x$ as determined by the network generation process.
@@
 
-```{=tex}
-\begin{align}
+$$
 \xi_{i,x}(t)=\frac{\sum_{y, y \ne x}E_{i,y}(t)exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}E_{i,y}(t)exp(-\theta d_{xy})}
-\end{align}
-```
+$$
+
 where $d_{xy}$ is the network distance between patch $x$ and $y$ (shortest path), and $\theta$ indicates the inverse of mean dispersal distance (i.e., larger numbers mean shorter dispersal distance).
@@
 
-```{=tex}
-\begin{align}
+$$
 TP_{P,x}(t) = \delta_{x}(t) + 2(1 - \delta_{x}(t)) + 1
-\end{align}
-```
+$$
+
 This definition accounts for the influence of omnivory.

### 2022-05-09 rmd/manuscript.Rmd
edit: maintext

@@
 We show independent, positive influences of ecosystem size and complexity on food chain length.
-However, these effects were contingent upon other proposed drivers, such as the disturbance regime.
+However, these effects were contingent upon other drivers.
+In particular, the effects of ecosystem size and complexity appeared more clearly in high disturbance regimes.
 Our results suggest that ecosystem complexity is an important yet overlooked driver of food chain length in branching ecosystems.
@@
 
-Among the three hypotheses, the ecosystem size hypothesis has been consistently supported in empirical explorations, likely because multiple mechanisms may underlie the positive association [@takimotoEnvironmentalDeterminantsFoodchain2013].
-For example, large ecosystems have increased species richness (species-area relationship [@lomolinoEcologyMostGeneral2000]), buffer against disturbances [@saboRoleDischargeVariation2010], and have greater numbers of immigrant sources in the ecosystem with stable metapopulations [@takimotoEnvironmentalDeterminantsFoodchain2013].
+Among the three hypotheses, the ecosystem size hypothesis has been consistently supported in empirical explorations, likely because it encompasses multiple mechanisms [@takimotoEnvironmentalDeterminantsFoodchain2013].
+For example, large ecosystems have increased species richness (species-area relationship [@lomolinoEcologyMostGeneral2000]), are buffered against disturbances [@saboRoleDischargeVariation2010], and have greater numbers of immigrant sources in the ecosystem with stable metapopulations [@takimotoEnvironmentalDeterminantsFoodchain2013].
 Meanwhile, the theoretical basis for the other two hypotheses rests on implicit assumptions of linear food chains, which are often violated in natural systems.
@@
 When omnivory is absent in three-species communities, the trophic structure assumes a simple linear food chain with a maximum length of three (**Figure 1D**).
-However, FCL decreases proportionally as the top predator eats lower in the food web as omnivory increases [@wardMechanisticTheoryAquatic2017; @woottonOmnivoryStabilityFreshwater2017].
-Theoretical work predicts that food webs are destabilized when omnivory is prevalent, leading to the loss of species and shorter food chains even in productive, large ecosystems.
+However, FCL decreases proportionally with increasing omnivory as the top predator eats lower in the food web [@wardMechanisticTheoryAquatic2017; @woottonOmnivoryStabilityFreshwater2017].
+Theoretical work predicts that food webs are destabilized when omnivory is too strong, leading to the loss of species and shorter food chains [@woottonOmnivoryStabilityFreshwater2017].
 Conversely, weak omnivory stabilizes the food web and increases the persistence of species within a community, particularly at intermediate levels of productivity and high disturbance levels [@woottonOmnivoryStabilityFreshwater2017; @woottonManyWeakInteractions2016].
-Due to the context-dependency of FCL to multiple drivers and the potential for interactive effects, a unified framework of multiple food chain drivers is needed to better understand when and where FCL will respond to these extrinsic and intrinsic factors.
+Due to the context-dependency of FCL to multiple drivers and the potential for interactive effects, a unified framework of multiple FCL drivers is needed to better understand when and where FCL will respond to these extrinsic and intrinsic factors.
 
@@
 
+In our simulation, disturbance occurs simultaneously with probability $p_m$ but habitat patches vary in their vulnerability.
 We simulated the spatial distribution of disturbance intensity (i.e., proportional mortality in the event of stochastic regional disturbance) as follows.
@@
 where $m_{down}$ and $m_{up}$ are the disturbance values at down- and upstream patches, and the local environmental noise at the downstream patch $\epsilon$ was drawn from $Normal(0, \sigma_{l}^2)$.
-This autoregressive process resembles a natural disturbance regime, where flood or drought disturbance cascades downstream with some local variation along the channel.
+This autoregressive process resembles a natural disturbance regime, where flood or drought disturbance cascades downstream with some local variation along the channel [@nakamuraDisturbanceRegimesStream2000].
 At confluence patches, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, $Q$ (the number of upstream contributing patches):
@@
 
-We also mimicked the downstream increase of local carrying capacity [@teruiThreeEcologicalFactors2016; @carraraComplexInteractionDendritic2014; @rahelFishAssemblagesHabitat1991].
+We also mimicked the downstream increase of local carrying capacity reflecting downstream increases in algal resource availability [@teruiThreeEcologicalFactors2016; @carraraComplexInteractionDendritic2014; @rahelFishAssemblagesHabitat1991].
 Carrying capacity in patch $x$, $K_x$, was assumed to increase as a power function of ecosystem area (i.e., watershed area) [@finlayStreamSizeHuman2011; @koenigEmergentProductivityRegimes2019]:
@@
 
-where $K_{base}$ is the carrying capacity at the upstream terminal patch, $N_{p,up}$ the number of patches upstream, and $z$ is the scaling exponent.
+where $K_{base}$ is the carrying capacity at the upstream terminal patch, $N_{p,up}$ the number of patches upstream, and $z$ the scaling exponent.
 We used $K_{base} = 100$ and $z = 0.54$ based on an empirical relationship of gross primary production and watershed area [@finlayStreamSizeHuman2011].
@@
 $$
-\hat{A}_{P,x}(t) =\left\{\begin{aligned}&s_{0,P} \left[e_{BP} \delta_{x}(t) W_{BP,x}(t) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) \right] &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\&s_{0,P} \left[e_{BP} \left(\frac{\delta_{x}(t)W_{BP,x}(t)}{W_{B,x}(t)}A_{B,x}\right) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) \right] &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
+\hat{A}_{P,x}(t) =\left\{\begin{aligned}&s_{0,P} \left[e_{BP} \delta_{x}(t) W_{BP,x}(t) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) \right] &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\&s_{0,P} \left[e_{BP} \delta_{x}(t)\left(\frac{W_{BP,x}(t)}{W_{B,x}(t)}A_{B,x}\right) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) \right] &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
 $$
@@
 
-Before the main simulation, we performed an extensive sensitivity analysis to identify influential parameters (see **Supplementary Text** and **Table S2-S3**).
+Before the main simulation, we performed an extensive sensitivity analysis to identify influential parameters (see **Supplementary Text** and **Table S1-S3**).
 Our sensitivity analysis revealed that ecosystem size and/or complexity effects were sensitive to disturbance, productivity, dispersal, and omnivory.
-Therefore, we ran the main simulation with `r n_scenario` parameter sets differing in these ecological contexts (two levels of disturbance, two levels of productivity, two levels of dispersal distance, and three levels of omnivory).
+Therefore, we ran the main simulation with `r n_scenario` parameter sets differing in these ecological contexts (two levels of disturbance, two levels of productivity, two levels of dispersal distance, and three levels of omnivory) and parameter values were summarized in **Table S1**.
 
 We controlled disturbance level with disturbance probability $p_{m}$ (low disturbance, `r min(sim_main_result$p_disturb)`; high disturbance, `r max(sim_main_result$p_disturb)`).
-We used the identical value of mean disturbance intensity at headwaters ($\mu_{m}=$ `r max(sim_main_result$mean_disturb_source)`), spatial variation in disturbance intensity at headwaters ($\sigma_{h}=$ `r min(sim_main_result$sd_disturb_source)`), and local disturbance variation ($\sigma_{l}=$ `r unique(sim_main_result$sd_disturb_lon)`) regardless of disturbance scenario.
+We used the identical value of mean disturbance intensity at headwaters ($\mu_{m}=$ `r max(sim_main_result$mean_disturb_source)`), spatial variation in disturbance intensity at headwaters ($\sigma_{h}=$ `r max(sim_main_result$sd_disturb_source)`), and local disturbance variation ($\sigma_{l}=$ `r min(sim_main_result$sd_disturb_lon)`) regardless of disturbance scenario.
+This setup ($\sigma_h > \sigma_l$) produces high spatial variation among tributaries with strong within-tributary downstream correlation, a typical pattern in real river networks [@nakamuraDisturbanceRegimesStream2000; @bendaNetworkDynamicsHypothesis2004].
 Productivity was varied by setting different values of basal species' growth rate ($r_{b}=$ `r unique(sim_main_result$r_b)`).
-For dispersal, we examined short- and long-distance dispersal scenarios (short-distance dispersal $\theta=$ `r max(sim_main_result$theta)`; long-distance dispersal $\theta=$ `r min(sim_main_result$theta)`), while using constant dispersal probability ($p_{d}=$ `r unique(sim_main_result$p_dispersal)`).
+For dispersal, we examined short- and long-distance dispersal scenarios (short-distance dispersal $\theta=$ `r max(sim_main_result$theta)`; long-distance dispersal $\theta=$ `r min(sim_main_result$theta)`) with constant dispersal probability ($p_{d}=$ `r unique(sim_main_result$p_dispersal)`).
 
@@
 The primary mechanism for the decreased FCL was different in low and high productivity scenarios.
-With low productivity, strong omnivory caused shorter FCL because it intensifies competition between consumer and predator species.
+With low productivity, strong omnivory caused shorter FCL because it intensified competition between consumer and predator species.
 We can see this effect in the decreased occupancies of consumer and predator species in a low productivity scenario (most left column in **Figure S1-S4**).
-However, when productivity was high, the occupancies of both species were \~1.0, indicating that the structural constraint of strong omnivory made FCL shorter (second left column in **Figure S1-S4**).
-Productivity played a secondary role in increasing FCL, particularly when omnivory was strong (compare left and right panels in **Figure 2**).
-Dispersal had no apparent effects on FCL (compare top and bottom panels in **Figure 2**).
-These trends were consistent across dispersal scenarios (**Figure 2**).
+However, when productivity was high, the occupancies of both species were \~1.0, indicating that the structural constraint of strong omnivory made FCL shorter (second column from left in **Figure S1-S4**).
+Productivity played a secondary role in increasing FCL, particularly when omnivory was strong (compare left and right panels in **Figure 2A and 2B**).
+Dispersal had no apparent effects on FCL (compare top and bottom panels in **Figure 2A and 2B**).
 
 Although the effects of ecosystem properties on FCL were minimal, there were noticeable effects on the occupancies of constituent species in some scenarios.
-For example, the occupancies of $C$ and $P$ increased with ecosystem size when omnivory was strong (see the most left column in **Figures S1 and S2**).
+For example, the occupancies of $C$ and $P$ increased with ecosystem size when omnivory was strong (see bottom two rows in left most column in **Figures S1 and S2**).
 However, these changes did not translate into FCL because $C$ and $P$ occupy a similar trophic position (**Figure 1D**).
@@
 
-On average, increased disturbance frequency decreased FCL, and omnivory had similar effects on FCL (**Figures 3**).
+On average, increased disturbance frequency decreased FCL, and omnivory had similar effects on FCL (**Figure 3**).
 More importantly, a high disturbance regime produced different responses of FCL to ecosystem properties.
@@
 The effect of ecosystem complexity was generally clearer than that of ecosystem size.
-It is important to note that the positive relationships emerged in a high disturbance regime because smaller and simpler ecosystems were disproportionately vulnerable to increased disturbance frequency (compare far left in **Figures 2 and 3**).
+It is important to note that the positive relationships emerged in a high disturbance regime because smaller and simpler ecosystems were disproportionately vulnerable to increased disturbance frequency (compare far left along the x-axis in each panel of **Figures 2 and 3**).
 
@@
 Different processes drove the positive effects of ecosystem size and complexity on FCL when strong omnivory exists in low-productivity systems.
-In this scenario, the occupancies of $B$ and $C$ decreased at larger ecosystem sizes, while $P$ increased its occupancy (third column in **Figure S1 and S2**).
+In this scenario, the occupancies of $B$ and $C$ decreased at larger ecosystem sizes, while $P$ increased its occupancy (third column from left in **Figure S1 and S2**).
 Therefore, $P$ compensated for the loss of $B$ and $C$ to maintain the positive relationship between FCL and ecosystem size.
@@
 In contrast, ecosystem complexity has not been incorporated in theoretical explorations of FCL to date.
-Here, we show that FCL responds positively to increasing ecosystem size and complexity.
-This supports our hypothesis that ecosystem complexity may represent an overlooked yet important extrinsic control of FCL. However, the relationship between FCL and ecosystem complexity varied based on additional extrinsic and intrinsic factors, supporting the emerging paradigm that there is no single driver of FCL in ecosystems [@mchughDualInfluencesEcosystem2010; @wardMechanisticTheoryAquatic2017].
-Although we focus on river networks here, we expect our results to be transferable to other branching structures, such as mountain ranges and cave systems.
+Here, we show that FCL responds positively to increasing ecosystem size and complexity, here assessed as the number of patches and branching probability in river networks.
+This supports our hypothesis that ecosystem complexity may represent an overlooked yet important extrinsic control of FCL. However, the relationship between FCL and ecosystem complexity varied with other extrinsic and intrinsic factors, supporting the emerging paradigm that FCL is regulated by multiple factors [@mchughDualInfluencesEcosystem2010; @wardMechanisticTheoryAquatic2017].
+Although we focus on river networks here, we expect our results to be transferable to other systems with branching structures, such as mountain ranges and cave systems.
+Our findings should be relevant to non-branching systems, where different metrics of ecosystem complexity would be necessary.
 
-While increased disturbance frequency consistently reduced FCL as predicted by the dynamical stability hypothesis [@pimmNumberTrophicLevels1977], more important findings in this study were evident from their synergistic effects on ecosystem structure.
+While increased disturbance frequency consistently reduced FCL as predicted by the dynamical stability hypothesis [@pimmNumberTrophicLevels1977], more important findings in this study were evident from their synergistic effects with ecosystem structure.
 Our low disturbance regime scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independently of ecosystem size and complexity.
@@
 
-Our findings add an important distinction in the discussion of the suitability of assigning protection or conservation status to regions.
-Current conservation schemes generally emphasize size, aiming to protect the largest areas possible.
-While this is certainly a useful attribute generally, designating networks with a high degree of variability, or many small patches, can be more beneficial than a large, homogenous area [@fahrigWhySeveralSmall2020].
+Our findings add an important distinction to work to assess and prioritize protection or conservation status within human impacted regions.
+Classical conservation schemes emphasize ecosystem size, aiming to protect the largest areas possible [@diamondIslandDilemmaLessons1975].
+While this is certainly a useful attribute generally, designating networks with a high degree of variability, or many small patches, can be more beneficial than a large, homogeneous area [@fahrigWhySeveralSmall2020].
 Accounting for ecosystem complexity is particularly suited to ecosystem architectures in which branches have variable environmental conditions [@koningNetworkGrassrootsReserves2020; @teruiEmergentDualScaling2021].
-Conservation and management activities should focus on activities that increase the habitat heterogeneity and connectivity of suitable habitat patches to increase resilience to predicted increases in disturbance frequency and magnitude.
+As suggested by modern conservation initiatives, conservation and management activities should focus on activities that increase the habitat heterogeneity and connectivity of suitable habitat patches to increase resilience to predicted increases in disturbance frequency and magnitude [@schindlerPopulationDiversityPortfolio2010; @schindlerPortfolioConceptEcology2015a; @andersonPortfolioConservationMetapopulations2015].
 
@@
 
-JP developed the theoretical model of trophic dynamics, ran the simulations, and made the figures with input from AT.
+JP developed the theoretical model of trophic dynamics with input from AT.
+JP and AT ran the simulations, and made the figures.
 JP wrote the first draft, while AT and JF contributed significantly to the final version.

### 2022-05-10 rmd/figures.Rmd
edit: manuscript
edit: figures

@@
 output: pdf_document
+knit: (function(inputFile, encoding) {
+        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
+      })
 ---
@@
 Shaded ribbons are 95% confidence intervals.
-Parameters for low disturbance regime: disturbance probability $p_d = 0.01$ and mean proportional mortality at headwaters $\mu_{m}=0.8$.
+Parameters for low disturbance regime: disturbance probability $p_m = 0.01$ and mean proportional mortality at headwaters $\mu_{m}=0.8$.
 
@@
 Shaded ribbons are 95% confidence intervals.
-Parameters for high disturbance regime: disturbance probability $p_{d} = 0.01$ and mean proportional mortality at headwaters $\mu_{m}=0.8$.
+Parameters for high disturbance regime: disturbance probability $p_{m} = 0.01$ and mean proportional mortality at headwaters $\mu_{m}=0.8$.

### 2022-05-10 rmd/suppl_info.Rmd
Update suppl_info.Rmd
edit: figures

@@
 # Sensitivity data
+source(here::here("code/analysis_stvy.R"))
 load(file = here::here("output/result_stvy.RData"))
+
 sim_sv <- sim_stvy_result
@@
 We performed a sensitivity analysis of the trophic community simulation to identify key simulation parameters that strongly affect the relationships between food chain length and ecosystem properties (the number of habitat patches $N_{p}$ and branching probability $P_{b}$.
-We generated 500 sets of parameter combinations by randomly drawing values of 16 simulation parameters from uniform distributions (**Table S1**).
-For each parameter combination, we generated `r max(sim_stvy_result$n_rep)` branching networks with the gradients of ecosystem size ($N_p \sim Unif(10, 150)$) and complexity ($P_b \sim Unif(0.01, 0.99)$).
-This results in a total of `r nrow(sim_stvy_result)` simulation replicates.
-In each simulation replicate, we ran `r n_sim_stvy` time steps of trophic community dynamics.
+We generated 500 sets of parameter combinations by randomly drawing values of `r length(attributes(fit_sense$fit[[1]]$terms)$term.label)` simulation parameters from uniform distributions (**Table S1**).
+For each parameter combination, we generated `r max(sim_sv$n_rep)` branching networks with the gradients of ecosystem size ($N_p \sim Unif(10, 150)$) and complexity ($P_b \sim Unif(0.01, 0.99)$).
+This results in a total of `r nrow(sim_sv)` simulation replicates.
+We ran `r n_sim_stvy` time steps of trophic community dynamics in each simulation replicate.
 We obtained temporal means of food chain length for the last `r unique(sim_sv$n_timestep)` time steps.
@@
 ```
-where $\zeta_k~(k = 0...14)$ are the intercept and regression coefficients, and $x_{k,n}$ are the simulation parameters drawn from uniform distributions (**Table S1**).
-Explanatory variables were standardized to a mean of zero and a standard deviation of one, so that regression coefficients are comparable.
+where $\zeta_k$ are the intercept and regression coefficients, and $x_{k,n}$ are the simulation parameters drawn from uniform distributions (**Table S1**).
+Response and explanatory variables were standardized to a mean of zero and a standard deviation of one so that regression coefficients are comparable.
 
@@
 
-      figure_legend <- paste0(str_to_sentence(c_text[i])," effect on food chain length and occupancies of constituent species.
+      figure_legend <- paste0(str_to_sentence(c_text[i])," effect on food chain length and occupancies of constituent species (proportion of habitat patches occupied).
 Rows are different response variables, and columns are different simulation scenarios.

### 2022-05-10 code/figure_si_fcl.R

@@ -36,14 +36,14 @@ df_sim <- sim_main_result %>%
                names_to = "y",
                values_to = "value") %>% 
   mutate(y = case_when(y == "fcl" ~ "Food~chain~length",
-                       y == "p_basal" ~ "Basal~species~occupancy",
-                       y == "p_igprey" ~ "IG-prey~occupancy",
-                       y == "p_igpred" ~ "IG-predator~occupancy"),
+                       y == "p_basal" ~ "Basal~occupancy~(B)",
+                       y == "p_igprey" ~ "Consumer~occupancy~(C)",
+                       y == "p_igpred" ~ "Predator~occupancy~(P)"),
          y = factor(y,
                     levels = c("Food~chain~length",
-                               "Basal~species~occupancy",
-                               "IG-prey~occupancy",
-                               "IG-predator~occupancy")))
+                               "Basal~occupancy~(B)",
+                               "Consumer~occupancy~(C)",
+                               "Predator~occupancy~(P)")))
 
 df_param <- expand.grid(theta = c(0.1, 1),
                         sd_disturb_source = c(0.1, 3),
@@ -72,12 +72,14 @@ list_g_np <- foreach(i = seq_len(nrow(df_param))) %do% {
     facet_grid(rows = vars(y),
                cols = vars(disturb, productivity),
                labeller = label_parsed,
-               scales = "free_y") +
+               scales = "free_y",
+               switch = "y") +
     labs(x = "Ecosystem size (number of patches)",
-         y = "Value",
+         y = "",
          linetype = "Omnivory") +
     MetBrewer::scale_color_met_d("Hiroshige") +
-    MetBrewer::scale_fill_met_d("Hiroshige")
+    MetBrewer::scale_fill_met_d("Hiroshige") +
+    theme(strip.placement = "outside")
   
   return(g_np)
 }
@@ -98,12 +100,14 @@ list_g_pb <- foreach(i = seq_len(nrow(df_param))) %do% {
     facet_grid(rows = vars(y),
                cols = vars(disturb, productivity),
                labeller = label_parsed,
-               scales = "free_y") +
+               scales = "free_y",
+               switch = "y") +
     labs(x = "Ecosystem complexity (branching prob.)",
-         y = "Value",
+         y = "",
          linetype = "Omnivory") +
     MetBrewer::scale_color_met_d("Hiroshige") +
-    MetBrewer::scale_fill_met_d("Hiroshige")
[…]

### 2022-05-10 rmd/cover_letter.Rmd
edit: author info

---
output: pdf_document
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
---

Dear Editor,

We are excited to submit our manuscript entitled "Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks" for consideration for publication in The American Naturalist.

The structure of food webs in natural communities has inspired ecological research for generations, and much attention has been paid to understanding the determinants of food chain length. Generally, ecosystem size is thought to be a universal control of food chain length. In spatially complex ecosystems, however, there may be an overlooked effect of 'scale-invariant' complexity in fractal ecosystem structures. Branching river networks are an outstanding example of such complex ecosystems, where the geometric structure (e.g., branching pattern) remains constant across size gradients.

In this research, we developed a novel simulation model to investigate the influences of ecosystem size and complexity on food chain length in branching habitat networks. We found that ecosystem size and complexity both regulate food chain length (the "dual control"), but the strength of these regulations is modulated by disturbance regimes - the dual control was more apparent when disturbance is strong and frequent. However, other factors - such as food web structure and primary productivity - did not influence the qualitative relationships of food chain length and ecosystem structures, raising the possibility that the dual control of food chain length may emerge across a wide range of productivity and in diverse food webs.

Although ecosystems are inherently complex in space, there have been very limited efforts to investigate the role of ecosystem complexity in regulating food chain length. Our findings will stimulate future research addressing this critical research topic. Thus, our manuscript represents a novel contribution to food web ecology. In addition, the connection between network complexity and food web structure will be of interest to an interdisciplinary audience, including scientists working in fluvial geomorphology, ecological networks, and macrosystems ecology.

We suggest the following leading scientists in the field as potential reviewers:

-   Dr. Gaku Takimoto (University of Tokyo; [gakut\@es.a.u-tokyo.ac.jp](mailto:gakut@es.a.u-tokyo.ac.jp){.email}): Works on food web theory

-   Dr. Colette Ward (NCEAS, [ward\@nceas.ucsb.edu](mailto:ward@nceas.ucsb.edu){.email}): Works on theoretical food web ecology

-   Dr. Justin Yeakel (University of California Merced; [jdyeakel\@gmail.com](mailto:jdyeakel@gmail.com){.email}): Develops theoretical models for ecological communities, including species interactions in a spatial context

-   Dr. Kevin McCann (University of Guelph, [ksmccann\@uoguelph.ca](mailto:ksmccann@uoguelph.ca){.email}): Works on food web theory

-   Dr. Jean Gibert (Duke University; [jean.gibert\@duke.edu](mailto:jean.gibert@duke.edu){.email}): Develops general theoretical models for ecological communities

This work is original and is not under consideration in any other journal. All authors agree with the content and with the submission for publication, and all appropriate ethical and other necessary approvals were obtained for this research. We look forward to your reply and hope that you find our article suitable for publication in your journal.

Sincerely,

Justin Pomeranz, Jacques Finlay, Akira Terui

University of North Carolina Greensboro

\pagebreak

# Author information

Author: Justin P.F. Pomeranz^1,2^, Jacques C. Finlay^3^, Akira Terui^1^

1 Department of Biology, University of North Carolina at Greensboro; 2 Department of Environmental Sciences, Colorado Mesa University; 3 Department of Ecology, Evolution and Behavior, University of Minnesota

\*Corresponding Author: JPF Pomeranz, [jfpomeranz\@gmail.com](mailto:jfpomeranz@gmail.com){.email}; A Terui, [hanabi0111\@gmail.com](mailto:hanabi0111@gmail.com){.email}

## Acknowledgements

We thank C Perovich for thoughtful discussions in the early stages of this project. This material is based upon work supported by the National Science Foundation through the Division of Environmental Biology (DEB 2015634).

## Author contributions

JP developed the theoretical model of trophic dynamics with input from AT. JP and AT ran the simulations, and made the figures. JP wrote the first draft, while AT and JF contributed significantly to the final version.

## Data availability

No data were used or produced in this research. R functions for simulations are provided as the R package mcbrnet version 1.2.3 (available at <https://github.com/aterui/mcbrnet>). Codes for simulations, statistical analysis, figures, and tables are available at <https://github.com/aterui/public-proj_fcl-river-theory>.

### 2022-05-10 rmd/manuscript.Rmd
edit: author info

@@
 ```
-
-**Author:** Justin P.F. Pomeranz^1,2,\*^, Jacques C. Finlay^3^, Akira Terui^1,\*^
-
-^1^ Department of Biology, University of North Carolina at Greensboro; ^2^ Department of Environmental Sciences, Colorado Mesa University; ^3^ Department of Ecology, Evolution and Behavior, University of Minnesota
-
-**\*Corresponding Author:** JPF Pomeranz, jfpomeranz\@gmail.com; A Terui, hanabi0111\@gmail.com
 
@@
 Note that simulation replicates are differ in ecosystem size ($N_{p}\sim Unif(10,150)$) and ecosystem complexity ($P_{b} \sim Unif(0.01, 0.99)$) so we can estimate their effects on FCL.
+All analysis was performed in R 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2021]
 
@@
 
-# **Acknowledgements**
-
-We thank C Perovich for thoughtful discussions in the early stages of this project.
-This material is based upon work supported by the National Science Foundation through the Division of Environmental Biology (DEB 2015634).
-
-# **Author contributions**
-
-JP developed the theoretical model of trophic dynamics with input from AT.
-JP and AT ran the simulations, and made the figures.
-JP wrote the first draft, while AT and JF contributed significantly to the final version.
-
-# **Data availability**
-
-No data were used or produced in this research.
-R functions for simulations are provided as the R package *mcbrnet* (available at <https://github.com/aterui/mcbrnet>).
-Codes for simulations, statistical analysis, figures, and tables are available at <https://github.com/aterui/public-proj_fcl-river-theory>.
-
 # References

### 2022-05-10 rmd/manuscript.Rmd
edit: misc

@@
 This autoregressive process resembles a natural disturbance regime, where flood or drought disturbance cascades downstream with some local variation along the channel [@nakamuraDisturbanceRegimesStream2000].
-At confluence patches, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, $Q$ (the number of upstream contributing patches):
+At confluence patches, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, $Q$ (the number of upstream contributing patches + 1):
 
@@
 Thus, the larger tributary has a greater influence on the downstream environment, as observed in natural river networks [@bendaNetworkDynamicsHypothesis2004].
-Disturbance value distributions had medians close to $m$ with long tails (**Figure 1B**).
+Disturbance value distributions had medians close to $\mu_m$ with long tails (**Figure 1B**).
 The low-magnitude patches can be viewed as refugia, which supply individuals to impacted sites in subsequent time steps.
@@
 $$
-K_{x} = K_{base}(N_{p,up} + 1)^z
+K_{x} = K_{base}Q_{x}^z
 $$
 
-where $K_{base}$ is the carrying capacity at the upstream terminal patch, $N_{p,up}$ the number of patches upstream, and $z$ the scaling exponent.
+where $K_{base}$ is the carrying capacity at the upstream terminal patch, and $z$ the scaling exponent.
+In this formulation, the most upstream patch, at which $Q_x = 1$ (the number of upstream patches + 1), has the carrying capacity of $K_{base}$.
 We used $K_{base} = 100$ and $z = 0.54$ based on an empirical relationship of gross primary production and watershed area [@finlayStreamSizeHuman2011].

### 2022-05-10 rmd/suppl_info.Rmd
edit: misc

@@
 source(here::here("code/analysis_stvy.R"))
+load(file = here::here("output/result_main.RData"))
 load(file = here::here("output/result_stvy.RData"))
@@
 
-Based on the results, we considered ecological scenarios relevant for putative controls of food chain: low and high productivity ($r_b = 4, 8$), low and high disturbance ($p_m = 0.01, 0.1$), and three food web architecture ("Chain" [$\{e_{BC}, e_{BP}, e_{CP}\} = \{4, 0, 4\}, \{a_{BC}, a_{BP}, a_{CP}\} = \{0.5, 0.1, 0.5\}$], "Weak omnivory" [$\{e_{BC}, e_{BP}, e_{CP}\} = \{4, 2, 4\}, \{a_{BC}, a_{BP}, a_{CP}\} = \{0.5, 0.1, 0.5\}$], and "Strong omnivory" [$\{e_{BC}, e_{BP}, e_{CP}\} = \{4, 4, 2\}, \{a_{BC}, a_{BP}, a_{CP}\} = \{0.5, 0.5, 0.1\}$]).
-Handling times were fixed at $0.5$ to reduce the number of parameter combinations.
+Based on the results, we considered ecological scenarios relevant for putative controls of food chain: low and high productivity ($r_b = 4, 8$), and low and high disturbance ($p_m = 0.01, 0.1$).
+We also investigated the role of omnivory by changing attack rate ($a_{BP}$ and $a_{CP}$) and conversion efficiency ($e_{BP}$ and $e_{CP}$) of two prey items for predator $P$.
+Specifically, we used the following combinations to represent three levels of omnivory: "Chain", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[1], max(sim_main_result$e_cp))` $\}$; "Weak omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[2], max(sim_main_result$e_cp))` $\}$; and "Strong omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(max(sim_main_result$a_bp), min(sim_main_result$a_cp), unique(sim_main_result$e_bp)[3], min(sim_main_result$e_cp))` $\}$.
+To simplify simulation setup, we fixed the following parameters: $e_{BC}=$ `r unique(sim_main_result$e_bc)` , $a_{BC}=$ `r unique(sim_main_result$a_bc)`, and $h_{ij}=$ `r unique(sim_main_result$h_bc)`.
 

### 2022-05-10 rmd/suppl_info.Rmd
Update suppl_info.Rmd

@@
 The response variable is the Spearman's rank correlation between food chain length and ", c_text[i],".
-Response and explanatory variables (i.e., simulation parameters) were standardized (mean = 0, SD = 1) before the analysis.")
+Response and explanatory variables (i.e., simulation parameters) were standardized (mean = 0, SD = 1) before the analysis.\n")
 
-  knit_chunk <- paste0("## Table S", i + j, " Sensitivity analysis for ", c_text[i]," \n",
+  knit_chunk <- paste0("## Table S", i + j, " Sensitivity analysis for ", c_text[i],
+                       "\n\n",
                        table_legend,
-                       "\n```{r table_s", i + j,", results='asis'}
+                       "```{r table_s", i + j,", results='asis'}
                        \n\n
                        table_stvy[[", i,"]]
-                       \n\n
-                       ```\n",
+                       \n\n",
+                       "```",
                        "\n\\pagebreak\n")
@@
 
-  knit_chunk <- paste0("## Figure S", K * (i - 1) + k + j, " ", str_to_sentence(c_text[i]), " effect on food chain length", " \n",
-                       "\n```{r figure_s", K * (i - 1) + k + j,", fig.height = 10, fig.width = 11.5}
-                       \n\n
+  knit_chunk <- paste0("## Figure S", K * (i - 1) + k + j, " ", str_to_sentence(c_text[i]), " effect on food chain length",
+                       "\n\n",
+                       "```{r figure_s", K * (i - 1) + k + j,", fig.height = 10, fig.width = 11.5}",
+                       "\n\n
                        list_g_fcl[[", i, "]][[", k, "]]
-                       \n\n
-                       ```\n",
+                       \n\n",
+                       "```",
+                       "\n\n",
                        figure_legend,

### 2022-05-10 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 In contrast, ecosystem complexity has not been incorporated in theoretical explorations of FCL to date.
-Here, we show that FCL responds positively to increasing ecosystem size and complexity, here assessed as the number of patches and branching probability in river networks.
+Here, we show that FCL responds positively to increasing ecosystem size and complexity, assessed as the number of patches and branching probability in river networks.
 This supports our hypothesis that ecosystem complexity may represent an overlooked yet important extrinsic control of FCL. However, the relationship between FCL and ecosystem complexity varied with other extrinsic and intrinsic factors, supporting the emerging paradigm that FCL is regulated by multiple factors [@mchughDualInfluencesEcosystem2010; @wardMechanisticTheoryAquatic2017].
 Although we focus on river networks here, we expect our results to be transferable to other systems with branching structures, such as mountain ranges and cave systems.
-Our findings should be relevant to non-branching systems, where different metrics of ecosystem complexity would be necessary.
+Our findings should also be relevant to non-branching systems, where different metrics of ecosystem complexity would be necessary.
 
@@
 However, under high disturbance regimes, only large and complex networks could provide diverse and heterogeneous habitats with potential refugia, allowing species to thrive and maintain long food chains.
-While empirical studies rarely looked at the influence of ecosystem structure on FCL, several studies provide some support for this theoretical prediction.
+While empirical studies rarely looked at the influence of ecosystem complexity on FCL, several studies provide some support for this theoretical prediction.
 For example, @sullivanEcosystemStructureEmerges2015 reported FCL is controlled primarily by the number of tributary confluences in Idaho streams.

### 2022-05-10 code/figure_si_fcl.R
Update figure_si_fcl.R

@@ -76,7 +76,8 @@ list_g_np <- foreach(i = seq_len(nrow(df_param))) %do% {
                switch = "y") +
     labs(x = "Ecosystem size (number of patches)",
          y = "",
-         linetype = "Omnivory") +
+         color = "Omnivory",
+         fill = "Omnivory") +
     MetBrewer::scale_color_met_d("Hiroshige") +
     MetBrewer::scale_fill_met_d("Hiroshige") +
     theme(strip.placement = "outside")
@@ -104,7 +105,8 @@ list_g_pb <- foreach(i = seq_len(nrow(df_param))) %do% {
                switch = "y") +
     labs(x = "Ecosystem complexity (branching prob.)",
          y = "",
-         linetype = "Omnivory") +
+         color = "Omnivory",
+         fill = "Omnivory") +
     MetBrewer::scale_color_met_d("Hiroshige") +
     MetBrewer::scale_fill_met_d("Hiroshige") +
     theme(strip.placement = "outside")

### 2022-05-22 rmd/cover_letter.Rmd
edit: resubmit to Ecosystems

@@
 University of North Carolina Greensboro
-
-\pagebreak
-
-# Author information
-
-Author: Justin P.F. Pomeranz^1,2^, Jacques C. Finlay^3^, Akira Terui^1^
-
-1 Department of Biology, University of North Carolina at Greensboro; 2 Department of Environmental Sciences, Colorado Mesa University; 3 Department of Ecology, Evolution and Behavior, University of Minnesota
-
-\*Corresponding Author: JPF Pomeranz, [jfpomeranz\@gmail.com](mailto:jfpomeranz@gmail.com){.email}; A Terui, [hanabi0111\@gmail.com](mailto:hanabi0111@gmail.com){.email}
-
-## Acknowledgements
-
-We thank C Perovich for thoughtful discussions in the early stages of this project.
-This material is based upon work supported by the National Science Foundation through the Division of Environmental Biology (DEB 2015634).
-
-## Author contributions
-
-JP developed the theoretical model of trophic dynamics with input from AT.
-JP and AT ran the simulations, and made the figures.
-JP wrote the first draft, while AT and JF contributed significantly to the final version.
-
-## Data availability
-
-No data were used or produced in this research.
-R functions for simulations are provided as the R package mcbrnet version 1.2.3 (available at <https://github.com/aterui/mcbrnet>).
-Codes for simulations, statistical analysis, figures, and tables are available at <https://github.com/aterui/public-proj_fcl-river-theory>.

### 2022-05-22 rmd/manuscript.Rmd
edit: resubmit to Ecosystems

@@
 bibliography: references.bib
-csl: the-american-naturalist.csl
+csl: ecosystems.csl
 header-includes:
@@
 ```
+
+Author: Justin P.F. Pomeranz^1,2^, Jacques C. Finlay^3^, Akira Terui^1^
+
+1 Department of Biology, University of North Carolina at Greensboro; 2 Department of Environmental Sciences, Colorado Mesa University; 3 Department of Ecology, Evolution and Behavior, University of Minnesota
+
+\*Corresponding Author: JPF Pomeranz, [jfpomeranz\@gmail.com](mailto:jfpomeranz@gmail.com){.email}; A Terui, [hanabi0111\@gmail.com](mailto:hanabi0111@gmail.com){.email}
 
@@
 
+## Acknowledgements
+
+We thank C Perovich for thoughtful discussions in the early stages of this project.
+This material is based upon work supported by the National Science Foundation through the Division of Environmental Biology (DEB 2015634).
+
+## Author contributions
+
+JP developed the theoretical model of trophic dynamics with input from AT.
+JP and AT ran the simulations, and made the figures.
+JP wrote the first draft, while AT and JF contributed significantly to the final version.
+
+## Data availability
+
+No data were used or produced in this research.
+R functions for simulations are provided as the R package mcbrnet version 1.2.3 (available at [<https://github.com/aterui/mcbrnet>](https://github.com/aterui/mcbrnet){.uri}).
+Codes for simulations, statistical analysis, figures, and tables are available at [<https://github.com/aterui/public-proj_fcl-river-theory>](https://github.com/aterui/public-proj_fcl-river-theory){.uri}.
+
 # References

### 2022-06-02 rmd/manuscript.Rmd
edit minor

@@
 No data were used or produced in this research.
-R functions for simulations are provided as the R package mcbrnet version 1.2.3 (available at [<https://github.com/aterui/mcbrnet>](https://github.com/aterui/mcbrnet){.uri}).
-Codes for simulations, statistical analysis, figures, and tables are available at [<https://github.com/aterui/public-proj_fcl-river-theory>](https://github.com/aterui/public-proj_fcl-river-theory){.uri}.
+R functions for simulations are provided as the R package mcbrnet version 1.2.3 (available at [\<https://github.com/aterui/mcbrnet\>](https://github.com/aterui/mcbrnet){.uri}).
+Codes for simulations, statistical analysis, figures, and tables are available at [\<https://github.com/aterui/public-proj_fcl-river-theory\>](https://github.com/aterui/public-proj_fcl-river-theory){.uri}.
 

### 2022-10-26 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 Branches (tributaries) are a series of connected patches between confluences (or upstream terminal patches).
-The number of patches within branch $q$, $n_{p,q}$, is a random variable drawn from a geometric distribution $n_{p,q} \sim Ge(P_{b})$ but conditional on $N_{p} = \sum_{q}^{N_{b}} n_{p,q}$ ($N_{b}$ is the number of branches in a network).
+The number of patches within branch $q$, $n_{p,q}$, is a random variable drawn from a geometric distribution $n_{p,q} \sim \text{Ge}(P_{b})$ but conditional on $N_{p} = \sum_{q}^{N_{b}} n_{p,q}$ ($N_{b}$ is the number of branches in a network).
 Full details of the network generation process is described in @teruiEmergentDualScaling2021.
@@
 We simulated the spatial distribution of disturbance intensity (i.e., proportional mortality in the event of stochastic regional disturbance) as follows.
-A disturbance value, $m_{h}$, was assigned randomly to upstream terminal patch $h$ as $logit~m_{h} \sim Normal(logit~\mu_m, \sigma_{h}^2)$, where $\mu_m$ is the average of disturbance intensity at headwaters in an ordinary scale.
+A disturbance value, $m_{h}$, was assigned randomly to upstream terminal patch $h$ as $\text{logit}~m_{h} \sim \text{Normal}(\text{logit}~\mu_m, \sigma_{h}^2)$, where $\mu_m$ is the average of disturbance intensity at headwaters in an ordinary scale.
 The disturbance values propagate downstream through a spatial autoregressive process as:
@@
 $$
-logit~m_{down} = logit~m_{up}+ \epsilon
+\text{logit}~m_{down} = \text{logit}~m_{up}+ \varepsilon
 $$
 
-where $m_{down}$ and $m_{up}$ are the disturbance values at down- and upstream patches, and the local environmental noise at the downstream patch $\epsilon$ was drawn from $Normal(0, \sigma_{l}^2)$.
+where $m_{down}$ and $m_{up}$ are the disturbance values at down- and upstream patches, and the local environmental noise at the downstream patch $\varepsilon$ was drawn from $\text{Normal}(0, \sigma_{l}^2)$.
 This autoregressive process resembles a natural disturbance regime, where flood or drought disturbance cascades downstream with some local variation along the channel [@nakamuraDisturbanceRegimesStream2000].
@@
 $$
-logit~m_{down} = \omega~(logit~m_{1,up} +\epsilon_1)+(1-\omega)~(logit~m_{2,up} +\epsilon_2)
+\text{logit}~m_{down} = \omega~(\text{logit}~m_{1,up} +\varepsilon_1)+(1-\omega)~(\text{logit}~m_{2,up} +\varepsilon_2)
 $$
 
-$m_{1,up}$ and $m_{2,up}$ are the disturbance values of upstream contributing tributaries, $\epsilon_1$ and $\epsilon_2$ are independent local noises, and $\omega$ the relative size of one tributary over another defined as $\omega = \frac{Q_{1}}{Q_{1} + Q_{2}}$.
+$m_{1,up}$ and $m_{2,up}$ are the disturbance values of upstream contributing tributaries, $\varepsilon_1$ and $\varepsilon_2$ are independent local noises, and $\omega$ the relative size of one tributary over another defined as $\omega = \frac{Q_{1}}{Q_{1} + Q_{2}}$.
 Thus, the larger tributary has a greater influence on the downstream environment, as observed in natural river networks [@bendaNetworkDynamicsHypothesis2004].
@@
 $$
-A_{i,x}(t+1) \sim Poisson(\bar{A}_{i,x}(t) + I_{i,x}(t) - E_{i,x}(t))
+A_{i,x}(t+1) \sim \text{Poisson}(\bar{A}_{i,x}(t) + I_{i,x}(t) - E_{i,x}(t))
 $$
@@
 $$
-\xi_{i,x}(t)=\frac{\sum_{y, y \ne x}E_{i,y}(t)exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}E_{i,y}(t)exp(-\theta d_{xy})}
+\xi_{i,x}(t)=\frac{\sum_{y, y \ne x}E_{i,y}(t)\exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}\text{E}_{i,y}(t)\exp(-\theta d_{xy})}
 $$

### 2022-10-31 rmd/manuscript.Rmd
edit

@@
 output:
-  #pdf_document:
-  #  latex_engine: xelatex
-  word_document:
-    reference_docx: word_format.docx
+  pdf_document:
+    latex_engine: xelatex
+  #word_document:
+  #  reference_docx: word_format.docx
 knit: (function(inputFile, encoding) {
@@
 A discrete model is suitable for organisms with predictable reproduction cycles, a common life history for many lotic species, including macroinvertebrates and fish.
-Our simulation model consists of three species communities; a basal species $B$ with positive logistic growth up to a carrying capacity, $K$; a primary consumer species or intraguild prey, $C$, which grows based on the consumption of $B$; and an intraguild predator $P$, which can consume both $B$ and $C$ (**Figure 1C**).
+Our simulation model consists of three species; a basal species $B$ with positive logistic growth up to a carrying capacity, $K$; a primary consumer species or intraguild prey, $C$, which grows based on the consumption of $B$; and an intraguild predator $P$, which can consume both $B$ and $C$ (**Figure 1C**).
 In our model, the realized species abundance $A_{i,x}(t+1)$ for species i (either $B,C,~or~P$) at patch $x$ and time $t+1$ is expressed as a composite of within-patch trophic interactions, immigration, and emigration:
@@
 
-where $\bar{A}_{i,x}(t+1)$ is the expected species abundance after within-patch trophic interactions, $I_{i,x}(t)$ the expected number of immigrants into patch $x$, and $E_{i,x}(t)$ the expected number of emigrants from patch $x$.
+where $\bar{A}_{i,x}(t)$ is the expected species abundance after within-patch trophic interactions, $I_{i,x}(t)$ the expected number of immigrants into patch $x$, and $E_{i,x}(t)$ the expected number of emigrants from patch $x$.
 The realized discrete number of individuals is drawn from a Poisson distribution to account for stochasticity in demographic and dispersal processes.
 
-***Internal dynamics.*** Each time step starts with internal dynamics. The basal species is exposed to predation as well as other sources of mortality:
+***Internal dynamics.*** Each time step starts with internal dynamics.
+We accounted for background mortality, predator-induced mortality, and disturbance-induced mortality before the density-dependent growth occurs:
 
 $$
-A_{B,x}'(t) = s_{0,B}[1-z_{B,x}(t)]A_{B,x}(t)
+\begin{aligned}
+A'_{i,x} &= s_{0,i}~\beta_{i,x}(t)~A_{i,x}(t)\\
+\bar{A}_{i,x}(t) &= g_i(A'_{i,x})~A'_{i,x}~\gamma_{x}(t)
+\end{aligned}
 $$
 
-where $s_{0,i}$ is the background survival probability (density independent) and $z_{B,x}(t)$ the proportional mortality caused by predation.
-Let $W_{B,x}(t)$ be the total number of basal species consumed by $C$ and $P$.
-The predator-induced mortality $z_{B,x}(t)$, the sum of proportional predation by $C$ ($z_{BC,x}(t)$) and $P$ ($z_{BP,x}(t)$), is then:
+where $s_{0,i}$ is the background survival probability (density independent), $\beta_{i,x}(t)$ the survival probability after predation, and $\gamma_{x}(t)$ the survival probability after stochastic disturbance.
+The function $g_i(\centerdot)$ defines the form of the density-dependent recruitment for species $i$.
+
+For the basal species, we considered the predation impacts by consumer and predator species:
 
 $$
-z_{B,x}(t) = z_{BC,x}(t) + z_{BP,x}(t) =\left\{\begin{aligned}&\frac{W_{B,x}(t)}{A_{B,x}(t)} &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\&1 &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
+\beta_{B,x}(t) = 1 - [z_{BC,x}(t) + z_{BP,x}(t)]
 $$
-
-and
+where $z_{BC,x}(t)$ and $z_{BP,x}(t)$ are the proportional mortality of basal species caused by consumer and predator species, respectively.
+The total mortality is calculated as the proportion of total individuals consumed ($W_{B,x}(t)=W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)$) relative to the current abundance of the basal species ($A_{B,x}(t)$):
 
 $$
-W_{B,x}(t) = W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)
+z_{BC,x}(t) + z_{BP,x}(t) =\left\{\begin{aligned}&\frac{W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)}{A_{B,x}(t)} &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\&1 &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
 $$
@@
 $$
-\hat{A}_{B,x}(t) = \frac{r_{b} A_{B,x}'(t)}{1 + \frac{r_b - 1}{K_{x}}A_{B,x}'(t)}
+g_B(A'_{B,x}) = \frac{r_{b}} {1 + \frac{r_b - 1}{K_{x}}A_{B,x}'}
 $$
@@
 
-The population growth of the consumer is controlled by conversion efficiency, number of prey captured, background mortality, and predation:
+For the consumer species, the predation-induced mortality is the proportion of individuals consumed by the predator species:
 
 $$
-\hat{A}_{C,x}(t) =\left\{\begin{aligned}&s_{0,C} (1 - z_{CP,x}) e_{BC}W_{BC,x}(t) &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\&s_{0,C} (1 - z_{CP,x}) e_{BC} \left( \frac{W_{BC,x}(t)}{W_{B,x}(t)} A_{B,x} \right) &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
+\beta_{CP,x}(t) = 1 - z_{CP,x} = 1 - \frac{W_{CP,x}}{A_{C,x}(t)}
+$$
+
+The population growth of the consumer is controlled by conversion efficiency and the number of prey captured:
+
+$$
+g_C(A'_{C,x}) =
+\left\{
+\begin{aligned}
+&e_{BC}W_{BC,x}(t) &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
+&e_{BC} \left( \frac{W_{BC,x}(t)}{W_{B,x}(t)} A_{B,x} \right) &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
 $$
@@
 
-The population of the predator is fueled by two prey species:
+The predator species is not exposed to predation, thus $\beta_{P,x}(t) = 1$. The population growth of the predator considers two prey species:
 
 $$
-\hat{A}_{P,x}(t) =\left\{\begin{aligned}&s_{0,P} \left[e_{BP} \delta_{x}(t) W_{BP,x}(t) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) \right] &&\text{if~$W_{B,x}(t) < A_{B,x}(t)$}\\&s_{0,P} \left[e_{BP} \delta_{x}(t)\left(\frac{W_{BP,x}(t)}{W_{B,x}(t)}A_{B,x}\right) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) \right] &&\text{if~$W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
+g_P(A'_{P,x}) =
+\left\{
+\begin{aligned}
+&e_{BP} \delta_{x}(t) W_{BP,x}(t) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
+&e_{BP} \delta_{x}(t)\left(\frac{W_{BP,x}(t)}{W_{B,x}(t)}A_{B,x}\right) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\
+\end{aligned}
+\right.
 $$
@@
 $$
-\bar{A}_{i,x}(t) = [1 - \psi(t) m_{x}] \hat{A}_{i,x}(t)
+\gamma_{i,x}(t) = [1 - \psi(t) m_{x}]
 $$
@@
 where $d_{xy}$ is the network distance between patch $x$ and $y$ (shortest path), and $\theta$ indicates the inverse of mean dispersal distance (i.e., larger numbers mean shorter dispersal distance).
-The immigration potential at patch $x$, $\sum_{y, y \ne x}E_{i,y}(t)exp(-\theta d_{xy})$, is scaled by the sum of immigration potential across patches so that $\sum_{x} \xi_{i,x}=1$.
+The immigration potential at patch $x$, $\sum_{y, y \ne x}E_{i,y}(t)\exp(-\theta d_{xy})$, is scaled by the sum of immigration potential across patches so that $\sum_{x} \xi_{i,x}=1$.
 The expected number of immigrants into patch $x$ is calculated as $I_{i,x}(t)=\xi_{i,x}(t) \sum_{x}E_{i,x}(t)$.

### 2022-10-31 rmd/manuscript.Rmd
edit

@@
 csl: ecosystems.csl
+mainfont: Times New Roman
 header-includes:
@@
 Our simulation model consists of three species; a basal species $B$ with positive logistic growth up to a carrying capacity, $K$; a primary consumer species or intraguild prey, $C$, which grows based on the consumption of $B$; and an intraguild predator $P$, which can consume both $B$ and $C$ (**Figure 1C**).
-In our model, the realized species abundance $A_{i,x}(t+1)$ for species i (either $B,C,~or~P$) at patch $x$ and time $t+1$ is expressed as a composite of within-patch trophic interactions, immigration, and emigration:
+In our model, the realized species abundance $A_{i,x}(t+1)$ for species $i$ (either $B,C,~or~P$) at patch $x$ and time $t+1$ is expressed as a composite of within-patch trophic interactions, immigration, and emigration:
 
@@
 The function $g_i(\centerdot)$ defines the form of the density-dependent recruitment for species $i$.
+Below, we explain how we described predation and recruitment processes for each species.
 
-For the basal species, we considered the predation impacts by consumer and predator species:
+For the basal species, we considered the predation impacts of consumer and predator species:
 
@@
 
-$r_b$ is the reproductive rate for survived individuals.
+where $r_b$ is the reproductive rate for survived individuals.
 
@@
 Although dispersal has received less attention in food chain research [but see @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012a; @wangMetapopulationCapacityDetermines2021], our simulation highlighted food webs in these networks are more vulnerable to habitat fragmentation because habitat patches are inherently isolated from one another.
-Dams and road crossings have altered the connectivity between suitable habitat patches or have completely blocked dispersal [@ishiyamaPredictingEcologicalImpacts2018; @moritaEffectsHabitatFragmentation2002; @perkinFragmentationAltersStream2012a].
+Dams and road crossings have altered the connectivity between suitable habitat patches or have completely blocked dispersal [@ishiyamaPredictingEcologicalImpacts2018; @moritaEffectsHabitatFragmentation2002; @perkinFragmentationAltersStream2012].
 Light pollution has also negatively affected the ability of adult aquatic insects to disperse between habitat patches [@perkinEffectsArtificialLighting2014].
@@
 Accounting for ecosystem complexity is particularly suited to ecosystem architectures in which branches have variable environmental conditions [@koningNetworkGrassrootsReserves2020; @teruiEmergentDualScaling2021].
-As suggested by modern conservation initiatives, conservation and management activities should focus on activities that increase the habitat heterogeneity and connectivity of suitable habitat patches to increase resilience to predicted increases in disturbance frequency and magnitude [@schindlerPopulationDiversityPortfolio2010; @schindlerPortfolioConceptEcology2015a; @andersonPortfolioConservationMetapopulations2015].
+As suggested by modern conservation initiatives, conservation and management activities should focus on activities that increase the habitat heterogeneity and connectivity of suitable habitat patches to increase resilience to predicted increases in disturbance frequency and magnitude [@schindlerPopulationDiversityPortfolio2010; @schindlerPortfolioConceptEcology2015; @andersonPortfolioConservationMetapopulations2015].
 

### 2022-10-31 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 Thus, habitats with frequent and/or severe disturbance should support fewer trophic levels [@pimmNumberTrophicLevels1977].
-Lastly, the ecosystem size hypothesis posits that larger ecosystems should harbor longer food chains [@takimotoEffectsProductivityDisturbance2012a; @teruiSpatialDisturbanceSynchrony2019; @postEcosystemSizeDetermines2000].
+Lastly, the ecosystem size hypothesis posits that larger ecosystems should harbor longer food chains [@takimotoEffectsProductivityDisturbance2012; @teruiSpatialDisturbanceSynchrony2019; @postEcosystemSizeDetermines2000].
 
@@
 Meanwhile, the theoretical basis for the other two hypotheses rests on implicit assumptions of linear food chains, which are often violated in natural systems.
-As such, the effects of disturbance and resource availability on FCL have been found to be highly variable among ecosystem types and geographic regions [@takimotoEffectsProductivityDisturbance2012a; @diehlIntraguildPreySuffer2001; @takimotoEnvironmentalDeterminantsFoodchain2013; @postLongShortFoodchain2002].
+As such, the effects of disturbance and resource availability on FCL have been found to be highly variable among ecosystem types and geographic regions [@takimotoEffectsProductivityDisturbance2012; @diehlIntraguildPreySuffer2001; @takimotoEnvironmentalDeterminantsFoodchain2013; @postLongShortFoodchain2002].
 
@@
 Such context-dependent responses to extrinsic factors seem to be tightly coupled with the arrangement of trophic links within a food web.
-For example, omnivory (feeding at multiple trophic levels) is a strong intrinsic driver of FCL [@takimotoEffectsProductivityDisturbance2012a; @postProximateStructuralMechanisms2007; @woottonOmnivoryStabilityFreshwater2017].
+For example, omnivory (feeding at multiple trophic levels) is a strong intrinsic driver of FCL [@takimotoEffectsProductivityDisturbance2012; @postProximateStructuralMechanisms2007; @woottonOmnivoryStabilityFreshwater2017].
 When omnivory is absent in three-species communities, the trophic structure assumes a simple linear food chain with a maximum length of three (**Figure 1D**).
@@
 Branching ecosystems, for example, are ubiquitous yet overlooked landscape structures (e.g., rivers, trees, caves, mountain ranges) [@grantLivingBranchesPopulation2007; @altermattDiversityRiverineMetacommunities2013].
-While we know that food web interactions are spatially influenced [@calcagnoConstraintsFoodChain2011; @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012a; @teruiSpatialDisturbanceSynchrony2019], the theory to account for it has lagged, particularly in branching habitat architectures.
+While we know that food web interactions are spatially influenced [@calcagnoConstraintsFoodChain2011; @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012; @teruiSpatialDisturbanceSynchrony2019], the theory to account for it has lagged, particularly in branching habitat architectures.
 Here, we theoretically explore drivers of FCL in rivers, an excellent example of branching ecosystems.
@@
 
-The ecosystem size hypothesis of FCL has received support from both theoretical and empirical research [@vanderzandenPatternsFoodChain1999; @postEcosystemSizeDetermines2000; @takimotoEffectsProductivityDisturbance2012a], although empirical work so far has rarely elaborated on the underlying drivers [but see @saboRoleDischargeVariation2010; @wardMechanisticTheoryAquatic2017].
+The ecosystem size hypothesis of FCL has received support from both theoretical and empirical research [@vanderzandenPatternsFoodChain1999; @postEcosystemSizeDetermines2000; @takimotoEffectsProductivityDisturbance2012], although empirical work so far has rarely elaborated on the underlying drivers [but see @saboRoleDischargeVariation2010; @wardMechanisticTheoryAquatic2017].
 In contrast, ecosystem complexity has not been incorporated in theoretical explorations of FCL to date.
@@
 In addition, maintaining habitat connectivity is particularly important for large, linearly-arranged networks.
-Although dispersal has received less attention in food chain research [but see @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012a; @wangMetapopulationCapacityDetermines2021], our simulation highlighted food webs in these networks are more vulnerable to habitat fragmentation because habitat patches are inherently isolated from one another.
+Although dispersal has received less attention in food chain research [but see @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012; @wangMetapopulationCapacityDetermines2021], our simulation highlighted food webs in these networks are more vulnerable to habitat fragmentation because habitat patches are inherently isolated from one another.
 Dams and road crossings have altered the connectivity between suitable habitat patches or have completely blocked dispersal [@ishiyamaPredictingEcologicalImpacts2018; @moritaEffectsHabitatFragmentation2002; @perkinFragmentationAltersStream2012].
@@
 It is not surprising that FCL declined with increasing omnivory because it effectively compresses a food web (**Figure 1C**) and/or intensifies competition between consumer and predator species.
-However, how omnivory influences the associations with ecosystem structure has rarely been explored [@takimotoEffectsProductivityDisturbance2012a; @wardMechanisticTheoryAquatic2017].
+However, how omnivory influences the associations with ecosystem structure has rarely been explored [@takimotoEffectsProductivityDisturbance2012; @wardMechanisticTheoryAquatic2017].
 An important result in our simulation is that omnivory does not change the qualitative relationships between FCL and ecosystem properties, as opposed to productivity effects.

### 2022-10-31 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 A'_{i,x} &= s_{0,i}~\beta_{i,x}(t)~A_{i,x}(t)\\
-\bar{A}_{i,x}(t) &= g_i(A'_{i,x})~A'_{i,x}~\gamma_{x}(t)
+\bar{A}_{i,x}(t) &= g_i(A)~A'_{i,x}~\gamma_{x}(t)
 \end{aligned}
@@
 where $z_{BC,x}(t)$ and $z_{BP,x}(t)$ are the proportional mortality of basal species caused by consumer and predator species, respectively.
-The total mortality is calculated as the proportion of total individuals consumed ($W_{B,x}(t)=W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)$) relative to the current abundance of the basal species ($A_{B,x}(t)$):
+The total mortality is calculated as the proportion of total individuals consumed ($W_{B,x}(t)$) relative to the current abundance of the basal species ($A_{B,x}(t)$):
 
 $$
-z_{BC,x}(t) + z_{BP,x}(t) =\left\{\begin{aligned}&\frac{W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)}{A_{B,x}(t)} &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\&1 &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
+z_{BC,x}(t) + z_{BP,x}(t) =
+\left\{
+\begin{aligned}
+&\frac{W_{B,x}(t)}{A_{B,x}(t)} &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
+&1 &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\
+\end{aligned}
+\right.
 $$
 
-where $W_{ij,x}(t)$ is the population-wide consumption of consumer $j$ on prey $i$ (number of individuals), and the parameter $\delta_{x}(t)$ is the relative preference of predator to basal species over consumer.
+where
+
+$$
+W_{B,x}(t)=W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)
+$$\
+$W_{ij,x}(t)$ is the population-wide consumption of consumer $j$ on prey $i$ (number of individuals), and the parameter $\delta_{x}(t)$ is the relative preference of predator to basal species over consumer.
 We assume the Holling Type-II functional response for predator-prey interactions:
@@
 $$
-g_B(A'_{B,x}) = \frac{r_{b}} {1 + \frac{r_b - 1}{K_{x}}A_{B,x}'}
+g_B(A) = \frac{r_{b}} {1 + \frac{r_b - 1}{K_{x}}A_{B,x}'}
 $$
@@
 $$
-g_C(A'_{C,x}) =
+g_C(A) =
 \left\{
@@
 &e_{BC}W_{BC,x}(t) &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
-&e_{BC} \left( \frac{W_{BC,x}(t)}{W_{B,x}(t)} A_{B,x} \right) &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
+&e_{BC} \left( \frac{W_{BC,x}(t)}{W_{B,x}(t)} A_{B,x}(t) \right) &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
 $$
@@
 $$
-g_P(A'_{P,x}) =
+g_P(A) =
 \left\{

### 2022-11-09 rmd/manuscript.Rmd
edit

@@
 ---
-title: "Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks"
+title: "Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems"
 output:
@@
 bibliography: references.bib
-csl: ecosystems.csl
+csl: https://www.zotero.org/styles/ecosphere
 mainfont: Times New Roman
@@
 
-While recent research efforts of food web ecologists have substantially advanced our understanding of food chain drivers, most of the previous work has been conducted in simple, two-dimensional systems, such as lakes [@postEcosystemSizeDetermines2000; @doiResourceAvailabilityEcosystem2009] and oceanic islands [@takimotoEcosystemSizeNot2008; @pringlePredatorinducedCollapseNiche2019].
+While recent research efforts of food web ecologists have substantially advanced our understanding of food chain drivers, most of the previous work has been conducted in simple, two-dimensional systems, such as lakes [@postEcosystemSizeDetermines2000; @doiResourceAvailabilityEcosystem2009] and oceanic islands [@takimotoEcosystemSizeNot2008].
 However, many natural habitats have a high degree of spatial complexity that cannot be represented simply by ecosystem size.
@@
 
-We generated branching networks as connected habitat patches (**Figure 1A**).
+We generated branching networks as connected habitat patches (**Figure 1A;** see **Table 1** for parameter notations).
 The geometric arrangement of networks was controlled by two parameters: ecosystem size, the number of patches ($N_{p}$); and ecosystem complexity, branching probability ($P_{b}$).
@@
 
-In our simulation, disturbance occurs simultaneously with probability $p_m$ but habitat patches vary in their vulnerability.
+In our simulation, disturbance occurs simultaneously across habitat patches with probability $p_m$ but the patches vary in their vulnerability.
 We simulated the spatial distribution of disturbance intensity (i.e., proportional mortality in the event of stochastic regional disturbance) as follows.
@@
 $$
-\text{logit}~m_{down} = \text{logit}~m_{up}+ \varepsilon
+\begin{aligned}
+\text{logit}~m_{down} &= \text{logit}~m_{up}+ \varepsilon &&\text{(1)}
+\end{aligned}
 $$
@@
 $$
-\text{logit}~m_{down} = \omega~(\text{logit}~m_{1,up} +\varepsilon_1)+(1-\omega)~(\text{logit}~m_{2,up} +\varepsilon_2)
+\begin{aligned}
+\text{logit}~m_{down} &= \omega~(\text{logit}~m_{1,up} +\varepsilon_1)+(1-\omega)~(\text{logit}~m_{2,up} +\varepsilon_2) &&\text{(2)}
+\end{aligned}
 $$
@@
 $$
-K_{x} = K_{base}Q_{x}^z
+\begin{aligned}
+K_{x} &= K_{base}Q_{x}^z &&\text{(3)}
+\end{aligned}
 $$
@@
 
-We employed a discrete-time model to describe tri-trophic dynamics in each habitat patch.
+We employed a discrete-time model to describe tri-trophic dynamics in each habitat patch (see **Table 1** for parameter notations).
 A discrete model is suitable for organisms with predictable reproduction cycles, a common life history for many lotic species, including macroinvertebrates and fish.
@@
 $$
-A_{i,x}(t+1) \sim \text{Poisson}(\bar{A}_{i,x}(t) + I_{i,x}(t) - E_{i,x}(t))
+\begin{aligned}
+A_{i,x}(t+1) &\sim \text{Poisson}(\bar{A}_{i,x}(t) + I_{i,x}(t) - E_{i,x}(t)) &&\text{(4)}
+\end{aligned}
 $$
@@
 
-***Internal dynamics.*** Each time step starts with internal dynamics.
-We accounted for background mortality, predator-induced mortality, and disturbance-induced mortality before the density-dependent growth occurs:
+***Internal dynamics.*** Each time step starts with internal dynamics. We modeled density-dependent growth after accounting for background and predator-induced mortality:
 
@@
 \begin{aligned}
-A'_{i,x} &= s_{0,i}~\beta_{i,x}(t)~A_{i,x}(t)\\
-\bar{A}_{i,x}(t) &= g_i(A)~A'_{i,x}~\gamma_{x}(t)
+\begin{aligned}
+A'_{i,x} &= s_{0,i} \beta_{i,x}(t) A_{i,x}(t)\\
+\bar{A}_{i,x}(t) &= g_i(A) A'_{i,x}
+\end{aligned}
+&&\text{(5)}
 \end{aligned}
@@
 
-where $s_{0,i}$ is the background survival probability (density independent), $\beta_{i,x}(t)$ the survival probability after predation, and $\gamma_{x}(t)$ the survival probability after stochastic disturbance.
-The function $g_i(\centerdot)$ defines the form of the density-dependent recruitment for species $i$.
+where $s_{0,i}$ is the background survival probability (density independent) and $\beta_{i,x}(t)$ the survival probability after predation.
+The function $g_i(\cdot)$ defines the form of the density-dependent recruitment for species $i$.
 Below, we explain how we described predation and recruitment processes for each species.
 
-For the basal species, we considered the predation impacts of consumer and predator species:
+*Basal species* -- We considered the predation impacts of consumer and predator species:
 
 $$
-\beta_{B,x}(t) = 1 - [z_{BC,x}(t) + z_{BP,x}(t)]
-$$
-where $z_{BC,x}(t)$ and $z_{BP,x}(t)$ are the proportional mortality of basal species caused by consumer and predator species, respectively.
+\begin{aligned}
+\beta_{B,x}(t) &= 1 - [z_{BC,x}(t) + z_{BP,x}(t)] &&\text{(6)}
+\end{aligned}
+$$ where $z_{BC,x}(t)$ and $z_{BP,x}(t)$ are the proportional mortality of basal species caused by consumer and predator species, respectively.
 The total mortality is calculated as the proportion of total individuals consumed ($W_{B,x}(t)$) relative to the current abundance of the basal species ($A_{B,x}(t)$):
@@
 $$
+\begin{aligned}
 z_{BC,x}(t) + z_{BP,x}(t) =
@@
 \right.
+&&\text{(7)}
+\end{aligned}
 $$
@@
 $$
-W_{B,x}(t)=W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)
+\begin{aligned}
+W_{B,x}(t) &= W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t) &&\text{(8)}
+\end{aligned}
 $$\
@@
 $$
-W_{ij,x}(t) = \frac{a_{ij}A_{i,x}(t)A_{j,x}(t)}{A_{j,x}(t) + a_{ij}h_{ij}A_{i,x}(t)}
+\begin{aligned}
+W_{ij,x}(t) &= \frac{a_{ij}A_{i,x}(t)A_{j,x}(t)}{A_{j,x}(t) + a_{ij}h_{ij}A_{i,x}(t)} &&\text{(9)}
+\end{aligned}
 $$
@@
 $$
+\begin{aligned}
 \delta_{x}(t) = \frac{e_{BP} f_{BP}(A_{B,x}(t))}{e_{BP} f_{BP}(A_{B,x}(t)) + e_{CP} f_{CP}(A_{C,x}(t))}
+&&\text{(10)}
+\end{aligned}
 $$
@@
 $$
-g_B(A) = \frac{r_{b}} {1 + \frac{r_b - 1}{K_{x}}A_{B,x}'}
+\begin{aligned}
+g_B(A) = \frac{r_{b}} {1 + \frac{r_b - 1}{K_{x}}A_{B,x}'} &&\text{(11)}
+\end{aligned}
 $$
@@
 
-For the consumer species, the predation-induced mortality is the proportion of individuals consumed by the predator species:
+*Consumer species* -- We modeled the predation-induced mortality as the proportion of individuals consumed by the predator species:
 
 $$
-\beta_{CP,x}(t) = 1 - z_{CP,x} = 1 - \frac{W_{CP,x}}{A_{C,x}(t)}
+\begin{aligned}
+\beta_{CP,x}(t) = 1 - z_{CP,x} = 1 - \frac{W_{CP,x}}{A_{C,x}(t)} &&\text{(12)}
+\end{aligned}
 $$
@@
 $$
+\begin{aligned}
 g_C(A) =
@@
 &e_{BC}W_{BC,x}(t) &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
-&e_{BC} \left( \frac{W_{BC,x}(t)}{W_{B,x}(t)} A_{B,x}(t) \right) &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
+&e_{BC} \left( \frac{W_{BC,x}(t)}{W_{B,x}(t)} A_{B,x}(t) \right) &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}
+\end{aligned}
+\right.
+&&\text{(13)}
+\end{aligned}
 $$
@@
 
-The predator species is not exposed to predation, thus $\beta_{P,x}(t) = 1$. The population growth of the predator considers two prey species:
+*Predator species* -- The predator species is not exposed to predation, thus $\beta_{P,x}(t) = 1$. The population growth of the predator considers two prey species:
 
 $$
+\begin{aligned}
 g_P(A) =
@@
 \right.
+&&\text{(14)}
+\end{aligned}
 $$
 
-After recruitment, the constituent species experience stochastic regional disturbance, which causes proportional loss of $m$ to populations:
+***Disturbance.*** After recruitment, all species experience proportional loss of $m$ if stochastic regional disturbance occurs, modifying equation (5) as:
 
 $$
-\gamma_{i,x}(t) = [1 - \psi(t) m_{x}]
+\begin{aligned}
+\bar{A}_{i,x}(t) &= g_i(A) A'_{i,x} (1 - m_{x}) &&\text{(15)}
+\end{aligned}
 $$
 
-The binary random variable $\psi(t)$ expresses the occurrence of regional disturbance (disturbance occurs if $\psi(t)=1$ otherwise $0$), and $m_x$ is the disturbance intensity at patch $x$ as determined by the network generation process.
-$\psi(t)$ is drawn from a Bernoulli distribution with disturbance probability $p_m$.
+where $m_x$ is the disturbance intensity at patch $x$ as determined by the network generation process. Regional disturbance occurs with probability $p_m$.
 
@@
 $$
-\xi_{i,x}(t)=\frac{\sum_{y, y \ne x}E_{i,y}(t)\exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}\text{E}_{i,y}(t)\exp(-\theta d_{xy})}
+\begin{aligned}
+\xi_{i,x}(t) &= \frac{\sum_{y, y \ne x}E_{i,y}(t)\exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}\text{E}_{i,y}(t)\exp(-\theta d_{xy})}
+&&\text{(16)}
+\end{aligned}
 $$
@@
 $$
-TP_{P,x}(t) = \delta_{x}(t) + 2(1 - \delta_{x}(t)) + 1
+\begin{aligned}
+TP_{P,x}(t) &= \delta_{x}(t) + 2(1 - \delta_{x}(t)) + 1
+&&\text{(17)}
+\end{aligned}
 $$
@@
 
-Before the main simulation, we performed an extensive sensitivity analysis to identify influential parameters (see **Supplementary Text** and **Table S1-S3**).
+Before the main simulation, we performed an extensive sensitivity analysis to identify influential parameters (see **Supplementary Text** and **Table S1 -- S3**).
 Our sensitivity analysis revealed that ecosystem size and/or complexity effects were sensitive to disturbance, productivity, dispersal, and omnivory.
@@
 During the warmup period (`r unique(sim_main_result$n_warmup)` time steps), we introduced propagules into habitat patches every 10 time steps to allow species to establish populations.
-Propagules are random draws from $Poisson(0.5)$.
+Propagules are random draws from $\text{Poisson}(0.5)$.
 Subsequently, we ran a burn-in period with no propagules for `r unique(sim_main_result$n_burnin)` time steps to remove the effects of initial conditions.
 We obtained the mean FCL across all habitat patches for the last `r unique(sim_main_result$n_timestep)` time steps.
-Note that simulation replicates are differ in ecosystem size ($N_{p}\sim Unif(10,150)$) and ecosystem complexity ($P_{b} \sim Unif(0.01, 0.99)$) so we can estimate their effects on FCL.
+Note that simulation replicates are differ in ecosystem size ($N_{p}\sim \text{Unif}(10,150)$) and ecosystem complexity ($P_{b} \sim \text{Unif}(0.01, 0.99)$) so we can estimate their effects on FCL.
 All analysis was performed in R 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2021]
@@
 
-Our simulation highlighted the interactive nature of putative controls of FCL.
-In particular, disturbance regimes strongly influenced the qualitative relationships between FCL and ecosystem properties.
+## Overall pattern
+
+There were three consistent patterns in the relationship between FCL and ecosystem properties.
+First, both ecosystem size and complexity had a stronger positive effect in high disturbance regime.
+Second, the effects of ecosystem complexity was stronger than ecosystem size in general.
+Third, omnivory had a negative influence on the FCL regardless of simulation scenarios.
 Below, we report our results for low and high disturbance regimes separately.
@@
 With low productivity, strong omnivory caused shorter FCL because it intensified competition between consumer and predator species.
-We can see this effect in the decreased occupancies of consumer and predator species in a low productivity scenario (most left column in **Figure S1-S4**).
-However, when productivity was high, the occupancies of both species were \~1.0, indicating that the structural constraint of strong omnivory made FCL shorter (second column from left in **Figure S1-S4**).
+We can see this effect in the decreased occupancies of consumer and predator species in a low productivity scenario (most left column in **Figure S1 -- S4**).
+However, when productivity was high, the occupancies of both species were \~1.0, indicating that the structural constraint of strong omnivory made FCL shorter (second column from left in **Figure S1 -- S4**).
 Productivity played a secondary role in increasing FCL, particularly when omnivory was strong (compare left and right panels in **Figure 2A and 2B**).
@@
 Dispersal buffered the disturbance effects on FCL and the magnitude of buffering differed by ecosystem structure.
-The greater buffering effect of dispersal emerged in larger ecosystems (compare top and bottom panels in **Figure 3**) because greater dispersal capability is required to reach distant patches.
-In the meantime, simpler, less branched ecosystems benefited more from increased dispersal because habitat patches are more isolated in linearly arranged networks (**Figure 3**).
+Larger ecosystems benefited more from long dispersal (compare top and bottom panels in **Figure 3**) because greater dispersal capability is required to reach distant patches.
+In particular, this pattern is clearer in simple, less branched ecosystems because habitat patches are more isolated in linearly arranged networks (**Figure 3**).
 Productivity increased FCL as observed in a low disturbance regime, but it did not modulate the qualitative relationships between FCL and ecosystem properties (compare left and right panels in **Figure 3**).
 
-Different processes drove the positive effects of ecosystem size and complexity on FCL when strong omnivory exists in low-productivity systems.
-In this scenario, the occupancies of $B$ and $C$ decreased at larger ecosystem sizes, while $P$ increased its occupancy (third column from left in **Figure S1 and S2**).
-Therefore, $P$ compensated for the loss of $B$ and $C$ to maintain the positive relationship between FCL and ecosystem size.
-In contrast, the increased occupancy of $C$ maintained the positive relationship between FCL and ecosystem complexity by offsetting the loss of $P$.
+When strong omnivory exists in low-productivity systems, different processes drove the positive effects of ecosystem size and complexity on FCL.
+The positive effects of ecosystem size are mainly explained by the increased occupancy of $P$ (third column from left in **Figure S1 and S2**).
+In contrast, the increased occupancy of $C$ maintained the positive relationship between FCL and ecosystem complexity (third column from left in **Figure S3 and S4**).
 

### 2022-11-09 rmd/suppl_info.Rmd
edit

@@
       before_body: title.tex
+mainfont: Times New Roman
 knit: (function(inputFile, encoding) {
@@
 We generated 500 sets of parameter combinations by randomly drawing values of `r length(attributes(fit_sense$fit[[1]]$terms)$term.label)` simulation parameters from uniform distributions (**Table S1**).
-For each parameter combination, we generated `r max(sim_sv$n_rep)` branching networks with the gradients of ecosystem size ($N_p \sim Unif(10, 150)$) and complexity ($P_b \sim Unif(0.01, 0.99)$).
+For each parameter combination, we generated `r max(sim_sv$n_rep)` branching networks with the gradients of ecosystem size ($N_p \sim \text{Unif}(10, 150)$) and complexity ($P_b \sim \text{Unif}(0.01, 0.99)$).
 This results in a total of `r nrow(sim_sv)` simulation replicates.
@@
 \begin{align}
-  u_{n} &\sim Normal(\mu_n, \sigma^2)\\
+  u_{n} &\sim \text{Normal}(\mu_n, \sigma^2)\\
   \mu_n &= \zeta_0 + \sum_{k}{\zeta_k x_{k, n}}

### 2022-11-09 code/table_param.R

@@ -30,15 +30,15 @@ df_value <- sim_main_result %>%
                names_to = "param",
                values_to = "Main") %>%
   mutate(Parameter = case_when(param == "r_b" ~ "$r_b$",
-                               param == "e_bc" ~"$e_{bc}$",
-                               param == "e_bp" ~"$e_{bp}$",
-                               param == "e_cp" ~"$e_{cp}$",
-                               param == "a_bc" ~"$a_{bc}$",
-                               param == "a_bp" ~"$a_{bp}$",
-                               param == "a_cp" ~"$a_{cp}$",
-                               param == "h_bc" ~"$h_{bc}$",
-                               param == "h_bp" ~"$h_{bp}$",
-                               param == "h_cp" ~"$h_{cp}$",
+                               param == "e_bc" ~"$e_{BC}$",
+                               param == "e_bp" ~"$e_{BP}$",
+                               param == "e_cp" ~"$e_{CP}$",
+                               param == "a_bc" ~"$a_{BC}$",
+                               param == "a_bp" ~"$a_{BP}$",
+                               param == "a_cp" ~"$a_{CP}$",
+                               param == "h_bc" ~"$h_{BC}$",
+                               param == "h_bp" ~"$h_{BP}$",
+                               param == "h_cp" ~"$h_{CP}$",
                                param == "theta" ~"$\\theta$",
                                param == "p_disturb" ~"$p_m$",
                                param == "mean_disturb_source" ~ "$\\mu_m$",
@@ -46,7 +46,7 @@ df_value <- sim_main_result %>%
                                param == "sd_disturb_lon" ~ "$\\sigma_l$",
                                param == "s0" ~ "$s_0$",
                                param == "p_dispersal" ~ "$p_d$"),
-         Group = case_when(param %in% c("r_b",
+         Model = case_when(param %in% c("r_b",
                                         "s0",
                                         "e_bc",
                                         "e_bp",
@@ -56,13 +56,13 @@ df_value <- sim_main_result %>%
                                         "a_cp",
                                         "h_bc",
                                         "h_bp",
-                                        "h_cp") ~ "Food web",
-                           param %in% c("theta",
-                                        "p_dispersal") ~ "Dispersal",
-                           param %in% c("p_disturb",
-                                        "mean_disturb_source",
-                                        "sd_disturb_source",
-                                        "sd_disturb_lon") ~ "Disturbance"),
[…]

### 2022-11-09 code/table_stvy.R

@@ -11,7 +11,7 @@ source(here::here("code/analysis_stvy.R"))
 table_stvy <- foreach(i = 1:2) %do% {
   fit_sense$fit[[i]] %>% 
     tidy() %>%
-    mutate(Group = case_when(term %in% c("r_b",
+    mutate(Model = case_when(term %in% c("r_b",
                                          "s0",
                                          "e_bc",
                                          "e_bp",
@@ -21,23 +21,23 @@ table_stvy <- foreach(i = 1:2) %do% {
                                          "a_cp",
                                          "h_bc",
                                          "h_bp",
-                                         "h_cp") ~ "Food web",
-                             term %in% c("theta",
-                                         "p_dispersal") ~ "Dispersal",
-                             term %in% c("p_disturb",
-                                         "mean_disturb_source",
+                                         "h_cp",
+                                         "theta",
+                                         "p_dispersal",
+                                         "p_disturb") ~ "Food web",
+                             term %in% c("mean_disturb_source",
                                          "sd_disturb_source",
-                                         "sd_disturb_lon") ~ "Disturbance"),
+                                         "sd_disturb_lon") ~ "Network"),
            Term = case_when(term == "r_b" ~ "$r_b$",
-                            term == "e_bc" ~"$e_{bc}$",
-                            term == "e_bp" ~"$e_{bp}$",
-                            term == "e_cp" ~"$e_{cp}$",
-                            term == "a_bc" ~"$a_{bc}$",
-                            term == "a_bp" ~"$a_{bp}$",
-                            term == "a_cp" ~"$a_{cp}$",
-                            term == "h_bc" ~"$h_{bc}$",
-                            term == "h_bp" ~"$h_{bp}$",
-                            term == "h_cp" ~"$h_{cp}$",
+                            term == "e_bc" ~"$e_{BC}$",
+                            term == "e_bp" ~"$e_{BP}$",
+                            term == "e_cp" ~"$e_{CP}$",
+                            term == "a_bc" ~"$a_{BC}$",
+                            term == "a_bp" ~"$a_{BP}$",
+                            term == "a_cp" ~"$a_{CP}$",
+                            term == "h_bc" ~"$h_{BC}$",
+                            term == "h_bp" ~"$h_{BP}$",
+                            term == "h_cp" ~"$h_{CP}$",
                             term == "theta" ~"$\\theta$",
[…]

### 2022-11-09 rmd/cover_letter.Rmd
edit

@@
 
-We are excited to submit our manuscript entitled "Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks" for consideration for publication in The American Naturalist.
+We are excited to submit our manuscript entitled "*Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems*" for consideration.
 

### 2022-11-09 rmd/manuscript.Rmd
edit

@@
     latex_engine: xelatex
-  #word_document:
-  #  reference_docx: word_format.docx
 knit: (function(inputFile, encoding) {
@@
 header-includes:
-  \usepackage{textgreek}
+  \usepackage{lineno, textgreek}
+  \linenumbers
 ---
@@
 
-Before the main simulation, we performed an extensive sensitivity analysis to identify influential parameters (see **Supplementary Text** and **Table S1 -- S3**).
+Before the main simulation, we performed an extensive sensitivity analysis to identify influential parameters (see **Table1**, **Supplementary Text,** and **Table S1 -- S2**).
 Our sensitivity analysis revealed that ecosystem size and/or complexity effects were sensitive to disturbance, productivity, dispersal, and omnivory.
-Therefore, we ran the main simulation with `r n_scenario` parameter sets differing in these ecological contexts (two levels of disturbance, two levels of productivity, two levels of dispersal distance, and three levels of omnivory) and parameter values were summarized in **Table S1**.
+Therefore, we ran the main simulation with `r n_scenario` parameter sets differing in these ecological contexts (two levels of disturbance, two levels of productivity, two levels of dispersal distance, and three levels of omnivory) and parameter values were summarized in **Table 1**.
 

### 2022-11-14 rmd/manuscript.Rmd
edit

@@
 **Keywords:** Food chain length, omnivory, branching networks, fractals, intraguild predation, complex network architecture
+
+## Data availability
+
+No data were used or produced in this research.
+R functions for simulations are provided as the R package mcbrnet version 1.2.3 (available at <https://github.com/aterui/mcbrnet>).
+Codes for simulations, statistical analysis, figures, and tables are available at <https://github.com/aterui/public-proj_fcl-river-theory>.
+
+\clearpage
 
@@
 
-## Data availability
-
-No data were used or produced in this research.
-R functions for simulations are provided as the R package mcbrnet version 1.2.3 (available at [\<https://github.com/aterui/mcbrnet\>](https://github.com/aterui/mcbrnet){.uri}).
-Codes for simulations, statistical analysis, figures, and tables are available at [\<https://github.com/aterui/public-proj_fcl-river-theory\>](https://github.com/aterui/public-proj_fcl-river-theory){.uri}.
-
 # References

### 2023-02-18 code/figure_network.R
edit: sim within net

@@ -64,34 +64,22 @@ g_net <- (ng[[1]] + labs(title = "A")) + ng[[2]] + ng[[3]] + plot_layout(guide =
 
 theme_set(plt_theme)
 
-df_assemble <- foreach(i = seq_len(nrow(para))) %do% {
-  
-  foreach(j = 1:10,
-          .combine = bind_rows) %do% {
-            set.seed(j)
-            x <- brnet(n_patch = para$n_patch[i],
-                       p_branch = para$p_branch[i],
-                       mean_disturb_source = 0.8,
-                       sd_disturb_source = para$sd_source[i],
-                       sd_disturb_lon = para$sd_lon[i],
-                       plot = FALSE)
-            
-            return(x$df_patch)
-          }  
-  
-}
+df_wn <- readRDS("output/sim_within_net.rds")
 
-g_disturb <- bind_rows(mutate(df_assemble[[1]], p_branch = 0.2),
-                       mutate(df_assemble[[2]], p_branch = 0.5),
-                       mutate(df_assemble[[3]], p_branch = 0.8)) %>% 
-  ggplot(aes(x = disturbance,
+labs <- c(`0.2` = "p[b]==0.2",
+          `0.5` = "p[b]==0.5",
+          `0.8` = "p[b]==0.8")
+
+df_wn %>% 
+  ggplot(aes(y = disturbance,
+             x = factor(n_patch),
              color = factor(p_branch),
              fill = factor(p_branch))) +
-  geom_density(alpha = 0.2) +
-  geom_vline(xintercept = 0.8,
-             color = grey(0.5),
-             linetype = "dashed") +
-  labs(x = "Disturbance intensity",
-       y = "Density",
-       fill = "Branching prob.",
-       color = "Branching prob.")
+  geom_violin(alpha = 0.2, draw_quantiles = 0.5) +
+  facet_wrap(facets = ~p_branch,
+             labeller = labeller(p_branch = as_labeller(labs,
+                                                        label_parsed))) +
+  labs(x = "Ecosystem size (number of patches)",
+       y = "Disturbance intensity (m)",
+       color = "Branching prob.",
+       fill = "Branching prob.")

### 2023-02-18 code/figure_si_within_net.R

@@ -1,0 +1,24 @@
+
+# setup -------------------------------------------------------------------
+
+rm(list = ls())
+
+lapply(list("code/library.R",
+            "code/set_figure_theme.R"),
+       source)
+
+
+# figure ------------------------------------------------------------------
+
+df_wn <- readRDS("output/sim_within_net.rds")
+
+df_wn %>% 
+  filter(n_patch == 150,
+         a_bp == 0.04) %>% 
+  ggplot(aes(x = carrying_capacity,
+             y = fcl,
+             color = factor(p_branch),
+             fill = factor(p_branch))) +
+  geom_point(alpha = 0.2) +
+  geom_smooth()# +
+  #facet_wrap(facets = ~factor(n_patch))

### 2023-02-18 code/run_sim_within_net.R

@@ -26,7 +26,7 @@ df_param <- expand.grid(mean_disturb_source = 0.8,
                         n_burnin = 400,
                         
                         # food web parameter
-                        r_b = 20,
+                        r_b = 8,
                         e = 1, # to conv_eff
                         a_bc = 0.5, # to attack_rate[1]
                         a_bp = c(0, 0.02, 0.04), # to attack_rate[2]

### 2023-02-18 rmd/manuscript.Rmd
edit: manuscript

@@
 In this formulation, the most upstream patch, at which $Q_x = 1$ (the number of upstream patches + 1), has the carrying capacity of $K_{base}$.
-We used $K_{base} = 100$ and $z = 0.54$ based on an empirical relationship of gross primary production and watershed area [@finlayStreamSizeHuman2011].
+We used $K_{base} = 500$ and $z = 0.54$ based on an empirical relationship of gross primary production and watershed area [@finlayStreamSizeHuman2011].
 
@@
 Our simulation model consists of three species; a basal species $B$ with positive logistic growth up to a carrying capacity, $K$; a primary consumer species or intraguild prey, $C$, which grows based on the consumption of $B$; and an intraguild predator $P$, which can consume both $B$ and $C$ (**Figure 1C**).
-In our model, the realized species abundance $A_{i,x}(t+1)$ for species $i$ (either $B,C,~or~P$) at patch $x$ and time $t+1$ is expressed as a composite of within-patch trophic interactions, immigration, and emigration:
+The realized species abundance $A_{ix,t+1}$ for species $i$ (either $B,C,~or~P$) at patch $x$ and time $t+1$ is expressed as a composite of within-patch trophic interactions, immigration, and emigration:
 
@@
 \begin{aligned}
-A_{i,x}(t+1) &\sim \text{Poisson}(\bar{A}_{i,x}(t) + I_{i,x}(t) - E_{i,x}(t)) &&\text{(4)}
+A_{ix,t+1} &\sim \text{Poisson}(\bar{A}_{ix,t} + I_{ix,t} - E_{ix,t}) &&\text{(4)}
 \end{aligned}
@@
 
-where $\bar{A}_{i,x}(t)$ is the expected species abundance after within-patch trophic interactions, $I_{i,x}(t)$ the expected number of immigrants into patch $x$, and $E_{i,x}(t)$ the expected number of emigrants from patch $x$.
+where $\bar{A}_{ix,t}$ is the expected species abundance after within-patch trophic interactions, $I_{ix,t}$ the expected number of immigrants into patch $x$, and $E_{ix,t}$ the expected number of emigrants from patch $x$.
 The realized discrete number of individuals is drawn from a Poisson distribution to account for stochasticity in demographic and dispersal processes.
 
-***Internal dynamics.*** Each time step starts with internal dynamics. We modeled density-dependent growth after accounting for background and predator-induced mortality:
+***Internal dynamics.*** Each time step starts with internal dynamics. The general formula is as follows:
+
+$$
+\begin{aligned}
+B_{x,t+1} &= B_{x,t}g(B_{x,t})f_1(B_{x,t}, C_{x,t})f_2(B_{x,t}, C_{x,t})\\
+C_{x,t+1} &= e_{BC}B_{x,t}[1 - f_1(B_{x,t}, C_{x,t})]f_3(C_{x,t}, P_{x,t})\\
+P_{x,t+1} &= e_{BP}B_{x,t}f_1(B_{x,t}, C_{x,t})[1 - f_2(B_{x,t}, P_{x,t})] + e_{CP}C_{x,t}[1 - f_3(C_{x,t}, P_{x,t})]
+\end{aligned}
+$$
+
+We modeled density-dependent growth after accounting for background and predator-induced mortality:
 

### 2023-02-18 code/figure_si_one_patch.R
edit: figure update

@@ -53,8 +53,8 @@ foreach(i = 1:length(a1_set)) %do% {
             axis.text = element_text(size = 10)) +
       ggtitle(paste0("K = ", k_set[j])) +
       geom_point(data = expand.grid(a2 = c(0, 0.02, 0.04),
-                                    a3 = 0.02,
-                                    h = 0.5,
+                                    a3 = 0.005,
+                                    h = 0.75,
                                     state = 0))
     
     return(g_one)

### 2023-02-19 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 
-***Internal dynamics.*** Each time step starts with internal dynamics. The general formula is as follows:
+***Internal dynamics.*** Each time step starts with internal dynamics. The general formula is given as follows:
 
@@
 \begin{aligned}
-B_{x,t+1} &= B_{x,t}g(B_{x,t})f_1(B_{x,t}, C_{x,t})f_2(B_{x,t}, C_{x,t})\\
-C_{x,t+1} &= e_{BC}B_{x,t}[1 - f_1(B_{x,t}, C_{x,t})]f_3(C_{x,t}, P_{x,t})\\
-P_{x,t+1} &= e_{BP}B_{x,t}f_1(B_{x,t}, C_{x,t})[1 - f_2(B_{x,t}, P_{x,t})] + e_{CP}C_{x,t}[1 - f_3(C_{x,t}, P_{x,t})]
-\end{aligned}
-$$
-
-We modeled density-dependent growth after accounting for background and predator-induced mortality:
-
-$$
-\begin{aligned}
-\begin{aligned}
-A'_{i,x} &= s_{0,i} \beta_{i,x}(t) A_{i,x}(t)\\
-\bar{A}_{i,x}(t) &= g_i(A) A'_{i,x}
-\end{aligned}
-&&\text{(5)}
-\end{aligned}
-$$
-
-where $s_{0,i}$ is the background survival probability (density independent) and $\beta_{i,x}(t)$ the survival probability after predation.
-The function $g_i(\cdot)$ defines the form of the density-dependent recruitment for species $i$.
-Below, we explain how we described predation and recruitment processes for each species.
-
-*Basal species* -- We considered the predation impacts of consumer and predator species:
-
-$$
-\begin{aligned}
-\beta_{B,x}(t) &= 1 - [z_{BC,x}(t) + z_{BP,x}(t)] &&\text{(6)}
-\end{aligned}
-$$ where $z_{BC,x}(t)$ and $z_{BP,x}(t)$ are the proportional mortality of basal species caused by consumer and predator species, respectively.
-The total mortality is calculated as the proportion of total individuals consumed ($W_{B,x}(t)$) relative to the current abundance of the basal species ($A_{B,x}(t)$):
-
-$$
-\begin{aligned}
-z_{BC,x}(t) + z_{BP,x}(t) =
-\left\{
-\begin{aligned}
-&\frac{W_{B,x}(t)}{A_{B,x}(t)} &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
-&1 &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\
-\end{aligned}
-\right.
-&&\text{(7)}
+B^+_{x,t} &= B_{x,t}g(B_{x,t})f_1(B'_{x,t}, C_{x,t})f_2(B'_{x,t}, P_{x,t})\\
+C^+_{x,t} &= e_{BC}B'_{x,t}[1 - f_1(B'_{x,t}, C_{x,t})]f_3(C'_{x,t}, P_{x,t})\\
+P^+_{x,t} &= e_{BP}B'_{x,t}f_1(B'_{x,t}, C_{x,t})[1 - f_2(B''_{x,t}, P_{x,t})] + e_{CP}C_{x,t}[1 - f_3(C'_{x,t}, P_{x,t})]
+~~\text{(5)}
 \end{aligned}
@@
 \begin{aligned}
-W_{B,x}(t) &= W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t) &&\text{(8)}
+B'_{x,t} &= B_{x,t}g(B_{x,t}) &&\text{after B's recruitment}\\
+B''_{x,t} &= B'_{x,t}f_1(B'_{x,t}, C_{x,t}) &&\text{after C's predation on B}\\
+C'_{x,t} &= e_{BC}B'_{x,t}[1 - f_1(B'_{x,t}, C_{x,t})] &&\text{after C's recruitment}
 \end{aligned}
-$$\
-$W_{ij,x}(t)$ is the population-wide consumption of consumer $j$ on prey $i$ (number of individuals), and the parameter $\delta_{x}(t)$ is the relative preference of predator to basal species over consumer.
-We assume the Holling Type-II functional response for predator-prey interactions:
+$$
+
+The function $g(\cdot)$ defines the density-dependent recruitment for the basal species, and the function $f_k(\cdot)$ dictates the proportion of survived prey individuals after predation.
+We used the Beverton-Holt function to describe the density-dependent growth of the basal species:
 
@@
 \begin{aligned}
-W_{ij,x}(t) &= \frac{a_{ij}A_{i,x}(t)A_{j,x}(t)}{A_{j,x}(t) + a_{ij}h_{ij}A_{i,x}(t)} &&\text{(9)}
+g(B_{x,t}) = \frac{r_{b}} {1 + \frac{r_b - 1}{K_{x}}B_{x,t}} &&\text{(11)}
 \end{aligned}
@@
 
-$a_{ij}$ is the attack rate and $h_{ij}$ the handling time. To be realistic, the preference parameter is assumed to depend on the relative potential amount of energy obtained from each prey:
+where $r_{B}$ is the intrinsic growth rate and $K_x$ is the carrying capacity at patch $x$. We assumed the Type-II function response in the survival function $f_k(\cdot)$ as follows:
 
@@
 \begin{aligned}
-\delta_{x}(t) = \frac{e_{BP} f_{BP}(A_{B,x}(t))}{e_{BP} f_{BP}(A_{B,x}(t)) + e_{CP} f_{CP}(A_{C,x}(t))}
-&&\text{(10)}
+f_1(B'_{x,t}, C_{x,t}) &= \exp \left(-\frac{a_{BC} C_{x,t}} {1 + a_{BC} h_{BC} B'_{x,t}} \right) &&\text{C on B}\\
+f_2(B''_{x,t}, P_{x,t}) &= \exp \left(-\frac{(1 + \phi) a_{BP} P_{x,t}} {1 + a_{BP} h_{BP} B''_{x,t}} \right) &&\text{P on B}\\
+f_3(C'_{x,t}, P_{x,t}) &= \exp \left(-\frac{(1 - \phi) a_{CP} P_{x,t}} {1 + a_{CP} h_{CP} C'_{x,t}} \right) &&\text{P on C}
 \end{aligned}
@@
 
-where $e_{ij}$ is the conversion efficiency from prey $i$ to consumer $j$, and $f_{ij}(A_{i,x}(t)) = \frac{a_{ij}A_{i,x}(t)}{1 + a_{ij}h_{ij}A_{i,x}(t)}$.
-An analogous approach has been applied to a model with multiple prey choices [@nishijimaRolesAlternativePrey2014].
-Subsequently, a density-dependent recruitment occurs:
+The parameter $a_{ij}$ defines the encounter rate between $j$ and $i$, $h_{ij}$ the handling time, and $\phi$ the switching function for the intraguild predator.
+We assumed that the degree of omnivory is dependent on the relative abundance of $B''_{x,t}$ and $C'_{x,t}$:
+
+$$
+\phi = \frac{s(B''_x(x,t) - C'_{x,t})}{B''_x(x,t) + C'_{x,t}}
+$$
+
+where the parameter $s$ determines the strength of switching ($0 \le s \le 1$).
+In this expression, the intraguild predator increases its attack rate to the basal species if $\phi > 0$, properly representing the switching behavior of the intraguild prey.
+
+***Disturbance.*** After recruitment, all species experience proportional loss if stochastic regional disturbance occurs:
 
@@
 \begin{aligned}
-g_B(A) = \frac{r_{b}} {1 + \frac{r_b - 1}{K_{x}}A_{B,x}'} &&\text{(11)}
+\bar{A}_{ix,t}(t) &= A^+_{ix,t} (1 - \nu_t m'_{x,t}) &&\text{(15)}
 \end{aligned}
@@
 
-where $r_b$ is the reproductive rate for survived individuals.
-
-*Consumer species* -- We modeled the predation-induced mortality as the proportion of individuals consumed by the predator species:
-
-$$
-\begin{aligned}
-\beta_{CP,x}(t) = 1 - z_{CP,x} = 1 - \frac{W_{CP,x}}{A_{C,x}(t)} &&\text{(12)}
-\end{aligned}
-$$
-
-The population growth of the consumer is controlled by conversion efficiency and the number of prey captured:
-
-$$
-\begin{aligned}
-g_C(A) =
-\left\{
-\begin{aligned}
-&e_{BC}W_{BC,x}(t) &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
-&e_{BC} \left( \frac{W_{BC,x}(t)}{W_{B,x}(t)} A_{B,x}(t) \right) &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}
-\end{aligned}
-\right.
-&&\text{(13)}
-\end{aligned}
-$$
-
-The above equation assumes competition between $C$ and $P$ when their total consumption $W_{B,x}(t)$ exceeds the abundance of basal species $A_{B,x}(t)$, and the relative consumption $\frac{W_{BC,x}(t)}{W_{B,x}(t)}$ rests on $a_{ij}$, $h_{ij}$, and $A_{j,x}(t)$.
-
-*Predator species* -- The predator species is not exposed to predation, thus $\beta_{P,x}(t) = 1$. The population growth of the predator considers two prey species:
-
-$$
-\begin{aligned}
-g_P(A) =
-\left\{
-\begin{aligned}
-&e_{BP} \delta_{x}(t) W_{BP,x}(t) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
-&e_{BP} \delta_{x}(t)\left(\frac{W_{BP,x}(t)}{W_{B,x}(t)}A_{B,x}\right) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\
-\end{aligned}
-\right.
-&&\text{(14)}
-\end{aligned}
-$$
-
-***Disturbance.*** After recruitment, all species experience proportional loss of $m$ if stochastic regional disturbance occurs, modifying equation (5) as:
-
-$$
-\begin{aligned}
-\bar{A}_{i,x}(t) &= g_i(A) A'_{i,x} (1 - m_{x}) &&\text{(15)}
-\end{aligned}
-$$
-
-where $m_x$ is the disturbance intensity at patch $x$ as determined by the network generation process. Regional disturbance occurs with probability $p_m$.
+where $m'_{x,t}$ is the disturbance intensity at patch $x$ and time $t$, which was drawn from a Beta distribution with mean $m_x$ and precision $1000$.
+The mean disturbance intensity $m_x$ is determined by the network generation process.
+The random variable $\nu_t$ represents the incidence of regional disturbance ($\nu_t = 1$ if disturbance occurs, otherwise zero), which is drawn from a Bernoulli distribution as $\nu_t \sim \text{Bernouilli}(p_m)$.
 

### 2023-02-19 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 
+The parameter $e_{ij}$ is the energetic conversion efficiency from $i$ to $j$.
 The function $g(\cdot)$ defines the density-dependent recruitment for the basal species, and the function $f_k(\cdot)$ dictates the proportion of survived prey individuals after predation.
@@
 
-where $r_{B}$ is the intrinsic growth rate and $K_x$ is the carrying capacity at patch $x$. We assumed the Type-II function response in the survival function $f_k(\cdot)$ as follows:
+where $r_{B}$ is the intrinsic growth rate and $K_x$ is the carrying capacity at patch $x$. We assumed the Type-II functional response in the survival function $f_k(\cdot)$ as follows:
 
@@
 
-Before the main simulation, we performed an extensive sensitivity analysis to identify influential parameters (see **Table1**, **Supplementary Text,** and **Table S1 -- S2**).
-Our sensitivity analysis revealed that ecosystem size and/or complexity effects were sensitive to disturbance, productivity, dispersal, and omnivory.
-Therefore, we ran the main simulation with `r n_scenario` parameter sets differing in these ecological contexts (two levels of disturbance, two levels of productivity, two levels of dispersal distance, and three levels of omnivory) and parameter values were summarized in **Table 1**.
+Prior to the main simulation, we performed a one-patch simulation to explore the predation parameter space ($a_{ij}$ and $h_{ij}$) that allows stable coexistence of the three species in the absence of disturbance.
+In this simulation, we assumed $h_{ij} = h$ to reduce the dimension of parameter space.
+This analysis revealed that the stable coexistence of the three species is possible with a certain combinations of predation parameters.
+Within this parameter space, we explored the three scenarios of food web structure: "chain" ($(a_{BC}, a_{BP}, a_{CP}, h) = (X, X, X, X)$), "weak omnivory" ($(a_{BC}, a_{BP}, a_{CP}, h) = (X, X, X)$), and "strong omnivory" ($(a_{BC}, a_{BP}, a_{CP}, h) = (X, X, X)$).
+We used $s = 1$ except for the chain scenario with $s = 0$.
+Without the loss of generality, we assumed $e_{ij} = e = 1$.
 
@@
 For dispersal, we examined short- and long-distance dispersal scenarios (short-distance dispersal $\theta=$ `r max(sim_main_result$theta)`; long-distance dispersal $\theta=$ `r min(sim_main_result$theta)`) with constant dispersal probability ($p_{d}=$ `r unique(sim_main_result$p_dispersal)`).
-
-Finally, we investigated three levels of omnivory (**Figure 1D**).
-Trophic interaction strengths can be controlled by varying the attack rate and/or conversion efficiency [@wardMechanisticTheoryAquatic2017].
-We changed attack rate ($a_{BP}$ and $a_{CP}$) and conversion efficiency ($e_{BP}$ and $e_{CP}$) of two prey items for predator $P$.
-Specifically, we used the following combinations to represent three levels of omnivory: "Chain", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[1], max(sim_main_result$e_cp))` $\}$; "Weak omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[2], max(sim_main_result$e_cp))` $\}$; and "Strong omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(max(sim_main_result$a_bp), min(sim_main_result$a_cp), unique(sim_main_result$e_bp)[3], min(sim_main_result$e_cp))` $\}$.
-To simplify simulation setup, we fixed the following parameters: $e_{BC}=$ `r unique(sim_main_result$e_bc)` , $a_{BC}=$ `r unique(sim_main_result$a_bc)`, and $h_{ij}=$ `r unique(sim_main_result$h_bc)`.
 

### 2023-02-19 code/figure_si_one_patch.R
edit: one-patch figure

@@ -53,7 +53,7 @@ foreach(i = 1:length(a1_set)) %do% {
             axis.text = element_text(size = 10)) +
       ggtitle(paste0("K = ", k_set[j])) +
       geom_point(data = expand.grid(a2 = c(0, 0.02, 0.04),
-                                    a3 = 0.005,
+                                    a3 = 0.0025,
                                     h = 0.75,
                                     state = 0))
     

### 2023-02-19 code/run_sim_one_patch.R
edit: one patch sim

@@ -13,7 +13,7 @@ registerDoSNOW(cl)
 df_param <- expand.grid(n_timestep = 1000,
                         n_warmup = 200,
                         n_burnin = 400,
-                        r_b = seq(2, 20, length = 4),
+                        r_b = seq(4, 16, length = 4),
                         k = c(100, 1000),
                         e = 1, # to conv_eff[1]
                         a1 = c(0.25, 0.5),

### 2023-02-19 code/figure_delta.R
edit: figure delta

@@ -15,9 +15,9 @@ m_x <- rbind(seq(0, 100, length = 100),
 
 ## simulation scenarios
 df_param <- expand.grid(e = 1, # to conv_eff
-                        a_bc = 0.5, # to attack_rate
+                        a_bc = 0.25, # to attack_rate
                         a_bp = c(0, 0.02, 0.04), # to attack_rate
-                        a_cp = 0.02, # to attack_rate
+                        a_cp = 0.0025, # to attack_rate
                         h = 0.5, # to handling_time
                         s = 1) %>% 
   as_tibble() %>% 

### 2023-02-19 code/run_sim.R
edit: run_sim parameter

@@ -26,12 +26,12 @@ df_param <- expand.grid(mean_disturb_source = c(0.2, 0.8),
                         n_burnin = 400,
                         
                         # food web parameter
-                        r_b = seq(2, 20, length = 4),
+                        r_b = seq(4, 16, length = 4),
                         e = 1, # to conv_eff
-                        a_bc = 0.5, # to attack_rate[1]
+                        a_bc = 0.25, # to attack_rate[1]
                         a_bp = c(0, 0.02, 0.04), # to attack_rate[2]
-                        a_cp = c(0.02), # to attack_rate[3]
-                        h = 0.5, # to handling_time
+                        a_cp = 0.0025, # to attack_rate[3]
+                        h = 0.75, # to handling_time
                         s = c(0, 1),
                         p_disturb = seq(0, 0.15, length = 4),
                         p_dispersal = 0.01,

### 2023-02-19 code/figure_si_one_patch.R
edit: one-patch sim

@@ -13,10 +13,10 @@ lab <- list(c(`0` = "h==0.0",
               `0.5` = "h==0.50",
               `0.75` = "h==0.75",
               `1` = "h==1.00"),
-            c(`2` = "r[B]==2",
+            c(`4` = "r[B]==4",
               `8` = "r[B]==8",
-              `14` = "r[B]==14",
-              `20` = "r[B]==20"))
+              `12` = "r[B]==12",
+              `16` = "r[B]==16"))
 
 state_lab <- c("No species", "B", "B + C", "B + P", "B + C + P")
 

### 2023-02-24 code/figure_main_sim.R
edit: figure main & si, fcl

@@ -8,10 +8,12 @@ lapply(list("code/library.R",
             "code/format_sim_data.R"),
        source)
 
+df_coef <- readRDS("output/df_coef.rds")
+
 ## filter s & mean_disturb_source for visualization
 s_set <- 1
 mu_disturb <- 0.8
-r_set <- c(4, 16)
+r_set <- c(8, 16)
 
 ## df for heatmap
 df_heat <- df_param %>% 
@@ -20,7 +22,7 @@ df_heat <- df_param %>%
   filter(mean_disturb_source == mu_disturb,
          (s == 0 & a_bp == 0) | s == s_set)
 
-## df for gam plot
+## df for loess plot
 df_plot <- df_sim %>% 
   filter(mean_disturb_source == mu_disturb,
          (s == 0 & a_bp == 0) | s == s_set,
@@ -48,7 +50,8 @@ g_size <- df_heat %>%
        x = expression("Productivity ("*r[b]*")"),
        fill = "Slope") +
   scale_x_continuous(breaks = sort(unique(df_sim$r_b))) +
-  scale_fill_gradient2(mid = 0) +
+  scale_fill_gradient2(mid = 0,
+                       limits = range(c(df_heat$rho1, df_heat$rho2))) +
   theme_classic() +
   theme(strip.background = element_blank()) +
   ggtitle("Ecosystem size")
@@ -66,7 +69,8 @@ g_branch <- df_heat %>%
        x = expression("Productivity ("*r[b]*")"),
        fill = "Slope") +
   scale_x_continuous(breaks = sort(unique(df_sim$r_b))) +
-  scale_fill_gradient2(mid = 0) +
+  scale_fill_gradient2(mid = 0,
+                       limits = range(c(df_heat$rho1, df_heat$rho2))) +
   theme_classic() +
   theme(strip.background = element_blank()) +
   ggtitle("Ecosystem complexity")
@@ -76,27 +80,27 @@ g_m <- (g_size + g_branch) + plot_annotation(tag_levels = "A")
 
 # plot: geometry effect ---------------------------------------------------
 
-lab <- c(`4` = "Low~productivity~(r[b]==4)",
+lab <- c(`8` = "Low~productivity~(r[b]==8)",
          `16` = "High~productivity~(r[b]==16)")
 
 ## ecosystem size effect
 g_np <-  df_plot %>% 
-    ggplot(aes(x = n_patch,
-               y = fcl,
-               color = factor(p_disturb),
[…]

### 2023-02-24 code/figure_si_sim.R

@@ -8,10 +8,12 @@ lapply(list("code/library.R",
             "code/format_sim_data.R"),
        source)
 
+df_coef <- readRDS("output/df_coef.rds")
+
 ## filter s & mean_disturb_source for visualization
 s_set <- 1
 mu_disturb <- 0.2
-r_set <- c(8, 20)
+r_set <- c(8, 16)
 
 ## df for heatmap
 df_heat <- df_param %>% 
@@ -20,14 +22,14 @@ df_heat <- df_param %>%
   filter(mean_disturb_source == mu_disturb | p_disturb == 0,
          (s == 0 & a_bp == 0) | s == s_set)
 
-## df for gam plot
+## df for loess plot
 df_plot <- df_sim %>% 
   filter(mean_disturb_source == mu_disturb | p_disturb == 0,
          (s == 0 & a_bp == 0) | s == s_set,
          r_b %in% r_set)
 
 
-# heatmap -----------------------------------------------------------------
+# plot: heatmap -----------------------------------------------------------
 
 theme_set(plt_theme)
 df_point <- expand.grid(r_b = r_set,
@@ -48,7 +50,8 @@ g_size <- df_heat %>%
        x = expression("Productivity ("*r[b]*")"),
        fill = "Slope") +
   scale_x_continuous(breaks = sort(unique(df_sim$r_b))) +
-  scale_fill_gradient2(mid = 0) +
+  scale_fill_gradient2(mid = 0,
+                       limits = range(c(df_heat$rho1, df_heat$rho2))) +
   theme_classic() +
   theme(strip.background = element_blank()) +
   ggtitle("Ecosystem size")
@@ -66,17 +69,19 @@ g_branch <- df_heat %>%
        x = expression("Productivity ("*r[b]*")"),
        fill = "Slope") +
   scale_x_continuous(breaks = sort(unique(df_sim$r_b))) +
-  scale_fill_gradient2(mid = 0) +
+  scale_fill_gradient2(mid = 0,
+                       limits = range(c(df_heat$rho1, df_heat$rho2))) +
   theme_classic() +
   theme(strip.background = element_blank()) +
   ggtitle("Ecosystem complexity")
 
 g_m <- (g_size + g_branch) + plot_annotation(tag_levels = "A")
 
-# gam plot ----------------------------------------------------------------
 
-lab <- c(`2` = "Low~productivity~(r[b]==8)",
-         `20` = "High~productivity~(r[b]==20)")
[…]

### 2023-02-24 code/figure_si_state.R
edit: state figure

@@ -11,7 +11,7 @@ lapply(list("code/library.R",
 ## filter s & mean_disturb_source for visualization
 s_set <- 1
 mu_disturb <- 0.8
-r_set <- c(4, 16)
+r_set <- c(8, 16)
 
 ## df for gam plot
 df_plot <- df_sim %>% 
@@ -23,11 +23,11 @@ df_plot <- df_sim %>%
                names_to = "state")
 
 
-# heatmap -----------------------------------------------------------------
+# plot: state transition --------------------------------------------------
 
 theme_set(plt_theme)
 
-lab <- c(`4` = "Low~productivity",
+lab <- c(`8` = "Low~productivity",
          `16` = "High~productivity")
 
 state_lab <- c(`s0` = "No species",

### 2023-02-24 code/figure_si_state.R
edit: figure state

@@ -112,19 +112,3 @@ ggsave(g_bp_all,
        filename = here::here(paste0("figure/si_figure_state_p_branch.pdf")),
        height = 14,
        width = 14)
-
-# omn_lab <- c("chain", "weak", "strong")
-# 
-# lapply(1:length(omn_lab), function(i) {
-#   
-#   ggsave(list_g_np[[i]],
-#          filename = here::here(paste0("figure/si_figure_state_n_patch_", omn_lab[i], ".pdf")),
-#          height = 7,
-#          width = 14)
-#   
-#   ggsave(list_g_bp[[i]],
-#          filename = here::here(paste0("figure/si_figure_state_p_branch_", omn_lab[i], ".pdf")),
-#          height = 7,
-#          width = 14)
-#   
-# })

### 2023-02-24 rmd/manuscript.Rmd
edit: manuscript, method

@@
 # Sensitivity data
-load(file = here::here("output/result_main.RData"))
+df_sim <- readRDS(here::here("output/sim_main.rds")) %>% 
+  filter(r_b %in% c(8, 16))
 
-n_rep <- max(sim_main_result$n_rep)
+n_rep <- max(df_sim$n_rep)
 
-n_scenario <- sim_main_result %>% 
+n_scenario <- df_sim %>% 
   filter(sd_disturb_source > sd_disturb_lon) %>% 
@@
 
-n_sim <- sim_main_result %>% 
+n_sim <- df_sim %>% 
   mutate(n_sim = n_timestep + n_burnin + n_warmup) %>% 
@@
 No data were used or produced in this research.
-R functions for simulations are provided as the R package mcbrnet version 1.2.3 (available at <https://github.com/aterui/mcbrnet>).
+R functions for simulations are provided as the R package mcbrnet version 1.3.0 (available at <https://github.com/aterui/mcbrnet>).
 Codes for simulations, statistical analysis, figures, and tables are available at <https://github.com/aterui/public-proj_fcl-river-theory>.
@@
 
-## **Tri-trophic dynamic simulation model**
+## **Tri-trophic model**
 
@@
 
-***Internal dynamics.*** Each time step starts with internal dynamics. The general formula is given as follows:
+***Internal dynamics.*** Each time step starts with internal dynamics. The general formula is the extension of the Nicholson-Bailey model, which is given as follows:
 
@@
 
-where
+where $e_{ij}$ is the energetic conversion efficiency from $i$ to $j$.
+The function $g(\cdot)$ defines the density-dependent recruitment for the basal species, and the function $f_k(\cdot)$ dictates the proportion of survived prey individuals after predation.
+In the above equations, we re-defined variables as follows to simplify the expression:
 
@@
 
-The parameter $e_{ij}$ is the energetic conversion efficiency from $i$ to $j$.
-The function $g(\cdot)$ defines the density-dependent recruitment for the basal species, and the function $f_k(\cdot)$ dictates the proportion of survived prey individuals after predation.
+The model assumes that ecological events occur in the following order: (1) $B$'s recruitment, (2) $C$'s predation on $B$ followed by its recruitment, and (3) $P$'s predation on $B$ and $C$ followed by its recruitment.
+We assumed that the intraguild prey $C$ consumes the basal species $B$ first because the species is specialized to prey on $B$.
+
 We used the Beverton-Holt function to describe the density-dependent growth of the basal species:
@@
 
-The parameter $a_{ij}$ defines the encounter rate between $j$ and $i$, $h_{ij}$ the handling time, and $\phi$ the switching function for the intraguild predator.
-We assumed that the degree of omnivory is dependent on the relative abundance of $B''_{x,t}$ and $C'_{x,t}$:
+The parameter $a_{ij}$ defines $j$'s attack rate to $i$, $h_{ij}$ the handling time, and $\phi$ the switching function for the intraguild predator.
+We assumed that switching is dependent on the relative abundance of $B''_{x,t}$ and $C'_{x,t}$:
 
 $$
-\phi = \frac{s(B''_x(x,t) - C'_{x,t})}{B''_x(x,t) + C'_{x,t}}
+\phi = \frac{s(B''_{x,t} - C'_{x,t})}{B''_{x,t} + C'_{x,t}}
 $$
@@
 ***Dispersal.*** We simulated distance-dependent dispersal along with a network.
-The number of emigrants from patch $x$ at time $t$, $E_{i,x}(t)$, is estimated as the product of dispersal probability $p_{d}$ and species abundance after internal dynamics $\bar{A}_{i,x}(t)$.
+The number of emigrants from patch $x$ at time $t$, $E_{ix,t}$, is estimated as the product of dispersal probability $p_{d}$ and species abundance after internal dynamics $\bar{A}_{ix,t}$.
 The dispersal probability was assumed to be constant for all three species.
-We calculated the relative immigration potential $\xi_{i,x}(t)$ considering network structure and species dispersal capability as follows:
+We calculated the relative immigration potential $\xi_{ix,t}$ considering network structure and species dispersal capability as follows:
 
@@
 \begin{aligned}
-\xi_{i,x}(t) &= \frac{\sum_{y, y \ne x}E_{i,y}(t)\exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}\text{E}_{i,y}(t)\exp(-\theta d_{xy})}
+\xi_{ix,t} &= \frac{\sum_{y, y \ne x}E_{iy,t}\exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}E_{iy,t}\exp(-\theta d_{xy})}
 &&\text{(16)}
@@
 where $d_{xy}$ is the network distance between patch $x$ and $y$ (shortest path), and $\theta$ indicates the inverse of mean dispersal distance (i.e., larger numbers mean shorter dispersal distance).
-The immigration potential at patch $x$, $\sum_{y, y \ne x}E_{i,y}(t)\exp(-\theta d_{xy})$, is scaled by the sum of immigration potential across patches so that $\sum_{x} \xi_{i,x}=1$.
-The expected number of immigrants into patch $x$ is calculated as $I_{i,x}(t)=\xi_{i,x}(t) \sum_{x}E_{i,x}(t)$.
+The immigration potential at patch $x$, $\sum_{y, y \ne x}E_{iy,t}\exp(-\theta d_{xy})$, is scaled by the sum of immigration potential across patches so that $\sum_{x} \xi_{i,x}=1$.
+The expected number of immigrants into patch $x$ is calculated as $I_{ix,t}=\xi_{ix,t} \sum_{x}E_{ix,t}$.
 
-***Food chain length.*** At each patch and time step, we determined FCL as the highest trophic position ($TP_{x}(t)$) of the existing species.
+***Food chain length.*** At each patch and time step, we determined local FCL as the highest trophic position ($\tau_{x,t}$) of the existing species.
 When the predator exists, the trophic position is determined as follows:
@@
 \begin{aligned}
-TP_{P,x}(t) &= \delta_{x}(t) + 2(1 - \delta_{x}(t)) + 1
+\tau_{x,t} &= \delta_{x,t} + 2(1 - \delta_{x,t}) + 1
 &&\text{(17)}
@@
 
-This definition accounts for the influence of omnivory.
+where $\delta_{x,t}$ is the derived parameter that represents the degree of omnivory and is defined as:
+
+$$
+\delta_{x,t} = \frac{e_{BP}B'_{x,t}f_1(B'_{x,t}, C_{x,t})[1 - f_2(B''_{x,t}, P_{x,t})]}{P^+_{x,t}}
+$$
+
+Thus, our definition accounts for the influence of omnivory.
+We defined the global FCL $\bar{\tau}$ across patches $N_p$ and time $T$ as $\bar{\tau} = \frac{\sum_x \sum_t \tau_{x,t}}{N_p T}$ (i.e., average over space and time).
 
@@
 
-Prior to the main simulation, we performed a one-patch simulation to explore the predation parameter space ($a_{ij}$ and $h_{ij}$) that allows stable coexistence of the three species in the absence of disturbance.
+Prior to the main simulation, we performed a one-patch simulation to explore the predation parameter space ($a_{ij}$ and $h_{ij}$) that allows stable coexistence of the three species without disturbance.
 In this simulation, we assumed $h_{ij} = h$ to reduce the dimension of parameter space.
 This analysis revealed that the stable coexistence of the three species is possible with a certain combinations of predation parameters.
-Within this parameter space, we explored the three scenarios of food web structure: "chain" ($(a_{BC}, a_{BP}, a_{CP}, h) = (X, X, X, X)$), "weak omnivory" ($(a_{BC}, a_{BP}, a_{CP}, h) = (X, X, X)$), and "strong omnivory" ($(a_{BC}, a_{BP}, a_{CP}, h) = (X, X, X)$).
+Within this parameter space, we used three parameter combinations to represent distinctive food webs with varied levels of omnivory: "chain (no omnivory)" [$(a_{BC}, a_{BP}, a_{CP}, h) = (0.5, 0, 0.0025, 0.5)$], "weak omnivory" ($0.5, 0.02, 0.025, 0.5$), and "strong omnivory" ($0.5, 0.04, 0.0025, 0.75$).
 We used $s = 1$ except for the chain scenario with $s = 0$.
@@
 
-We controlled disturbance level with disturbance probability $p_{m}$ (low disturbance, `r min(sim_main_result$p_disturb)`; high disturbance, `r max(sim_main_result$p_disturb)`).
-We used the identical value of mean disturbance intensity at headwaters ($\mu_{m}=$ `r max(sim_main_result$mean_disturb_source)`), spatial variation in disturbance intensity at headwaters ($\sigma_{h}=$ `r max(sim_main_result$sd_disturb_source)`), and local disturbance variation ($\sigma_{l}=$ `r min(sim_main_result$sd_disturb_lon)`) regardless of disturbance scenario.
-This setup ($\sigma_h > \sigma_l$) produces high spatial variation among tributaries with strong within-tributary downstream correlation, a typical pattern in real river networks [@nakamuraDisturbanceRegimesStream2000; @bendaNetworkDynamicsHypothesis2004].
-Productivity was varied by setting different values of basal species' growth rate ($r_{b}=$ `r unique(sim_main_result$r_b)`).
-For dispersal, we examined short- and long-distance dispersal scenarios (short-distance dispersal $\theta=$ `r max(sim_main_result$theta)`; long-distance dispersal $\theta=$ `r min(sim_main_result$theta)`) with constant dispersal probability ($p_{d}=$ `r unique(sim_main_result$p_dispersal)`).
+Under each scenario of omnivory, we crossed disturbance (frequency and intensity), productivity, and dispersal.
+Disturbance frequency and intensity were controlled via $p_m$ ( `r unique(df_sim$p_disturb) %>% sort()`) and $\mu_m$ (`r unique(df_sim$mean_disturb_source) %>% sort()`), respectively.
+Spatial variation in disturbance intensity at headwaters ($\sigma_{h}=$ `r max(df_sim$sd_disturb_source)`), and local disturbance variation ($\sigma_{l}=$ `r min(df_sim$sd_disturb_lon)`).
+This setup ($\sigma_h > \sigma_l$) produces high spatial variation among tributaries while allowing strong within-tributary downstream correlation, a typical pattern in real river networks [@nakamuraDisturbanceRegimesStream2000; @bendaNetworkDynamicsHypothesis2004].
+We used the basal species' growth rate $r_b$ (`r unique(df_sim$r_b)`) as a proxy for resource availability as it resembles the gross primary production.
+For dispersal, we examined short- and long-distance dispersal scenarios (short-distance dispersal $\theta=$ `r max(df_sim$theta)`; long-distance dispersal $\theta=$ `r min(df_sim$theta)`) with constant dispersal probability ($p_{d}=$ `r unique(df_sim$p_dispersal)`).
 
-Under each scenario, we ran `r max(sim_main_result$n_rep)` simulation replicates with `r n_sim` timesteps of trophic dynamics (including `r n_sim - unique(sim_main_result$n_timestep)` steps of warmup and burn-in).
-During the warmup period (`r unique(sim_main_result$n_warmup)` time steps), we introduced propagules into habitat patches every 10 time steps to allow species to establish populations.
-Propagules are random draws from $\text{Poisson}(0.5)$.
-Subsequently, we ran a burn-in period with no propagules for `r unique(sim_main_result$n_burnin)` time steps to remove the effects of initial conditions.
-We obtained the mean FCL across all habitat patches for the last `r unique(sim_main_result$n_timestep)` time steps.
+We ran `r max(df_sim$n_rep)` simulation replicates with `r n_sim` timesteps of trophic dynamics (including `r n_sim - unique(df_sim$n_timestep)` steps of warmup and burn-in).
+During the warmup period (`r unique(df_sim$n_warmup)` time steps), we introduced propagules into habitat patches every 10 time steps to allow species to establish populations.
+Propagules are random draws from $\text{Poisson}(\lambda)$ ($\lambda=100, 10,~\text{and}~1$ for $B$, $C$, and $P$, respectively).
+Subsequently, we ran a burn-in period with no propagules for `r unique(df_sim$n_burnin)` time steps to remove the effects of initial conditions.
+We obtained the global FCL $\bar{\tau}$ for the last `r unique(df_sim$n_timestep)` time steps.
 Note that simulation replicates are differ in ecosystem size ($N_{p}\sim \text{Unif}(10,150)$) and ecosystem complexity ($P_{b} \sim \text{Unif}(0.01, 0.99)$) so we can estimate their effects on FCL.

### 2023-02-26 rmd/manuscript.Rmd
edit: manuscript

@@
 
-The origin of variation in food chain length (FCL), the number of feeding links from the basal species to top predators, has intrigued ecologists for decades due to its importance in regulating top-down effects [@paceTrophicCascadesRevealed1999], energy flow [@wangBiodiversityEcosystemFunctioning2018], and contaminant concentrations in top predators [@kiddEffectsTrophicPosition1998].
+The origin of variation in food chain length (FCL) has intrigued ecologists for decades due to its importance in regulating top-down effects [@paceTrophicCascadesRevealed1999], energy flow [@wangBiodiversityEcosystemFunctioning2018], and contaminant concentrations in top predators [@kiddEffectsTrophicPosition1998].
 Hypotheses regarding controls of FCL can be organized into three broad categories.

### 2023-02-26 rmd/manuscript.Rmd
edit: manuscript
result, one paragraph

@@
 
-We simulate trophic community dynamics within branching habitats to investigate the effects of extrinsic controls of FCL, such as ecosystem size (number of patches), complexity (branching probability), and disturbance regime.
+We simulate trophic community dynamics within branching habitats to investigate the effects of extrinsic controls of FCL, such as ecosystem geometry [size (number of patches) and complexity (branching probability] and disturbance regime.
 In addition, our simulation model incorporates intrinsic drivers of FCL, including the degree of omnivory, productivity level, and species' dispersal ability, allowing us to assess the context dependency on drivers of FCL across extrinsic factors.
@@
 
-## Overall pattern
+## Overall effects of ecosystem geometry
 
-There were three consistent patterns in the relationship between FCL and ecosystem properties.
-First, both ecosystem size and complexity had a stronger positive effect in high disturbance regime.
-Second, the effects of ecosystem complexity was stronger than ecosystem size in general.
-Third, omnivory had a negative influence on the FCL regardless of simulation scenarios.
-Below, we report our results for low and high disturbance regimes separately.
+There were three consistent patterns in the relationship between FCL and ecosystem geometry (size and complexity).
+First, positive effects of ecosystem size and complexity were generally found when both disturbance frequency and intensity were high.
+Second, omnivory plays a central role in modulating the effects of ecosystem geometry on the FCL.
+Third, resource availability and dispersal did not change the qualitative relationship between the FCL and ecosystem geometry, although they had important direct effects on the FCL.
 
-## **Low disturbance regime**
+## Ecosystem geometry interacts with disturbance
+
+Ecosystem geometry interacted with disturbance to dictate FCL in branching ecosystems.
+With no disturbance, both geometric features (size and complexity) had limited influence on FCL.
+However, as disturbance frequency increases, the relationships between FCL and ecosystem geometry became positive.
+The exception is the "chain" scenario (i.e., no omnivory).
+When this food web structure was combined with no disturbance, ecosystem size had a clear positive effect on FCL while ecosystem complexity had a negative influence.
+These opposing influences of ecosystem size and complexity may reflect their contrasting impacts on the total carrying capacity within a network (i.e., carrying capacity summed across all patches): ecosystem size increases the total carrying capacity whereas ecosystem complexity decreases it.
 
@@
 Dispersal had no apparent effects on FCL (compare top and bottom panels in **Figure 2A and 2B**).
+
+## Ecosystem geometry interacts with omnivory
+
+## Productivity and dispersal
+
+##
 

### 2023-02-27 rmd/manuscript.Rmd
edit: manuscript

@@
 \begin{aligned}
-B'_{x,t} &= B_{x,t}g(B_{x,t}) &&\text{after B's recruitment}\\
-B''_{x,t} &= B'_{x,t}f_1(B'_{x,t}, C_{x,t}) &&\text{after C's predation on B}\\
-C'_{x,t} &= e_{BC}B'_{x,t}[1 - f_1(B'_{x,t}, C_{x,t})] &&\text{after C's recruitment}
+B'_{x,t} &= B_{x,t}g(B_{x,t}) &&\text{B's recruitment}\\
+B''_{x,t} &= B'_{x,t}f_1(B'_{x,t}, C_{x,t}) &&\text{C's predation on B}\\
+C'_{x,t} &= e_{BC}B'_{x,t}[1 - f_1(B'_{x,t}, C_{x,t})] &&\text{C's recruitment}
 \end{aligned}
@@
 There were three consistent patterns in the relationship between FCL and ecosystem geometry (size and complexity).
-First, positive effects of ecosystem size and complexity were generally found when both disturbance frequency and intensity were high.
-Second, omnivory plays a central role in modulating the effects of ecosystem geometry on the FCL.
-Third, resource availability and dispersal did not change the qualitative relationship between the FCL and ecosystem geometry, although they had important direct effects on the FCL.
+First, the effects of ecosystem size and complexity were modulated by disturbance and omnivory.
+Second, resource availability and dispersal had limited influences on the qualitative relationship between FCL and ecosystem geometry, although they had important direct effects on the FCL.
+Below we describe these results in detail.
 
-## Ecosystem geometry interacts with disturbance
+## Interactive effects of ecosystem geometry, disturbance, and omnivory
 
@@
 However, as disturbance frequency increases, the relationships between FCL and ecosystem geometry became positive.
-The exception is the "chain" scenario (i.e., no omnivory).
-When this food web structure was combined with no disturbance, ecosystem size had a clear positive effect on FCL while ecosystem complexity had a negative influence.
+This interactive effect emerged because the sensitivity of FCL to disturbance depended on ecosystem geometry.
+Although increased disturbance frequency reduced overall FCL, smaller and/or simpler ecosystems exhibited the greater vulnerability.
+Therefore, disturbance interact with ecosystem geometry, producing the interdependency of these factors in determining FCL.
+The interactive effects, however, were not observed with low disturbance intensity (proportion of individuals removed per disturbance event); this suggests that disturbance intensity must be strong enough to be influential.
+
+The exception is the "chain" scenario (i.e., no omnivory) with no disturbance.
+In this scenario, ecosystem size had a clear positive effect on FCL while ecosystem complexity had a negative influence.
 These opposing influences of ecosystem size and complexity may reflect their contrasting impacts on the total carrying capacity within a network (i.e., carrying capacity summed across all patches): ecosystem size increases the total carrying capacity whereas ecosystem complexity decreases it.
-
-In ecosystems with a low disturbance regime, omnivory largely determined FCL (colored lines in **Figure 2**), with no-omnivory food chains ($P$ only consumes $C$) having the longest FCL, and strong omnivory ($P$ consumes very little $C$) having the shortest FCL.
-The primary mechanism for the decreased FCL was different in low and high productivity scenarios.
-With low productivity, strong omnivory caused shorter FCL because it intensified competition between consumer and predator species.
-We can see this effect in the decreased occupancies of consumer and predator species in a low productivity scenario (most left column in **Figure S1 -- S4**).
-However, when productivity was high, the occupancies of both species were \~1.0, indicating that the structural constraint of strong omnivory made FCL shorter (second column from left in **Figure S1 -- S4**).
-Productivity played a secondary role in increasing FCL, particularly when omnivory was strong (compare left and right panels in **Figure 2A and 2B**).
-Dispersal had no apparent effects on FCL (compare top and bottom panels in **Figure 2A and 2B**).
-
-## Ecosystem geometry interacts with omnivory
+However, these patterns characteristic to the "chain" scenario diminished when considering some levels of omnivory because food webs with omnivory were less reliant on the carrying capacity of the basal species.
+In addition, omnivory gave rise to more robustness/resilience to increased disturbance frequency.
 
 ## Productivity and dispersal
-
-##
-
-Although the effects of ecosystem properties on FCL were minimal, there were noticeable effects on the occupancies of constituent species in some scenarios.
-For example, the occupancies of $C$ and $P$ increased with ecosystem size when omnivory was strong (see bottom two rows in left most column in **Figures S1 and S2**).
-However, these changes did not translate into FCL because $C$ and $P$ occupy a similar trophic position (**Figure 1D**).
-Meanwhile, the occupancy of $P$ decreased with ecosystem complexity when omnivory was strong, but FCL remained almost the same as the increase of $C$ could compensate for the loss of $P$ (**Figure S3 and S4**).
-
-## **High disturbance regime**
-
-On average, increased disturbance frequency decreased FCL, and omnivory had similar effects on FCL (**Figure 3**).
-More importantly, a high disturbance regime produced different responses of FCL to ecosystem properties.
-FCL increased across gradients of ecosystem size and complexity under all productivity, omnivory, and dispersal scenarios (**Figure 3**).
-The effect of ecosystem complexity was generally clearer than that of ecosystem size.
-It is important to note that the positive relationships emerged in a high disturbance regime because smaller and simpler ecosystems were disproportionately vulnerable to increased disturbance frequency (compare far left along the x-axis in each panel of **Figures 2 and 3**).
-
-Dispersal buffered the disturbance effects on FCL and the magnitude of buffering differed by ecosystem structure.
-Larger ecosystems benefited more from long dispersal (compare top and bottom panels in **Figure 3**) because greater dispersal capability is required to reach distant patches.
-In particular, this pattern is clearer in simple, less branched ecosystems because habitat patches are more isolated in linearly arranged networks (**Figure 3**).
-Productivity increased FCL as observed in a low disturbance regime, but it did not modulate the qualitative relationships between FCL and ecosystem properties (compare left and right panels in **Figure 3**).
-
-When strong omnivory exists in low-productivity systems, different processes drove the positive effects of ecosystem size and complexity on FCL.
-The positive effects of ecosystem size are mainly explained by the increased occupancy of $P$ (third column from left in **Figure S1 and S2**).
-In contrast, the increased occupancy of $C$ maintained the positive relationship between FCL and ecosystem complexity (third column from left in **Figure S3 and S4**).
 

### 2023-03-02 rmd/manuscript.Rmd
edit: manuscript

@@
 Subsequently, we ran a burn-in period with no propagules for `r unique(df_sim$n_burnin)` time steps to remove the effects of initial conditions.
+Note that simulation replicates are differ in ecosystem size ($N_{p}\sim \text{Unif}(10,150)$) and ecosystem complexity ($P_{b} \sim \text{Unif}(0.01, 0.99)$) so we can estimate their effects on FCL.
+
 We obtained the global FCL $\bar{\tau}$ for the last `r unique(df_sim$n_timestep)` time steps.
-Note that simulation replicates are differ in ecosystem size ($N_{p}\sim \text{Unif}(10,150)$) and ecosystem complexity ($P_{b} \sim \text{Unif}(0.01, 0.99)$) so we can estimate their effects on FCL.
-All analysis was performed in R 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2021]
+We also calculated the proportion of the patch state ("no species", "B", "B + C", "B + P", and "B + C + P") to infer mechanisms that drive changes in FCL.
+All analysis was performed in R 4.2.1 [@rcoreteamLanguageEnvironmentStatistical2021]
 
@@
 In this scenario, ecosystem size had a clear positive effect on FCL while ecosystem complexity had a negative influence.
-These opposing influences of ecosystem size and complexity may reflect their contrasting impacts on the total carrying capacity within a network (i.e., carrying capacity summed across all patches): ecosystem size increases the total carrying capacity whereas ecosystem complexity decreases it.
+These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increases the total carrying capacity whereas ecosystem complexity decreases it.
 However, these patterns characteristic to the "chain" scenario diminished when considering some levels of omnivory because food webs with omnivory were less reliant on the carrying capacity of the basal species.
-In addition, omnivory gave rise to more robustness/resilience to increased disturbance frequency.
 
 ## Productivity and dispersal
+
+Productivity and dispersal had limited influences on the qualitative relationship between FCL and ecosystem geometry.
+However, they had important direct effects on FCL.
+Productivity influences were dependent on omnivory.
+Overall, greater productivity increased (chain & strong omnivory) or decreased FCL (weak omnivory), although the effect size varied with ecosystem geometry and disturbance levels.
+Dispersal distance had a positive effect on FCL, particularly in scenarios with frequent and intense disturbance.
+
+## Underlying state changes
 
@@
 This supports our hypothesis that ecosystem complexity may represent an overlooked yet important extrinsic control of FCL. However, the relationship between FCL and ecosystem complexity varied with other extrinsic and intrinsic factors, supporting the emerging paradigm that FCL is regulated by multiple factors [@mchughDualInfluencesEcosystem2010; @wardMechanisticTheoryAquatic2017].
-Although we focus on river networks here, we expect our results to be transferable to other systems with branching structures, such as mountain ranges and cave systems.
-Our findings should also be relevant to non-branching systems, where different metrics of ecosystem complexity would be necessary.
+Although we focus on river networks here, we expect our results to be transferable to other systems with non-branching structures.
+Ecosystems are intrinsically complex -- forest stands possess complex vertical structures, which may provide heterogeneous habitats with different levels of vulnerability to wind and/or fire disturbance.
+Topographic variation can also mediate spatial disturbance dynamics by providing local refugia in a landscape.
+An explicit consideration of ecosystem complexity may therefore increase our ability to predict FCL.
 
-While increased disturbance frequency consistently reduced FCL as predicted by the dynamical stability hypothesis [@pimmNumberTrophicLevels1977], more important findings in this study were evident from their synergistic effects with ecosystem structure.
-Our low disturbance regime scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independently of ecosystem size and complexity.
-Consequently, the food web structure largely controlled FCL.
-However, under high disturbance regimes, only large and complex networks could provide diverse and heterogeneous habitats with potential refugia, allowing species to thrive and maintain long food chains.
+The most important finding in this study is the synergistic effects between ecosystem geometry and disturbance.
+Our no disturbance scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independently of ecosystem size and complexity (except the chain scenario).
+However, as the frequency of intense disturbance increases, small and/or simple networks could not maintain long food chains because habitats were too homogeneous with limited disturbance refugia.
 While empirical studies rarely looked at the influence of ecosystem complexity on FCL, several studies provide some support for this theoretical prediction.
@@
 
-The interactive effect of disturbance and ecosystem structure has important implications for ecosystem management because anthropogenic climate change is predicted to increase the frequency and magnitude of disturbances [@brunnerFutureStreamflowRegime2020; @mittalImpactHumanIntervention2016].
+The synergy between ecosystem geometry and disturbance has also an implication for why the dynamical stability hypothesis [@pimmNumberTrophicLevels1977] has been least supported by the literature.
+While increased disturbance frequency consistently reduced FCL, its magnitude depended strongly on ecosystem geometry.
+This observation should provide a possible explanation for the limited support for the dynamical stability hypothesis, because such an interaction has been rarely assessed in previous studies.
+Although FCL research often assumes that the effects of environmental controls are independent of one another, exploring potential interactions among them could provide useful insights into what controls FCL in nature.
+
+Our results are crucial given that anthropogenic climate change is predicted to increase the frequency and magnitude of disturbances [@brunnerFutureStreamflowRegime2020; @mittalImpactHumanIntervention2016].
 River networks are especially vulnerable due to their propensity to go dry during droughts [@zipperPervasiveChangesStream2021] as well as susceptibility to more intense flood disturbance with increasing storm intensity [@schiermeierIncreasedFloodRisk2011].
 Therefore, the preservation of habitat diversity conferred by ecosystem size and complexity is critical to biodiversity conservation under rapid environmental changes.
-In addition, maintaining habitat connectivity is particularly important for large, linearly-arranged networks.
-Although dispersal has received less attention in food chain research [but see @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012; @wangMetapopulationCapacityDetermines2021], our simulation highlighted food webs in these networks are more vulnerable to habitat fragmentation because habitat patches are inherently isolated from one another.
+In addition, maintaining habitat connectivity is important because the effect of spatial refuges can emerge only through recolonization after disturbance.
 Dams and road crossings have altered the connectivity between suitable habitat patches or have completely blocked dispersal [@ishiyamaPredictingEcologicalImpacts2018; @moritaEffectsHabitatFragmentation2002; @perkinFragmentationAltersStream2012].
 Light pollution has also negatively affected the ability of adult aquatic insects to disperse between habitat patches [@perkinEffectsArtificialLighting2014].
-These human activities may disproportionately impact food webs in large watersheds with simple network structures.
+These human activities may increase the vulnerability of food webs to rapid environmental changes.
 
-It is not surprising that FCL declined with increasing omnivory because it effectively compresses a food web (**Figure 1C**) and/or intensifies competition between consumer and predator species.
-However, how omnivory influences the associations with ecosystem structure has rarely been explored [@takimotoEffectsProductivityDisturbance2012; @wardMechanisticTheoryAquatic2017].
-An important result in our simulation is that omnivory does not change the qualitative relationships between FCL and ecosystem properties, as opposed to productivity effects.
+Another interesting result of our simulation is the interaction with omnivory.
+In particular, when the chain motif was combined with the no-disturbance scenario, ecosystem size and complexity had contrasting influences on FCL.
+It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity; therefore, the total carrying capacity within a network dictates FCL.
+The effect of ecosystem size is understandable because it simply increases the total carrying capacity.
+In contrast, ecosystem complexity, here assessed as branching probability, decreases the total carrying capacity as more branched networks have disproportionate numbers of headwater habitats with small local carrying capacities.
+However, the dependence of FCL on local carrying capacity diminished in weak and strong omnivory scenarios, explaining the lack of positive (ecosystem size) or negative FCL-geomtery relationship (ecosystem complexity) in these motifs with no disturbance.
+
+Importantly, however, disturbance overwhelmed the effect of omnivory on the FCL-geometry relationships: as the frequency of intense disturbance increases, the positive relationships between FCL and ecosystem geometry emerged regardless of omnivory levels.
 This result provides a possible explanation for why the ecosystem size hypothesis has been consistently supported in various ecosystems [@takimotoEnvironmentalDeterminantsFoodchain2013] and raises the possibility of ecosystem complexity as an important control of FCL in branching networks.
-In addition, our results imply that the size and complexity effects are robust to food web reorganization, which is often associated with species extirpations or introductions.
+Our results also imply that, if disturbance is strong and frequent, the effects of ecosystem geometry are robust to food web reorganization, which is often associated with species extirpations or introductions.
 The strength of omnivory in these novel communities can be increased or decreased, depending on the introduced species and the extant species pool or community.
 For example, non-indigenous fishes in the Eastern Mediterranean Sea outcompete native species for high-quality prey items, causing native species to increase their consumption of non-preferred prey items [@fanelliDepictingNovelEastern2015].
-However, such substantial changes in food web structure may have limited influences on how ecosystem structure dictates FCL.
+However, such substantial food web reorganization may have limited influences on how ecosystem geometry dictates FCL.
 

### 2023-03-03 rmd/manuscript.Rmd
edit: manuscript

@@
 For example, large ecosystems have increased species richness (species-area relationship [@lomolinoEcologyMostGeneral2000]), are buffered against disturbances [@saboRoleDischargeVariation2010], and have greater numbers of immigrant sources in the ecosystem with stable metapopulations [@takimotoEnvironmentalDeterminantsFoodchain2013].
-Meanwhile, the theoretical basis for the other two hypotheses rests on implicit assumptions of linear food chains, which are often violated in natural systems.
+Meanwhile, the productivity and disturbance effects are highly sensitive to the internal structure of food web motifs.
 As such, the effects of disturbance and resource availability on FCL have been found to be highly variable among ecosystem types and geographic regions [@takimotoEffectsProductivityDisturbance2012; @diehlIntraguildPreySuffer2001; @takimotoEnvironmentalDeterminantsFoodchain2013; @postLongShortFoodchain2002].
@@
 
+State transition shed lights on how FCL changes occurred.
+In the chain motif, the "additive" mechanism (i.e., the addition of higher trophic levels) is always responsible for the changes in FCL.
+With disturbance, state transition occurred from "B" to "B + C" as ecosystem size or complexity increases.
+With no disturbance, state transition was different between size and complexity.
+Ecosystem size facilitated the transition from "B + C" to "B + C + P", while ecosystem complexity drove the opposite transition ("B + C + P" $\rightarrow$ "B + C").
+
+In the weak and strong omnivory motifs, ecosystem size and complexity increased FCL through different mechanisms.
+Ecosystem size increased FCL primarily through the additive mechanism, facilitating the state transition from "B" to "B + P", particularly with frequent disturbance.
+However, the "insertion" mechanism (the transition from "B + P" to "B + C + P") also took place in the strong omnivory motif with rare disturbance ($p_m = 0.05$).
+In contrast, the insertion mechanism was responsible for the effect of ecosystem complexity on FCL.
+In the weak omnivory scenario, the transition from "B + P" to "B + C + P" was predominant.
+In the strong omnivory scenario, transition patterns depended on disturbance frequency.
+When disturbance was rare ($p_m = 0.05$), the insertion mechanism ("B + P" $\rightarrow$ "B + C + P") drove the FCL change.
+However, as disturbance frequency increased, the primary process switched to the additive mechanism ("B" $\rightarrow$ "B + P").
+
 # Discussion
@@
 Another interesting result of our simulation is the interaction with omnivory.
-In particular, when the chain motif was combined with the no-disturbance scenario, ecosystem size and complexity had contrasting influences on FCL.
+In particular, when the chain motif was combined with no-disturbance, ecosystem size and complexity had contrasting influences on FCL.
 It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity; therefore, the total carrying capacity within a network dictates FCL.
@@
 However, such substantial food web reorganization may have limited influences on how ecosystem geometry dictates FCL.
+
+The effect of productivity was context dependent, supporting previous theoretical predictions.
+While productivity clearly increased FCL in the chain motif, this effect was vague (strong omnivory) or even reversed (weak omnivory).
+High productivity effectively increases the abundance of the basal species, stimulating the prey switching of the intraguild predator from the intraguild prey to the basal species.
+As such, increased productivity does not necessarily increase FCL.
+Our results suggest that this prediction from non-spatial models can be applied to spatially complex ecosystems.
+
+Ecosystem size and complexity led to the distinct patterns of state transitions.
+The state transitions with increased ecosystem size are likely associated with increased immigration.
+It is clear that the total number of dispersing propagules increases with increasing ecosystem size, providing opportunities to maintain longer food chains.
+The exception was the strong omnivory motif with rare disturbance, in which the insertion mechanism operated.
+This is likely because rare disturbance inhibited the exclusion of the intraguild prey by suppressing the abundance of the intraguild predator.
+Similarly, increased immigration may be generally responsible for the state transitions with increasing ecosystem complexity.
+Habitat patches are closer in more branched networks as it compresses the network configuration.
+However, unlike ecosystem size, ecosystem complexity does not increase the total number of dispersing propagules - rather, it enhances connectivity among habitat patches.
+This difference may explain the distinct patterns of state transitions along the gradients of ecosystem size and complexity.
 

### 2023-03-06 rmd/reply1.Rmd
add: response letter

---
title: "RE: Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems"
output:
  pdf_document:
    latex_engine: xelatex
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
mainfont: Times New Roman
---

## Response to Reviewer #1:

### Major comment

This ms uses a simulation model to study how environmental variables affect food-chain length (FCL) in river networks. Environmental determinants of FCL are a classic topic in ecology. Productivity, disturbance and ecosystem size have been considered as major potential determinants. The ecosystem size effect of increasing FCL has received strongest support from empirical studies so far. Productivity has received less support, and the disturbance effect has been supported the least frequently. A novel contribution of the current ms is to consider another potential determinant of FCL, ecosystem complexity, in addition to the other three. The ms is focused on examining the following environmental variables: ecosystem size (the number of habitat patches in a whole river network), ecosystem complexity (the probability of branching, which is a model parameter controlling the proportion of habitat patches serving as confluences in a river network), disturbance frequency, and productivity (intrinsic growth rate of basal species).

A significant finding of this ms is that ecosystem size and complexity can interact with the disturbance effect. This is because ecosystem size and complexity increase spatial heterogeneity in disturbance. In small or simple ecosystems, disturbance is spatially uniform, strongly limiting FCL. In large or complex ecosystems, heterogeneity in disturbance creates spatial refuges and weakens the negative effect of disturbance. The way that ecosystem size and complexity create disturbance heterogeneity is nicely modeled.

> Thank you!

The ms contains potentially important findings for FCL research but underlying modeling exercises are poor. I have two main concerns. First, more in depth analyses and interpretations are necessary about interacting effects between ecosystem size and complexity and disturbance. Second, the way that trophic interactions are modeled seems incorrect in some respects. A thorough reconstruction of a model may be necessary, although I feel that a correct modeling would not alter the basic conclusions of the current ms. Below, I illustrate major points. I will also list some minor points.

> Thank you for your careful review - your review improved our manuscript substantially. We followed your advise -- please see our response below.

(1-1) Based on the obtained results, the ms emphasizes that ecosystem size and complexity have positive effects on FCL but their effects depend on disturbance. This is an important message. Another important message is that the disturbance effect depends on ecosystem size and complexity. Although both messages mean the same literally, the latter could explain why empirical research has found variable disturbance effects on FCL. Since the present results may provide an answer, the ms could emphasize the latter message, too.

> This is a nice comment! Yes, we have clarified this point in the revised manuscript by adding a paragraph in the discussion (XXX). Also, we changed the language in Results (XXX).

(1-2) Given that a major finding is the interacting effects between disturbance and ecosystem size/complexity, more in depth analysis on the interacting effects would greatly improve the ms. In particular, it would be worthwhile to provide results illustrating how FCL responds along a disturbance gradient and how the responses differ between small vs. large ecosystems and between simple vs. complex ecosystems. In addition, while the current ms looks at disturbance frequencies (p_m), effects of disturbance intensities (mu_m) should also be explored because frequency and intensity are both important characteristics of disturbance. Because disturbance propagates through a river network, effects of frequency and intensity might be different, which could add interesting and useful insights.

> Agreed. We simulated a wider range of $p_m$ values to see how disturbance frequency changes FCL-geometry relationships (please see revised figures). We also included a different level of disturbance intensity ($\mu_m = 0.2$) - the lower intensity weakened the interactive effect of disturbance frequency. We noted this result in the revised MS (XXX).

(1-3) Mechanisms underlying the interacting effects should also be explored in some more detail. A potential mechanism, as discussed in the ms, is greater spatial heterogeneity in disturbance in larger or more complex ecosystems. Figure 1B is presented to show that spatial refuges from disturbance are more likely to occur in more complex ecosystems (i.e., with higher branching probabilities). To show that refuges are a potential mechanism, figures like fig. 1B must be compared between frequent and infrequent disturbance. I would expect that the occurrence of spatial refuges is more certain with more frequent disturbance (i.e., clearer longer tails on the low intensity side).

> XXX

(2-1) The way that trophic interactions are modeled is unnatural and probably wrong. One of the problems is that the amount of a species being consumed can exceed its abundance. When the consumed amount exceeds the abundance, surplus consumption vanishes (equations 7). There seems no biological justification for this peculiar assumption. Rather, it seems that the assumption was made up to keep the survival probability positive (equation 6). In general, it is not straightforward to formulate type II functional responses into discrete-time predator-prey population models. Detailed accounts are given in Hassell (1978, The dynamics of arthropod predator-prey systems, Chap. 2 and Appendix I) and in a recent work by Weide et al. (2019, <https://doi.org/10.1016/j.mbs.2018.12.010>). Another solution might be to model within-time step events (consumption and predation-induced death) as continuous time processes and between-time step events (reproduction and disturbance-induced death) as discrete time processes.

> Thank you for this very important comment! We have completely rewrote the model as an extension of the Nicholson-Bailey model. Important results were similar, but there were some qualitative changes as well. We have fully re-structured the Method section based on this revision.

(2-2) Although it is claimed that type II functional responses are assumed, the formulation having predator abundance in the denominator and numerator (equation 9) is incorrect. The formulation of f_ij (line 144) is a correct formulation of type II responses. Moreover, once this error is corrected, it should be checked if beta\_(CP, x) is positive (equation 12).

> Thank you, we have corrected this point as well.

(2-3) The preference function for predators is introduced (equation 10). The logic underlying this preference function (Nishijima et al. 2004) would be that relative preference reflects relative gains of predators feeding exclusively on resource and consumer species (e_BP*f_BP vs. e_CP*f_CP). However, these resource gains do not closely translate into predator's population growth rate (equation 14), breaking the logic underlying the preference function.

> We appreciate your insights. We revisited the switching function to properly represent this in a discrete form.

### Minor comment

Line 39+: "the theoretical basis for the other two hypotheses (resource availability and disturbance hypotheses) rests on implicit assumptions of linear food chains." This seems incorrect. Classic literature on the two hypotheses have considered non-linear food webs.

> The reviewer is correct. We have address this point.

Line 49+: "a maximum length of three": FCL may be defined as the number of trophic links or trophic levels. Here the ms messes up the two definitions. Please be consistent.

> Thank you, we have corrected.

Line 59: Lakes and islands are not necessarily either simple or two-dimensional. They can be viewed as having complex vertical structures of water columns or vegetations.

> Thank you, we have revised as follows: "...theoretical conceptualization has implicitly assumed two-dimensional systems, which may be a fair approximation of lakes."

Lines 110+: What is the relationship between ecosystem size/complexity and the total carrying capacity of a whole ecosystem? It appears that total carrying capacity depends on ecosystem size and complexity. What parts of the ecosystem size/complexity effects are explained by total carrying capacity?

> Thank you for raising this point. Ecosystem size increases the total carrying capacity, while ecosystem complexity decreases it (branching disproportionately increases the number of headwaters with low carrying capacities). Our additional analysis indicated that the "chain" motif is particularly influenced by local carrying capacity; in the meantime, omnivory scenarios were insensitive to local carrying capacity. We have clarified these points in Results and Discussion.

Although both intrinsic growth rates and carrying capacities of basal species should affect local productivity, the ms considers only intrinsic growth rates as local productivity.

> XXX

Line 167+: The definition of food-chain length. FCL at the whole ecosystem level should be defined here (instead of line 194).

> Thank you, we followed this.

Line 184+: More explanations may be necessary about why both attack rates and conversion efficiencies are controlled to change omnivory.

> After some rethinking, we decided to change only attack rates because it is enough to represent scenarios we are tying to create. Thank you for this comment!

Line 184+: In case of "chain", it is wired that a_BP=0.1 despite basal resource does not contribute to predator abundance at all (e_BP=0).

> True, we have corrected this.

Line 184+: Since the model describes abundance dynamics, it does not seem to make biological sense that e_BP, e_CP, or e_BC (conversion efficiencies) are greater than 1 unless these predators and consumers are parasites.

> Thank you, we have fixed the values as one for simplicity.

Line 221: "similar", does it mean "similar to the case of low disturbance"? Please clarify.

> We have restructured Results in light of the reanalysis. Your input is appreciated though!

Line 247+: Here, the ms suggests implications beyond river networks. More elaborate inferences would be helpful. The current results obtained for river networks depend on how disturbance propagates through a network. Are there similar mechanisms in other types of "branching ecosystems" such as mountain ranges and caves? It seems that ecosystem complexity, but not branching structures, is an integral part of the mechanism that create spatial refuges from disturbance. Too much emphasis on branching structures might mislead implications of the current results.

> This is a very thoughtful comment -- we have revised the first paragraph of Discussion following this advise. Thank you!

Figure 1B: More detailed explanations are necessary about how the figure is created. The X axis should have a larger maximum value to show the whole range of the disturbance intensity distributions. What does the dashed line represent?

> XXX

------------------------------------------------------------------------

## Response to Reviewer #2:

Food chain length (FCL) is important in the movement of energy and material through ecosystems and is an indicator used in ecosystem management. I have read the manuscript "Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems". The manuscript takes a simulation approach to examining multiple drivers of food chain length related to branching ecosystems like river ecosystems. I enjoyed reading this manuscript and found that is generally well written. I think that the topic is of interest to general ecological audience. The manuscript is motivated by the claim that there is a need for a unified framework that includes multiple drivers of FCL, including that there is little research on the relationship between complexity (spatial complexity) and heterogeneity.

The results that relate FCL to ecosystem complexity are interesting particularly as they are considered explicitly in a simulation with other factors that have received more attention (ecosystem size, productivity, disturbance). Some of the other result like the general importance of omnivory to FCL are not new but are treated reasonably and have value combined with other factors. However, note that previous research has addressed some of the spatial/heterogeneity effects of the availability of resources with omnivory and FCL (e.g., Holt 1996; McCann 2005; Tunney et al. 2012 as a few examples). Generally, the manuscript could benefit from a little more explaining and understanding of what was generating the observed patterns in FCL since this is possible with a simulation.

In what follows my review focuses on questions and concerns related to concepts and discussion in the current version of the manuscript. I have added a couple editorial comments, but this is not exhaustive.

Expectedly, omnivory is a major determinate of food chain length. I think it is reasonable to set parameters to show a few cases (chain, weak, strong) where the preference for omnivory by the predator in the model changes. But I wonder in interpreting how omnivory influences FCL - how the preference for omnivory (set in parameters) is related to the amount of prey eaten from different trophic levels? In addition to the parameter values, I expect omnivory should also depend on the availability of the prey populations at B,C and P? Can you discuss the degree of omnivory in the simulations (these would determine FCL) ? It is not clear to me that ominvory that is determined by the amount of material moving through the linkages reflects the kind of changes being referred to as structural constraints in the manuscript. Or if these refer to parameters.

At low disturbance it is stated that the level of omnivory increased competition between P and C for B and reduce FCL. It seems this is because one trophic level went extinct. But at high productivity a different mechanism for changes in FCL is suggested. In this case the structural constraint of omnivory determined FCL. I think a little more explanation on what you are defining as competition and occupancy changes versus structural constraints could be helpful.

I was expecting that high productivity you might also see some occupancy changes but maybe for different reasons than low productivity. I wondered if at high productivity you could get increased trophic control where P controls C because it gets biomass from C and B. Also, as P grows it reduces the amount of B consumed by C (which should be increasingly dynamically unstable)? Here I was wondering if the more structural change without occupancy at high productivity might be a parameter limitation based on the levels of r that were used, also the way that you use r as a surrogate of productivity. Often carrying capacity is implicated as a factor in productivity as it sets the level for the amount of material that moves around in the modelled system. I raise these questions because I am used to thinking about the dynamical implications of food webs in terms of the parameters but also the populations as these both determine the strength of interactions (in terms of flux of material in linkages between species). And K can limit the potential for population sizes and has known implications on stability etc..

I think understanding these results associated with branching ecosystems would benefit from a clear description of results and discussion for non-branching networks and then connect the results to the branching complexity ideas. What is happening in the branching network model that governs the results and is providing a different (or similar) response as with the other factors explored here (size, productivity etc.) ?

"The effect of ecosystem complexity was generally clearer than that of ecosystem size" I found that the idea that ecosystem complexity seemed to be more influential than ecosystem size was very subtle. I agree that in a few cases there is a larger initial effect for complexity, but it doesn't seem to be much different overall. Is there a metric that can be provided to show that there is a clearer effect of ecosystem complexity compared to ecosystem size across the different treatments? Is there an explanation that would help to justify the statement? Does this have to do with the dampening of the disturbance moving down stream? As presented it seems like this answer is better aligned with the motivation for doing the study rather than a clear result of the simulations.

One of the results that wasn't discussed is that the changes in FCL in several cases occurred at low levels of ecosystem complexity. Does this mean that you need some complexity, but you don't need a lot of complexity to support longer food chains? Or that above a small amount of complexity it is not that important? Another aspect of this pattern to discuss is whether this is parameter dependent?

"Productivity did not modulate the relationship between FCL and ecosystem size and ecosystem complexity." -- I see difference between high and low productivity simulations for the strong omnivory particularly in the low disturbance case. Perhaps the text ought to be changed to reflect that difference? Can you explain why these differences occur?

Figure 1. a) For clarity, is there supposed to be a change in the darkness of the arrows between B and C for the (chain) and the omnivory cases. Since the caption suggests B-C is fixed? b) I am having a hard time understanding that the "strength of the B-C interaction is fixed". I see that it is the same in the Figure but does this happen in the simulations? I would expect that the strength of that link is influenced by changes in the other two linkages (P-B and P-C) and the densities of the populations (along the same lines as comment 1)?

Line 255: "under high disturbance regimes, only large and complex networks could provide diverse and heterogenous habitats...and maintain long food chains". I am not sure that this is supported as written. In many cases the changes in FCL are at low or moderate levels of size and complexity. Also, it would be helpful to use model simulations to show where FCL these proposed mechanisms are operating i.e., that large and complex .

Line 254: "Consequently food web structure largely controlled FCL." This is a strange way to describe what is happening in the low disturbance systems as FCL is food web structure. Can you provide a more specific mechanism here?

Line 276: "An important result in our simulation is that omnivory does not change the qualitative relationships between FCL and ecosystem properties, as opposed to productivity effects." I find this sentence to

Line 287 and the paragraph: How do the result of this study contribute to the ideas in this paragraph in particular that many small patches is better than large homogenous patches etc.? I find that this discussion is interesting but the text doesn't make a strong connection to the findings of this research.

Ensure that terms are clearly defined. For example, ecosystem structure versus ecosystem feature versus ecosystem property versus food web structure. Ecosystem structure seems to not refer to food web maybe habitat? Anyway as a general comment it would be good to define terms so that the usage is understood.

Line 45: Ecosystem Size?

Line 194: "are differ". Suggest remove "are".

### 2023-03-06 rmd/manuscript.Rmd
edit: manuscript

@@
 For example, omnivory (feeding at multiple trophic levels) is a strong intrinsic driver of FCL [@takimotoEffectsProductivityDisturbance2012; @postProximateStructuralMechanisms2007; @woottonOmnivoryStabilityFreshwater2017].
-When omnivory is absent in three-species communities, the trophic structure assumes a simple linear food chain with a maximum length of three (**Figure 1D**).
-However, FCL decreases proportionally with increasing omnivory as the top predator eats lower in the food web [@wardMechanisticTheoryAquatic2017; @woottonOmnivoryStabilityFreshwater2017].
+When omnivory is absent in three-species communities, the trophic structure or "motif" assumes a simple linear food chain with a maximum trophic position of three (**Figure 1D**).
+However, FCL decreases with increasing omnivory as the top predator eats lower in the food web [@wardMechanisticTheoryAquatic2017; @woottonOmnivoryStabilityFreshwater2017].
 Theoretical work predicts that food webs are destabilized when omnivory is too strong, leading to the loss of species and shorter food chains [@woottonOmnivoryStabilityFreshwater2017].
@@
 
-While recent research efforts of food web ecologists have substantially advanced our understanding of food chain drivers, most of the previous work has been conducted in simple, two-dimensional systems, such as lakes [@postEcosystemSizeDetermines2000; @doiResourceAvailabilityEcosystem2009] and oceanic islands [@takimotoEcosystemSizeNot2008].
+While recent research efforts of food web ecologists have substantially advanced our understanding of food chain drivers, theoretical conceptualization has implicitly assumed two-dimensional systems, which may be a fair approximation of lakes [@postEcosystemSizeDetermines2000; @doiResourceAvailabilityEcosystem2009] and oceanic islands [@takimotoEcosystemSizeNot2008].
 However, many natural habitats have a high degree of spatial complexity that cannot be represented simply by ecosystem size.
@@
 
-There were three consistent patterns in the relationship between FCL and ecosystem geometry (size and complexity).
+There were two consistent patterns in the relationship between FCL and ecosystem geometry (size and complexity).
 First, the effects of ecosystem size and complexity were modulated by disturbance and omnivory.
@@
 
-Ecosystem geometry interacted with disturbance to dictate FCL in branching ecosystems.
+Ecosystem geometry interacted with disturbance to dictate FCL in branching ecosystems (**Figures 2 and 3**).
 With no disturbance, both geometric features (size and complexity) had limited influence on FCL.
@@
 This interactive effect emerged because the sensitivity of FCL to disturbance depended on ecosystem geometry.
-Although increased disturbance frequency reduced overall FCL, smaller and/or simpler ecosystems exhibited the greater vulnerability.
-Therefore, disturbance interact with ecosystem geometry, producing the interdependency of these factors in determining FCL.
-The interactive effects, however, were not observed with low disturbance intensity (proportion of individuals removed per disturbance event); this suggests that disturbance intensity must be strong enough to be influential.
+Although increased disturbance frequency reduced overall FCL, smaller and/or simpler ecosystems exhibited the greater vulnerability (**Figures 2 and 3**).
+Therefore, disturbance interacted with ecosystem geometry to produce the interdependency of these factors in determining FCL.
+However, the maximum levels of FCL was strongly regulated by disturbance frequency, highlighting the important direct effect of disturbance.
+It is important to note that the interactive effects were not observed with low disturbance intensity (proportion of individuals removed per disturbance event); this suggests that disturbance intensity must be strong enough to be influential.
 
 The exception is the "chain" scenario (i.e., no omnivory) with no disturbance.
-In this scenario, ecosystem size had a clear positive effect on FCL while ecosystem complexity had a negative influence.
-These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increases the total carrying capacity whereas ecosystem complexity decreases it.
-However, these patterns characteristic to the "chain" scenario diminished when considering some levels of omnivory because food webs with omnivory were less reliant on the carrying capacity of the basal species.
+In this scenario, ecosystem size had a clear positive effect on FCL (**Figure 2**) while ecosystem complexity had a negative influence (**Figure 3**).
+These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increases the total carrying capacity whereas ecosystem complexity decreases it (**Figure XXX**).
+However, these patterns characteristic to the "chain" scenario diminished when considering some levels of omnivory because FCLs with omnivory were less reliant on the carrying capacity of the basal species.
 
@@
 It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity; therefore, the total carrying capacity within a network dictates FCL.
-The effect of ecosystem size is understandable because it simply increases the total carrying capacity.
-In contrast, ecosystem complexity, here assessed as branching probability, decreases the total carrying capacity as more branched networks have disproportionate numbers of headwater habitats with small local carrying capacities.
-However, the dependence of FCL on local carrying capacity diminished in weak and strong omnivory scenarios, explaining the lack of positive (ecosystem size) or negative FCL-geomtery relationship (ecosystem complexity) in these motifs with no disturbance.
+This effect of ecosystem size is understandable because it simply increases the total carrying capacity.
+In contrast, ecosystem complexity, here assessed as branching probability, decreased the total carrying capacity as more branched networks have disproportionate numbers of headwater habitats with small local carrying capacities.
+Thus, ecosystem complexity had an opposite effect on FCL with the chain motif.
+However, the dependence of FCL on local carrying capacity diminished in weak and strong omnivory scenarios.
+This fact explains the lack of positive (ecosystem size) or negative FCL-geomtery relationship (ecosystem complexity) in these motifs with no disturbance.
 
@@
 
-The effect of productivity was context dependent, supporting previous theoretical predictions.
-While productivity clearly increased FCL in the chain motif, this effect was vague (strong omnivory) or even reversed (weak omnivory).
-High productivity effectively increases the abundance of the basal species, stimulating the prey switching of the intraguild predator from the intraguild prey to the basal species.
+The effect of productivity was context-dependent, supporting previous theoretical predictions [@postProximateStructuralMechanisms2007; @kondohFoodchainLengthAdaptive2009; @wardMechanisticTheoryAquatic2017].
+While productivity clearly increased FCL in the chain motif, this effect was vague (strong omnivory) or even reversed (weak omnivory) in omnivory motifs.
+With omnivory, high productivity effectively increases the abundance of the basal species, stimulating the prey switching of the intraguild predator from the intraguild prey to the basal species.
 As such, increased productivity does not necessarily increase FCL.
@@
 The state transitions with increased ecosystem size are likely associated with increased immigration.
-It is clear that the total number of dispersing propagules increases with increasing ecosystem size, providing opportunities to maintain longer food chains.
+The total number of dispersing propagules increases with increasing ecosystem size, providing opportunities to maintain longer food chains in connected patches.
 The exception was the strong omnivory motif with rare disturbance, in which the insertion mechanism operated.
@@
 Similarly, increased immigration may be generally responsible for the state transitions with increasing ecosystem complexity.
-Habitat patches are closer in more branched networks as it compresses the network configuration.
+Habitat patches are closer in more branched networks as branching compresses the network configuration.
 However, unlike ecosystem size, ecosystem complexity does not increase the total number of dispersing propagules - rather, it enhances connectivity among habitat patches.

### 2023-03-06 rmd/reply1.Rmd
Update reply1.Rmd

@@
 
+> Thank you, please see our response below.
+
 In what follows my review focuses on questions and concerns related to concepts and discussion in the current version of the manuscript.
@@
 
+> We have revisited our model in light of comments from Reviewer 1.
+In the revised version, the parameter $a_{BP}$ changes how quickly the intraguild predator shifts its prey from intraguild prey to basal species as B increases.
+We have clarified this point and explanations for the "structural constraint" (we have avoided to use this vague term throughout the ms).
+
 At low disturbance it is stated that the level of omnivory increased competition between P and C for B and reduce FCL.
@@
 
+> In light of comments from Reviewer 1, we have reanalyzed and fully re-structured Results and Discussion according to the new results.
+The above statement does not hold in the revised MS, so they have been removed.
+
 I was expecting that high productivity you might also see some occupancy changes but maybe for different reasons than low productivity.
@@
 Also, as P grows it reduces the amount of B consumed by C (which should be increasingly dynamically unstable)?
+
 Here I was wondering if the more structural change without occupancy at high productivity might be a parameter limitation based on the levels of r that were used, also the way that you use r as a surrogate of productivity.

### 2023-03-06 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 
-## Productivity and dispersal
-
-Productivity and dispersal had limited influences on the qualitative relationship between FCL and ecosystem geometry.
-However, they had important direct effects on FCL.
-Productivity influences were dependent on omnivory.
-Overall, greater productivity increased (chain & strong omnivory) or decreased FCL (weak omnivory), although the effect size varied with ecosystem geometry and disturbance levels.
-Dispersal distance had a positive effect on FCL, particularly in scenarios with frequent and intense disturbance.
-
-## Underlying state changes
-
 State transition shed lights on how FCL changes occurred.
@@
 
+## Productivity and dispersal
+
+Productivity and dispersal had limited influences on the qualitative relationship between FCL and ecosystem geometry.
+However, they had important direct effects on FCL.
+Productivity influences were dependent on omnivory.
+Overall, greater productivity increased (chain & strong omnivory) or decreased FCL (weak omnivory), although the effect size varied with ecosystem geometry and disturbance levels.
+Dispersal distance had a positive effect on FCL, particularly in scenarios with frequent and intense disturbance.
+
+xxx
+
 # Discussion
@@
 
-The most important finding in this study is the synergistic effects between ecosystem geometry and disturbance.
+The most significant finding in this study is the synergistic effects between ecosystem geometry and disturbance.
 Our no disturbance scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independently of ecosystem size and complexity (except the chain scenario).
@@
 It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity; therefore, the total carrying capacity within a network dictates FCL.
-This effect of ecosystem size is understandable because it simply increases the total carrying capacity.
+This effect of ecosystem size is understandable because it increases the total carrying capacity.
 In contrast, ecosystem complexity, here assessed as branching probability, decreased the total carrying capacity as more branched networks have disproportionate numbers of headwater habitats with small local carrying capacities.

### 2023-03-06 code/figure_si_r_disp.R
edit: figure productivity & dispersal

@@ -1,0 +1,147 @@
+
+# setup -------------------------------------------------------------------
+
+rm(list = ls())
+
+lapply(list("code/library.R",
+            "code/set_figure_theme.R",
+            "code/format_sim_data.R"),
+       source)
+
+df_coef <- readRDS("output/df_coef.rds")
+
+## filter s & mean_disturb_source for visualization
+s_set <- 1
+mu_disturb <- 0.8
+r_set <- c(8, 16)
+
+## df for heatmap
+df_heat <- df_param %>% 
+  left_join(df_coef,
+            by = "param_set") %>% 
+  filter(mean_disturb_source == mu_disturb | p_disturb == 0,
+         (s == 0 & a_bp == 0) | s == s_set)
+
+## df for loess plot
+df_plot <- df_sim %>% 
+  filter(mean_disturb_source == mu_disturb | p_disturb == 0,
+         (s == 0 & a_bp == 0) | s == s_set,
+         r_b %in% r_set) %>% 
+  mutate(disp = fct_relevel(disp, "Short dispersal"))
+
+
+# plot: fcl boxplot -------------------------------------------------------
+
+theme_set(plt_theme)
+
+g_r <- df_plot %>% 
+  ggplot(aes(x = factor(p_disturb),
+             y = fcl,
+             color = factor(r_b),
+             fill = factor(r_b))) +
+  geom_boxplot(alpha = 0.2,
+               linewidth = 0.1,
+               outlier.size = 0.5) +
+  facet_wrap(facets =~ omn,
+             ncol = 1) +
+  MetBrewer::scale_color_met_d("VanGogh3") +
+  MetBrewer::scale_fill_met_d("VanGogh3") +
+  labs(x = "Disturbance probability",
+       y = "Food chain length",
+       color = expression("Productivity"~"("*r[b]*")"),
+       fill = expression("Productivity"~"("*r[b]*")"))
+
+g_disp <- df_plot %>% 
+  ggplot(aes(x = factor(p_disturb),
+             y = fcl,
+             color = factor(disp),
+             fill = factor(disp))) +
+  geom_boxplot(alpha = 0.2,
+               linewidth = 0.1,
+               outlier.size = 0.5) +
+  facet_wrap(facets =~ omn,
+             ncol = 1) +
+  labs(x = "Disturbance probability",
+       y = "",
+       color = "Dispersal",
+       fill = "Dispersal")
+
[…]

### 2023-03-06 code/run_sim.R
edit: simulation

@@ -17,7 +17,7 @@ df_param <- expand.grid(mean_disturb_source = c(0.2, 0.8),
                         sd_disturb_lon = c(0.01, 1),
                         
                         # carring capacity
-                        base_k = 100,
+                        base_k = c(100, 1000),
                         z = 0.54, # Finlay 2011 Ecosphere
                         
                         # timestep
@@ -26,7 +26,7 @@ df_param <- expand.grid(mean_disturb_source = c(0.2, 0.8),
                         n_burnin = 400,
                         
                         # food web parameter
-                        r_b = seq(4, 16, length = 4),
+                        r_b = c(8, 16),
                         e = 1, # to conv_eff
                         a_bc = 0.5, # to attack_rate[1]
                         a_bp = c(0, 0.02, 0.04), # to attack_rate[2]
@@ -43,7 +43,9 @@ df_param <- expand.grid(mean_disturb_source = c(0.2, 0.8),
          (a_bp == 0.04 & a_cp == 0.0025 & h == 0.75), # strong omnivory
          !(s == 1 & a_bp == 0), # remove s = 1 for chain scenario
          !(s == 0 & a_bp != 0), # remove s = 0 for omnivory scenarios
-         !(p_disturb == 0 & mean_disturb_source == 0.2)) %>% 
+         !(p_disturb == 0 & mean_disturb_source == 0.2),
+         !(base_k == 1000 & r_b == 8),
+         !(base_k == 100 & r_b == 16)) %>% 
   arrange(p_disturb,
           theta) %>% 
   mutate(param_set = seq_len(nrow(.))) %>% 

### 2023-03-06 code/run_sim_stvy.R

@@ -1,119 +1,0 @@
-
-# setup -------------------------------------------------------------------
-
-rm(list = ls(all.names = TRUE))
-source(here::here("code/library.R"))
-
-cl <- makeCluster(detectCores())
-registerDoSNOW(cl)
-
-# set parameters ----------------------------------------------------------
-
-# igpsim parameters
-n_para <- 500
-df_param <- tibble(# parameters for brnet
-                   mean_disturb_source = 0.8,
-                   sd_disturb_source = runif(n_para, 0.05, 5),
-                   sd_disturb_lon = runif(n_para, 0.05, 5),
-                   
-                   # carring capacity
-                   base_k = 100,
-                   z = 0.54, # Finlay 2011 Ecosphere
-                   
-                   # parameters for igpsim
-                   n_timestep = 1000,
-                   n_warmup = 200,
-                   n_burnin = 400,
-                   r_b = runif(n_para, 1, 10),
-                   e_bc = runif(n_para, 1, 10), # to conv_eff[1]
-                   e_bp = runif(n_para, 1, 10), # to conv_eff[2]
-                   e_cp = runif(n_para, 1, 10), # to conv_eff[3]
-                   a_bc = runif(n_para, 0.05, 0.5), # to attack_rate[1]
-                   a_bp = runif(n_para, 0.05, 0.5), # to attack_rate[2]
-                   a_cp = runif(n_para, 0.05, 0.5), # to attack_rate[3]
-                   h_bc = runif(n_para, 0.5, 5), # to handling_time[1]
-                   h_bp = runif(n_para, 0.5, 5), # to handling_time[2]
-                   h_cp = runif(n_para, 0.5, 5), # to handling_time[3]
-                   s0 = runif(n_para, 0.5, 1),
-                   p_disturb = runif(n_para, 0, 0.2),
-                   p_dispersal = runif(n_para, 0, 0.1),
-                   theta = runif(n_para, 0.01, 1)) %>% 
-  mutate(param_set = seq_len(nrow(.)))
-
-# geometry parameters
-n_rep <- 100
-repeat {
-  n_patch <- round(runif(n_rep, 10, 150))
-  p_branch <- runif(n_rep, 0.01, 0.99)
-  if(min(n_patch) < 15 & 
-     max(n_patch) > 145 &
-     min(p_branch) < 0.05 &
-     max(p_branch) > 0.95) break    
-}
-
[…]

### 2023-03-07 code/figure_main_sim.R
edit: figures

@@ -29,59 +29,10 @@ df_plot <- df_sim %>%
          r_b %in% r_set)
 
 
-# plot: heatmap -----------------------------------------------------------
-
-theme_set(plt_theme)
-df_point <- expand.grid(r_b = r_set,
-                        p_disturb = unique(df_sim$p_disturb),
-                        rho1 = 0,
-                        rho2 = 0)
-
-## Ecosystem size
-g_size <- df_heat %>% 
-  ggplot(aes(x = r_b,
-             y = p_disturb,
-             fill = rho1)) +
-  geom_raster() +
-  geom_point(data = df_point) +
-  facet_grid(cols = vars(disp),
-             rows = vars(omn)) +
-  labs(y = expression("Disturbance prob. ("*p[m]*")"),
-       x = expression("Productivity ("*r[b]*")"),
-       fill = "Slope") +
-  scale_x_continuous(breaks = sort(unique(df_sim$r_b))) +
-  scale_fill_gradient2(mid = 0,
-                       limits = range(c(df_heat$rho1, df_heat$rho2))) +
-  theme_classic() +
-  theme(strip.background = element_blank()) +
-  ggtitle("Ecosystem size")
-
-## Ecosystem complexity
-g_branch <- df_heat %>% 
-  ggplot(aes(x = r_b,
-             y = p_disturb,
-             fill = rho2)) +
-  geom_raster() +
-  geom_point(data = df_point) +
-  facet_grid(cols = vars(disp),
-             rows = vars(omn)) +
-  labs(y = expression("Disturbance prob. ("*p[m]*")"),
-       x = expression("Productivity ("*r[b]*")"),
-       fill = "Slope") +
-  scale_x_continuous(breaks = sort(unique(df_sim$r_b))) +
-  scale_fill_gradient2(mid = 0,
-                       limits = range(c(df_heat$rho1, df_heat$rho2))) +
-  theme_classic() +
-  theme(strip.background = element_blank()) +
-  ggtitle("Ecosystem complexity")
-
-g_m <- (g_size + g_branch) + plot_annotation(tag_levels = "A")
-
-
 # plot: geometry effect ---------------------------------------------------
 
-lab <- c(`8` = "Low~productivity~(r[b]==8)",
-         `16` = "High~productivity~(r[b]==16)")
+lab <- c(`8` = "Low~productivity",
+         `16` = "High~productivity")
 
 ## ecosystem size effect
 g_np <-  df_plot %>% 
@@ -93,7 +44,7 @@ g_np <-  df_plot %>%
[…]

### 2023-03-07 code/figure_si_r_disp.R

@@ -13,7 +13,6 @@ df_coef <- readRDS("output/df_coef.rds")
 ## filter s & mean_disturb_source for visualization
 s_set <- 1
 mu_disturb <- 0.8
-r_set <- c(8, 16)
 
 ## df for heatmap
 df_heat <- df_param %>% 
@@ -25,9 +24,8 @@ df_heat <- df_param %>%
 ## df for loess plot
 df_plot <- df_sim %>% 
   filter(mean_disturb_source == mu_disturb | p_disturb == 0,
-         (s == 0 & a_bp == 0) | s == s_set,
-         r_b %in% r_set) %>% 
-  mutate(disp = fct_relevel(disp, "Short dispersal"))
+         (s == 0 & a_bp == 0) | s == s_set) %>% 
+  mutate(disp = ifelse(disp == "Short dispersal", "Short", "Long"))
 
 
 # plot: fcl boxplot -------------------------------------------------------
@@ -37,8 +35,8 @@ theme_set(plt_theme)
 g_r <- df_plot %>% 
   ggplot(aes(x = factor(p_disturb),
              y = fcl,
-             color = factor(r_b),
-             fill = factor(r_b))) +
+             color = factor(prod),
+             fill = factor(prod))) +
   geom_boxplot(alpha = 0.2,
                linewidth = 0.1,
                outlier.size = 0.5) +
@@ -48,8 +46,8 @@ g_r <- df_plot %>%
   MetBrewer::scale_fill_met_d("VanGogh3") +
   labs(x = "Disturbance probability",
        y = "Food chain length",
-       color = expression("Productivity"~"("*r[b]*")"),
-       fill = expression("Productivity"~"("*r[b]*")"))
+       color = "Productivity",
+       fill = "Productivity")
 
 g_disp <- df_plot %>% 
   ggplot(aes(x = factor(p_disturb),
@@ -94,8 +92,8 @@ df_plot_st <- df_plot %>%
 g_r_state <- df_plot_st %>% 
   ggplot(aes(x = factor(state),
              y = value,
-             color = factor(r_b),
-             fill = factor(r_b))) +
+             color = factor(prod),
+             fill = factor(prod))) +
   geom_boxplot(alpha = 0.2,
                linewidth = 0.1,
                outlier.size = 0.5) +
@@ -107,8 +105,8 @@ g_r_state <- df_plot_st %>%
                                                          label_parsed))) +
   labs(x = "State",
        y = "Occupancy",
[…]

### 2023-03-07 code/format_sim_data.R

@@ -11,11 +11,15 @@ source(here::here("code/set_figure_theme.R"))
 df_sim <- readRDS(file = here::here("output/sim_main.rds")) %>% 
   #filter(!(a_bp == 0 & s != 0)) %>% 
   mutate(omn = case_when(a_bp == 0 ~ "Chain",
-                         a_bp == 0.02 ~ "Weak",
-                         a_bp == 0.04 ~ "Strong"),
-         omn = fct_relevel(omn, c("Chain", "Weak", "Strong")),
+                         a_bp == 0.02 ~ "Weak omnivory",
+                         a_bp == 0.04 ~ "Strong omnivory"),
+         omn = fct_relevel(omn, c("Chain", "Weak omnivory", "Strong omnivory")),
          disp = case_when(theta == 1 ~ "Short dispersal",
-                          theta == 0.1 ~ "Long dispersal"))
+                          theta == 0.1 ~ "Long dispersal"),
+         prod = case_when(r_b == 8 ~ "Low",
+                          r_b == 16 ~ "High"),
+         disp = fct_relevel(disp, "Short dispersal"),
+         prod = fct_relevel(prod, "Low"))
 
 df_param <- df_sim %>% 
   select(-c(fcl,

### 2023-03-07 code/figure_si_geom.R
edit: figures

@@ -25,63 +25,13 @@ df_heat <- df_param %>%
 ## df for loess plot
 df_plot <- df_sim %>% 
   filter(mean_disturb_source == mu_disturb | p_disturb == 0,
-         (s == 0 & a_bp == 0) | s == s_set,
-         r_b %in% r_set)
-
-
-# plot: heatmap -----------------------------------------------------------
-
-theme_set(plt_theme)
-df_point <- expand.grid(r_b = r_set,
-                        p_disturb = unique(df_sim$p_disturb),
-                        rho1 = 0,
-                        rho2 = 0)
-
-## Ecosystem size
-g_size <- df_heat %>% 
-  ggplot(aes(x = r_b,
-             y = p_disturb,
-             fill = rho1)) +
-  geom_raster() +
-  geom_point(data = df_point) +
-  facet_grid(cols = vars(disp),
-             rows = vars(omn)) +
-  labs(y = expression("Disturbance prob. ("*p[m]*")"),
-       x = expression("Productivity ("*r[b]*")"),
-       fill = "Slope") +
-  scale_x_continuous(breaks = sort(unique(df_sim$r_b))) +
-  scale_fill_gradient2(mid = 0,
-                       limits = range(c(df_heat$rho1, df_heat$rho2))) +
-  theme_classic() +
-  theme(strip.background = element_blank()) +
-  ggtitle("Ecosystem size")
-
-## Ecosystem complexity
-g_branch <- df_heat %>% 
-  ggplot(aes(x = r_b,
-             y = p_disturb,
-             fill = rho2)) +
-  geom_raster() +
-  geom_point(data = df_point) +
-  facet_grid(cols = vars(disp),
-             rows = vars(omn)) +
-  labs(y = expression("Disturbance prob. ("*p[m]*")"),
-       x = expression("Productivity ("*r[b]*")"),
-       fill = "Slope") +
-  scale_x_continuous(breaks = sort(unique(df_sim$r_b))) +
-  scale_fill_gradient2(mid = 0,
-                       limits = range(c(df_heat$rho1, df_heat$rho2))) +
-  theme_classic() +
-  theme(strip.background = element_blank()) +
-  ggtitle("Ecosystem complexity")
-
-g_m <- (g_size + g_branch) + plot_annotation(tag_levels = "A")
+         (s == 0 & a_bp == 0) | s == s_set)
 
 
 # plot: geometry effect ---------------------------------------------------
 
[…]

### 2023-03-07 code/format_sim_data.R

@@ -31,20 +31,3 @@ df_param <- df_sim %>%
             n_patch,
             n_rep)) %>% 
   distinct()
-
-# df_coef <- df_sim %>% 
-#   group_by(param_set) %>% 
-#   do(param_set = unique(.$param_set),
-#      rho1 = ppcor::pcor(tibble(.$fcl, .$n_patch, .$p_branch),
-#                         method = "spearman")$estimate[1, 2],
-#      rho2 = ppcor::pcor(tibble(.$fcl, .$n_patch, .$p_branch),
-#                         method = "spearman")$estimate[1, 3]) %>% 
-#   summarize(param_set = param_set,
-#             rho1 = unlist(rho1),
-#             rho2 = unlist(rho2))
-# 
-# 
-# # export ------------------------------------------------------------------
-# 
-# saveRDS(df_coef,
-#         "output/df_coef.rds")

### 2023-03-07 code/figure_main_geom.R
edit: figures

@@ -15,13 +15,6 @@ s_set <- 1
 mu_disturb <- 0.8
 r_set <- c(8, 16)
 
-## df for heatmap
-df_heat <- df_param %>% 
-  left_join(df_coef,
-            by = "param_set") %>% 
-  filter(mean_disturb_source == mu_disturb,
-         (s == 0 & a_bp == 0) | s == s_set)
-
 ## df for loess plot
 df_plot <- df_sim %>% 
   filter(mean_disturb_source == mu_disturb,

### 2023-03-07 code/figure_si_geom.R

@@ -15,13 +15,6 @@ s_set <- 1
 mu_disturb <- 0.2
 r_set <- c(8, 16)
 
-## df for heatmap
-df_heat <- df_param %>% 
-  left_join(df_coef,
-            by = "param_set") %>% 
-  filter(mean_disturb_source == mu_disturb | p_disturb == 0,
-         (s == 0 & a_bp == 0) | s == s_set)
-
 ## df for loess plot
 df_plot <- df_sim %>% 
   filter(mean_disturb_source == mu_disturb | p_disturb == 0,

### 2023-03-07 code/figure_main_geom.R
edit: figure

@@ -13,13 +13,11 @@ df_coef <- readRDS("output/df_coef.rds")
 ## filter s & mean_disturb_source for visualization
 s_set <- 1
 mu_disturb <- 0.8
-r_set <- c(8, 16)
 
 ## df for loess plot
 df_plot <- df_sim %>% 
   filter(mean_disturb_source == mu_disturb,
-         (s == 0 & a_bp == 0) | s == s_set,
-         r_b %in% r_set)
+         (s == 0 & a_bp == 0) | s == s_set)
 
 
 # plot: geometry effect ---------------------------------------------------

### 2023-03-07 code/figure_si_r_disp.R
edit: figures

@@ -14,13 +14,6 @@ df_coef <- readRDS("output/df_coef.rds")
 s_set <- 1
 mu_disturb <- 0.8
 
-## df for heatmap
-df_heat <- df_param %>% 
-  left_join(df_coef,
-            by = "param_set") %>% 
-  filter(mean_disturb_source == mu_disturb | p_disturb == 0,
-         (s == 0 & a_bp == 0) | s == s_set)
-
 ## df for loess plot
 df_plot <- df_sim %>% 
   filter(mean_disturb_source == mu_disturb | p_disturb == 0,

### 2023-03-07 code/figure_si_geom.R
edit: figures

@@ -13,7 +13,6 @@ df_coef <- readRDS("output/df_coef.rds")
 ## filter s & mean_disturb_source for visualization
 s_set <- 1
 mu_disturb <- 0.2
-r_set <- c(8, 16)
 
 ## df for loess plot
 df_plot <- df_sim %>% 

### 2023-03-07 code/figure_si_state.R
edit: figures

@@ -11,13 +11,11 @@ lapply(list("code/library.R",
 ## filter s & mean_disturb_source for visualization
 s_set <- 1
 mu_disturb <- 0.8
-r_set <- c(8, 16)
 
 ## df for gam plot
 df_plot <- df_sim %>% 
   filter(mean_disturb_source == mu_disturb,
-         (s == 0 & a_bp == 0) | s == s_set,
-         r_b %in% r_set) %>% 
+         (s == 0 & a_bp == 0) | s == s_set) %>% 
   pivot_longer(cols = paste0("s", 0:4),
                values_to = "p_state",
                names_to = "state")

### 2023-03-07 code/figure_si_one_patch.R
edit: figures

@@ -40,8 +40,9 @@ foreach(i = 1:length(a1_set)) %do% {
       geom_raster(alpha = 0.8) +
       facet_grid(rows = vars(r_b),
                  cols = vars(h), labeller = labeller(h = as_labeller(lab[[1]], label_parsed),
-                                                        r_b = as_labeller(lab[[2]], label_parsed))) +
-      scale_fill_viridis_d(limits = factor(0:4), labels = state_lab) +
+                                                     r_b = as_labeller(lab[[2]], label_parsed))) +
+      scale_fill_viridis_d(limits = factor(0:4),
+                           labels = state_lab) +
       scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
       labs(x = expression("Attack rate ("*a[BP]*")"),
            y = expression("Attack rate ("*a[CP]*")"),

### 2023-03-07 code/figure_si_r_disp.R

@@ -18,7 +18,8 @@ mu_disturb <- 0.8
 df_plot <- df_sim %>% 
   filter(mean_disturb_source == mu_disturb | p_disturb == 0,
          (s == 0 & a_bp == 0) | s == s_set) %>% 
-  mutate(disp = ifelse(disp == "Short dispersal", "Short", "Long"))
+  mutate(disp = ifelse(disp == "Short dispersal", "Short", "Long"),
+         disp = fct_relevel(disp, "Short"))
 
 
 # plot: fcl boxplot -------------------------------------------------------

### 2023-03-07 rmd/manuscript.Rmd
update: mansucript

@@
 This setup ($\sigma_h > \sigma_l$) produces high spatial variation among tributaries while allowing strong within-tributary downstream correlation, a typical pattern in real river networks [@nakamuraDisturbanceRegimesStream2000; @bendaNetworkDynamicsHypothesis2004].
-We used the basal species' growth rate $r_b$ (`r unique(df_sim$r_b)`) as a proxy for resource availability as it resembles the gross primary production.
+We used a combination of basal species' growth rate $r_b$ and base carrying capacity $K_{base}$ (low productivity: $r_b =$ `r min(df_sim$r_b)`, $K_{base} =$ `r min(df_sim$base_k)` ; high productivity = $r_b =$ `r max(df_sim$r_b)`, $K_{base} =$ `r max(df_sim$base_k)` ) as a proxy for resource availability.
 For dispersal, we examined short- and long-distance dispersal scenarios (short-distance dispersal $\theta=$ `r max(df_sim$theta)`; long-distance dispersal $\theta=$ `r min(df_sim$theta)`) with constant dispersal probability ($p_{d}=$ `r unique(df_sim$p_dispersal)`).
@@
 There were two consistent patterns in the relationship between FCL and ecosystem geometry (size and complexity).
-First, the effects of ecosystem size and complexity were modulated by disturbance and omnivory.
-Second, resource availability and dispersal had limited influences on the qualitative relationship between FCL and ecosystem geometry, although they had important direct effects on the FCL.
+First, the effects of ecosystem size and complexity were modulated by disturbance.
+Second, productivity strengthened the relationship between FCL and ecosystem geometry with important direct effects that were motif dependent.
+Third, dispersal had a minimal but non-negligible positive effect on FCL.
 Below we describe these results in detail.
 
-## Interactive effects of ecosystem geometry, disturbance, and omnivory
+## Interactive effect of ecosystem geometry and disturbance
 
@@
 
-The exception is the "chain" scenario (i.e., no omnivory) with no disturbance.
-In this scenario, ecosystem size had a clear positive effect on FCL (**Figure 2**) while ecosystem complexity had a negative influence (**Figure 3**).
-These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increases the total carrying capacity whereas ecosystem complexity decreases it (**Figure XXX**).
-However, these patterns characteristic to the "chain" scenario diminished when considering some levels of omnivory because FCLs with omnivory were less reliant on the carrying capacity of the basal species.
+The exception is the "chain" scenario (i.e., no omnivory) with no disturbance and low productivity.
+In this scenario, ecosystem size had a positive effect on FCL (**Figure 2**) while ecosystem complexity had a negative influence (**Figure 3**).
+These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increased the total carrying capacity whereas ecosystem complexity decreased it (**Figure XXX**).
+However, these patterns characteristic to the "chain" scenario diminished when considering some levels of omnivory.
 
-State transition shed lights on how FCL changes occurred.
-In the chain motif, the "additive" mechanism (i.e., the addition of higher trophic levels) is always responsible for the changes in FCL.
-With disturbance, state transition occurred from "B" to "B + C" as ecosystem size or complexity increases.
-With no disturbance, state transition was different between size and complexity.
-Ecosystem size facilitated the transition from "B + C" to "B + C + P", while ecosystem complexity drove the opposite transition ("B + C + P" $\rightarrow$ "B + C").
+State transitions shed lights on how FCL changes occurred.
+In the chain motif, the "additive" mechanism (i.e., the addition of higher trophic levels) was always responsible for the changes in FCL.
+With disturbance, state transition occurred from "B" to "B + C" (low productivity) or "B + C" to "B + C + P" (high productivity) as ecosystem size or complexity increases.
+Exceptionally, the state transition of "B + C + P" $\rightarrow$ "B + C" occurred as ecosystem complexity increased with no disturbance and low productivity, a scenario in which ecosystem complexity decreased FCL.
 
 In the weak and strong omnivory motifs, ecosystem size and complexity increased FCL through different mechanisms.
-Ecosystem size increased FCL primarily through the additive mechanism, facilitating the state transition from "B" to "B + P", particularly with frequent disturbance.
-However, the "insertion" mechanism (the transition from "B + P" to "B + C + P") also took place in the strong omnivory motif with rare disturbance ($p_m = 0.05$).
-In contrast, the insertion mechanism was responsible for the effect of ecosystem complexity on FCL.
-In the weak omnivory scenario, the transition from "B + P" to "B + C + P" was predominant.
-In the strong omnivory scenario, transition patterns depended on disturbance frequency.
-When disturbance was rare ($p_m = 0.05$), the insertion mechanism ("B + P" $\rightarrow$ "B + C + P") drove the FCL change.
-However, as disturbance frequency increased, the primary process switched to the additive mechanism ("B" $\rightarrow$ "B + P").
+Ecosystem size increased FCL primarily through the additive mechanism, facilitating the state transition of "B" $\rightarrow$ "B + P" (low productivity) or "No species" $\rightarrow$ "B + P" (high productivity), particularly with frequent disturbance (Figure XXX).
+In contrast, the insertion and additive mechanisms were responsible for the effect of ecosystem complexity on FCL, depending on simulation scenarios.
+With weak omnivory, the transition from "B + P" to "B + C + P" (insertion) was predominant when productivity was low (Figure XXX).
+However, the additive mechanism ("No species" $\rightarrow$ "B + P") seems to be dominant in the high productivity scenario (Figure XXX).
+In the strong omnivory motif, transition patterns depended on disturbance frequency.
+When disturbance was rare ($p_m = 0.05$), the insertion mechanism ("B + P" $\rightarrow$ "B + C + P") drove the FCL change (Figure XXX).
+As disturbance frequency increased, the primary process switched to the additive mechanism ("B" $\rightarrow$ "B + P" or "No species" $\rightarrow$ "B + P"; Figure XXX).
 
-## Productivity and dispersal
+## Productivity effect
 
-Productivity and dispersal had limited influences on the qualitative relationship between FCL and ecosystem geometry.
-However, they had important direct effects on FCL.
-Productivity influences were dependent on omnivory.
-Overall, greater productivity increased (chain & strong omnivory) or decreased FCL (weak omnivory), although the effect size varied with ecosystem geometry and disturbance levels.
-Dispersal distance had a positive effect on FCL, particularly in scenarios with frequent and intense disturbance.
+High productivity strengthened the relationship between FCL and ecosystem geometry.
+In the chain motif, productivity greatly increased the overall FCL (Figure XXX), thereby shifting the upper bound of FCL (**Figures 2 and 3**).
+In contrast, increased productivity greatly decreased the minimum levels of FCL with weak and strong omnivory with little influence on the maximum FCL (**Figures 2, 3, and XXX**).
+This negative effect of increased productivity was apparent in smaller and/or simpler ecosystems, steepening the slopes between FCL and ecosystem geometry (**Figures 2 and 3**).
+Such patterns were particularly clear with long-distance dispersal (**Figures 2 and 3**).
 
-xxx
+State transitions associated with increased productivity depended on the motif structure.
+In the chain motif, the transition from "B + C" to "B + C + P" (i.e., additive) was dominant regrardless of disturbance frequency.
+The introduction of omnivory allowed the transition between "B + P" and "B + C + P", the primary direction of the transition differed between weak and strong omnivory.
+With weak omnivory, the transition of "B + C + P" $\rightarrow$ "B + P" was prevalent and shortened FCL.
+However, this transition was marginal in the strong omnivory motif.
+Instead, productivity significantly increased the proportion of "No species" state (Figure XXX), especially in small and/or simple ecosystems (Figure XXX).
+
+## Dispersal effect
+
+Dispersal distance had a positive influence on FCL, regardless of simulation scenarios. However, this direct effect was rather weak in comparison to other factors.
+
+State transitions occurred through either additive (chain) or insertion mechanisms (omnivory).
+In the chain motif, the transitions were either "B" $\rightarrow$ "B + C" or "B + C" $\rightarrow$ "B + C+ P", depending on disturbance frequency.
+In the omnivory motifs, the predominant transition was from "B + P" to "B + C + P."
 
@@
 The most significant finding in this study is the synergistic effects between ecosystem geometry and disturbance.
-Our no disturbance scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independently of ecosystem size and complexity (except the chain scenario).
-However, as the frequency of intense disturbance increases, small and/or simple networks could not maintain long food chains because habitats were too homogeneous with limited disturbance refugia.
+Our no disturbance scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independently of ecosystem size and complexity (except the chain motif with low productivity).
+However, as the frequency of intense disturbance increases, small and/or simple networks became unable to maintain long food chains.
+Here, increased ecosystem size and complexity may give greater resistance to disturbance through different mechanisms.
+Ecosystem size simply increases the "total amount" of disturbance refugia, from which dispersing propagules recolonize impacted habitat patches.
+In the meantime, ecosystem complexity may increases the connectivity to disturbance refugia.
+Indeed, branching shortens the mean distance between patches and therefore may enhance immigration from less undisturbed patches.
+It is also conceivable that the proportion of disturbance refugia increased in more branched networks.
+However, we did not observe a clear proportional increase in disturbance refugia with increasing branching probability.
+Therefore, each of geometric features support long food chains through distinct mechanisms.
 While empirical studies rarely looked at the influence of ecosystem complexity on FCL, several studies provide some support for this theoretical prediction.
@@
 
-The synergy between ecosystem geometry and disturbance has also an implication for why the dynamical stability hypothesis [@pimmNumberTrophicLevels1977] has been least supported by the literature.
-While increased disturbance frequency consistently reduced FCL, its magnitude depended strongly on ecosystem geometry.
-This observation should provide a possible explanation for the limited support for the dynamical stability hypothesis, because such an interaction has been rarely assessed in previous studies.
+The synergy between ecosystem geometry and disturbance has also an important implication for why the dynamical stability hypothesis [@pimmNumberTrophicLevels1977] has been least supported by the literature.
+Although increased disturbance frequency consistently reduced FCL, its magnitude depended strongly on ecosystem geometry.
+This observation should provide a possible explanation for the limited support for the dynamical stability hypothesis in field studies.
 Although FCL research often assumes that the effects of environmental controls are independent of one another, exploring potential interactions among them could provide useful insights into what controls FCL in nature.
@@
 Another interesting result of our simulation is the interaction with omnivory.
-In particular, when the chain motif was combined with no-disturbance, ecosystem size and complexity had contrasting influences on FCL.
-It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity; therefore, the total carrying capacity within a network dictates FCL.
-This effect of ecosystem size is understandable because it increases the total carrying capacity.
-In contrast, ecosystem complexity, here assessed as branching probability, decreased the total carrying capacity as more branched networks have disproportionate numbers of headwater habitats with small local carrying capacities.
-Thus, ecosystem complexity had an opposite effect on FCL with the chain motif.
-However, the dependence of FCL on local carrying capacity diminished in weak and strong omnivory scenarios.
-This fact explains the lack of positive (ecosystem size) or negative FCL-geomtery relationship (ecosystem complexity) in these motifs with no disturbance.
-
+In particular, with no disturbance and low productivity, FCL of the chain motif had contrasting relationships with ecosystem size (positive) and complexity (negative).
+It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity; therefore, FCL followed the changes in the total carrying capacity along the gradients of ecosystem size and complexity.
+However, these relationships diminished in omnivory motifs as the dependence of FCL on local carrying capacity decreased (**Figure XXX**).
 Importantly, however, disturbance overwhelmed the effect of omnivory on the FCL-geometry relationships: as the frequency of intense disturbance increases, the positive relationships between FCL and ecosystem geometry emerged regardless of omnivory levels.
@@
 While productivity clearly increased FCL in the chain motif, this effect was vague (strong omnivory) or even reversed (weak omnivory) in omnivory motifs.
-With omnivory, high productivity effectively increases the abundance of the basal species, stimulating the prey switching of the intraguild predator from the intraguild prey to the basal species.
-As such, increased productivity does not necessarily increase FCL.
-Our results suggest that this prediction from non-spatial models can be applied to spatially complex ecosystems.
+Intraguild predation plays an important role in explaining this context dependent effect.
+In the high productivity scenario with omnivory, the increased abundance of the basal species enhances the intraguild predator.
+This may subsequently facilitate the local exclusion of the intraguild prey through increased predation pressure, if the attack rate is sufficiently high [@postProximateStructuralMechanisms2007; @kondohFoodchainLengthAdaptive2009; @wardMechanisticTheoryAquatic2017].
+Indeed, in the weak omnivory motif (the high value of the attack rate of P on C), we observed the state transition from "B + C + P" to "B + P" that shortened FCL.
+However, in the strong omnivory motif, the transition was
 

### 2023-03-07 rmd/manuscript.Rmd
update: manuscript

@@
 State transitions shed lights on how FCL changes occurred.
-In the chain motif, the "additive" mechanism (i.e., the addition of higher trophic levels) was always responsible for the changes in FCL.
+In the chain motif, the "additive" mechanism (i.e., the addition of higher trophic species) was always responsible for the changes in FCL.
 With disturbance, state transition occurred from "B" to "B + C" (low productivity) or "B + C" to "B + C + P" (high productivity) as ecosystem size or complexity increases.
@@
 Ecosystem size increased FCL primarily through the additive mechanism, facilitating the state transition of "B" $\rightarrow$ "B + P" (low productivity) or "No species" $\rightarrow$ "B + P" (high productivity), particularly with frequent disturbance (Figure XXX).
-In contrast, the insertion and additive mechanisms were responsible for the effect of ecosystem complexity on FCL, depending on simulation scenarios.
+In contrast, the insertion (i.e., the insertion of intermediate trophic species) and additive mechanisms were responsible for the effect of ecosystem complexity on FCL, depending on simulation scenarios.
 With weak omnivory, the transition from "B + P" to "B + C + P" (insertion) was predominant when productivity was low (Figure XXX).
@@
 State transitions associated with increased productivity depended on the motif structure.
-In the chain motif, the transition from "B + C" to "B + C + P" (i.e., additive) was dominant regrardless of disturbance frequency.
+In the chain motif, the transition from "B + C" to "B + C + P" (i.e., additive) was dominant regardless of disturbance frequency.
 The introduction of omnivory allowed the transition between "B + P" and "B + C + P", the primary direction of the transition differed between weak and strong omnivory.
-With weak omnivory, the transition of "B + C + P" $\rightarrow$ "B + P" was prevalent and shortened FCL.
+With weak omnivory, the transition of "B + C + P" $\rightarrow$ "B + P" was prevalent and shortened FCL regardless of disturbance frequency (Figure XXX).
 However, this transition was marginal in the strong omnivory motif.
-Instead, productivity significantly increased the proportion of "No species" state (Figure XXX), especially in small and/or simple ecosystems (Figure XXX).
+Instead, productivity significantly increased the proportion of "No species" state (Figure XXX) in disturbed scenarios, especially in small and/or simple ecosystems (Figure XXX).
 
@@
 Here, we show that FCL responds positively to increasing ecosystem size and complexity, assessed as the number of patches and branching probability in river networks.
-This supports our hypothesis that ecosystem complexity may represent an overlooked yet important extrinsic control of FCL. However, the relationship between FCL and ecosystem complexity varied with other extrinsic and intrinsic factors, supporting the emerging paradigm that FCL is regulated by multiple factors [@mchughDualInfluencesEcosystem2010; @wardMechanisticTheoryAquatic2017].
+This supports our hypothesis that ecosystem complexity may represent an overlooked yet important extrinsic control of FCL.
+However, the relationship between FCL and ecosystem complexity varied with other extrinsic and intrinsic factors, supporting the emerging paradigm that FCL is regulated by multiple factors [@mchughDualInfluencesEcosystem2010; @wardMechanisticTheoryAquatic2017].
 Although we focus on river networks here, we expect our results to be transferable to other systems with non-branching structures.
@@
 Ecosystem size simply increases the "total amount" of disturbance refugia, from which dispersing propagules recolonize impacted habitat patches.
-In the meantime, ecosystem complexity may increases the connectivity to disturbance refugia.
-Indeed, branching shortens the mean distance between patches and therefore may enhance immigration from less undisturbed patches.
-It is also conceivable that the proportion of disturbance refugia increased in more branched networks.
+In the meantime, ecosystem complexity may increase the connectivity to disturbance refugia.
+Indeed, branching shortens the mean distance between patches and therefore may enhance immigration from less disturbed patches.
+It is possible that the proportion of disturbance refugia increased in more branched networks.
 However, we did not observe a clear proportional increase in disturbance refugia with increasing branching probability.
-Therefore, each of geometric features support long food chains through distinct mechanisms.
+Collectively, each of geometric features likely support long food chains through distinct mechanisms, which may also explain the different patterns of state transitions (additive vs.
+insertion mechanisms) [@postProximateStructuralMechanisms2007].
+
 While empirical studies rarely looked at the influence of ecosystem complexity on FCL, several studies provide some support for this theoretical prediction.
@@
 
-The synergy between ecosystem geometry and disturbance has also an important implication for why the dynamical stability hypothesis [@pimmNumberTrophicLevels1977] has been least supported by the literature.
+The synergy between ecosystem geometry and disturbance has also an important implication for why the dynamical stability hypothesis has been least supported by the literature [@takimotoEnvironmentalDeterminantsFoodchain2013].
 Although increased disturbance frequency consistently reduced FCL, its magnitude depended strongly on ecosystem geometry.
 This observation should provide a possible explanation for the limited support for the dynamical stability hypothesis in field studies.
-Although FCL research often assumes that the effects of environmental controls are independent of one another, exploring potential interactions among them could provide useful insights into what controls FCL in nature.
+Although FCL research often assumes that the effects of environmental controls are independent, exploring potential interactions among them could provide useful insights into environmental controls of FCL in nature.
 
@@
 
-The effect of productivity was context-dependent, supporting previous theoretical predictions [@postProximateStructuralMechanisms2007; @kondohFoodchainLengthAdaptive2009; @wardMechanisticTheoryAquatic2017].
-While productivity clearly increased FCL in the chain motif, this effect was vague (strong omnivory) or even reversed (weak omnivory) in omnivory motifs.
-Intraguild predation plays an important role in explaining this context dependent effect.
+The effect of productivity was motif-dependent, supporting previous theoretical predictions [@postProximateStructuralMechanisms2007; @kondohFoodchainLengthAdaptive2009; @wardMechanisticTheoryAquatic2017].
+While productivity clearly increased FCL in the chain motif through an additive mechanism, the productivity effect was vague (strong omnivory) or even reversed (weak omnivory) in omnivory motifs.
+Intraguild predation plays an important role in explaining this context dependency.
 In the high productivity scenario with omnivory, the increased abundance of the basal species enhances the intraguild predator.
 This may subsequently facilitate the local exclusion of the intraguild prey through increased predation pressure, if the attack rate is sufficiently high [@postProximateStructuralMechanisms2007; @kondohFoodchainLengthAdaptive2009; @wardMechanisticTheoryAquatic2017].
-Indeed, in the weak omnivory motif (the high value of the attack rate of P on C), we observed the state transition from "B + C + P" to "B + P" that shortened FCL.
-However, in the strong omnivory motif, the transition was
+Indeed, in the weak omnivory motif (the high value of the attack rate of P on C), we observed the state transition from "B + C + P" to "B + P" with shortened FCL regardless of disturbance frequency.
+However, this was not the case in the strong omnivory motif; the increased proportion of "No species" state had a strong impact on the minimum FCL, and this transition was mediated by disturbance.
+This result suggests that, although the strong omnivory motif is innately stable with high productivity (see one-patch simulations; Figure XXX), external disturbance undermines the sensitive balance of predator-prey interactions with the abrupt increase of "No species" state, which must be countered by immigration from less disturbed habitats to maintain food chains.
 
-Ecosystem size and complexity led to the distinct patterns of state transitions.
-The state transitions with increased ecosystem size are likely associated with increased immigration.
-The total number of dispersing propagules increases with increasing ecosystem size, providing opportunities to maintain longer food chains in connected patches.
-The exception was the strong omnivory motif with rare disturbance, in which the insertion mechanism operated.
-This is likely because rare disturbance inhibited the exclusion of the intraguild prey by suppressing the abundance of the intraguild predator.
-Similarly, increased immigration may be generally responsible for the state transitions with increasing ecosystem complexity.
-Habitat patches are closer in more branched networks as branching compresses the network configuration.
-However, unlike ecosystem size, ecosystem complexity does not increase the total number of dispersing propagules - rather, it enhances connectivity among habitat patches.
-This difference may explain the distinct patterns of state transitions along the gradients of ecosystem size and complexity.
+It is therefore probable that different processes underlie the productivity-driven enforcement of the FCL-geometry relationships among food web motifs.
+In the chain and strong omnivory motifs, increased productivity magnified the disproportionate impact of disturbance on small and/or simple ecosystems because they seem less stable with limited supply of immigrants and external perturbations.
+With the weak omnivory motif, productivity facilitated the spatial spread of the intraguild predator in small ecosystems, where dispersing propagules can reach wider areas of a network with local extirpation of the intraguild prey.
 

### 2023-03-07 rmd/manuscript.Rmd
update: manuscript
equation numbers

@@
 \begin{aligned}
-B^+_{x,t} &= B_{x,t}g(B_{x,t})f_1(B'_{x,t}, C_{x,t})f_2(B'_{x,t}, P_{x,t})\\
-C^+_{x,t} &= e_{BC}B'_{x,t}[1 - f_1(B'_{x,t}, C_{x,t})]f_3(C'_{x,t}, P_{x,t})\\
+B^+_{x,t} &= B_{x,t}g(B_{x,t})f_1(B'_{x,t}, C_{x,t})f_2(B'_{x,t}, P_{x,t}) &&\text{(5a)}\\
+C^+_{x,t} &= e_{BC}B'_{x,t}[1 - f_1(B'_{x,t}, C_{x,t})]f_3(C'_{x,t}, P_{x,t}) &&\text{(5b)}\\
 P^+_{x,t} &= e_{BP}B'_{x,t}f_1(B'_{x,t}, C_{x,t})[1 - f_2(B''_{x,t}, P_{x,t})] + e_{CP}C_{x,t}[1 - f_3(C'_{x,t}, P_{x,t})]
-~~\text{(5)}
+&&\text{(5c)}
 \end{aligned}
@@
 \begin{aligned}
-B'_{x,t} &= B_{x,t}g(B_{x,t}) &&\text{B's recruitment}\\
-B''_{x,t} &= B'_{x,t}f_1(B'_{x,t}, C_{x,t}) &&\text{C's predation on B}\\
-C'_{x,t} &= e_{BC}B'_{x,t}[1 - f_1(B'_{x,t}, C_{x,t})] &&\text{C's recruitment}
+B'_{x,t} &= B_{x,t}g(B_{x,t}) &&\text{B's recruitment} &&&\text{(6a)}\\
+B''_{x,t} &= B'_{x,t}f_1(B'_{x,t}, C_{x,t}) &&\text{C's predation on B} &&&\text{(6b)}\\
+C'_{x,t} &= e_{BC}B'_{x,t}[1 - f_1(B'_{x,t}, C_{x,t})] &&\text{C's recruitment} &&&\text{(6c)}
 \end{aligned}
@@
 \begin{aligned}
-g(B_{x,t}) = \frac{r_{b}} {1 + \frac{r_b - 1}{K_{x}}B_{x,t}} &&\text{(11)}
+g(B_{x,t}) = \frac{r_{b}} {1 + \frac{r_b - 1}{K_{x}}B_{x,t}} &&\text{(7)}
 \end{aligned}
@@
 \begin{aligned}
-f_1(B'_{x,t}, C_{x,t}) &= \exp \left(-\frac{a_{BC} C_{x,t}} {1 + a_{BC} h_{BC} B'_{x,t}} \right) &&\text{C on B}\\
-f_2(B''_{x,t}, P_{x,t}) &= \exp \left(-\frac{(1 + \phi) a_{BP} P_{x,t}} {1 + a_{BP} h_{BP} B''_{x,t}} \right) &&\text{P on B}\\
-f_3(C'_{x,t}, P_{x,t}) &= \exp \left(-\frac{(1 - \phi) a_{CP} P_{x,t}} {1 + a_{CP} h_{CP} C'_{x,t}} \right) &&\text{P on C}
+f_1(B'_{x,t}, C_{x,t}) &= \exp \left(-\frac{a_{BC} C_{x,t}} {1 + a_{BC} h_{BC} B'_{x,t}} \right) &&\text{C on B} &&&\text{(8a)}\\
+f_2(B''_{x,t}, P_{x,t}) &= \exp \left(-\frac{(1 + \phi) a_{BP} P_{x,t}} {1 + a_{BP} h_{BP} B''_{x,t}} \right) &&\text{P on B} &&&\text{(8b)}\\
+f_3(C'_{x,t}, P_{x,t}) &= \exp \left(-\frac{(1 - \phi) a_{CP} P_{x,t}} {1 + a_{CP} h_{CP} C'_{x,t}} \right) &&\text{P on C} &&&\text{(8c)}
 \end{aligned}
@@
 $$
-\phi = \frac{s(B''_{x,t} - C'_{x,t})}{B''_{x,t} + C'_{x,t}}
+\begin{aligned}
+\phi = \frac{s(B''_{x,t} - C'_{x,t})}{B''_{x,t} + C'_{x,t}} &&\text{(9)}
+\end{aligned}
 $$
@@
 \begin{aligned}
-\bar{A}_{ix,t}(t) &= A^+_{ix,t} (1 - \nu_t m'_{x,t}) &&\text{(15)}
+\bar{A}_{ix,t}(t) &= A^+_{ix,t} (1 - \nu_t m'_{x,t}) &&\text{(10)}
 \end{aligned}
@@
 \xi_{ix,t} &= \frac{\sum_{y, y \ne x}E_{iy,t}\exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}E_{iy,t}\exp(-\theta d_{xy})}
-&&\text{(16)}
+&&\text{(11)}
 \end{aligned}
@@
 \tau_{x,t} &= \delta_{x,t} + 2(1 - \delta_{x,t}) + 1
-&&\text{(17)}
+&&\text{(12)}
 \end{aligned}
@@
 $$
-\delta_{x,t} = \frac{e_{BP}B'_{x,t}f_1(B'_{x,t}, C_{x,t})[1 - f_2(B''_{x,t}, P_{x,t})]}{P^+_{x,t}}
+\begin{aligned}
+\delta_{x,t} = \frac{e_{BP}B'_{x,t}f_1(B'_{x,t}, C_{x,t})[1 - f_2(B''_{x,t}, P_{x,t})]}{P^+_{x,t}} &&\text{(13)}
+\end{aligned}
 $$

### 2023-03-07 rmd/manuscript.Rmd
update: manuscript

@@
 In this formulation, the most upstream patch, at which $Q_x = 1$ (the number of upstream patches + 1), has the carrying capacity of $K_{base}$.
-We used $K_{base} = 500$ and $z = 0.54$ based on an empirical relationship of gross primary production and watershed area [@finlayStreamSizeHuman2011].
+We used $z = 0.54$ based on an empirical relationship of gross primary production and watershed area [@finlayStreamSizeHuman2011].
 
@@
 where $d_{xy}$ is the network distance between patch $x$ and $y$ (shortest path), and $\theta$ indicates the inverse of mean dispersal distance (i.e., larger numbers mean shorter dispersal distance).
-The immigration potential at patch $x$, $\sum_{y, y \ne x}E_{iy,t}\exp(-\theta d_{xy})$, is scaled by the sum of immigration potential across patches so that $\sum_{x} \xi_{i,x}=1$.
+The immigration potential at patch $x$, $\sum_{y, y \ne x}E_{iy,t}\exp(-\theta d_{xy})$, is scaled by the sum of immigration potential across patches so that $\sum_{x} \xi_{ix,t}=1$.
 The expected number of immigrants into patch $x$ is calculated as $I_{ix,t}=\xi_{ix,t} \sum_{x}E_{ix,t}$.
@@
 In the chain motif, the transition from "B + C" to "B + C + P" (i.e., additive) was dominant regardless of disturbance frequency.
-The introduction of omnivory allowed the transition between "B + P" and "B + C + P", the primary direction of the transition differed between weak and strong omnivory.
+The introduction of omnivory allowed the transition between "B + P" and "B + C + P", and the primary direction of the transition differed between weak and strong omnivory.
 With weak omnivory, the transition of "B + C + P" $\rightarrow$ "B + P" was prevalent and shortened FCL regardless of disturbance frequency (Figure XXX).
@@
 However, we did not observe a clear proportional increase in disturbance refugia with increasing branching probability.
-Collectively, each of geometric features likely support long food chains through distinct mechanisms, which may also explain the different patterns of state transitions (additive vs.
+Collectively, each of geometric features likely supports long food chains through distinct mechanisms, which may also explain the different patterns of state transitions (additive vs.
 insertion mechanisms) [@postProximateStructuralMechanisms2007].
-
 While empirical studies rarely looked at the influence of ecosystem complexity on FCL, several studies provide some support for this theoretical prediction.

### 2023-03-08 rmd/manuscript.Rmd
update: si document

@@
 This analysis revealed that the stable coexistence of the three species is possible with a certain combinations of predation parameters.
-Within this parameter space, we used three parameter combinations to represent distinctive food webs with varied levels of omnivory: "chain (no omnivory)" [$(a_{BC}, a_{BP}, a_{CP}, h) = (0.5, 0, 0.0025, 0.5)$], "weak omnivory" ($0.5, 0.02, 0.025, 0.5$), and "strong omnivory" ($0.5, 0.04, 0.0025, 0.75$).
+Within this parameter space, we used three parameter combinations to represent distinctive food webs with varied levels of omnivory: "chain (no omnivory)" $(a_{BC}, a_{BP}, a_{CP}, h) = (0.5, 0, 0.0025, 0.5)$, "weak omnivory" ($0.5, 0.02, 0.025, 0.5$), and "strong omnivory" ($0.5, 0.04, 0.0025, 0.75$).
 We used $s = 1$ except for the chain scenario with $s = 0$.

### 2023-03-08 rmd/suppl_info.Rmd
edit
update: si document

--- 2022-05-10 rmd/suppl_info.Rmd
@@
       })
+fontsize: 11pt
 header-includes:
@@
 
+library(tidyverse)
+
 knitr::opts_chunk$set(echo = FALSE,
@@
 
-# Sensitivity data
-source(here::here("code/analysis_stvy.R"))
-load(file = here::here("output/result_main.RData"))
-load(file = here::here("output/result_stvy.RData"))
-
-sim_sv <- sim_stvy_result
-n_sim_stvy <- unique(sim_sv$n_warmup) + unique(sim_sv$n_burnin) + unique(sim_sv$n_timestep)
-
+df_sim1 <- readRDS(here::here("output/sim_one_patch.rds"))
+n_sim <- df_sim1 %>% 
+  mutate(x = n_timestep + n_warmup + n_burnin,
+         y = n_warmup,
+         z = n_burnin) %>% 
+  dplyr::select(x, y, z) %>% 
+  distinct() %>% 
+  unlist()
 ```
@@
 
-## Sensitivity analysis
+## One-patch simulation
 
-We performed a sensitivity analysis of the trophic community simulation to identify key simulation parameters that strongly affect the relationships between food chain length and ecosystem properties (the number of habitat patches $N_{p}$ and branching probability $P_{b}$.
-We generated 500 sets of parameter combinations by randomly drawing values of `r length(attributes(fit_sense$fit[[1]]$terms)$term.label)` simulation parameters from uniform distributions (**Table S1**).
-For each parameter combination, we generated `r max(sim_sv$n_rep)` branching networks with the gradients of ecosystem size ($N_p \sim Unif(10, 150)$) and complexity ($P_b \sim Unif(0.01, 0.99)$).
-This results in a total of `r nrow(sim_sv)` simulation replicates.
-We ran `r n_sim_stvy` time steps of trophic community dynamics in each simulation replicate.
-We obtained temporal means of food chain length for the last `r unique(sim_sv$n_timestep)` time steps.
-The first `r n_sim_stvy - unique(sim_sv$n_timestep)` time steps were discarded as initialization and burn-in periods.
+We performed one-patch simulations to identify parameter combinations that allow stable coexistence of the three species without dispersal and disturbance.
+Specifically, we varied attack rates $a_{ij}$, handling time $h_{ij}$, intrinsic growth rate of the basal species $r_b$, and carrying capacity $K$.
+First, we considered `r nrow(dplyr::distinct(df_sim1, r_b, k, a1, h))` discrete combinations of $a_{BC}$, $r_b$, $K$, and $h$ (assumed $h_{ij} = h$).
+We further crossed `r nrow(dplyr::distinct(df_sim1, a2))` values of $a_{BP}$ (`r min(df_sim1$a2)` -- `r max(df_sim1$a2)` with an equal interval) and $a_{CP}$ (`r min(df_sim1$a3)` -- `r max(df_sim1$a3)`).
+Consequently, we had a total of `r nrow(dplyr::distinct(df_sim1, r_b, k, a1, a2, a3, h))` combinations for one-patch simulations.
 
-For each parameter combination, we estimated the Spearman's rank correlation between food chain length and ecosystem properties ($\psi_1$ for $N_{p}$ and $\psi_2$ for $P_{b}$).
-To examine influences of simulation parameters (**Table S1**), we developed the following regression model taking $\psi_1$ or $\psi_2$ as a response variable $u_n$ (parameter combination $n$):
+For each parameter combination, we ran 10 replications of `r n_sim[1]` time steps, which include `r n_sim[2]` warmup and `r n_sim[3]` burn-in periods.
+We recorded the patch state as numeric values ("No species" = 0, "B" = 1, "B + C" = 2, "B + P" =3, and "B + C + P" = 4) at $t = 1600$ and retrieved the maximum value of 10 replicates.
 
-```{=tex}
-\begin{align}
-  u_{n} &\sim Normal(\mu_n, \sigma^2)\\
-  \mu_n &= \zeta_0 + \sum_{k}{\zeta_k x_{k, n}}
-\end{align}
-```
-where $\zeta_k$ are the intercept and regression coefficients, and $x_{k,n}$ are the simulation parameters drawn from uniform distributions (**Table S1**).
-Response and explanatory variables were standardized to a mean of zero and a standard deviation of one so that regression coefficients are comparable.
+Results were summarized in **Figures S1 and S2**.
+We identified the following three parameter combinations that could maintain the state of "B + C + P" with distinct food web motifs: "chain (no omnivory)" $(a_{BP}, a_{CP}, h) = (0, 0.0025, 0.5)$, "weak omnivory" ($0.02, 0.025, 0.5$), and "strong omnivory" ($0.04, 0.0025, 0.75$) (see symbols in **Figures S1 and S2**).
+In the maintext, we reported the results with $a_{BC} = 0.5$ because qualitative patterns did not depend on $a_{BC}$.
 
-The sensitivity analysis revealed key simulation parameters.
-For ecosystem size, dispersal distance $\theta$, local disturbance variation $\sigma_{l}$, and the reproductive rate of basal species $r_b$ were influential.
-However, parameters that influence food web structure were less influential (attack rate, conversion efficiency, and handling time).
-For ecosystem complexity, dispersal distance $\theta$ and probability $p_d$, local disturbance variation $\sigma_{l}$, and the reproductive rate of basal species $r_b$ were influential.
-Food web parameters were also important (attack rate, conversion efficiency, and handling time).
-
-Based on the results, we considered ecological scenarios relevant for putative controls of food chain: low and high productivity ($r_b = 4, 8$), and low and high disturbance ($p_m = 0.01, 0.1$).
-We also investigated the role of omnivory by changing attack rate ($a_{BP}$ and $a_{CP}$) and conversion efficiency ($e_{BP}$ and $e_{CP}$) of two prey items for predator $P$.
-Specifically, we used the following combinations to represent three levels of omnivory: "Chain", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[1], max(sim_main_result$e_cp))` $\}$; "Weak omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[2], max(sim_main_result$e_cp))` $\}$; and "Strong omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(max(sim_main_result$a_bp), min(sim_main_result$a_cp), unique(sim_main_result$e_bp)[3], min(sim_main_result$e_cp))` $\}$.
-To simplify simulation setup, we fixed the following parameters: $e_{BC}=$ `r unique(sim_main_result$e_bc)` , $a_{BC}=$ `r unique(sim_main_result$a_bc)`, and $h_{ij}=$ `r unique(sim_main_result$h_bc)`.
-
-\pagebreak
-
-# Tables
-
-## Table S1 Parameter values
-
-Parameter values used for the main and sensitivity analysis.
-
-```{r table_param}
-
-source(here::here("code/table_param.R"))
-knitr::kable(df_value, format = "markdown")
-
-```
-
-\pagebreak
-
-```{r table_sensitivity}
-
-source(here::here("code/table_stvy.R"))
-
-j <- 1
-
-## conditional text
-c_text <- c("ecosystem size",
-            "ecosystem complexity")
-
-table_stvy_chunk <- knit_chunk <- NULL
-
-for(i in 1:length(table_stvy)) {
-
-  table_legend <- paste0("Sensitivity analysis of the ", c_text[i]," effect on food chain length.
-Parameter estimates of linear regression models are shown.
-The response variable is the Spearman's rank correlation between food chain length and ", c_text[i],".
-Response and explanatory variables (i.e., simulation parameters) were standardized (mean = 0, SD = 1) before the analysis.")
-
-  knit_chunk <- paste0("## Table S", i + j, " Sensitivity analysis for ", c_text[i]," \n",
-                       table_legend,
-                       "\n```{r table_s", i + j,", results='asis'}
-                       \n\n
-                       table_stvy[[", i,"]]
-                       \n\n
-                       ```\n",
-                       "\n\\pagebreak\n")
-
-  table_stvy_chunk <- c(table_stvy_chunk, knit_chunk)
-
-}
-
-```
-
-`r paste(knit(text = table_stvy_chunk), collapse = '\n')`
-
-\pagebreak
+\newpage
 
@@
 
-```{r si_figure_fcl}
-
-source(here::here("code/figure_si_fcl.R"))
-
-j <- 0
-
-## conditional text
-c_text <- c("ecosystem size",
-            "ecosystem complexity")
-
-fig_fcl_chunk <- knit_chunk <- NULL
-
-for(i in 1:length(list_g_fcl)) {
-
-  K <- length(list_g_fcl[[i]])
-
-  for(k in 1:K) {
-
-      figure_legend <- paste0(str_to_sentence(c_text[i])," effect on food chain length and occupancies of constituent species (proportion of habitat patches occupied).
-Rows are different response variables, and columns are different simulation scenarios.
-Lines and shades represent loess curves fitted to the simulated data and associated 95% confidence intervals.
-In this simulation, the following values of parameters were used: $\\theta$ = ", df_param$theta[k], ", $\\sigma_l$ = ", df_param$sd_disturb_lon[k], ", and $\\sigma_h$ = ", df_param$sd_disturb_source[k], ".")
-
-  knit_chunk <- paste0("## Figure S", K * (i - 1) + k + j, " ", str_to_sentence(c_text[i]), " effect on food chain length", " \n",
-                       "\n```{r figure_s", K * (i - 1) + k + j,", fig.height = 10, fig.width = 11.5}
-                       \n\n
-                       list_g_fcl[[", i, "]][[", k, "]]
-                       \n\n
-                       ```\n",
-                       figure_legend,
-                       "\n\\pagebreak\n")
-
-  fig_fcl_chunk <- c(fig_fcl_chunk, knit_chunk)
-
-  }
-
-}
-
+```{r fig1, fig.cap="Results of one-patch simulation with $a_{BC} = 0.25$.
+Colors indicate the patch state, and points with different symbols identify parameter combinations used in the main simulation.
+Panels distinguishh the value of carrying capacity $K$.
+In both panels, rows represent different values of intrinsic growth rate $r_b$, and columns identify handling time $h$."}
+knitr::include_graphics(here::here("figure/figure_si_one_patch_a1_25.pdf"))
 ```
 
-`r paste(knit(text = fig_fcl_chunk), collapse = '\n')`
+\newpage
+
+```{r fig2, fig.cap="Results of one-patch simulation with $a_{BC} = 0.50$.
+Colors indicate the patch state, and points with different symbols identify parameter combinations used in the main simulation.
+Panels distinguishh the value of carrying capacity $K$.
+In both panels, rows represent different values of intrinsic growth rate $r_b$, and columns identify handling time $h$."}
+knitr::include_graphics(here::here("figure/figure_si_one_patch_a1_50.pdf"))
+```
+
+\newpage
+
+```{r fig3, fig.cap="Relationships between food chain length and ecosystem size with low disturbance intensity $\\mu_m = 0.2$.
+Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios."}
+knitr::include_graphics(here::here("figure/figure_si_n_patch.pdf"))
+```
+
+\newpage
+
+```{r fig4, fig.cap="Relationships between food chain length and ecosystem complexity with low disturbance intensity $\\mu_m = 0.2$.
+Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios."}
+knitr::include_graphics(here::here("figure/figure_si_p_branch.pdf"))
+```
+
+\newpage
+
+```{r fig5, fig.cap="Relationships between patch state occupancy and ecosystem size with high disturbance intensity $\\mu_m = 0.8$.
+Panels indicate different food web motifs.
+In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios."}
+knitr::include_graphics(here::here("figure/figure_si_state_n_patch.pdf"))
+```
+
+\newpage
+
+```{r fig6, fig.cap="Relationships between patch state occupancy and ecosystem complexity with high disturbance intensity $\\mu_m = 0.8$.
+Panels indicate different food web motifs.
+In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios."}
+knitr::include_graphics(here::here("figure/figure_si_state_p_branch.pdf"))
+```
+
+\newpage
+
+```{r fig7, fig.cap="Effects of productivity (left column) and dispersal (right column) on food chain length grouped by disturbance frequency ($p_m$; x-axis).
+Colors distinguish productivity or dispersal scenarios, and rows identify food web motifs.
+Disturbance intensity $\\mu_m = 0.8$."}
+knitr::include_graphics(here::here("figure/figure_si_fcl_r_disp.pdf"))
+```
+
+\newpage
+
+```{r fig8, fig.cap="Effects of productivity on state transitions. Colors distinguish productivity scenarios, and rows identify food web motifs. Disturbance intensity $\\mu_m = 0.8$."}
+knitr::include_graphics(here::here("figure/figure_si_state_r.pdf"))
+```
+
+\newpage
+
+```{r fig9, fig.cap="Effects of dispersal on state transitions. Colors distinguish distinguish scenarios, and rows identify food web motifs. Disturbance intensity $\\mu_m = 0.8$."}
+knitr::include_graphics(here::here("figure/figure_si_state_disp.pdf"))
+```

### 2023-03-08 code/figure_si_geom.R

@@ -35,7 +35,7 @@ g_np <-  df_plot %>%
   geom_smooth(method = "loess") +
   facet_grid(rows = vars(omn),
              cols = vars(disp, r_b),
-             #scales = "free",
+             scales = "free",
              labeller = labeller(r_b = as_labeller(lab, label_parsed))) +
   labs(y = "Food chain length",
        x = "Ecosystem size (number of habitat patches)",
@@ -54,7 +54,7 @@ g_pb <-  df_plot %>%
   geom_smooth(method = "loess") +
   facet_grid(rows = vars(omn),
              cols = vars(disp, r_b),
-             #scales = "free",
+             scales = "free",
              labeller = labeller(r_b = as_labeller(lab, label_parsed))) +
   labs(y = "Food chain length",
        x = "Ecosystem complexity (branching probability)",
@@ -70,13 +70,13 @@ g_pb <-  df_plot %>%
 ## ecosystem size with low productivity
 ggsave(g_np,
        filename = here::here("figure/figure_si_n_patch.pdf"),
-       height = 9,
-       width = 15)
+       height = 6,
+       width = 10)
 
 ## ecosystem size with low productivity
 ggsave(g_pb,
        filename = here::here("figure/figure_si_p_branch.pdf"),
-       height = 9,
-       width = 15)
+       height = 6,
+       width = 10)
 
 

### 2023-03-08 code/figure_si_one_patch.R

@@ -20,6 +20,20 @@ lab <- list(c(`0` = "h==0.0",
 
 state_lab <- c("No species", "B", "B + C", "B + P", "B + C + P")
 
+df_point <- tibble(a1 = 0.5,
+                   a2 = c(0, 0.02, 0.04),
+                   a3 = c(0.0025, 0.025, 0.0025),
+                   h = c(0.5, 0.5, 0.75),
+                   motif = c("Chain",
+                             "Weak omnivory",
+                             "Strong omnivory"),
+                   state = 0)
+
+df_point <- bind_rows(df_point, df_point) %>% 
+  mutate(r_b = rep(c(8, 16), each = 3),
+         motif = fct_relevel(motif, "Chain", "Weak omnivory"))
+  
+
 df_fcl <- readRDS(here::here("output/sim_one_patch.rds"))
 k_set <- unique(df_fcl$k)
 a1_set <- unique(df_fcl$a1)
@@ -44,20 +58,19 @@ foreach(i = 1:length(a1_set)) %do% {
       scale_fill_viridis_d(limits = factor(0:4),
                            labels = state_lab) +
       scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
-      labs(x = expression("Attack rate ("*a[BP]*")"),
-           y = expression("Attack rate ("*a[CP]*")"),
-           fill = "State") +
       theme_classic() +
       theme(strip.background = element_blank(),
             strip.text = element_text(size = 12), 
             axis.title = element_text(size = 14),
             axis.text = element_text(size = 10)) +
       ggtitle(paste0("K = ", k_set[j])) +
-      geom_point(data = tibble(a2 = c(0, 0.02, 0.04),
-                               a3 = c(0.0025, 0.025, 0.0025),
-                               h = c(0.5, 0.5, 0.75),
-                               state = 0),
-                 color = "salmon")
+      geom_point(data = df_point,
+                 size = 1.5,
+                 aes(shape = motif)) +
+      labs(x = expression("Attack rate ("*a[BP]*")"),
+           y = expression("Attack rate ("*a[CP]*")"),
+           fill = "State",
+           shape = "Motif")
     
     return(g_one)
   }
@@ -69,11 +82,11 @@ foreach(i = 1:length(a1_set)) %do% {
   # export ------------------------------------------------------------------
   
   ggsave(g_one_joint, 
-         filename = here::here(paste0("figure/si_figure_one_patch_a1_",
[…]

### 2023-03-08 code/figure_si_r_disp.R

@@ -124,16 +124,16 @@ g_disp_state <- df_plot_st %>%
 # export ------------------------------------------------------------------
 
 ggsave(g_rd,
-       filename = here::here("figure/figure_si_rb_disp.pdf"),
+       filename = here::here("figure/figure_si_fcl_r_disp.pdf"),
        height = 8,
        width = 7)
 
 ggsave(g_r_state,
-       filename = here::here("figure/figure_si_r_state.pdf"),
+       filename = here::here("figure/figure_si_state_r.pdf"),
        height = 6,
        width = 10)
 
 ggsave(g_disp_state,
-       filename = here::here("figure/figure_si_disp_state.pdf"),
+       filename = here::here("figure/figure_si_state_disp.pdf"),
        height = 6,
        width = 10)

### 2023-03-08 code/figure_si_state.R

@@ -56,6 +56,8 @@ list_g_np <- foreach(x = unique(df_plot$omn)) %do% {
          fill = "Disturbance prob.") +
     scale_color_met_d("Hiroshige", direction = -1) +
     scale_fill_met_d("Hiroshige", direction = -1) +
+    theme(legend.title = element_text(size = 20),
+          legend.text =  element_text(size = 18)) +
     ggtitle(x)
 }
 
@@ -89,6 +91,8 @@ list_g_bp <- foreach(x = unique(df_plot$omn)) %do% {
     scale_x_continuous(breaks = c(0.1, 0.4, 0.7, 1.0)) +
     scale_color_met_d("Hiroshige", direction = -1) +
     scale_fill_met_d("Hiroshige", direction = -1) +
+    theme(legend.title = element_text(size = 20),
+          legend.text =  element_text(size = 18)) +
     ggtitle(x)
 }
 
@@ -102,11 +106,11 @@ g_bp_all <- list_g_bp[[1]] + list_g_bp[[2]] +
 # export ------------------------------------------------------------------
 
 ggsave(g_np_all,
-       filename = here::here(paste0("figure/si_figure_state_n_patch.pdf")),
+       filename = here::here(paste0("figure/figure_si_state_n_patch.pdf")),
        height = 14,
        width = 14)
 
 ggsave(g_bp_all,
-       filename = here::here(paste0("figure/si_figure_state_p_branch.pdf")),
+       filename = here::here(paste0("figure/figure_si_state_p_branch.pdf")),
        height = 14,
        width = 14)

### 2023-03-08 rmd/suppl_info.Rmd
update: si

@@
 First, we considered `r nrow(dplyr::distinct(df_sim1, r_b, k, a1, h))` discrete combinations of $a_{BC}$, $r_b$, $K$, and $h$ (assumed $h_{ij} = h$).
-We further crossed `r nrow(dplyr::distinct(df_sim1, a2))` values of $a_{BP}$ (`r min(df_sim1$a2)` -- `r max(df_sim1$a2)` with an equal interval) and $a_{CP}$ (`r min(df_sim1$a3)` -- `r max(df_sim1$a3)`).
-Consequently, we had a total of `r nrow(dplyr::distinct(df_sim1, r_b, k, a1, a2, a3, h))` combinations for one-patch simulations.
+We further crossed `r nrow(dplyr::distinct(df_sim1, a2))` values of $a_{BP}$ (`r min(df_sim1$a2)` -- `r max(df_sim1$a2)` with an equal interval) and $a_{CP}$ (`r min(df_sim1$a3)` -- `r max(df_sim1$a3)`), resulting in `r nrow(dplyr::distinct(df_sim1, a2, a3))` combinations.
+Consequently, we had a total of `r nrow(dplyr::distinct(df_sim1, r_b, k, a1, a2, a3, h))` combinations (`r nrow(dplyr::distinct(df_sim1, r_b, k, a1, h))` $\times$ `r nrow(dplyr::distinct(df_sim1, a2, a3))`) for one-patch simulations.
 
@@
 
-```{r fig8, fig.cap="Effects of productivity on state transitions. Colors distinguish productivity scenarios, and rows identify food web motifs. Disturbance intensity $\\mu_m = 0.8$."}
+```{r fig8, fig.cap="Effects of productivity on state transitions with high disturbance intensity $\\mu_m = 0.8$.
+Colors distinguish productivity scenarios, columns represent different disturbance probability $p_m$, and rows identify food web motifs."}
 knitr::include_graphics(here::here("figure/figure_si_state_r.pdf"))
@@
 
-```{r fig9, fig.cap="Effects of dispersal on state transitions. Colors distinguish distinguish scenarios, and rows identify food web motifs. Disturbance intensity $\\mu_m = 0.8$."}
+```{r fig9, fig.cap="Effects of dispersal on state transitions with high disturbance intensity $\\mu_m = 0.8$.
+Colors distinguish dispersal scenarios, columns represent different disturbance probability $p_m$, and rows identify food web motifs."}
 knitr::include_graphics(here::here("figure/figure_si_state_disp.pdf"))
 ```
+
+## Add
+
+- total carrying capacity
+- mean distance

### 2023-03-08 rmd/suppl_info.Rmd
update: si

@@
 header-includes:
-  \pagenumbering{gobble}
-  \usepackage{amsmath}
+  - \usepackage{amsmath, fancyhdr, caption, booktabs}
+  - \pagestyle{fancy}
+  - \renewcommand{\headrulewidth}{0pt}
+  - \pagenumbering{gobble}
 ---
@@
 \newpage
+\pagestyle{fancy}
+\fancyhead[L]{Pomeranz et al.}
+\fancyhead[C]{}
+\renewcommand{\headrulewidth}{0.5pt}
 \pagenumbering{arabic}
+\def\figurename{Figure S\hspace{-.4ex}}
+\def\tablename{Table S\hspace{-.4ex}}
 ```
@@
 In the maintext, we reported the results with $a_{BC} = 0.5$ because qualitative patterns did not depend on $a_{BC}$.
+
+\newpage
+
+# Tables
+
+```{r}
+source(here::here("code/table_param.R"))
+
+kable(df_value,
+      escape = FALSE,
+      booktabs = TRUE,
+      caption = "Parameter values used in the main simulation and one-patch simulation") %>% 
+  kableExtra::kable_styling(latex_options = 'HOLD_position')
+```
 
@@
 
-## Add
+\newpage
 
-- total carrying capacity
-- mean distance
+```{r fig10, fig.cap="Ecosystem size and complexity dictate total carrying capacity (= $\\sum_x K_x$) and mean distance between a given pair of habitat patches within a network."}
+knitr::include_graphics(here::here("figure/figure_si_k_distance.pdf"))
+```

### 2023-03-08 code/figure_si_within_net.R

@@ -7,18 +7,60 @@ lapply(list("code/library.R",
             "code/set_figure_theme.R"),
        source)
 
+theme_set(plt_theme)
 
-# figure ------------------------------------------------------------------
 
-df_wn <- readRDS("output/sim_within_net.rds")
+# data --------------------------------------------------------------------
 
-df_wn %>% 
-  filter(n_patch == 150,
-         a_bp == 0.04) %>% 
-  ggplot(aes(x = carrying_capacity,
-             y = fcl,
-             color = factor(p_branch),
-             fill = factor(p_branch))) +
-  geom_point(alpha = 0.2) +
-  geom_smooth()# +
-  #facet_wrap(facets = ~factor(n_patch))
+df_sim <- readRDS(here::here("output/sim_main.rds")) %>% 
+  filter(param_set == 1)
+
+n_patch <- df_sim$n_patch
+p_branch <- df_sim$p_branch
+
+df_dist <- foreach(i = seq_len(length(p_branch)),
+                   .combine = bind_rows) %do% {
+                     net <- brnet(n_patch = n_patch[i],
+                                  p_branch = p_branch[i])
+                     
+                     diag(net$distance_matrix) <- NA
+                     df_y <- tibble(n_patch = n_patch[i],
+                                    p_branch = p_branch[i],
+                                    mu_distance = mean(net$distance_matrix, na.rm = T))
+                     return(df_y)
+                   }
+
+df_plot <- left_join(df_sim, df_dist,
+                    by = c("n_patch", "p_branch")) %>% 
+  pivot_longer(cols = c(mc_capacity, mu_distance),
+               values_to = "value",
+               names_to = "y")
+
+
+# plot --------------------------------------------------------------------
+
+label <- c(`mc_capacity` = "Total carrying capacity",
+           `mu_distance` = "Mean distance")
+
+g_net <- df_plot %>% 
+  ggplot(aes(x = n_patch,
+             y = value,
+             color = p_branch)) +
+  geom_point(alpha = 0.5) +
+  scale_x_continuous(trans = "log10") +
+  scale_y_continuous(trans = "log10") +
+  facet_wrap(facets =~ y,
+             ncol = 1,
+             scales = "free",
+             switch = "y",
+             labeller = labeller(y = label)) +
+  labs(x = "Ecosystem size (number of patches)",
+       color = "Branching prob.") +
+  theme(strip.placement = "outside",
+        axis.title.y = element_blank())
+
+
[…]

### 2023-03-08 code/run_sim_one_patch.R

@@ -20,7 +20,9 @@ df_param <- expand.grid(n_timestep = 1000,
                         a2 = seq(0, 0.1, length = 20),
                         a3 = seq(0, 0.05, length = 20),
                         h = seq(0, 1, by = 0.25),
-                        s = 1) %>%
+                        s = c(0, 1)) %>%
+  filter(!(s == 1 & a2 == 0), # remove s = 1 for chain scenario
+         !(s == 0 & a2 != 0)) # remove s = 0 for omnivory scenarios
   mutate(param_set = seq_len(nrow(.))) %>% 
   as_tibble()
 

### 2023-03-08 code/table_param.R

@@ -7,25 +7,22 @@ source(here::here("code/library.R"))
 
 # table -------------------------------------------------------------------
 
-load(file = here::here("output/result_main.RData"))
+df_sim <- readRDS(file = here::here("output/sim_main.rds"))
 
-df_value <- sim_main_result %>% 
-  dplyr::select(-n_rep,
-                -mc_capacity,
-                -n_patch,
-                -p_branch,
-                -param_set,
-                -base_k,
-                -z,
-                -n_timestep,
-                -n_warmup,
-                -n_burnin,
-                -fcl,
-                -p_basal,
-                -p_igprey,
-                -p_igpred) %>% 
-  filter(sd_disturb_source > sd_disturb_lon) %>% 
-  summarize(across(.fns = function(x) list(sort(unique(x))))) %>% 
+df_value <- df_sim %>% 
+  dplyr::select(r_b,
+                starts_with("e_"),
+                starts_with("a_"),
+                h,
+                s,
+                theta,
+                p_disturb,
+                base_k,
+                mean_disturb_source,
+                starts_with("sd_"),
+                p_dispersal) %>% 
+  summarize(across(.cols = everything(),
+                   .fns = function(x) list(sort(unique(x))))) %>% 
   pivot_longer(cols = everything(),
                names_to = "param",
                values_to = "Main") %>%
@@ -36,67 +33,63 @@ df_value <- sim_main_result %>%
                                param == "a_bc" ~"$a_{BC}$",
                                param == "a_bp" ~"$a_{BP}$",
                                param == "a_cp" ~"$a_{CP}$",
-                               param == "h_bc" ~"$h_{BC}$",
-                               param == "h_bp" ~"$h_{BP}$",
-                               param == "h_cp" ~"$h_{CP}$",
+                               param == "h" ~"$h$",
+                               param == "s" ~ "$s$",
                                param == "theta" ~"$\\theta$",
                                param == "p_disturb" ~"$p_m$",
+                               param == "base_k" ~ "$K_{base}$",
                                param == "mean_disturb_source" ~ "$\\mu_m$",
                                param == "sd_disturb_source" ~ "$\\sigma_h$",
                                param == "sd_disturb_lon" ~ "$\\sigma_l$",
-                               param == "s0" ~ "$s_0$",
                                param == "p_dispersal" ~ "$p_d$"),
          Model = case_when(param %in% c("r_b",
-                                        "s0",
+                                        "base_k",
                                         "e_bc",
                                         "e_bp",
                                         "e_cp",
[…]

### 2023-03-08 code/run_sim_one_patch.R
Update run_sim_one_patch.R

@@ -22,7 +22,7 @@ df_param <- expand.grid(n_timestep = 1000,
                         h = seq(0, 1, by = 0.25),
                         s = c(0, 1)) %>%
   filter(!(s == 1 & a2 == 0), # remove s = 1 for chain scenario
-         !(s == 0 & a2 != 0)) # remove s = 0 for omnivory scenarios
+         !(s == 0 & a2 != 0)) %>%  # remove s = 0 for omnivory scenarios
   mutate(param_set = seq_len(nrow(.))) %>% 
   as_tibble()
 

### 2023-03-08 rmd/suppl_info.Rmd
edit: one-patch sim

@@
   pdf_document:
-    toc: true
+    toc: false
     latex_engine: xelatex
@@
   - \pagenumbering{gobble}
+  - \setlength{\headheight}{14pt}
 ---
@@
 Consequently, we had a total of `r nrow(dplyr::distinct(df_sim1, r_b, k, a1, a2, a3, h))` combinations (`r nrow(dplyr::distinct(df_sim1, r_b, k, a1, h))` $\times$ `r nrow(dplyr::distinct(df_sim1, a2, a3))`) for one-patch simulations.
+Note that we set the switching parmeter $s = 0$ if $a_{BP} = 0$, otherwise $s = 1$.
+Parameter values were summarized in **Table S1**.
 

### 2023-03-08 code/figure_si_one_patch.R

@@ -57,7 +57,8 @@ foreach(i = 1:length(a1_set)) %do% {
                                                      r_b = as_labeller(lab[[2]], label_parsed))) +
       scale_fill_viridis_d(limits = factor(0:4),
                            labels = state_lab) +
-      scale_x_continuous(breaks = c(0, 0.05, 0.1)) +
+      scale_x_continuous(breaks = c(0.01, 0.05, 0.09)) +
+      scale_y_continuous(breaks = c(0, 0.025, 0.05)) +
       theme_classic() +
       theme(strip.background = element_blank(),
             strip.text = element_text(size = 12), 

### 2023-03-08 rmd/manuscript.Rmd
update: maintext

@@
 However, these effects were contingent upon other drivers.
-In particular, the effects of ecosystem size and complexity appeared more clearly in high disturbance regimes.
+In particular, the effects of ecosystem size and complexity appeared more clearly in high-disturbance regimes.
 Our results suggest that ecosystem complexity is an important yet overlooked driver of food chain length in branching ecosystems.
@@
 
-## **Network generation**
+## Network generation
 
@@
 The number of patches within branch $q$, $n_{p,q}$, is a random variable drawn from a geometric distribution $n_{p,q} \sim \text{Ge}(P_{b})$ but conditional on $N_{p} = \sum_{q}^{N_{b}} n_{p,q}$ ($N_{b}$ is the number of branches in a network).
-Full details of the network generation process is described in @teruiEmergentDualScaling2021.
+Full details of the network generation process are described in @teruiEmergentDualScaling2021.
 
-In our simulation, disturbance occurs simultaneously across habitat patches with probability $p_m$ but the patches vary in their vulnerability.
+In our simulation, a disturbance occurs simultaneously across habitat patches with probability $p_m$, but the patches vary in their vulnerability.
 We simulated the spatial distribution of disturbance intensity (i.e., proportional mortality in the event of stochastic regional disturbance) as follows.
@@
 
-where $K_{base}$ is the carrying capacity at the upstream terminal patch, and $z$ the scaling exponent.
+where $K_{base}$ is the carrying capacity at the upstream terminal patch, and $z$ is the scaling exponent.
 In this formulation, the most upstream patch, at which $Q_x = 1$ (the number of upstream patches + 1), has the carrying capacity of $K_{base}$.
-We used $z = 0.54$ based on an empirical relationship of gross primary production and watershed area [@finlayStreamSizeHuman2011].
+We used $z = 0.54$ based on an empirical relationship between gross primary production and watershed area [@finlayStreamSizeHuman2011].
 
-## **Tri-trophic model**
+## Tri-trophic model
 
@@
 
-***Internal dynamics.*** Each time step starts with internal dynamics. The general formula is the extension of the Nicholson-Bailey model, which is given as follows:
+***Internal dynamics.*** Each time step starts with internal dynamics.
+The general formula is the extension of the Nicholson-Bailey model [@hassellSpatialTemporalDynamics2000], which is given as follows:
 
@@
 
-The model assumes that ecological events occur in the following order: (1) $B$'s recruitment, (2) $C$'s predation on $B$ followed by its recruitment, and (3) $P$'s predation on $B$ and $C$ followed by its recruitment.
+The model assumes that ecological events occur in the following order: (1) $B$'s recruitment, (2) $C$'s predation on $B$ followed by $C$'s recruitment, and (3) $P$'s predation on $B$ and $C$ followed by $P$'s recruitment.
 We assumed that the intraguild prey $C$ consumes the basal species $B$ first because the species is specialized to prey on $B$.
@@
 
-where $r_{B}$ is the intrinsic growth rate and $K_x$ is the carrying capacity at patch $x$. We assumed the Type-II functional response in the survival function $f_k(\cdot)$ as follows:
+where $r_{B}$ is the intrinsic growth rate and $K_x$ is the carrying capacity at patch $x$.
+We assumed the Type-II functional response in the survival function $f_k(\cdot)$ as follows [@hassellSpatialTemporalDynamics2000; @weideHydraEffectParadox2019]:
 
@@
 The parameter $a_{ij}$ defines $j$'s attack rate to $i$, $h_{ij}$ the handling time, and $\phi$ the switching function for the intraguild predator.
-We assumed that switching is dependent on the relative abundance of $B''_{x,t}$ and $C'_{x,t}$:
+We assumed that switching is dependent on the relative abundance of $B''_{x,t}$ and $C'_{x,t}$ [@hassellSpatialTemporalDynamics2000]:
 
@@
 
-***Disturbance.*** After recruitment, all species experience proportional loss if stochastic regional disturbance occurs:
+***Disturbance.*** After recruitment, all species experience proportional loss if a stochastic regional disturbance occurs:
 
@@
 The mean disturbance intensity $m_x$ is determined by the network generation process.
-The random variable $\nu_t$ represents the incidence of regional disturbance ($\nu_t = 1$ if disturbance occurs, otherwise zero), which is drawn from a Bernoulli distribution as $\nu_t \sim \text{Bernouilli}(p_m)$.
+The random variable $\nu_t$ represents the incidence of regional disturbance ($\nu_t = 1$ if a disturbance occurs, otherwise zero), which is drawn from a Bernoulli distribution as $\nu_t \sim \text{Bernouilli}(p_m)$.
 
@@
 
-Prior to the main simulation, we performed a one-patch simulation to explore the predation parameter space ($a_{ij}$ and $h_{ij}$) that allows stable coexistence of the three species without disturbance.
+Prior to the main simulation, we performed one-patch simulations to explore the predation parameter space ($a_{ij}$ and $h_{ij}$) that allows stable coexistence of the three species without disturbance and dispersal (see **Supporting Information**).
 In this simulation, we assumed $h_{ij} = h$ to reduce the dimension of parameter space.
-This analysis revealed that the stable coexistence of the three species is possible with a certain combinations of predation parameters.
+This analysis revealed that the stable coexistence of the three species is possible with certain combinations of predation parameters.
 Within this parameter space, we used three parameter combinations to represent distinctive food webs with varied levels of omnivory: "chain (no omnivory)" $(a_{BC}, a_{BP}, a_{CP}, h) = (0.5, 0, 0.0025, 0.5)$, "weak omnivory" ($0.5, 0.02, 0.025, 0.5$), and "strong omnivory" ($0.5, 0.04, 0.0025, 0.75$).
@@
 
-Under each scenario of omnivory, we crossed disturbance (frequency and intensity), productivity, and dispersal.
+Under each motif, we crossed values of disturbance (frequency and intensity), productivity, and dispersal.
 Disturbance frequency and intensity were controlled via $p_m$ ( `r unique(df_sim$p_disturb) %>% sort()`) and $\mu_m$ (`r unique(df_sim$mean_disturb_source) %>% sort()`), respectively.
-Spatial variation in disturbance intensity at headwaters ($\sigma_{h}=$ `r max(df_sim$sd_disturb_source)`), and local disturbance variation ($\sigma_{l}=$ `r min(df_sim$sd_disturb_lon)`).
-This setup ($\sigma_h > \sigma_l$) produces high spatial variation among tributaries while allowing strong within-tributary downstream correlation, a typical pattern in real river networks [@nakamuraDisturbanceRegimesStream2000; @bendaNetworkDynamicsHypothesis2004].
+Spatial disturbance patterns were regulated by variation in disturbance intensity at headwaters ($\sigma_{h}=$ `r max(df_sim$sd_disturb_source)`), and local variation ($\sigma_{l}=$ `r min(df_sim$sd_disturb_lon)`).
+This setup ($\sigma_h > \sigma_l$) produces high spatial variation among tributaries while allowing strong within-tributary downstream correlation [@teruiEmergentDualScaling2021], a typical pattern in real river networks [@nakamuraDisturbanceRegimesStream2000; @bendaNetworkDynamicsHypothesis2004].
 We used a combination of basal species' growth rate $r_b$ and base carrying capacity $K_{base}$ (low productivity: $r_b =$ `r min(df_sim$r_b)`, $K_{base} =$ `r min(df_sim$base_k)` ; high productivity = $r_b =$ `r max(df_sim$r_b)`, $K_{base} =$ `r max(df_sim$base_k)` ) as a proxy for resource availability.
@@
 Subsequently, we ran a burn-in period with no propagules for `r unique(df_sim$n_burnin)` time steps to remove the effects of initial conditions.
-Note that simulation replicates are differ in ecosystem size ($N_{p}\sim \text{Unif}(10,150)$) and ecosystem complexity ($P_{b} \sim \text{Unif}(0.01, 0.99)$) so we can estimate their effects on FCL.
+Note that simulation replicates differ in ecosystem size ($N_{p}\sim \text{Unif}(10,150)$) and ecosystem complexity ($P_{b} \sim \text{Unif}(0.01, 0.99)$) so we can estimate their effects on FCL.
 
@@
 With no disturbance, both geometric features (size and complexity) had limited influence on FCL.
-However, as disturbance frequency increases, the relationships between FCL and ecosystem geometry became positive.
+However, as disturbance frequency increases, the relationships between FCL and ecosystem geometry become positive.
 This interactive effect emerged because the sensitivity of FCL to disturbance depended on ecosystem geometry.
-Although increased disturbance frequency reduced overall FCL, smaller and/or simpler ecosystems exhibited the greater vulnerability (**Figures 2 and 3**).
-Therefore, disturbance interacted with ecosystem geometry to produce the interdependency of these factors in determining FCL.
-However, the maximum levels of FCL was strongly regulated by disturbance frequency, highlighting the important direct effect of disturbance.
-It is important to note that the interactive effects were not observed with low disturbance intensity (proportion of individuals removed per disturbance event); this suggests that disturbance intensity must be strong enough to be influential.
+Although increased disturbance frequency reduced overall FCL, smaller and/or simpler ecosystems exhibited greater vulnerability (**Figures 2 and 3**).
+Therefore, disturbance interacted with ecosystem geometry to produce the interdependence of these factors in determining FCL.
+However, the maximum levels of FCL were strongly regulated by disturbance frequency, highlighting the important direct effect of disturbance.
+It is important to note that the interactive effects were not observed with low disturbance intensity (**Figures S3 and S4**); this suggests that disturbance intensity must be strong enough to be influential.
 
-The exception is the "chain" scenario (i.e., no omnivory) with no disturbance and low productivity.
-In this scenario, ecosystem size had a positive effect on FCL (**Figure 2**) while ecosystem complexity had a negative influence (**Figure 3**).
-These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increased the total carrying capacity whereas ecosystem complexity decreased it (**Figure XXX**).
-However, these patterns characteristic to the "chain" scenario diminished when considering some levels of omnivory.
+The exception is the chain scenario (no omnivory) with no disturbance and low productivity.
+In this scenario, ecosystem size had a positive effect on FCL (**Figure 2**), while ecosystem complexity had a negative influence (**Figure 3**).
+These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increased the total carrying capacity, whereas ecosystem complexity decreased it (**Figure S10**).
+However, these patterns characteristic of the chain scenario diminished when considering some levels of omnivory.
 
-State transitions shed lights on how FCL changes occurred.
-In the chain motif, the "additive" mechanism (i.e., the addition of higher trophic species) was always responsible for the changes in FCL.
-With disturbance, state transition occurred from "B" to "B + C" (low productivity) or "B + C" to "B + C + P" (high productivity) as ecosystem size or complexity increases.
+State transitions shed light on how FCL changes occurred.
+In the chain motif, an additive mechanism (i.e., the addition of higher trophic species) was always responsible for the changes in FCL.
+With disturbance, state transition occurs from "B" to "B + C" (low productivity) or "B + C" to "B + C + P" (high productivity) as ecosystem size or complexity increases.
 Exceptionally, the state transition of "B + C + P" $\rightarrow$ "B + C" occurred as ecosystem complexity increased with no disturbance and low productivity, a scenario in which ecosystem complexity decreased FCL.
@@
 In the weak and strong omnivory motifs, ecosystem size and complexity increased FCL through different mechanisms.
-Ecosystem size increased FCL primarily through the additive mechanism, facilitating the state transition of "B" $\rightarrow$ "B + P" (low productivity) or "No species" $\rightarrow$ "B + P" (high productivity), particularly with frequent disturbance (Figure XXX).
-In contrast, the insertion (i.e., the insertion of intermediate trophic species) and additive mechanisms were responsible for the effect of ecosystem complexity on FCL, depending on simulation scenarios.
-With weak omnivory, the transition from "B + P" to "B + C + P" (insertion) was predominant when productivity was low (Figure XXX).
-However, the additive mechanism ("No species" $\rightarrow$ "B + P") seems to be dominant in the high productivity scenario (Figure XXX).
+Ecosystem size increased FCL primarily through additive mechanisms, facilitating the state transition of "B" $\rightarrow$ "B + P" (low productivity) or "No species" $\rightarrow$ "B + P" (high productivity), particularly with frequent disturbance (**Figure S5**).
+In contrast, insertion (i.e., the insertion of intermediate trophic species) and additive mechanisms were responsible for the effect of ecosystem complexity.
+With weak omnivory, the transition from "B + P" to "B + C + P" (insertion) was predominant when productivity was low (**Figure S6B**).
+However, the additive mechanism ("No species" $\rightarrow$ "B + P") took primacy in the high productivity scenario (**Figure S6B**).
 In the strong omnivory motif, transition patterns depended on disturbance frequency.
-When disturbance was rare ($p_m = 0.05$), the insertion mechanism ("B + P" $\rightarrow$ "B + C + P") drove the FCL change (Figure XXX).
-As disturbance frequency increased, the primary process switched to the additive mechanism ("B" $\rightarrow$ "B + P" or "No species" $\rightarrow$ "B + P"; Figure XXX).
+When disturbance was rare ($p_m = 0.05$), the insertion mechanism ("B + P" $\rightarrow$ "B + C + P") drove the FCL change (**Figure S6C**).
+As disturbance frequency increased, the primary process switched to an additive mechanism ("B" $\rightarrow$ "B + P" or "No species" $\rightarrow$ "B + P"; **Figure S6C**).
 
@@
 High productivity strengthened the relationship between FCL and ecosystem geometry.
-In the chain motif, productivity greatly increased the overall FCL (Figure XXX), thereby shifting the upper bound of FCL (**Figures 2 and 3**).
-In contrast, increased productivity greatly decreased the minimum levels of FCL with weak and strong omnivory with little influence on the maximum FCL (**Figures 2, 3, and XXX**).
+In the chain motif, productivity greatly increased the overall FCL (**Figure S7**), thereby shifting the upper bound of FCL (**Figures 2 and 3**).
+In contrast, increased productivity greatly decreased the minimum levels of FCL with weak and strong omnivory with little influence on the maximum FCL (**Figures 2, 3, and S7**).
 This negative effect of increased productivity was apparent in smaller and/or simpler ecosystems, steepening the slopes between FCL and ecosystem geometry (**Figures 2 and 3**).
-Such patterns were particularly clear with long-distance dispersal (**Figures 2 and 3**).
 
 State transitions associated with increased productivity depended on the motif structure.
-In the chain motif, the transition from "B + C" to "B + C + P" (i.e., additive) was dominant regardless of disturbance frequency.
-The introduction of omnivory allowed the transition between "B + P" and "B + C + P", and the primary direction of the transition differed between weak and strong omnivory.
-With weak omnivory, the transition of "B + C + P" $\rightarrow$ "B + P" was prevalent and shortened FCL regardless of disturbance frequency (Figure XXX).
+In the chain motif, the transition from "B + C" to "B + C + P" was dominant regardless of disturbance frequency (**Figure S8**).
+The introduction of omnivory allowed the transition between "B + P" and "B + C + P".
+With weak omnivory, the transition of "B + C + P" $\rightarrow$ "B + P" was prevalent and shortened FCL across disturbance regimes (**Figure S8**).
 However, this transition was marginal in the strong omnivory motif.
-Instead, productivity significantly increased the proportion of "No species" state (Figure XXX) in disturbed scenarios, especially in small and/or simple ecosystems (Figure XXX).
+Instead, productivity significantly increased the proportion of the "No species" state (**Figure S8**) in disturbed scenarios, especially in small and/or simple ecosystems (**Figures S5 and S6**).
 
@@
 
-Dispersal distance had a positive influence on FCL, regardless of simulation scenarios. However, this direct effect was rather weak in comparison to other factors.
+Dispersal distance had a positive influence on FCL regardless of simulation scenarios (**Figure S7**). However, this direct effect was weak in comparison to other factors.
 
-State transitions occurred through either additive (chain) or insertion mechanisms (omnivory).
-In the chain motif, the transitions were either "B" $\rightarrow$ "B + C" or "B + C" $\rightarrow$ "B + C+ P", depending on disturbance frequency.
-In the omnivory motifs, the predominant transition was from "B + P" to "B + C + P."
+State transitions occurred through an additive (chain) or insertion mechanism (omnivory).
+In the chain motif, the transitions were either "B" $\rightarrow$ "B + C" or "B + C" $\rightarrow$ "B + C+ P", depending on disturbance frequency (**Figure S9**).
+In the omnivory motifs, the predominant transition was from "B + P" to "B + C + P" (**Figure S9**).
 
@@
 Although we focus on river networks here, we expect our results to be transferable to other systems with non-branching structures.
-Ecosystems are intrinsically complex -- forest stands possess complex vertical structures, which may provide heterogeneous habitats with different levels of vulnerability to wind and/or fire disturbance.
-Topographic variation can also mediate spatial disturbance dynamics by providing local refugia in a landscape.
-An explicit consideration of ecosystem complexity may therefore increase our ability to predict FCL.
+Ecosystems are intrinsically complex -- forest stands possess complex vertical and horizontal structures, which may provide heterogeneous habitats with different levels of vulnerability to wind and/or fire disturbances [@canslerClimateFireSize2014; @petersonContagiousDisturbanceEcological2002; @rommeHistoricalPerspectiveYellowstone1989].
+Topographic variation can also mediate spatial disturbance dynamics by providing local refugia in a landscape [@connell30YearStudy1997].
+Explicit consideration of ecosystem complexity may therefore increase our ability to predict FCL.
 
 The most significant finding in this study is the synergistic effects between ecosystem geometry and disturbance.
-Our no disturbance scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independently of ecosystem size and complexity (except the chain motif with low productivity).
-However, as the frequency of intense disturbance increases, small and/or simple networks became unable to maintain long food chains.
+Our no-disturbance scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independently of ecosystem size and complexity (except the chain motif with low productivity).
+However, as the frequency of intense disturbance increases, small and/or simple networks become unable to maintain long food chains.
 Here, increased ecosystem size and complexity may give greater resistance to disturbance through different mechanisms.
-Ecosystem size simply increases the "total amount" of disturbance refugia, from which dispersing propagules recolonize impacted habitat patches.
+Ecosystem size increases the "total amount" of disturbance refugia, from which dispersing propagules recolonize impacted habitat patches.
 In the meantime, ecosystem complexity may increase the connectivity to disturbance refugia.
-Indeed, branching shortens the mean distance between patches and therefore may enhance immigration from less disturbed patches.
-It is possible that the proportion of disturbance refugia increased in more branched networks.
-However, we did not observe a clear proportional increase in disturbance refugia with increasing branching probability.
-Collectively, each of geometric features likely supports long food chains through distinct mechanisms, which may also explain the different patterns of state transitions (additive vs.
+Branching shortens the mean distance between patches and, therefore, may enhance immigration from less disturbed patches (**Figure S10**).
+The proportion of disturbance refugia may have increased in more branched networks.
+However, we did not observe a proportional increase in disturbance refugia with increasing branching probability.
+Collectively, each of the geometric features likely supports long food chains through distinct mechanisms, which may also explain the different patterns of state transitions (additive vs.
 insertion mechanisms) [@postProximateStructuralMechanisms2007].
@@
 
-The synergy between ecosystem geometry and disturbance has also an important implication for why the dynamical stability hypothesis has been least supported by the literature [@takimotoEnvironmentalDeterminantsFoodchain2013].
+The synergy between ecosystem geometry and disturbance also has an important implication for why the disturbance hypothesis has been least supported by the literature [@takimotoEnvironmentalDeterminantsFoodchain2013].
 Although increased disturbance frequency consistently reduced FCL, its magnitude depended strongly on ecosystem geometry.
-This observation should provide a possible explanation for the limited support for the dynamical stability hypothesis in field studies.
+This observation should provide a possible explanation for the mixed effects of disturbance in field studies.
 Although FCL research often assumes that the effects of environmental controls are independent, exploring potential interactions among them could provide useful insights into environmental controls of FCL in nature.
@@
 Therefore, the preservation of habitat diversity conferred by ecosystem size and complexity is critical to biodiversity conservation under rapid environmental changes.
-In addition, maintaining habitat connectivity is important because the effect of spatial refuges can emerge only through recolonization after disturbance.
+In addition, maintaining dispersal corridors is important because the effect of spatial refuges can emerge only through recolonization after disturbance.
 Dams and road crossings have altered the connectivity between suitable habitat patches or have completely blocked dispersal [@ishiyamaPredictingEcologicalImpacts2018; @moritaEffectsHabitatFragmentation2002; @perkinFragmentationAltersStream2012].
@@
 Another interesting result of our simulation is the interaction with omnivory.
-In particular, with no disturbance and low productivity, FCL of the chain motif had contrasting relationships with ecosystem size (positive) and complexity (negative).
-It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity; therefore, FCL followed the changes in the total carrying capacity along the gradients of ecosystem size and complexity.
-However, these relationships diminished in omnivory motifs as the dependence of FCL on local carrying capacity decreased (**Figure XXX**).
-Importantly, however, disturbance overwhelmed the effect of omnivory on the FCL-geometry relationships: as the frequency of intense disturbance increases, the positive relationships between FCL and ecosystem geometry emerged regardless of omnivory levels.
+In particular, with no disturbance and low productivity, the FCL of the chain motif had contrasting relationships with ecosystem size (positive) and complexity (negative).
+It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity (**Figures S1 and S2**); therefore, FCL followed the changes in the total carrying capacity along the gradients of ecosystem size and complexity.
+These relationships diminished in omnivory motifs as the dependence of FCL on local carrying capacity weakened (**Figures S1 and S2**).
+Importantly, however, disturbance overwhelmed the effect of omnivory on the FCL-geometry relationships: as the frequency of intense disturbance increases, the positive relationships between FCL and ecosystem geometry emerge regardless of omnivory levels.
 This result provides a possible explanation for why the ecosystem size hypothesis has been consistently supported in various ecosystems [@takimotoEnvironmentalDeterminantsFoodchain2013] and raises the possibility of ecosystem complexity as an important control of FCL in branching networks.
-Our results also imply that, if disturbance is strong and frequent, the effects of ecosystem geometry are robust to food web reorganization, which is often associated with species extirpations or introductions.
+Our results also imply that if disturbance is strong and frequent, the effects of ecosystem geometry are robust to food web reorganization, which is often associated with species extirpations or introductions.
 The strength of omnivory in these novel communities can be increased or decreased, depending on the introduced species and the extant species pool or community.
 For example, non-indigenous fishes in the Eastern Mediterranean Sea outcompete native species for high-quality prey items, causing native species to increase their consumption of non-preferred prey items [@fanelliDepictingNovelEastern2015].
-However, such substantial food web reorganization may have limited influences on how ecosystem geometry dictates FCL.
+However, such substantial food web reorganization may not alter qualitative relationships between FCL and ecosystem geometry.
 
 The effect of productivity was motif-dependent, supporting previous theoretical predictions [@postProximateStructuralMechanisms2007; @kondohFoodchainLengthAdaptive2009; @wardMechanisticTheoryAquatic2017].
-While productivity clearly increased FCL in the chain motif through an additive mechanism, the productivity effect was vague (strong omnivory) or even reversed (weak omnivory) in omnivory motifs.
+While productivity clearly increased FCL in the chain motif, the productivity effect was negative in omnivory motifs.
 Intraguild predation plays an important role in explaining this context dependency.
 In the high productivity scenario with omnivory, the increased abundance of the basal species enhances the intraguild predator.
-This may subsequently facilitate the local exclusion of the intraguild prey through increased predation pressure, if the attack rate is sufficiently high [@postProximateStructuralMechanisms2007; @kondohFoodchainLengthAdaptive2009; @wardMechanisticTheoryAquatic2017].
+This may subsequently facilitate the local exclusion of the intraguild prey through increased predation pressure if the attack rate is sufficiently high [@postProximateStructuralMechanisms2007].
 Indeed, in the weak omnivory motif (the high value of the attack rate of P on C), we observed the state transition from "B + C + P" to "B + P" with shortened FCL regardless of disturbance frequency.
-However, this was not the case in the strong omnivory motif; the increased proportion of "No species" state had a strong impact on the minimum FCL, and this transition was mediated by disturbance.
-This result suggests that, although the strong omnivory motif is innately stable with high productivity (see one-patch simulations; Figure XXX), external disturbance undermines the sensitive balance of predator-prey interactions with the abrupt increase of "No species" state, which must be countered by immigration from less disturbed habitats to maintain food chains.
+However, this was not the case in the strong omnivory motif; the increased proportion of the "No species" state strongly impacted the minimum FCL, and this transition was mediated by disturbance.
+This result suggests that, although the strong omnivory motif is innately stable with high productivity (see one-patch simulations; **Figures S1 and S2**), external disturbance undermines the sensitive balance of predator-prey interactions with the abrupt increase of "No species" state, which must be countered by immigration from less disturbed habitats to maintain food chains.
 
-It is therefore probable that different processes underlie the productivity-driven enforcement of the FCL-geometry relationships among food web motifs.
-In the chain and strong omnivory motifs, increased productivity magnified the disproportionate impact of disturbance on small and/or simple ecosystems because they seem less stable with limited supply of immigrants and external perturbations.
-With the weak omnivory motif, productivity facilitated the spatial spread of the intraguild predator in small ecosystems, where dispersing propagules can reach wider areas of a network with local extirpation of the intraguild prey.
+It is therefore probable that different processes underlie the productivity-driven enforcement of the FCL-geometry relationships even within omnivory motifs.
+In the strong omnivory motif, increased productivity magnified the disproportionate impact of disturbance on small and/or simple ecosystems because the motif seems highly susceptible to disturbance with a limited supply of immigrants.
+With the weak omnivory motif, productivity facilitated the spatial spread of the intraguild predator in small ecosystems, where dispersing propagules can reach most areas of a network with local extirpation of the intraguild prey.
 
-Our findings add an important distinction to work to assess and prioritize protection or conservation status within human impacted regions.
+Our findings add an important distinction to work to assess and prioritize protection or conservation status within human-impacted regions.
 Classical conservation schemes emphasize ecosystem size, aiming to protect the largest areas possible [@diamondIslandDilemmaLessons1975].

### 2023-03-09 rmd/figures.Rmd
edit: figure caption
edit: manuscript

--- 2022-03-29 rmd/figures.Rmd
@@
 ---
-title: "Figures"
 output: pdf_document
+knit: (function(inputFile, encoding) {
+        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
+      })
 ---
@@
                       message = FALSE)
+
+options(knitr.kable.NA = '')
 ```
 
+# Tables
+
+```{r}
+source(here::here("code/table_param.R"))
+
+knitr::kable(df_value %>%
+               dplyr::select(-"One-patch") %>% 
+               rename(Value = "Main"),
+             escape = FALSE,
+             booktabs = TRUE,
+             caption = "Parameter values used in the main simulation") %>% 
+  kableExtra::kable_styling(latex_options = 'HOLD_position')
+```
+
+\pagebreak
+
+# Figures
+
 ```{r figure1, fig.width=12, fig.height=6}
-source(here::here("code/figure_concept.R"))
-print(g_concept)
+knitr::include_graphics(here::here("figure/figure_concept.pdf"))
 ```
@@
 Node color is reflective of the disturbance value calculated in network generation.
-(B) Distribution of disturbance values in branching networks.
+(B) Density distributions of disturbance intensity (i.e., proportional reduction of a population).
+Three networks of 50 patches were produced with varied branching probabilities, which were shown in different colors.
+The vertical line indicates the mean disturbance intensity at headwaters ($\mu_m = 0.8$).
 (C) Trophic interactions in a three-species community.
@@
 The structure and strength of these trophic interactions determined by functional responses.
-The strength of trophic interactions involving $P$ is represented by the darkness of the arrows (strength of the $B$-$C$ interaction is fixed), and food chain length varies in response to the proportion that each resource makes to the diet of $P$.
 On the far left, a simple linear food chain exists when the predator does not consume the basal species.
 Moving to the right, basal species' biomass is converted to predator biomass at higher efficiencies with greater attack rates.
-(D) Preference of predator $P$ to basal species $B$ over consumer $C$ under different omnivory scenarios.
-The preference changes with the relative density of $B$ to $C$ ($C=50$ in this example), and the rate of change is determined by attack rate and handling time.
+(D) Proportional contribution of basal species $B$ to predator $P$ under different omnivory scenarios.
+The contribution changes with the relative density of $B$ to $C$ ($C=50$ in this example), and the rate of change is determined by attack rate and handling time.
 
@@
 ```{r figure2, fig.width=12, fig.height=5.5}
-source(here::here("code/figure_main_sim.R"))
-print(g_ld)
+knitr::include_graphics(here::here("figure/figure_main_n_patch.pdf"))
 ```
 
-**Figure 2** Ecosystem size (A) and complexity (B) have little influence on food chain length under low disturbance regime.
-Lines are loess curves fitted to simulated data and are colored based on omnivory.
-Shaded ribbons are 95% confidence intervals.
-Parameters for low disturbance regime: disturbance probability $p_d = 0.01$ and mean proportional mortality at headwaters $\mu_{m}=0.8$.
+**Figure 2** Relationships between food chain length and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
+Lines indicate loess curves, and shades are associated 95% confidence intervals.
+Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
+Different line colors indicate different disturbance probabilities.
 
@@
 ```{r figure3, fig.width=12, fig.height=5.5}
-source(here::here("code/figure_main_sim.R"))
-print(g_hd)
+knitr::include_graphics(here::here("figure/figure_main_p_branch.pdf"))
 ```
 
-**Figure 3** Ecosystem size (A) and complexity (B) influence food chain length under high disturbance regime.
-Lines are loess curves fitted to simulated data and are colored based on omnivory.
-Shaded ribbons are 95% confidence intervals.
-Parameters for high disturbance regime: disturbance probability $p_{d} = 0.01$ and mean proportional mortality at headwaters $\mu_{m}=0.8$.
+**Figure 3** Relationships between food chain length and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
+Lines indicate loess curves, and shades are associated 95% confidence intervals.
+Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
+Different line colors indicate different disturbance probabilities.

### 2023-03-09 rmd/manuscript.Rmd
edit: manuscript

@@
 However, these effects were contingent upon other drivers.
-In particular, the effects of ecosystem size and complexity appeared more clearly in high-disturbance regimes.
+In particular, the effects of ecosystem size and complexity appeared more clearly in high-disturbance and high-productivity regimes.
 Our results suggest that ecosystem complexity is an important yet overlooked driver of food chain length in branching ecosystems.
@@
 For example, large ecosystems have increased species richness (species-area relationship [@lomolinoEcologyMostGeneral2000]), are buffered against disturbances [@saboRoleDischargeVariation2010], and have greater numbers of immigrant sources in the ecosystem with stable metapopulations [@takimotoEnvironmentalDeterminantsFoodchain2013].
-Meanwhile, the productivity and disturbance effects are highly sensitive to the internal structure of food web motifs.
+Meanwhile, the productivity and disturbance effects are highly sensitive to the internal structure of food webs [@postProximateStructuralMechanisms2007; @takimotoEffectsProductivityDisturbance2012].
 As such, the effects of disturbance and resource availability on FCL have been found to be highly variable among ecosystem types and geographic regions [@takimotoEffectsProductivityDisturbance2012; @diehlIntraguildPreySuffer2001; @takimotoEnvironmentalDeterminantsFoodchain2013; @postLongShortFoodchain2002].
@@
 For dispersal, we examined short- and long-distance dispersal scenarios (short-distance dispersal $\theta=$ `r max(df_sim$theta)`; long-distance dispersal $\theta=$ `r min(df_sim$theta)`) with constant dispersal probability ($p_{d}=$ `r unique(df_sim$p_dispersal)`).
+Parameter values were summarized in **Table 1**.
 
@@
 Branching shortens the mean distance between patches and, therefore, may enhance immigration from less disturbed patches (**Figure S10**).
-The proportion of disturbance refugia may have increased in more branched networks.
-However, we did not observe a proportional increase in disturbance refugia with increasing branching probability.
+The proportion of disturbance refugia may also have increased in more branched networks.
+However, we did not observe a clear proportional increase in disturbance refugia with increasing branching probability (**Figure 1B**).
 Collectively, each of the geometric features likely supports long food chains through distinct mechanisms, which may also explain the different patterns of state transitions (additive vs.

### 2023-03-09 rmd/reply1.Rmd
edit: manuscript

@@
 
-> Thank you for your careful review - your review improved our manuscript substantially. We followed your advise -- please see our response below.
+> Thank you for your careful review - this reviewer's comments improved our manuscript substantially. We followed your advise -- please see our response below.
 
@@
 
-> This is a nice comment! Yes, we have clarified this point in the revised manuscript by adding a paragraph in the discussion (XXX). Also, we changed the language in Results (XXX).
+> This is a nice comment! Yes, we have clarified this point in the revised manuscript by adding a paragraph in Discussion (3rd paragraph). Also, we changed the language in Results.
 
@@
 We simulated a wider range of $p_m$ values to see how disturbance frequency changes FCL-geometry relationships (please see revised figures).
-We also included a different level of disturbance intensity ($\mu_m = 0.2$) - the lower intensity weakened the interactive effect of disturbance frequency.
-We noted this result in the revised MS (XXX).
+We also included a different level of disturbance intensity ($\mu_m = 0.2$) - the lower intensity removed the interactive effect of disturbance frequency.
+We noted this result in the revised manuscript (Results, section *Interactive effect of ecosystem geometry and disturbance*).
 
@@
 
-> XXX
+> Thank you for your comment.
+We have given a deep thought on this.
+After the re-analysis, our interpretation is the following: ecosystem size increases the total "amount" of disturbance refuges, while branching increases the "connectivity" to refuges; branching shortens mean distance between a given pair of patches (F**igure S10**).
+We also looked at the proportional distribution of disturbance intensity within a network -- however, branching did not increase the proportional amount of refuges substantially (some increases in the left tail, but marginal; note we changed parameters of $\sigma_h$ and $\sigma_l$ in the revised MS).
+Therefore, we believe that increased connectivity, which facilitates the recolonization from less disturbed patches, was the primary mechanism that increased the resistance to disturbance.
+>
+> In terms of the frequency, the "disturbance intensity" refers to the proportional reduction of a population if a regional disturbance occurs (all habitat patches are affected) - therefore, the frequency does not change the relative frequency of disturbance intensity (the temporal mean of disturbance intensity at each patch is fixed).
+Thank you for your insights though!
 
@@
 > Thank you for this very important comment!
-We have completely rewrote the model as an extension of the Nicholson-Bailey model.
-Important results were similar, but there were some qualitative changes as well.
+We have fully rewrote the model as an extension of the Nicholson-Bailey model, following Hassel 2000 and Weide et al.
+2019.
+Important results were similar.
 We have fully re-structured the Method section based on this revision.
@@
 
-> Thank you, we have corrected this point as well.
+> Thank you, we have corrected this point as well. We followed the guidance in Hassel 2000.
 
@@
 
-> We appreciate your insights. We revisited the switching function to properly represent this in a discrete form.
+> We appreciate your insights. We revisited the switching function to properly represent this in a discrete form (used a function in Hassel 2000; please see the main text).
 
@@
 
-> The reviewer is correct. We have address this point.
+> The reviewer is correct. We have addressed this point.
 
@@
 
-> Thank you, we have revised as follows: "...theoretical conceptualization has implicitly assumed two-dimensional systems, which may be a fair approximation of lakes."
+> Thank you, we have revised as follows: "...theoretical conceptualization has implicitly assumed two-dimensional systems, which may be a fair approximation of lakes and..."
 
@@
 > Thank you for raising this point.
-Ecosystem size increases the total carrying capacity, while ecosystem complexity decreases it (branching disproportionately increases the number of headwaters with low carrying capacities).
-Our additional analysis indicated that the "chain" motif is particularly influenced by local carrying capacity; in the meantime, omnivory scenarios were insensitive to local carrying capacity.
+Ecosystem size increases the total carrying capacity, while ecosystem complexity decreases it (branching disproportionately increases the number of headwaters with low carrying capacities; see **Figure S10**).
+Our additional analysis indicated that the "chain" motif is particularly influenced by local carrying capacity; in the meantime, omnivory scenarios were less sensitive to local carrying capacity.
 We have clarified these points in Results and Discussion.
@@
 
-> XXX
+> In the revised MS, we used a combination of $r_b$ and $K_{base}$ as a proxy for productivity.
 
@@
 
-> XXX
+> Our unclear explanation seems to cause confusion - disturbance intensity was expressed as the proportional reduction of a population.
+So, it ranges from 0 to 1.
+The vertical line is the mean disturbance intensity at headwaters.
+We have revised the figure legends carefully.
 
@@
 
-> In light of comments from Reviewer 1, we have reanalyzed and fully re-structured Results and Discussion according to the new results.
-The above statement does not hold in the revised MS, so they have been removed.
+> In light of comments from Reviewer 1, we have reanalyzed and fully re-structured Results and Discussion according to the new results. Please see our response to your next comment for details.
 
@@
 Also, as P grows it reduces the amount of B consumed by C (which should be increasingly dynamically unstable)?
+
+> We agree.
+We have changed the figures to better visualize "state transitions" (e.g., from "B + C + P" to "B + P").
+As you suggested, we did observed the exclusion of C with high productivity, but only with the weak omnivory motif.
+This is probably because in this motif, the contribution of B to P is sufficiently high yet maintains strong predation impact on C.
+We have included this statement in Results (see *Productivity effect*) and expanded the discussion.
 
@@
 
+> We appreciate your insight.
+In the revised MS, we used a combination of $r_b$ and carrying capacity to represent low and high productivity scenarios.
+Indeed.
+it has greatly changed the results.
+Please see the revised results in the main text.
+
 I think understanding these results associated with branching ecosystems would benefit from a clear description of results and discussion for non-branching networks and then connect the results to the branching complexity ideas.
 What is happening in the branching network model that governs the results and is providing a different (or similar) response as with the other factors explored here (size, productivity etc.) ?
+
+> Agreed, it will certainly help improve the understanding of how branching affects FCL.
+However, we are afraid that the inclusion of this aspect is beyond the scope of this manuscript.
+The current manuscript contains rich amount of results, and adding a new component would make the manuscript less accessible.
+We appreciate your input.
 
@@
 
+> True. We have removed this statement in the revised MS. Although it might be possible to quantify effect sizes and compare them, this result is not a significant portion of the manuscript.
+
 One of the results that wasn't discussed is that the changes in FCL in several cases occurred at low levels of ecosystem complexity.
@@
 
+> XXX
+
 "Productivity did not modulate the relationship between FCL and ecosystem size and ecosystem complexity." -- I see difference between high and low productivity simulations for the strong omnivory particularly in the low disturbance case.
@@
 Can you explain why these differences occur?
+
+> In the re-analysis, we did observe stronger effects of productivity (see *Productivity effect* in Results). We have fully restructured it and expended discussion according to this change.
 
@@
 
+> We are glad the reviewer picked it!
+Although the darkness was supposed to represent the strength of trophic interactions, we removed this because it might cause confusion.
+This figure is just a conceptual visualization and does not quantitatively represent simulation setups.
+
 Line 255: "under high disturbance regimes, only large and complex networks could provide diverse and heterogenous habitats...and maintain long food chains".
@@
 
+> Thank you. We have reworded. Smaller/simpler ecosystems were more vulnerable to disturbance.
+
 Line 254: "Consequently food web structure largely controlled FCL." This is a strange way to describe what is happening in the low disturbance systems as FCL is food web structure.
@@
 
+> We have removed this part of text during the current revision.
+
 Line 276: "An important result in our simulation is that omnivory does not change the qualitative relationships between FCL and ecosystem properties, as opposed to productivity effects." I find this sentence to
+
+> The comment ended in the middle, so we were not able to correctly understand the meaning of this.
+However, we suspect the referenced sentence needs to be improved.
+We have fully restructured this part in the revised MS because ecosystem geometry effects were now dependent on omnivory.
 
@@
 
+> Thank you, we have used another term "ecosystem geometry" (size and complexity) and clearly defined in Introduction. This eliminated the confusion between ecosystem vs. food web structure.
+
 Line 45: Ecosystem Size?
 
+> Corrected.
+
 Line 194: "are differ". Suggest remove "are".
+
+> Corrected.

### 2023-03-09 rmd/suppl_info.Rmd
edit: manuscript

@@
 Consequently, we had a total of `r nrow(dplyr::distinct(df_sim1, r_b, k, a1, a2, a3, h))` combinations (`r nrow(dplyr::distinct(df_sim1, r_b, k, a1, h))` $\times$ `r nrow(dplyr::distinct(df_sim1, a2, a3))`) for one-patch simulations.
-Note that we set the switching parmeter $s = 0$ if $a_{BP} = 0$, otherwise $s = 1$.
+Note that we set the switching parameter $s = 0$ if $a_{BP} = 0$; otherwise $s = 1$.
 Parameter values were summarized in **Table S1**.
 
-For each parameter combination, we ran 10 replications of `r n_sim[1]` time steps, which include `r n_sim[2]` warmup and `r n_sim[3]` burn-in periods.
-We recorded the patch state as numeric values ("No species" = 0, "B" = 1, "B + C" = 2, "B + P" =3, and "B + C + P" = 4) at $t = 1600$ and retrieved the maximum value of 10 replicates.
+For each parameter combination, we ran ten replications of `r n_sim[1]` time steps, which include `r n_sim[2]` warmup and `r n_sim[3]` burn-in periods.
+We recorded the patch state as numeric values ("No species" = 0, "B" = 1, "B + C" = 2, "B + P" =3, and "B + C + P" = 4) at $t = 1600$ and retrieved the maximum value of ten replicates.
 
@@
 We identified the following three parameter combinations that could maintain the state of "B + C + P" with distinct food web motifs: "chain (no omnivory)" $(a_{BP}, a_{CP}, h) = (0, 0.0025, 0.5)$, "weak omnivory" ($0.02, 0.025, 0.5$), and "strong omnivory" ($0.04, 0.0025, 0.75$) (see symbols in **Figures S1 and S2**).
-In the maintext, we reported the results with $a_{BC} = 0.5$ because qualitative patterns did not depend on $a_{BC}$.
+In the main text, we reported the results with $a_{BC} = 0.5$ because qualitative patterns did not depend on $a_{BC}$.
 
@@
       booktabs = TRUE,
-      caption = "Parameter values used in the main simulation and one-patch simulation") %>% 
+      caption = "Parameter values used in the main and one-patch simulation") %>% 
   kableExtra::kable_styling(latex_options = 'HOLD_position')
@@
 Colors indicate the patch state, and points with different symbols identify parameter combinations used in the main simulation.
-Panels distinguishh the value of carrying capacity $K$.
+Panels distinguish the value of carrying capacity $K$.
 In both panels, rows represent different values of intrinsic growth rate $r_b$, and columns identify handling time $h$."}
@@
 ```{r fig3, fig.cap="Relationships between food chain length and ecosystem size with low disturbance intensity $\\mu_m = 0.2$.
-Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios."}
+Lines indicate loess curves, and shades are associated 95% confidence intervals.
+Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
+Different line colors indicate different disturbance probabilities."}
 knitr::include_graphics(here::here("figure/figure_si_n_patch.pdf"))

### 2023-03-09 code/figure_delta.R

@@ -14,12 +14,12 @@ m_x <- rbind(seq(0, 100, length = 100),
              rep(10, 100))
 
 ## simulation scenarios
-df_param <- expand.grid(e = 1, # to conv_eff
-                        a_bc = 0.25, # to attack_rate
-                        a_bp = c(0, 0.02, 0.04), # to attack_rate
-                        a_cp = 0.0025, # to attack_rate
-                        h = 0.5, # to handling_time
-                        s = 1) %>% 
+df_param <- tibble(e = 1, # to conv_eff
+                   a_bc = 0.5, # to attack_rate
+                   a_bp = c(0, 0.02, 0.04), # to attack_rate
+                   a_cp = c(0.0025, 0.025, 0.0025), # to attack_rate
+                   h = c(0.5, 0.5, 0.75), # to handling_time
+                   s = c(0, 1, 1)) %>% 
   as_tibble() %>% 
   mutate(omn = case_when(a_bp == 0 ~ "Chain",
                          a_bp == 0.02 ~ "Weak",

### 2023-03-09 code/figure_food_web.R

@@ -56,9 +56,24 @@ g_fw <- foreach(i = 1:3) %do% {
                    start_cap = circle(0.07, 'npc'),
                    end_cap = circle(0.07, 'npc'),
                    show.legend = T) +
-    scale_edge_color_continuous(low = "lightgrey",
-                                high = "black",
-                                guide = "none") +
+    geom_node_label(aes(label = label), 
+                    color = grey(0.25),
+                    size = 6,
+                    label.size = 0,
+                    fill = NA) +
+    labs(subtitle = title_fw[i]) +
+    theme(rect = element_blank(),
+          panel.border = element_blank(),
+          panel.background = element_blank())
+  
+  g <- ggraph(layout) + 
+    geom_edge_link(arrow = arrow(type = "closed",
+                                 ends = "last",
+                                 length = unit(0.05, "npc"),
+                                 angle = 15),
+                   start_cap = circle(0.07, 'npc'),
+                   end_cap = circle(0.07, 'npc'),
+                   show.legend = T) +
     geom_node_label(aes(label = label), 
                     color = grey(0.25),
                     size = 6,

### 2023-03-09 code/figure_network.R

@@ -64,22 +64,34 @@ g_net <- (ng[[1]] + labs(title = "A")) + ng[[2]] + ng[[3]] + plot_layout(guide =
 
 theme_set(plt_theme)
 
-df_wn <- readRDS("output/sim_within_net.rds")
+p_branch <- c(0.2, 0.5, 0.8)
+
+df_wn <- foreach(i = seq_len(length(p_branch)),
+                 .combine = bind_rows) %do% {
+                   
+                   set.seed(11)
+                   
+                   net <- brnet(p_branch = p_branch[i])
+                   net$df_patch %>% 
+                     mutate(n_patch = 50,
+                            p_branch = p_branch[i]) %>% 
+                     return()
+                 }
+
 
 labs <- c(`0.2` = "p[b]==0.2",
           `0.5` = "p[b]==0.5",
           `0.8` = "p[b]==0.8")
 
-df_wn %>% 
-  ggplot(aes(y = disturbance,
-             x = factor(n_patch),
+g_disturb <- df_wn %>% 
+  ggplot(aes(x = disturbance,
              color = factor(p_branch),
              fill = factor(p_branch))) +
-  geom_violin(alpha = 0.2, draw_quantiles = 0.5) +
-  facet_wrap(facets = ~p_branch,
-             labeller = labeller(p_branch = as_labeller(labs,
-                                                        label_parsed))) +
-  labs(x = "Ecosystem size (number of patches)",
-       y = "Disturbance intensity (m)",
+  geom_density(alpha = 0.2) +
+  labs(y = "Density",
+       x = "Disturbance intensity",
        color = "Branching prob.",
-       fill = "Branching prob.")
+       fill = "Branching prob.") +
+  geom_vline(xintercept = 0.8,
+             color = "grey",
+             linetype = "dashed")

### 2023-03-09 rmd/manuscript.Rmd
edit rmd

@@
 output: 
-  pdf_document:
-    latex_engine: xelatex
+  word_document:
+    reference_docx: "word_format.docx"
+  #pdf_document:
+  #  latex_engine: xelatex
 knit: (function(inputFile, encoding) {
@@
 Although we focus on river networks here, we expect our results to be transferable to other systems with non-branching structures.
-Ecosystems are intrinsically complex -- forest stands possess complex vertical and horizontal structures, which may provide heterogeneous habitats with different levels of vulnerability to wind and/or fire disturbances [@canslerClimateFireSize2014; @petersonContagiousDisturbanceEcological2002; @rommeHistoricalPerspectiveYellowstone1989].
-Topographic variation can also mediate spatial disturbance dynamics by providing local refugia in a landscape [@connell30YearStudy1997].
-Explicit consideration of ecosystem complexity may therefore increase our ability to predict FCL.
+Ecosystems are intrinsically complex -- forests possess complex vertical and horizontal structures, which may provide heterogeneous microhabitats with different levels of vulnerability to wind and/or fire disturbances [@canslerClimateFireSize2014; @petersonContagiousDisturbanceEcological2002; @rommeHistoricalPerspectiveYellowstone1989].
+Topographic and terrain variation can also mediate spatial disturbance patterns in a landscape [@connell30YearStudy1997; @stueveSpatialPatternsIce2007].
+Explicit consideration of ecosystem complexity may therefore increase our ability to predict FCL and complements spatial food web theory that has emerged over the past two decades [@holtFoodWebsSpace2002a; @mccannDynamicsSpatiallyCoupled2005; @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012; @andersonEffectsDispersalRiver2018; @teruiSpatialDisturbanceSynchrony2019].
 
@@
 insertion mechanisms) [@postProximateStructuralMechanisms2007].
+We must note that, however, the effects of ecosystem size and complexity were  limited to small to intermediate levels.
+In large or complex ecosystems, dispersal propagules may not be a limiting factor of FCL as the supply is plenty; therefore, productivity or omnivory more strongly regulated FCL (**Figures 2 and 3**) as predicted by previous theory [@takimotoEffectsProductivityDisturbance2012]
+
 While empirical studies rarely looked at the influence of ecosystem complexity on FCL, several studies provide some support for this theoretical prediction.
@@
 It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity (**Figures S1 and S2**); therefore, FCL followed the changes in the total carrying capacity along the gradients of ecosystem size and complexity.
-These relationships diminished in omnivory motifs as the dependence of FCL on local carrying capacity weakened (**Figures S1 and S2**).
-Importantly, however, disturbance overwhelmed the effect of omnivory on the FCL-geometry relationships: as the frequency of intense disturbance increases, the positive relationships between FCL and ecosystem geometry emerge regardless of omnivory levels.
-This result provides a possible explanation for why the ecosystem size hypothesis has been consistently supported in various ecosystems [@takimotoEnvironmentalDeterminantsFoodchain2013] and raises the possibility of ecosystem complexity as an important control of FCL in branching networks.
+These relationships diminished in omnivory motifs as the dependence of FCL on local carrying capacity weakened or reversed (**Figures S1 and S2**).
+Importantly, however, disturbance overwhelmed the role of omnivory in modulating the FCL-geometry relationships: as the frequency of intense disturbance increases, the positive relationships between FCL and ecosystem geometry emerged regardless of omnivory levels.
+This result could explain the consistent support for the ecosystem size hypothesis in various ecosystems [@takimotoEnvironmentalDeterminantsFoodchain2013] and raises the possibility of ecosystem complexity as an important control of FCL in branching networks.
 Our results also imply that if disturbance is strong and frequent, the effects of ecosystem geometry are robust to food web reorganization, which is often associated with species extirpations or introductions.
@@
 For example, non-indigenous fishes in the Eastern Mediterranean Sea outcompete native species for high-quality prey items, causing native species to increase their consumption of non-preferred prey items [@fanelliDepictingNovelEastern2015].
-However, such substantial food web reorganization may not alter qualitative relationships between FCL and ecosystem geometry.
+Yet, such substantial food web reorganization may not alter qualitative relationships between FCL and ecosystem geometry.
 
@@
 It is therefore probable that different processes underlie the productivity-driven enforcement of the FCL-geometry relationships even within omnivory motifs.
+With the weak omnivory motif, productivity facilitated the spatial spread of the intraguild predator in small ecosystems, where dispersing propagules can reach most areas of a network with local extirpation of the intraguild prey.
 In the strong omnivory motif, increased productivity magnified the disproportionate impact of disturbance on small and/or simple ecosystems because the motif seems highly susceptible to disturbance with a limited supply of immigrants.
-With the weak omnivory motif, productivity facilitated the spatial spread of the intraguild predator in small ecosystems, where dispersing propagules can reach most areas of a network with local extirpation of the intraguild prey.
 
@@
 While this is certainly a useful attribute generally, designating networks with a high degree of variability, or many small patches, can be more beneficial than a large, homogeneous area [@fahrigWhySeveralSmall2020].
-Accounting for ecosystem complexity is particularly suited to ecosystem architectures in which branches have variable environmental conditions [@koningNetworkGrassrootsReserves2020; @teruiEmergentDualScaling2021].
-As suggested by modern conservation initiatives, conservation and management activities should focus on activities that increase the habitat heterogeneity and connectivity of suitable habitat patches to increase resilience to predicted increases in disturbance frequency and magnitude [@schindlerPopulationDiversityPortfolio2010; @schindlerPortfolioConceptEcology2015; @andersonPortfolioConservationMetapopulations2015].
+In the context of river networks, our results suggest that branching architectures may guide the spatial design of conservation actions.
+For example, regional food webs may be better protected if small reserves are placed widely across different tributaries with unique environmental conditions [@koningNetworkGrassrootsReserves2020; @teruiEmergentDualScaling2021].
+This implication aligns with modern conservation initiatives, in which conservation and management should focus on activities that increase the habitat heterogeneity and connectivity of suitable habitat patches to increase resilience to predicted increases in disturbance frequency and magnitude [@schindlerPopulationDiversityPortfolio2010; @schindlerPortfolioConceptEcology2015; @andersonPortfolioConservationMetapopulations2015].
+Considering ecosystem complexity is useful especially when direct information for spatial environmental heterogeneity is lacking because ecosystem complexity, such as branching probability, can be estimated from widely-available GIS layers [@teruiMetapopulationStabilityBranching2018; @teruiEmergentDualScaling2021].
 

### 2023-03-09 rmd/reply1.Rmd
edit rmd

@@
 
+Dear Editor,
+
+This is a re-submission of a manuscript: ECS22-0646 "*Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems.*" Thank you for allowing us to revise our manuscript.
+The comments were very helpful in improving our manuscript.
+When revising the manuscript, we have put a particular emphasis on the following points:
+
+1.  We revisited our model and fully restructured our manuscript based on the new results.
+2.  We substantially expanded Results and Discussion based on reviewer's comments.
+3.  Where appropriate, we re-worded text throughout the manuscript.
+
+We appended a point-by-point response below (reviewer comments are shown in *italics*). We hope our manuscript is now suitable for publication in your journal.
+
+Sincerely,
+
+Akira Terui on behalf of the authors
+
 ## Response to Reviewer #1:
@@
 
-This ms uses a simulation model to study how environmental variables affect food-chain length (FCL) in river networks.
+*This ms uses a simulation model to study how environmental variables affect food-chain length (FCL) in river networks.
 Environmental determinants of FCL are a classic topic in ecology.
@@
 A novel contribution of the current ms is to consider another potential determinant of FCL, ecosystem complexity, in addition to the other three.
-The ms is focused on examining the following environmental variables: ecosystem size (the number of habitat patches in a whole river network), ecosystem complexity (the probability of branching, which is a model parameter controlling the proportion of habitat patches serving as confluences in a river network), disturbance frequency, and productivity (intrinsic growth rate of basal species).
+The ms is focused on examining the following environmental variables: ecosystem size (the number of habitat patches in a whole river network), ecosystem complexity (the probability of branching, which is a model parameter controlling the proportion of habitat patches serving as confluences in a river network), disturbance frequency, and productivity (intrinsic growth rate of basal species).*
 
-A significant finding of this ms is that ecosystem size and complexity can interact with the disturbance effect.
+*A significant finding of this ms is that ecosystem size and complexity can interact with the disturbance effect.
 This is because ecosystem size and complexity increase spatial heterogeneity in disturbance.
@@
 In large or complex ecosystems, heterogeneity in disturbance creates spatial refuges and weakens the negative effect of disturbance.
-The way that ecosystem size and complexity create disturbance heterogeneity is nicely modeled.
+The way that ecosystem size and complexity create disturbance heterogeneity is nicely modeled.*
 
@@
 
-The ms contains potentially important findings for FCL research but underlying modeling exercises are poor.
+*The ms contains potentially important findings for FCL research but underlying modeling exercises are poor.
 I have two main concerns.
@@
 Below, I illustrate major points.
-I will also list some minor points.
+I will also list some minor points.*
 
@@
 
-(1-1) Based on the obtained results, the ms emphasizes that ecosystem size and complexity have positive effects on FCL but their effects depend on disturbance.
+*(1-1) Based on the obtained results, the ms emphasizes that ecosystem size and complexity have positive effects on FCL but their effects depend on disturbance.
 This is an important message.
@@
 Although both messages mean the same literally, the latter could explain why empirical research has found variable disturbance effects on FCL.
-Since the present results may provide an answer, the ms could emphasize the latter message, too.
+Since the present results may provide an answer, the ms could emphasize the latter message, too.*
 
@@
 
-(1-2) Given that a major finding is the interacting effects between disturbance and ecosystem size/complexity, more in depth analysis on the interacting effects would greatly improve the ms.
+*(1-2) Given that a major finding is the interacting effects between disturbance and ecosystem size/complexity, more in depth analysis on the interacting effects would greatly improve the ms.
 In particular, it would be worthwhile to provide results illustrating how FCL responds along a disturbance gradient and how the responses differ between small vs.
@@
 In addition, while the current ms looks at disturbance frequencies (p_m), effects of disturbance intensities (mu_m) should also be explored because frequency and intensity are both important characteristics of disturbance.
-Because disturbance propagates through a river network, effects of frequency and intensity might be different, which could add interesting and useful insights.
+Because disturbance propagates through a river network, effects of frequency and intensity might be different, which could add interesting and useful insights.*
 
@@
 
-(1-3) Mechanisms underlying the interacting effects should also be explored in some more detail.
+*(1-3) Mechanisms underlying the interacting effects should also be explored in some more detail.
 A potential mechanism, as discussed in the ms, is greater spatial heterogeneity in disturbance in larger or more complex ecosystems.
@@
 1B must be compared between frequent and infrequent disturbance.
-I would expect that the occurrence of spatial refuges is more certain with more frequent disturbance (i.e., clearer longer tails on the low intensity side).
+I would expect that the occurrence of spatial refuges is more certain with more frequent disturbance (i.e., clearer longer tails on the low intensity side).*
 
@@
 
-(2-1) The way that trophic interactions are modeled is unnatural and probably wrong.
+*(2-1) The way that trophic interactions are modeled is unnatural and probably wrong.
 One of the problems is that the amount of a species being consumed can exceed its abundance.
@@
 (2019, <https://doi.org/10.1016/j.mbs.2018.12.010>).
-Another solution might be to model within-time step events (consumption and predation-induced death) as continuous time processes and between-time step events (reproduction and disturbance-induced death) as discrete time processes.
+Another solution might be to model within-time step events (consumption and predation-induced death) as continuous time processes and between-time step events (reproduction and disturbance-induced death) as discrete time processes.*
 
@@
 
-(2-2) Although it is claimed that type II functional responses are assumed, the formulation having predator abundance in the denominator and numerator (equation 9) is incorrect.
+*(2-2) Although it is claimed that type II functional responses are assumed, the formulation having predator abundance in the denominator and numerator (equation 9) is incorrect.
 The formulation of f_ij (line 144) is a correct formulation of type II responses.
-Moreover, once this error is corrected, it should be checked if beta\_(CP, x) is positive (equation 12).
+Moreover, once this error is corrected, it should be checked if beta\_(CP, x) is positive (equation 12).*
 
@@
 
-(2-3) The preference function for predators is introduced (equation 10).
+*(2-3) The preference function for predators is introduced (equation 10).
 The logic underlying this preference function (Nishijima et al.
-2004) would be that relative preference reflects relative gains of predators feeding exclusively on resource and consumer species (e_BP*f_BP vs.
-e_CP*f_CP).
-However, these resource gains do not closely translate into predator's population growth rate (equation 14), breaking the logic underlying the preference function.
+2004) would be that relative preference reflects relative gains of predators feeding exclusively on resource and consumer species (e_BPf_BP vs.
+e_CPf_CP).
+However, these resource gains do not closely translate into predator's population growth rate (equation 14), breaking the logic underlying the preference function.*
 
@@
 
-Line 39+: "the theoretical basis for the other two hypotheses (resource availability and disturbance hypotheses) rests on implicit assumptions of linear food chains." This seems incorrect.
-Classic literature on the two hypotheses have considered non-linear food webs.
+*Line 39+: "the theoretical basis for the other two hypotheses (resource availability and disturbance hypotheses) rests on implicit assumptions of linear food chains." This seems incorrect.
+Classic literature on the two hypotheses have considered non-linear food webs.*
 
@@
 
-Line 49+: "a maximum length of three": FCL may be defined as the number of trophic links or trophic levels. Here the ms messes up the two definitions. Please be consistent.
+*Line 49+: "a maximum length of three": FCL may be defined as the number of trophic links or trophic levels. Here the ms messes up the two definitions. Please be consistent.*
 
@@
 
-Line 59: Lakes and islands are not necessarily either simple or two-dimensional. They can be viewed as having complex vertical structures of water columns or vegetations.
+*Line 59: Lakes and islands are not necessarily either simple or two-dimensional. They can be viewed as having complex vertical structures of water columns or vegetations.*
 
@@
 
-Lines 110+: What is the relationship between ecosystem size/complexity and the total carrying capacity of a whole ecosystem?
+*Lines 110+: What is the relationship between ecosystem size/complexity and the total carrying capacity of a whole ecosystem?
 It appears that total carrying capacity depends on ecosystem size and complexity.
-What parts of the ecosystem size/complexity effects are explained by total carrying capacity?
+What parts of the ecosystem size/complexity effects are explained by total carrying capacity?*
 
@@
 
-Although both intrinsic growth rates and carrying capacities of basal species should affect local productivity, the ms considers only intrinsic growth rates as local productivity.
+*Although both intrinsic growth rates and carrying capacities of basal species should affect local productivity, the ms considers only intrinsic growth rates as local productivity.*
 
@@
 
-Line 167+: The definition of food-chain length. FCL at the whole ecosystem level should be defined here (instead of line 194).
+*Line 167+: The definition of food-chain length. FCL at the whole ecosystem level should be defined here (instead of line 194).*
 
@@
 
-Line 184+: More explanations may be necessary about why both attack rates and conversion efficiencies are controlled to change omnivory.
+*Line 184+: More explanations may be necessary about why both attack rates and conversion efficiencies are controlled to change omnivory.*
 
@@
 
-Line 184+: In case of "chain", it is wired that a_BP=0.1 despite basal resource does not contribute to predator abundance at all (e_BP=0).
+*Line 184+: In case of "chain", it is wired that a_BP=0.1 despite basal resource does not contribute to predator abundance at all (e_BP=0).*
 
@@
 
-Line 184+: Since the model describes abundance dynamics, it does not seem to make biological sense that e_BP, e_CP, or e_BC (conversion efficiencies) are greater than 1 unless these predators and consumers are parasites.
+*Line 184+: Since the model describes abundance dynamics, it does not seem to make biological sense that e_BP, e_CP, or e_BC (conversion efficiencies) are greater than 1 unless these predators and consumers are parasites.*
 
@@
 
-Line 221: "similar", does it mean "similar to the case of low disturbance"? Please clarify.
+*Line 221: "similar", does it mean "similar to the case of low disturbance"? Please clarify.*
 
@@
 
-Line 247+: Here, the ms suggests implications beyond river networks.
+*Line 247+: Here, the ms suggests implications beyond river networks.
 More elaborate inferences would be helpful.
@@
 It seems that ecosystem complexity, but not branching structures, is an integral part of the mechanism that create spatial refuges from disturbance.
-Too much emphasis on branching structures might mislead implications of the current results.
+Too much emphasis on branching structures might mislead implications of the current results.*
 
@@
 
-Figure 1B: More detailed explanations are necessary about how the figure is created.
+*Figure 1B: More detailed explanations are necessary about how the figure is created.
 The X axis should have a larger maximum value to show the whole range of the disturbance intensity distributions.
-What does the dashed line represent?
+What does the dashed line represent?*
 
@@
 
-Food chain length (FCL) is important in the movement of energy and material through ecosystems and is an indicator used in ecosystem management.
+*Food chain length (FCL) is important in the movement of energy and material through ecosystems and is an indicator used in ecosystem management.
 I have read the manuscript "Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems".
@@
 I think that the topic is of interest to general ecological audience.
-The manuscript is motivated by the claim that there is a need for a unified framework that includes multiple drivers of FCL, including that there is little research on the relationship between complexity (spatial complexity) and heterogeneity.
+The manuscript is motivated by the claim that there is a need for a unified framework that includes multiple drivers of FCL, including that there is little research on the relationship between complexity (spatial complexity) and heterogeneity.*
 
-The results that relate FCL to ecosystem complexity are interesting particularly as they are considered explicitly in a simulation with other factors that have received more attention (ecosystem size, productivity, disturbance).
+*The results that relate FCL to ecosystem complexity are interesting particularly as they are considered explicitly in a simulation with other factors that have received more attention (ecosystem size, productivity, disturbance).
 Some of the other result like the general importance of omnivory to FCL are not new but are treated reasonably and have value combined with other factors.
 However, note that previous research has addressed some of the spatial/heterogeneity effects of the availability of resources with omnivory and FCL (e.g., Holt 1996; McCann 2005; Tunney et al.
-2012 as a few examples).
-Generally, the manuscript could benefit from a little more explaining and understanding of what was generating the observed patterns in FCL since this is possible with a simulation.
+2012 as a few examples).*
 
-> Thank you, please see our response below.
+> Thank you! We have related our work to spatial food web theory that has emerged over the past two decades (1st paragraph in Discussion).
 
-In what follows my review focuses on questions and concerns related to concepts and discussion in the current version of the manuscript.
-I have added a couple editorial comments, but this is not exhaustive.
+*Generally, the manuscript could benefit from a little more explaining and understanding of what was generating the observed patterns in FCL since this is possible with a simulation.*
 
-Expectedly, omnivory is a major determinate of food chain length.
+> Agreed, see our response below.
+
+*In what follows my review focuses on questions and concerns related to concepts and discussion in the current version of the manuscript.
+I have added a couple editorial comments, but this is not exhaustive.*
+
+*Expectedly, omnivory is a major determinate of food chain length.
 I think it is reasonable to set parameters to show a few cases (chain, weak, strong) where the preference for omnivory by the predator in the model changes.
@@
 It is not clear to me that ominvory that is determined by the amount of material moving through the linkages reflects the kind of changes being referred to as structural constraints in the manuscript.
-Or if these refer to parameters.
+Or if these refer to parameters.*
 
 > We have revisited our model in light of comments from Reviewer 1.
-In the revised version, the parameter $a_{BP}$ changes how quickly the intraguild predator shifts its prey from intraguild prey to basal species as B increases.
-We have clarified this point and explanations for the "structural constraint" (we have avoided to use this vague term throughout the ms).
+In the revised version, $\delta$ is the proportional contribution of $B$ to the predator.
+We have clarified this point in Methods.
+Also, we have avoided the term "structural constraint" as it causes confusion.
 
-At low disturbance it is stated that the level of omnivory increased competition between P and C for B and reduce FCL.
+*At low disturbance it is stated that the level of omnivory increased competition between P and C for B and reduce FCL.
 It seems this is because one trophic level went extinct.
@@
 In this case the structural constraint of omnivory determined FCL.
-I think a little more explanation on what you are defining as competition and occupancy changes versus structural constraints could be helpful.
+I think a little more explanation on what you are defining as competition and occupancy changes versus structural constraints could be helpful.*
 
-> In light of comments from Reviewer 1, we have reanalyzed and fully re-structured Results and Discussion according to the new results. Please see our response to your next comment for details.
+> In light of comments from Reviewer 1, we have reanalyzed and fully re-structured Results and Discussion according to the new results.
+Competition was an improper explanation; please see our response to your next comment for details.
 
-I was expecting that high productivity you might also see some occupancy changes but maybe for different reasons than low productivity.
+*I was expecting that high productivity you might also see some occupancy changes but maybe for different reasons than low productivity.
 I wondered if at high productivity you could get increased trophic control where P controls C because it gets biomass from C and B.
-Also, as P grows it reduces the amount of B consumed by C (which should be increasingly dynamically unstable)?
+Also, as P grows it reduces the amount of B consumed by C (which should be increasingly dynamically unstable)?*
 
@@
 We have changed the figures to better visualize "state transitions" (e.g., from "B + C + P" to "B + P").
-As you suggested, we did observed the exclusion of C with high productivity, but only with the weak omnivory motif.
+As you suggested, we did observe the exclusion of C from the B + C + P state with high productivity, but only with the weak omnivory motif.
 This is probably because in this motif, the contribution of B to P is sufficiently high yet maintains strong predation impact on C.
-We have included this statement in Results (see *Productivity effect*) and expanded the discussion.
+We have included this statement in Results (see *Productivity effect*).
+According to this result, we have expanded the discussion too (7th paragraph in Discussion).
 
-Here I was wondering if the more structural change without occupancy at high productivity might be a parameter limitation based on the levels of r that were used, also the way that you use r as a surrogate of productivity.
+*Here I was wondering if the more structural change without occupancy at high productivity might be a parameter limitation based on the levels of r that were used, also the way that you use r as a surrogate of productivity.
 Often carrying capacity is implicated as a factor in productivity as it sets the level for the amount of material that moves around in the modelled system.
 I raise these questions because I am used to thinking about the dynamical implications of food webs in terms of the parameters but also the populations as these both determine the strength of interactions (in terms of flux of material in linkages between species).
-And K can limit the potential for population sizes and has known implications on stability etc..
+And K can limit the potential for population sizes and has known implications on stability etc..*
 
 > We appreciate your insight.
-In the revised MS, we used a combination of $r_b$ and carrying capacity to represent low and high productivity scenarios.
+In the revised MS, we used a combination of $r_b$ and carrying capacity $K$ to represent low and high productivity scenarios.
 Indeed.
@@
 
-I think understanding these results associated with branching ecosystems would benefit from a clear description of results and discussion for non-branching networks and then connect the results to the branching complexity ideas.
-What is happening in the branching network model that governs the results and is providing a different (or similar) response as with the other factors explored here (size, productivity etc.) ?
+*I think understanding these results associated with branching ecosystems would benefit from a clear description of results and discussion for non-branching networks and then connect the results to the branching complexity ideas.
+What is happening in the branching network model that governs the results and is providing a different (or similar) response as with the other factors explored here (size, productivity etc.) ?*
 
@@
 However, we are afraid that the inclusion of this aspect is beyond the scope of this manuscript.
-The current manuscript contains rich amount of results, and adding a new component would make the manuscript less accessible.
-We appreciate your input.
+The current manuscript contains rich test and figures already -- adding a new component would make the manuscript less accessible.
+We hope the reviewer will understand our decision.
 
-"The effect of ecosystem complexity was generally clearer than that of ecosystem size" I found that the idea that ecosystem complexity seemed to be more influential than ecosystem size was very subtle.
+*"The effect of ecosystem complexity was generally clearer than that of ecosystem size" I found that the idea that ecosystem complexity seemed to be more influential than ecosystem size was very subtle.
 I agree that in a few cases there is a larger initial effect for complexity, but it doesn't seem to be much different overall.
@@
 Does this have to do with the dampening of the disturbance moving down stream?
-As presented it seems like this answer is better aligned with the motivation for doing the study rather than a clear result of the simulations.
+As presented it seems like this answer is better aligned with the motivation for doing the study rather than a clear result of the simulations.*
 
-> True. We have removed this statement in the revised MS. Although it might be possible to quantify effect sizes and compare them, this result is not a significant portion of the manuscript.
+> True.
+We have removed this statement in the revised MS.
+Although it might be possible to quantify the effect sizes and compare them, this result is not a significant portion of the manuscript.
+We appreciate your suggestion though!
 
-One of the results that wasn't discussed is that the changes in FCL in several cases occurred at low levels of ecosystem complexity.
+*One of the results that wasn't discussed is that the changes in FCL in several cases occurred at low levels of ecosystem complexity.
 Does this mean that you need some complexity, but you don't need a lot of complexity to support longer food chains?
 Or that above a small amount of complexity it is not that important?
-Another aspect of this pattern to discuss is whether this is parameter dependent?
+Another aspect of this pattern to discuss is whether this is parameter dependent?*
 
-> XXX
+> We believe it is linked to the saturation of dispersing propagules -- ecosystem size and complexity are likely to increase FCL by increasing immigrants from less disturbed patches.
+However, in large or complex ecosystems, dispersal propagules may not be a limiting factor of FCL as the supply is plenty; therefore, productivity or omnivory more strongly regulated FCL.
+We have clarified this point in Discussion (2nd paragraph).
 
-"Productivity did not modulate the relationship between FCL and ecosystem size and ecosystem complexity." -- I see difference between high and low productivity simulations for the strong omnivory particularly in the low disturbance case.
+*"Productivity did not modulate the relationship between FCL and ecosystem size and ecosystem complexity." -- I see difference between high and low productivity simulations for the strong omnivory particularly in the low disturbance case.
 Perhaps the text ought to be changed to reflect that difference?
-Can you explain why these differences occur?
+Can you explain why these differences occur?*
 
-> In the re-analysis, we did observe stronger effects of productivity (see *Productivity effect* in Results). We have fully restructured it and expended discussion according to this change.
+> In the re-analysis, we did observe stronger effects of productivity (see *Productivity effect* in Results).
+We have fully restructured it and expended discussion according to this change.
+Our supplementary analysis of one-patch simulations highlighted the strong dependence of FCL on local productivity.
+We have clarified this point in the revised MS.
 
-Figure 1.
+*Figure 1.
 a) For clarity, is there supposed to be a change in the darkness of the arrows between B and C for the (chain) and the omnivory cases.
@@
 I see that it is the same in the Figure but does this happen in the simulations?
-I would expect that the strength of that link is influenced by changes in the other two linkages (P-B and P-C) and the densities of the populations (along the same lines as comment 1)?
+I would expect that the strength of that link is influenced by changes in the other two linkages (P-B and P-C) and the densities of the populations (along the same lines as comment 1)?*
 
-> We are glad the reviewer picked it!
+> We are glad the reviewer picked this!
 Although the darkness was supposed to represent the strength of trophic interactions, we removed this because it might cause confusion.
@@
 
-Line 255: "under high disturbance regimes, only large and complex networks could provide diverse and heterogenous habitats...and maintain long food chains".
+*Line 255: "under high disturbance regimes, only large and complex networks could provide diverse and heterogenous habitats...and maintain long food chains".
 I am not sure that this is supported as written.
 In many cases the changes in FCL are at low or moderate levels of size and complexity.
-Also, it would be helpful to use model simulations to show where FCL these proposed mechanisms are operating i.e., that large and complex .
+Also, it would be helpful to use model simulations to show where FCL these proposed mechanisms are operating i.e., that large and complex.*
 
-> Thank you. We have reworded. Smaller/simpler ecosystems were more vulnerable to disturbance.
+> Thank you.
+We have reworded.
+Smaller/simpler ecosystems were more vulnerable to disturbance.
+We also added text about the saturation of immigration effects at high levels of size and complexity (see our response above).
 
-Line 254: "Consequently food web structure largely controlled FCL." This is a strange way to describe what is happening in the low disturbance systems as FCL is food web structure.
-Can you provide a more specific mechanism here?
+*Line 254: "Consequently food web structure largely controlled FCL." This is a strange way to describe what is happening in the low disturbance systems as FCL is food web structure.
+Can you provide a more specific mechanism here?*
 
-> We have removed this part of text during the current revision.
+> True. We have removed this part of text during the current revision. Discussion has been fully restructured.
 
-Line 276: "An important result in our simulation is that omnivory does not change the qualitative relationships between FCL and ecosystem properties, as opposed to productivity effects." I find this sentence to
+*Line 276: "An important result in our simulation is that omnivory does not change the qualitative relationships between FCL and ecosystem properties, as opposed to productivity effects." I find this sentence to*
 
-> The comment ended in the middle, so we were not able to correctly understand the meaning of this.
+> The comment ended in the middle, so we were not able to correctly understand what it was supposed to mean.
 However, we suspect the referenced sentence needs to be improved.
@@
 
-Line 287 and the paragraph: How do the result of this study contribute to the ideas in this paragraph in particular that many small patches is better than large homogenous patches etc.?
-I find that this discussion is interesting but the text doesn't make a strong connection to the findings of this research.
+*Line 287 and the paragraph: How do the result of this study contribute to the ideas in this paragraph in particular that many small patches is better than large homogenous patches etc.?
+I find that this discussion is interesting but the text doesn't make a strong connection to the findings of this research.*
 
-Ensure that terms are clearly defined.
+*Ensure that terms are clearly defined.
 For example, ecosystem structure versus ecosystem feature versus ecosystem property versus food web structure.
 Ecosystem structure seems to not refer to food web maybe habitat?
-Anyway as a general comment it would be good to define terms so that the usage is understood.
+Anyway as a general comment it would be good to define terms so that the usage is understood.*
 
@@
 
-Line 45: Ecosystem Size?
+*Line 45: Ecosystem Size?*
 
@@
 
-Line 194: "are differ". Suggest remove "are".
+*Line 194: "are differ". Suggest remove "are".*
 

### 2023-03-09 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 output: 
-  word_document:
-    reference_docx: "word_format.docx"
-  #pdf_document:
-  #  latex_engine: xelatex
+  # word_document:
+  #   reference_docx: "word_format.docx"
+  pdf_document:
+   latex_engine: xelatex
 knit: (function(inputFile, encoding) {

### 2023-03-23 rmd/manuscript.Rmd
edit: address Jacques' comments

--- 2023-03-09 rmd/manuscript.Rmd
@@
     reference_docx: "word_format.docx"
-  #pdf_document:
+  # pdf_document:
   #  latex_engine: xelatex
@@
 We show independent, positive influences of ecosystem size and complexity on food chain length.
-However, these effects were contingent upon other drivers.
-In particular, the effects of ecosystem size and complexity appeared more clearly in high-disturbance and high-productivity regimes.
-Our results suggest that ecosystem complexity is an important yet overlooked driver of food chain length in branching ecosystems.
+However, the effects of ecosystem size and complexity were contingent upon other factors, appearing more clearly in high-disturbance and high-productivity regimes.
+Our results suggest that ecosystem complexity is an important yet overlooked driver of food chain length that may increase the resilience to anthropogenic environmental changes.
 
@@
 
-No data were used or produced in this research.
+No primary data were used or produced in this research.
 R functions for simulations are provided as the R package mcbrnet version 1.3.0 (available at <https://github.com/aterui/mcbrnet>).
@@
 
-The origin of variation in food chain length (FCL) has intrigued ecologists for decades due to its importance in regulating top-down effects [@paceTrophicCascadesRevealed1999], energy flow [@wangBiodiversityEcosystemFunctioning2018], and contaminant concentrations in top predators [@kiddEffectsTrophicPosition1998].
+The origin of variation in food chain length (FCL) has intrigued ecologists for many decades due to its importance in regulating top-down effects [@paceTrophicCascadesRevealed1999], energy flow [@wangBiodiversityEcosystemFunctioning2018], and contaminant concentrations in top predators [@kiddEffectsTrophicPosition1998].
 Hypotheses regarding controls of FCL can be organized into three broad categories.
@@
 
-More recently, food web ecologists have begun to recognize that FCL rarely responds to a single driver, but instead is determined by interactions of multiple factors.
+More recently, food web ecologists have begun to recognize that FCL rarely responds to a single driver, but instead is determined by interactions of extrinsic (external environmental conditions) and intrinsic factors (species and/or food web attributes).
 For example, @wardMechanisticTheoryAquatic2017 showed that FCL was more responsive to increasing productivity in large ecosystems than small ones, whereas FCL was more sensitive to size in low-productivity systems.
@@
 Confluences are merging points where tributary dynamics aggregate to form distinct environmental conditions in downstream habitat patches.
-Hence, habitat heterogeneity increases with branching complexity (**Figure 1A**), which we define here as the probability of branching per unit river distance.
+Hence, habitat heterogeneity increases with branching probability (**Figure 1A**), a proxy for ecosystem complexity in rivers [@teruiMetapopulationStabilityBranching2018; @teruiEmergentDualScaling2021].
 Highly-branched river networks provide diverse habitats that buffer the impact of large-scale environmental fluctuations, broadening opportunities for species to recolonize habitat patches post-disturbance [@teruiMetapopulationStabilityBranching2018].
@@
 We simulate trophic community dynamics within branching habitats to investigate the effects of extrinsic controls of FCL, such as ecosystem geometry [size (number of patches) and complexity (branching probability] and disturbance regime.
-In addition, our simulation model incorporates intrinsic drivers of FCL, including the degree of omnivory, productivity level, and species' dispersal ability, allowing us to assess the context dependency on drivers of FCL across extrinsic factors.
+In addition, our simulation model incorporates established intrinsic drivers of FCL, including the degree of omnivory, productivity level, and species' dispersal ability, allowing us to assess the context dependency on drivers of FCL.
 We find that ecosystem size and complexity are both positively related to FCL, but the form of this relationship is dependent on other extrinsic and intrinsic factors.
@@
 
-In our simulation, a disturbance occurs simultaneously across habitat patches with probability $p_m$, but the patches vary in their vulnerability.
+In our simulation, a disturbance occurs simultaneously across habitat patches with probability $p_m$, but disturbance intensity varies across patches as described below.
 We simulated the spatial distribution of disturbance intensity (i.e., proportional mortality in the event of stochastic regional disturbance) as follows.
@@
 We obtained the global FCL $\bar{\tau}$ for the last `r unique(df_sim$n_timestep)` time steps.
-We also calculated the proportion of the patch state ("no species", "B", "B + C", "B + P", and "B + C + P") to infer mechanisms that drive changes in FCL.
+We also calculated the proportion of the patch state ("No species", "B", "B + C", "B + P", and "B + C + P") to infer mechanisms that drive changes in FCL.
 All analysis was performed in R 4.2.1 [@rcoreteamLanguageEnvironmentStatistical2021]
@@
 
-There were two consistent patterns in the relationship between FCL and ecosystem geometry (size and complexity).
+There were three consistent patterns in the relationship between FCL and ecosystem geometry (size and complexity).
 First, the effects of ecosystem size and complexity were modulated by disturbance.
@@
 
-While empirical studies rarely looked at the influence of ecosystem complexity on FCL, several studies provide some support for this theoretical prediction.
+While empirical studies have rarely examined the influence of ecosystem complexity on FCL, several studies provide some support for this theoretical prediction.
 For example, @sullivanEcosystemStructureEmerges2015 reported FCL is controlled primarily by the number of tributary confluences in Idaho streams.
@@
 
-Another interesting result of our simulation is the interaction with omnivory.
+Another interesting result of our simulation was the interactive influence of omnivory.
 In particular, with no disturbance and low productivity, the FCL of the chain motif had contrasting relationships with ecosystem size (positive) and complexity (negative).

### 2023-03-27 rmd/manuscript.Rmd
update: ms

@@
 
-***Food chain length.*** At each patch and time step, we determined local FCL as the highest trophic position ($\tau_{x,t}$) of the existing species.
+***Food chain length and patch state.*** At each patch and time step, we determined local FCL as the highest trophic position ($\tau_{x,t}$) of the existing species.
 When the predator exists, the trophic position is determined as follows:
@@
 We defined the global FCL $\bar{\tau}$ across patches $N_p$ and time $T$ as $\bar{\tau} = \frac{\sum_x \sum_t \tau_{x,t}}{N_p T}$ (i.e., average over space and time).
+
+We also calculated the proportion of the patch state ("No species", "B", "B + C", "B + P", and "B + C + P") at each time step to infer mechanisms that drive changes in FCL.
+The time-averaged proportion of patch state $\bar{s}$ was obtained as $\bar{s} = \frac{s_t}{N_p T}$, where $s_t$ is the number of patches in a given patch state at time step $t$.
 
@@
 
-We obtained the global FCL $\bar{\tau}$ for the last `r unique(df_sim$n_timestep)` time steps.
-We also calculated the proportion of the patch state ("No species", "B", "B + C", "B + P", and "B + C + P") to infer mechanisms that drive changes in FCL.
-All analysis was performed in R 4.2.1 [@rcoreteamLanguageEnvironmentStatistical2021]
+We obtained the global FCL $\bar{\tau}$ and $\bar{s}$ for the last `r unique(df_sim$n_timestep)` time steps. All analysis was performed in R 4.2.1 [@rcoreteamLanguageEnvironmentStatistical2021]
 

### 2023-03-27 rmd/manuscript.Rmd
edit

@@
 where $e_{ij}$ is the energetic conversion efficiency from $i$ to $j$.
-The function $g(\cdot)$ defines the density-dependent recruitment for the basal species, and the function $f_k(\cdot)$ dictates the proportion of survived prey individuals after predation.
+The function $g(\cdot)$ defines the density-dependent recruitment for the basal species, and the function $f_k(\cdot)$ dictates the proportion of survived prey individuals after predation (see equation (8) below).
 In the above equations, we re-defined variables as follows to simplify the expression:
@@
 We also calculated the proportion of the patch state ("No species", "B", "B + C", "B + P", and "B + C + P") at each time step to infer mechanisms that drive changes in FCL.
-The time-averaged proportion of patch state $\bar{s}$ was obtained as $\bar{s} = \frac{s_t}{N_p T}$, where $s_t$ is the number of patches in a given patch state at time step $t$.
+The time-averaged proportion of patch state $\bar{\psi}$ was obtained as $\bar{\psi} = \frac{q_t}{N_p T}$, where $q_t$ is the number of patches in a given patch state at time step $t$.
 
@@
 
-We obtained the global FCL $\bar{\tau}$ and $\bar{s}$ for the last `r unique(df_sim$n_timestep)` time steps. All analysis was performed in R 4.2.1 [@rcoreteamLanguageEnvironmentStatistical2021]
+We obtained the global FCL $\bar{\tau}$ and the proportion of the patch state $\bar{\psi}$ for the last `r unique(df_sim$n_timestep)` time steps.
+All analysis was performed in R 4.2.1 [@rcoreteamLanguageEnvironmentStatistical2021]
 

### 2023-03-27 code/figure_si_one_patch.R

@@ -67,6 +67,7 @@ foreach(i = 1:length(a1_set)) %do% {
       ggtitle(paste0("K = ", k_set[j])) +
       geom_point(data = df_point,
                  size = 1.5,
+                 color = "salmon",
                  aes(shape = motif)) +
       labs(x = expression("Attack rate ("*a[BP]*")"),
            y = expression("Attack rate ("*a[CP]*")"),

### 2023-03-28 rmd/manuscript.Rmd
edit: simulation update

@@
 In the chain motif, an additive mechanism (i.e., the addition of higher trophic species) was always responsible for the changes in FCL.
-With disturbance, state transition occurs from "B" to "B + C" (low productivity) or "B + C" to "B + C + P" (high productivity) as ecosystem size or complexity increases.
+With disturbance, state transition occurs from "B" to "B + C" (low productivity) or "B + C" to "B + C + P" (high productivity) as ecosystem size or complexity increases (**Figures S5 and S6**).
 Exceptionally, the state transition of "B + C + P" $\rightarrow$ "B + C" occurred as ecosystem complexity increased with no disturbance and low productivity, a scenario in which ecosystem complexity decreased FCL.
@@
 In the chain motif, productivity greatly increased the overall FCL (**Figure S7**), thereby shifting the upper bound of FCL (**Figures 2 and 3**).
-In contrast, increased productivity greatly decreased the minimum levels of FCL with weak and strong omnivory with little influence on the maximum FCL (**Figures 2, 3, and S7**).
+In contrast, increased productivity decreased FCL with weak and strong omnivory (**Figures 2, 3, and S7**).
 This negative effect of increased productivity was apparent in smaller and/or simpler ecosystems, steepening the slopes between FCL and ecosystem geometry (**Figures 2 and 3**).
@@
 Topographic and terrain variation can also mediate spatial disturbance patterns in a landscape [@connell30YearStudy1997; @stueveSpatialPatternsIce2007].
-Explicit consideration of ecosystem complexity may therefore increase our ability to predict FCL and complements spatial food web theory that has emerged over the past two decades [@holtFoodWebsSpace2002a; @mccannDynamicsSpatiallyCoupled2005; @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012; @andersonEffectsDispersalRiver2018; @teruiSpatialDisturbanceSynchrony2019].
+Explicit consideration of ecosystem complexity may therefore increase our ability to predict FCL and complements spatial food web theory that has emerged over the past two decades [@holtFoodWebsSpace2002; @mccannDynamicsSpatiallyCoupled2005; @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012; @andersonEffectsDispersalRiver2018; @teruiSpatialDisturbanceSynchrony2019].
 

### 2023-03-28 code/figure_si_within_net.R

@@ -31,7 +31,7 @@ df_dist <- foreach(i = seq_len(length(p_branch)),
                    }
 
 df_plot <- left_join(df_sim, df_dist,
-                    by = c("n_patch", "p_branch")) %>% 
+                     by = c("n_patch", "p_branch")) %>% 
   pivot_longer(cols = c(mc_capacity, mu_distance),
                values_to = "value",
                names_to = "y")
@@ -63,4 +63,6 @@ g_net <- df_plot %>%
 # export ------------------------------------------------------------------
 
 ggsave(g_net,
-       filename = "figure/figure_si_k_distance.pdf")
+       filename = "figure/figure_si_k_distance.pdf",
+       width = 8,
+       height = 11)

### 2023-03-28 rmd/suppl_info.Rmd
edit: simulation update
edit: fig si size

@@
 ```{r fig4, fig.cap="Relationships between food chain length and ecosystem complexity with low disturbance intensity $\\mu_m = 0.2$.
-Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios."}
+Lines indicate loess curves, and shades are associated 95% confidence intervals.
+Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
+Different line colors indicate different disturbance probabilities."}
 knitr::include_graphics(here::here("figure/figure_si_p_branch.pdf"))
@@
 ```{r fig5, fig.cap="Relationships between patch state occupancy and ecosystem size with high disturbance intensity $\\mu_m = 0.8$.
+Lines indicate loess curves, and shades are associated 95% confidence intervals.
 Panels indicate different food web motifs.
-In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios."}
+In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios.
+Different line colors indicate different disturbance probabilities."}
 knitr::include_graphics(here::here("figure/figure_si_state_n_patch.pdf"))
@@
 ```{r fig6, fig.cap="Relationships between patch state occupancy and ecosystem complexity with high disturbance intensity $\\mu_m = 0.8$.
+Lines indicate loess curves, and shades are associated 95% confidence intervals.
 Panels indicate different food web motifs.
-In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios."}
+In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios.
+Different line colors indicate different disturbance probabilities."}
 knitr::include_graphics(here::here("figure/figure_si_state_p_branch.pdf"))

### 2023-03-28 code/figure_si_within_net.R

@@ -64,5 +64,5 @@ g_net <- df_plot %>%
 
 ggsave(g_net,
        filename = "figure/figure_si_k_distance.pdf",
-       width = 8,
-       height = 11)
+       width = 6,
+       height = 8)

### 2023-03-28 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 output: 
-  word_document:
-    reference_docx: "word_format.docx"
-  # pdf_document:
-  #  latex_engine: xelatex
+  # word_document:
+  #   reference_docx: "word_format.docx"
+  pdf_document:
+   latex_engine: xelatex
 knit: (function(inputFile, encoding) {
@@
 No primary data were used or produced in this research.
-R functions for simulations are provided as the R package mcbrnet version 1.3.0 (available at <https://github.com/aterui/mcbrnet>).
+R functions for simulations are provided as the R package mcbrnet version 1.3.1 (available at <https://github.com/aterui/mcbrnet>).
 Codes for simulations, statistical analysis, figures, and tables are available at <https://github.com/aterui/public-proj_fcl-river-theory>.

### 2023-05-15 rmd/reply2.Rmd
edit

---
title: "RE: Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems"
output:
  pdf_document:
    latex_engine: xelatex
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
mainfont: Times New Roman
---

Dear Editor,

This is a re-submission of a manuscript: ECS22-0646 "*Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems.*" Thank you for allowing us to revise our manuscript. The comments were very helpful in improving our manuscript. When revising the manuscript, we have put a particular emphasis on the following points:

XXX

We appended a point-by-point response below (reviewer comments are shown in *italics*). We hope our manuscript is now suitable for publication in your journal.

Sincerely,

Akira Terui on behalf of the authors

## Response to Reviewer \#1

*This is a thorough revision of the previous manuscript (ms) and has taken into account all of my previous criticisms. In particular, I highly appreciate the full reformulation and analysis of the simulation model, which now seems theoretically sound. One thing that I noticed is that the new model assumes non-overlapping generations of consumer (C) and predator (P). Because all consumer and predator individuals die each time step, they go extinct locally (without immigration) in the next time step if their resource species are totally absent to preclude their reproduction. Although the assumption of non-overlapping generations is theoretically fine, I suspect that this might be unrealistic for the riverine food webs that are modelled. Correcting this assumption would require another reformulation and analysis of the model (such that some surviving fractions of C and P abundances are carried over to the next time steps). At least, the ms should acknowledge this assumption and provide a justification for it.*

> XXX

## Response to Reviewer \#2

*This is my second time reviewing the manuscript “Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems”. Food chain length (FCL) is important in the movement of energy and material through ecosystems and is an indicator used in ecosystem management. The manuscript takes a simulation approach to examining multiple drivers of food chain length related to branching ecosystems like river ecosystems. This is a broadly interesting topic and there are some interesting results presented. Since the manuscript is treated as a new submission, I do not have the ability to easily reference changes in the manuscript that specifically address comments of the first review. My review focuses on some general suggestions to improve the clarity of the manuscript.*

*My main comments are that the manuscript would benefit from a clearer focus on the influence of complexity on food chain length. This could include some more synthetic or summary display items. A revised manuscript should still include presenting the material and trends related to other hypotheses and drivers, but getting into all of the different patterns should be done logically and with the goal of understanding the new general findings related to complexity and branching ecosystems. You may want to provide a new Figure or Table so that the different layers can be understood. Right now, there are different trends in figure 2 and 3 and there is a lot to look at in the supplement. So it is not clear that there is a general increasing relationship between FCL and complexity. The supplement figures are heavily relied on in the main results. Figure captions need to be clear and complete. Summary display items could for example, include a table listing the different simulation experiments in rows with a column that includes the key the takeaway for understanding the relationship between complexity and FCL.*

*A question that remains unclear to me is how omnivory is affecting the FCL in the simulations. It seems that all trophic species (B,C,P) are not supported in the simulation under all conditions. So basal species only or the loss of the consumer or predator. Consequently, extinction in the simulations can reduce FCL (correct me if wrong but the simulations start with three species, so species are not added or inserted). Is this the only way that food chain length can change? Or is food chain length a product of the dynamic interactions that can be influenced by the magnitude of the trophic contributions from different trophic levels. This would be consistent with a continuous measure of food chain length rather than another measure like the number of distinct trophic levels that occupy a patch? There is a lot of occupancy information added to the supplement but I had a hard time understanding from the text.*

*For the ecosystem size plot, when you have the straight chain the max FCL is 3 but when you add a some omnivory the max is 2, so a whole trophic level is lost. But when you go from weak omnivory to strong omnivory the max FCL stays the same. Looking at the figure 1 c (and from experience) the expectation is that as omnivory increases the FCL decreases. Now I am doubting that I understand the way the FCL is being determined. Also, in the supplement there are figures that show some FCL between 2 and 3 (Figure S3 and S4) but not sure what is happening.*

*Generally, I found that the figure captions require more information. For example, it is not clear to me, for example in figure 2 and figure 3 what level of complexity is being shown in the ecosystem size plots and what level of size is being shown in the complexity plot.*

*There are a lot of supplementary figures, but these figures are referenced in the main text without much text in the supplement. I felt like this left a lot of work for me to figure out what was going on. For example, in Figure S5c it is not clear to me how you have a change in occupancy with ecosystem size for the first row that is labelled no species. To me it doesn’t make sense that there is patch occupancy if there aren’t any species and the caption is not providing definitions for me to understand this. There are similarly confusing examples for many supplementary figures.*

*Figure 1 c should be better labelled to confirm that the vertical position in the figure of P is showing the length of the food chain. You could add an axis to the figure labelled FCL or equivalent. Also, I still think that the figures showing weak omnivory and strong omnivory are incomplete. To me the figure shows a change in the vertical position of P, but it is not clear from the diagram that this is due to an increase in ominivory i.e., no indication of a increased reliance of P on B over P on C. So, is that what is intended?*

*I think the manuscript requires some more clarification about how patch level productivity and the whole system productivity are changing and how this affects the relationship between FCL and complexity. It seems that the new simulations productivity is being manipulated through the parameters r and K in each patch (compared to only r previously) but then there is a small section of text that points out the ecosystem size is increasing the overall biomass produced or K (which makes sense) and an increase in FCL and complexity is reducing the overall K and thus FCL decreases. Ok, K of the whole system is changing with size and complexity. But K is also being manipulated at the patch level as part of productivity. To me it seems like there are productivity changes at two scales. How do these patch level versus whole system level productivity changes relate to FCL is being calculated at the whole system level?*

*From Line 275: “however, these patterns characteristic of the chain scenario diminished when considering some levels of omnivory”. Some of the reasons for why this is happening follow in next paragraphs in the results section, but I found these sections to be written more as discussion than results. The figures that accompany the explanations are placed in the supplement which can be fine. But in this case, it may be better to have these as results and alter the text to describe the figures saving the discussion more for the discussion section. This is an important part of understanding what is happening in the manuscript.*

*The term ecosystem geometry now appears in the results section and includes size and complexity. It would we better to introduce the term in the introduction or just use size and complexity.*

### 2023-05-15 code/figure_summary.R
edit

@@ -1,0 +1,19 @@
+
+# setup -------------------------------------------------------------------
+
+rm(list = ls(all.names = TRUE))
+source(here::here("code/library.R"))
+source(here::here("code/set_figure_theme.R"))
+source(here::here("code/format_sim_data.R"))
+
+
+# data --------------------------------------------------------------------
+
+df_rho <- df_sim %>% 
+  group_by(param_set) %>% 
+  do(rho_n = cor(.$fcl, .$n_patch, method = "spearman"),
+     rho_pb = cor(.$fcl, .$p_branch, method = "spearman")) %>% 
+  mutate(across(.cols = where(is.list), .fns = unlist)) %>% 
+  left_join(df_param,
+            by = "param_set") %>% 
+  filter(mean_disturb_source == 0.8)

### 2023-05-15 code/format_sim_data.R

@@ -29,5 +29,6 @@ df_param <- df_sim %>%
             mc_capacity,
             p_branch,
             n_patch,
-            n_rep)) %>% 
+            n_rep,
+            starts_with("s"))) %>% 
   distinct()

### 2023-05-16 code/figure_gridplot.R
edit

@@ -1,0 +1,106 @@
+
+# setup -------------------------------------------------------------------
+
+rm(list = ls(all.names = TRUE))
+source(here::here("code/library.R"))
+source(here::here("code/set_figure_theme.R"))
+source(here::here("code/format_sim_data.R"))
+
+
+# figure ------------------------------------------------------------------
+
+## size effect
+m1 <- matrix(0, 12, 12)
+m1[1, c(2, 3)] <- 1 # disturbance frequency
+m1[2, 4:6] <- 1 # no distb x omn
+m1[3, 7:8] <- 1 # disturbance intensity
+m1[4, 9:10] <- 1 # freq:no x omn:chain x prod
+m1[5, 11:12] <- 1 # freq:no x omn:weak x prod
+
+adj <- igraph::graph.adjacency(m1)
+
+V(adj)$label[1] <- "Disturbance frequency"
+V(adj)$label[c(2, 8)] <- "Omnivory"
+V(adj)$label[3] <- "Disturbance intensity"
+V(adj)$label[4:5] <- "Productivity"
+
+V(adj)$label[c(6, 10, 11)] <- '"No effect"'
+V(adj)$label[7] <- '"Very weak effect" \n (Fig. S3)'
+V(adj)$label[c(8:9, 12)] <- '"Positive"'
+
+V(adj)$color[c(1:dim(m1)[1])] <- "node"
+V(adj)$color[c(6, 7, 8:12)] <- "pattern"
+
+E(adj)$label[1] <- "Absent"; E(adj)$label[2] <- "Present"
+E(adj)$label[3:5] <- c("None (Chain)", "Weak", "Strong")
+E(adj)$label[6:7] <- c("Weak", "Strong")
+E(adj)$label[8:11] <- rep(c("Low", "High"), 2)
+
+g_tree1 <- ggraph(adj, layout = 'tree') +
+  geom_edge_link(aes(label = label),
+                 color = grey(0, 0.25),
+                 linetype = "dashed",
+                 arrow = arrow(length = unit(2,'mm'),
+                               type = "closed"),
+                 start_cap = circle(8,'mm'),
+                 end_cap = circle(8,'mm')) +
+  geom_node_label(aes(label = label,
+                      color = color),
+                  label.size = 0) +
+  scale_color_manual(values = c("black", "steelblue")) +
+  theme_void() +
+  guides(color = "none")
+
+
+## Complexity effect
[…]

### 2023-05-16 code/figure_summary.R

@@ -1,19 +1,0 @@
-
-# setup -------------------------------------------------------------------
-
-rm(list = ls(all.names = TRUE))
-source(here::here("code/library.R"))
-source(here::here("code/set_figure_theme.R"))
-source(here::here("code/format_sim_data.R"))
-
-
-# data --------------------------------------------------------------------
-
-df_rho <- df_sim %>% 
-  group_by(param_set) %>% 
-  do(rho_n = cor(.$fcl, .$n_patch, method = "spearman"),
-     rho_pb = cor(.$fcl, .$p_branch, method = "spearman")) %>% 
-  mutate(across(.cols = where(is.list), .fns = unlist)) %>% 
-  left_join(df_param,
-            by = "param_set") %>% 
-  filter(mean_disturb_source == 0.8)

### 2023-05-16 code/figure_food_web.R
edit: figures

@@ -7,7 +7,7 @@ source(here::here("code/library.R"))
 
 # figure ------------------------------------------------------------------
 
-delta <- c(0, 0.31, 0.73)
+delta <- c(0, 0.30, 0.70)
 title_fw <- c("Chain", "Weak omnivory", "Strong omnivory")
 
 g_fw <- foreach(i = 1:3) %do% {
@@ -35,6 +35,24 @@ g_fw <- foreach(i = 1:3) %do% {
     
     layout$x <- rep(0.5, 5)
     
+    g <- ggraph(layout) + 
+      geom_edge_link(arrow = arrow(type = "closed",
+                                   ends = "last",
+                                   length = unit(0.1, "npc"),
+                                   angle = 20),
+                     start_cap = circle(0.07, 'npc'),
+                     end_cap = circle(0.07, 'npc'),
+                     show.legend = T) +
+      geom_node_label(aes(label = label), 
+                      color = grey(0.25),
+                      size = 6,
+                      label.size = 0,
+                      fill = NA) +
+      labs(subtitle = title_fw[i]) +
+      theme(rect = element_blank(),
+            panel.border = element_blank(),
+            panel.background = element_blank())
+    
   } else {
     
     layout$x <- c(0.5, # Basal
@@ -42,47 +60,31 @@ g_fw <- foreach(i = 1:3) %do% {
                   0.5 + 1/layout$y[3], # Intraguild predator
                   0, # dummy 1
                   1 # dummy 2
-                  )
+    )
+    
+    g <- ggraph(layout) + 
+      geom_edge_link(arrow = arrow(type = "closed",
+                                   ends = "last",
+                                   length = unit(0.1, "npc"),
+                                   angle = 20),
+                     aes(edge_color = weight),
+                     start_cap = circle(0.07, 'npc'),
+                     end_cap = circle(0.07, 'npc'),
+                     show.legend = T) +
+      geom_node_label(aes(label = label), 
+                      color = grey(0.25),
+                      size = 6,
+                      label.size = 0,
+                      fill = NA) +
+      scale_edge_color_gradient(low = "grey",
+                                high = "black") +
+      labs(subtitle = title_fw[i]) +
+      theme(rect = element_blank(),
+            panel.border = element_blank(),
+            panel.background = element_blank()) +
+      guides(edge_color = "none")
     
   }
-  
-  # graph
-  g <- ggraph(layout) + 
-    geom_edge_link(arrow = arrow(type = "closed",
-                                 ends = "last",
[…]

### 2023-05-16 code/figure_si_state.R

@@ -50,7 +50,7 @@ list_g_np <- foreach(x = unique(df_plot$omn)) %do% {
                #scales = "free",
                labeller = labeller(r_b = as_labeller(lab, label_parsed),
                                    state = state_lab)) +
-    labs(y = "Occupancy",
+    labs(y = "Proportion of patch state",
          x = "Ecosystem size (number of habitat patches)",
          color = "Disturbance prob.",
          fill = "Disturbance prob.") +
@@ -84,7 +84,7 @@ list_g_bp <- foreach(x = unique(df_plot$omn)) %do% {
                #scales = "free",
                labeller = labeller(r_b = as_labeller(lab, label_parsed),
                                    state = state_lab)) +
-    labs(y = "Occupancy",
+    labs(y = "Proportion of patch state",
          x = "Ecosystem complexity (branching prob.)",
          color = "Disturbance prob.",
          fill = "Disturbance prob.") +
@@ -106,11 +106,11 @@ g_bp_all <- list_g_bp[[1]] + list_g_bp[[2]] +
 # export ------------------------------------------------------------------
 
 ggsave(g_np_all,
-       filename = here::here(paste0("figure/figure_si_state_n_patch.pdf")),
+       filename = here::here(paste0("figure/figure_state_n_patch.pdf")),
        height = 14,
        width = 14)
 
 ggsave(g_bp_all,
-       filename = here::here(paste0("figure/figure_si_state_p_branch.pdf")),
+       filename = here::here(paste0("figure/figure_state_p_branch.pdf")),
        height = 14,
        width = 14)

### 2023-05-16 rmd/reply2.Rmd
Update reply2.Rmd

@@
 
+> Thank you for your valuable feedback! Please see our point-by-point response below.
+
 *My main comments are that the manuscript would benefit from a clearer focus on the influence of complexity on food chain length.
@@
 
+> This is a great idea! We included a new hierarchical tree figure to show how different factors detemine the FCL response to ecosystem size/complexity (Figure 2 in the revised MS).
+Also, we revised the figure captions to better guide readers.
+
 *A question that remains unclear to me is how omnivory is affecting the FCL in the simulations.
@@
 Or is food chain length a product of the dynamic interactions that can be influenced by the magnitude of the trophic contributions from different trophic levels.
-This would be consistent with a continuous measure of food chain length rather than another measure like the number of distinct trophic levels that occupy a patch?
-There is a lot of occupancy information added to the supplement but I had a hard time understanding from the text.*
+This would be consistent with a continuous measure of food chain length rather than another measure like the number of distinct trophic levels that occupy a patch?*
 
-*For the ecosystem size plot, when you have the straight chain the max FCL is 3 but when you add a some omnivory the max is 2, so a whole trophic level is lost.
-But when you go from weak omnivory to strong omnivory the max FCL stays the same.
+> Our unclear explanation may have confused the reviewer.
+There are a number of cases where all species (B, C, P) coexist.
+"*...is food chain length a product of the dynamic interactions that can be influenced by the magnitude of the trophic contributions from different trophic levels.*" This is how we defined FCL in our model, which was articulated in equations 12 & 13.
+
+*There is a lot of occupancy information added to the supplement but I had a hard time understanding from the text.*
+
+> We appreciate your feedback. The figure captions were editted so the readers can better understand what those figures mean.
+In particular, we have reworded the confusing term "occupancy" to "proportion of patch state" to clarify the y-axis.
+
+*For the ecosystem size plot, when you have the straight chain the max FCL is 3 but when you add a some omnivory the max is 2, so a whole trophic level is lost.*
+
+> Thank you, please let us clarify.
+First, we defined FCL as trophic position, not as trophic level. Thus, the FCL can take continuous values.
+With omnivory, the max is greater than 2 becuase of the coexistence of C and P.
+P's trophic level depends on the relative abundance of B.
+Thus, with omnivory, the FCL can be shorter than the chain motif even if the same set of species coexist.
+
+*But when you go from weak omnivory to strong omnivory the max FCL stays the same.
 Looking at the figure 1 c (and from experience) the expectation is that as omnivory increases the FCL decreases.
@@
 
+> Thank you for your input.
+"*Looking at the figure 1 c (and from experience) the expectation is that as omnivory increases the FCL decreases.*" -- The response of FCL to increasing omnivory can be more complex.
+Weak omnivory can cause similar or even lower FCL than strong omnivory because the intraguild predator has a greater predation pressure on intraguild prey.
+For the definition of FCL in our study, please see equations 12 & 13 -- it accounts for relative availability of B and C in a given habitat patch.
+
 *Generally, I found that the figure captions require more information.
 For example, it is not clear to me, for example in figure 2 and figure 3 what level of complexity is being shown in the ecosystem size plots and what level of size is being shown in the complexity plot.*
+
+> Thank you, we have revised figure captions.
+Also, we clarified the values of ecosystem size or complexity in the main figures (both are randomly drawn from uniform distributions).
 
@@
 
+> Thank you, we have revised figure captions.
+
 *Figure 1 c should be better labelled to confirm that the vertical position in the figure of P is showing the length of the food chain.
@@
 So, is that what is intended?*
+
+> Thank you! We have now added a color scale that indicate the relative reliance on B and C.
+Meanwhile, we avoided to include FCL in the figure; this is because the relative reliance per se may not necessarily determine the FCL.
+For example, weak omnivory could enhance the extirpation of intraguild prey, potentially resulting in a shorter FCL than that in strong omnivory.
 
@@
 
+> Agreed -- there are two levels of productivity in our model, and this set up was unavoidable.
+In the meantime, we feel that too much emphasis on this part may over-complicate the content, which is complex enough already (as you pointed out earlier in this round of review).
+In looking at the one-patch simulation, the FCL response to increased productivity can be well understood based on the single-patch analog.
+Thus, we decided not to delve into the specifics on this issue.
+
 *From Line 275: “however, these patterns characteristic of the chain scenario diminished when considering some levels of omnivory”.
@@
 
+> We are not sure if we correctly understood this comment, but we interpreted this comment suggesting to move the patch state figures to the maintext.
+We have now moved those figures to the maintext.
+In the meantime, we kept the text to explain those figures in the result section, since it seems inappropriate to mention those findings in the discussion section without mentioning them in the result.
+
 *The term ecosystem geometry now appears in the results section and includes size and complexity. It would we better to introduce the term in the introduction or just use size and complexity.*
+
+> We have defined the term in the Intro.

### 2023-05-16 code/figure_treeplot.R
edit: figure

@@ -24,9 +24,9 @@ V(adj)$label[c(2, 8)] <- "Omnivory"
 V(adj)$label[3] <- "Disturbance intensity"
 V(adj)$label[4:5] <- "Productivity"
 
-V(adj)$label[c(6, 10, 11)] <- '"No effect"'
-V(adj)$label[7] <- '"Very weak effect" \n (Fig. S3)'
-V(adj)$label[c(8:9, 12)] <- '"Positive"'
+V(adj)$label[c(6, 10, 11)] <- "No effect"
+V(adj)$label[7] <- "Very weak effect \n (Fig. S3)"
+V(adj)$label[c(8:9, 12)] <- "Positive"
 
 V(adj)$color[c(1:dim(m1)[1])] <- "node"
 V(adj)$color[c(6, 7, 8:12)] <- "pattern"
@@ -38,7 +38,7 @@ E(adj)$label[8:11] <- rep(c("Low", "High"), 2)
 
 g_tree1 <- ggraph(adj, layout = 'tree') +
   geom_edge_link(aes(label = label),
-                 color = grey(0, 0.25),
+                 color = grey(0.7),
                  linetype = "dashed",
                  arrow = arrow(length = unit(2,'mm'),
                                type = "closed"),
@@ -66,10 +66,10 @@ V(adj)$label[2] <- "Omnivory"
 V(adj)$label[3] <- "Disturbance intensity"
 V(adj)$label[4] <- "Productivity"
 
-V(adj)$label[c(5, 8)] <- '"Positive"'
-V(adj)$label[c(6, 10)] <- '"No effect"'
-V(adj)$label[7] <- '"Very weak effect" \n (Fig. S4)'
-V(adj)$label[9] <- '"Negative"'
+V(adj)$label[c(5, 8)] <- "Positive"
+V(adj)$label[c(6, 10)] <- "No effect"
+V(adj)$label[7] <- "Very weak effect \n (Fig. S4)"
+V(adj)$label[9] <- "Negative"
 
 V(adj)$color[1:10] <- "node"
 V(adj)$color[c(5:10)] <- "pattern"
@@ -82,7 +82,7 @@ E(adj)$label[8:9] <- c("Low", "High")
 
 g_tree2 <- ggraph(adj, layout = 'tree') +
   geom_edge_link(aes(label = label),
-                 color = grey(0, 0.25),
+                 color = grey(0.7),
                  linetype = "dashed",
                  arrow = arrow(length = unit(2,'mm'),
                                type = "closed"),

### 2023-05-16 rmd/manuscript.Rmd
edit: text

@@
 mainfont: Times New Roman
+fontsize: 12pt
 header-includes:
@@
                       message = FALSE)
+
+options(knitr.kable.NA = '')
 
@@
 When omnivory is absent in three-species communities, the trophic structure or "motif" assumes a simple linear food chain with a maximum trophic position of three (**Figure 1D**).
-However, FCL decreases with increasing omnivory as the top predator eats lower in the food web [@wardMechanisticTheoryAquatic2017; @woottonOmnivoryStabilityFreshwater2017].
-Theoretical work predicts that food webs are destabilized when omnivory is too strong, leading to the loss of species and shorter food chains [@woottonOmnivoryStabilityFreshwater2017].
-Conversely, weak omnivory stabilizes the food web and increases the persistence of species within a community, particularly at intermediate levels of productivity and high disturbance levels [@woottonOmnivoryStabilityFreshwater2017; @woottonManyWeakInteractions2016].
+However, omnivory changes the trophic structure with important implications for food web persistence.
+For example, theoretical work predicts that food webs are destabilized when weak omnivory is coupled with strong intraguild predation, leading to the loss of intraguild prey and shorter food chains at high productivity levels [@woottonOmnivoryStabilityFreshwater2017; @takimotoEffectsProductivityDisturbance2012; @postProximateStructuralMechanisms2007].
 Due to the context-dependency of FCL to multiple drivers and the potential for interactive effects, a unified framework of multiple FCL drivers is needed to better understand when and where FCL will respond to these extrinsic and intrinsic factors.
@@
 
-We simulate trophic community dynamics within branching habitats to investigate the effects of extrinsic controls of FCL, such as ecosystem geometry [size (number of patches) and complexity (branching probability] and disturbance regime.
+We simulate trophic community dynamics within branching habitats to investigate the effects of extrinsic controls of FCL.
+In our study, extrinsic controls include ecosystem geometry [size (number of patches) and complexity (branching probability] and disturbance regime.
 In addition, our simulation model incorporates established intrinsic drivers of FCL, including the degree of omnivory, productivity level, and species' dispersal ability, allowing us to assess the context dependency on drivers of FCL.
@@
 Thus, our definition accounts for the influence of omnivory.
+As a result, the FCL in a single patch can change due to the P's degree of omnivory and/or species extinction.
 We defined the global FCL $\bar{\tau}$ across patches $N_p$ and time $T$ as $\bar{\tau} = \frac{\sum_x \sum_t \tau_{x,t}}{N_p T}$ (i.e., average over space and time).
@@
 There were three consistent patterns in the relationship between FCL and ecosystem geometry (size and complexity).
-First, the effects of ecosystem size and complexity were modulated by disturbance.
-Second, productivity strengthened the relationship between FCL and ecosystem geometry with important direct effects that were motif dependent.
-Third, dispersal had a minimal but non-negligible positive effect on FCL.
+First, both ecosystem size and complexity had positive effects on FCL when disturbance was strong and frequent (**Figure 2**).
+Second, in the absence of disturbance, their effects were highly dependent on motifs and productivity levels (**Figure 2**).
+Third, dispersal did not qualitatively alter the effects of ecosystem geometry, although it had non-negligible direct positive effect on FCL.
 Below we describe these results in detail.
@@
 
-Ecosystem geometry interacted with disturbance to dictate FCL in branching ecosystems (**Figures 2 and 3**).
-With no disturbance, both geometric features (size and complexity) had limited influence on FCL.
-However, as disturbance frequency increases, the relationships between FCL and ecosystem geometry become positive.
+Ecosystem geometry interacted with disturbance to dictate FCL in branching ecosystems (**Figures 3 and 4**).
+In the presence of disturbance, both geometric features (size and complexity) had noticeable influences on FCL.
+As disturbance frequency increased, the relationships between FCL and ecosystem geometry became clearer and positive.
 This interactive effect emerged because the sensitivity of FCL to disturbance depended on ecosystem geometry.
-Although increased disturbance frequency reduced overall FCL, smaller and/or simpler ecosystems exhibited greater vulnerability (**Figures 2 and 3**).
+Although increased disturbance frequency reduced overall FCL, smaller and/or simpler ecosystems exhibited greater vulnerability (**Figures 3 and 4**).
 Therefore, disturbance interacted with ecosystem geometry to produce the interdependence of these factors in determining FCL.
-However, the maximum levels of FCL were strongly regulated by disturbance frequency, highlighting the important direct effect of disturbance.
-It is important to note that the interactive effects were not observed with low disturbance intensity (**Figures S3 and S4**); this suggests that disturbance intensity must be strong enough to be influential.
+Yet, the interactive effects were not observed with low disturbance intensity (**Figures S3 and S4**); this suggests that disturbance intensity must be strong enough to be influential.
 
-The exception is the chain scenario (no omnivory) with no disturbance and low productivity.
-In this scenario, ecosystem size had a positive effect on FCL (**Figure 2**), while ecosystem complexity had a negative influence (**Figure 3**).
+With no disturbance, the effects of ecosystem geometry were highly context dependent (**Figure 2**).
+In the chain scenario (no omnivory) with low productivity, ecosystem size had a positive effect on FCL (**Figure 3**), while ecosystem complexity had a negative influence (**Figure 4**).
 These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increased the total carrying capacity, whereas ecosystem complexity decreased it (**Figure S10**).
-However, these patterns characteristic of the chain scenario diminished when considering some levels of omnivory.
+However, when considering some levels of omnivory, these patterns characteristic of the chain scenario diminished.
+With weak omnivory (the second rows of **Figure 3 and 4**), ecosystem complexity had consistent positive effects on FCL, except the scenario with low productivity and short dispersal (**Figure 4**).
+In contrast, the effect of ecosystem size was more productivity dependent (**Figure 4**).
+Neither of ecosystem size and complexity influenced FCL with strong omnivory.
 
-State transitions shed light on how FCL changes occurred.
+## State transitions
+
+State transitions largely depended on motif structure and shed light on how FCL changes occurred.
 In the chain motif, an additive mechanism (i.e., the addition of higher trophic species) was always responsible for the changes in FCL.
-With disturbance, state transition occurs from "B" to "B + C" (low productivity) or "B + C" to "B + C + P" (high productivity) as ecosystem size or complexity increases (**Figures S5 and S6**).
+With disturbance, state transition occurs from "B" to "B + C" (low productivity) or "B + C" to "B + C + P" (high productivity) as ecosystem size or complexity increases (**Figures 5 and 6**).
 Exceptionally, the state transition of "B + C + P" $\rightarrow$ "B + C" occurred as ecosystem complexity increased with no disturbance and low productivity, a scenario in which ecosystem complexity decreased FCL.
 
-In the weak and strong omnivory motifs, ecosystem size and complexity increased FCL through different mechanisms.
-Ecosystem size increased FCL primarily through additive mechanisms, facilitating the state transition of "B" $\rightarrow$ "B + P" (low productivity) or "No species" $\rightarrow$ "B + P" (high productivity), particularly with frequent disturbance (**Figure S5**).
+In the weak and strong omnivory motifs, different state transitions were associated with ecosystem size and complexity.
+Ecosystem size increased FCL primarily through additive mechanisms, facilitating the state transition of "B" $\rightarrow$ "B + P" (low productivity) or "No species" $\rightarrow$ "B + P" (high productivity), particularly with frequent disturbance (**Figure 5**).
+
 In contrast, insertion (i.e., the insertion of intermediate trophic species) and additive mechanisms were responsible for the effect of ecosystem complexity.
-With weak omnivory, the transition from "B + P" to "B + C + P" (insertion) was predominant when productivity was low (**Figure S6B**).
-However, the additive mechanism ("No species" $\rightarrow$ "B + P") took primacy in the high productivity scenario (**Figure S6B**).
+With weak omnivory, the transition from "B + P" to "B + C + P" (insertion) was predominant when productivity was low (**Figure 6B**).
+However, the additive mechanism ("No species" $\rightarrow$ "B + P") took primacy in the high productivity scenario (**Figure 6B**).
 In the strong omnivory motif, transition patterns depended on disturbance frequency.
-When disturbance was rare ($p_m = 0.05$), the insertion mechanism ("B + P" $\rightarrow$ "B + C + P") drove the FCL change (**Figure S6C**).
-As disturbance frequency increased, the primary process switched to an additive mechanism ("B" $\rightarrow$ "B + P" or "No species" $\rightarrow$ "B + P"; **Figure S6C**).
+When disturbance was rare ($p_m = 0.05$), the insertion mechanism ("B + P" $\rightarrow$ "B + C + P") drove the FCL change (**Figure 6C**).
+As disturbance frequency increased, the primary process switched to an additive mechanism ("B" $\rightarrow$ "B + P" or "No species" $\rightarrow$ "B + P"; **Figure 6C**).
 
-## Productivity effect
+<!-- ## Productivity effect -->
 
-High productivity strengthened the relationship between FCL and ecosystem geometry.
+<!-- High productivity strengthened the relationship between FCL and ecosystem geometry.
 In the chain motif, productivity greatly increased the overall FCL (**Figure S7**), thereby shifting the upper bound of FCL (**Figures 2 and 3**).
@@
 This negative effect of increased productivity was apparent in smaller and/or simpler ecosystems, steepening the slopes between FCL and ecosystem geometry (**Figures 2 and 3**).
+-->
 
-State transitions associated with increased productivity depended on the motif structure.
+<!-- State transitions associated with increased productivity depended on the motif structure.
 In the chain motif, the transition from "B + C" to "B + C + P" was dominant regardless of disturbance frequency (**Figure S8**).
@@
 Instead, productivity significantly increased the proportion of the "No species" state (**Figure S8**) in disturbed scenarios, especially in small and/or simple ecosystems (**Figures S5 and S6**).
+-->
 
-## Dispersal effect
+<!-- ## Dispersal effect -->
 
-Dispersal distance had a positive influence on FCL regardless of simulation scenarios (**Figure S7**). However, this direct effect was weak in comparison to other factors.
+<!-- Dispersal distance had a positive influence on FCL regardless of simulation scenarios (**Figure S7**). However, this direct effect was weak in comparison to other factors. -->
 
-State transitions occurred through an additive (chain) or insertion mechanism (omnivory).
+<!-- State transitions occurred through an additive (chain) or insertion mechanism (omnivory).
 In the chain motif, the transitions were either "B" $\rightarrow$ "B + C" or "B + C" $\rightarrow$ "B + C+ P", depending on disturbance frequency (**Figure S9**).
 In the omnivory motifs, the predominant transition was from "B + P" to "B + C + P" (**Figure S9**).
+-->
 
@@
 
-## Acknowledgements
+# Acknowledgements
 
@@
 
-## Author contributions
+# Author contributions
 
@@
 
-# References
+\pagebreak
+
+# References{-}
+
+<div id="refs"></div>
+
+\pagebreak
+
+# Tables
+
+```{r}
+source(here::here("code/table_param.R"))
+
+knitr::kable(df_value %>%
+               dplyr::select(-"One-patch") %>% 
+               rename(Value = "Main"),
+             escape = FALSE,
+             booktabs = TRUE,
+             caption = "Parameter values used in the main simulation") %>% 
+  kableExtra::kable_styling(latex_options = 'HOLD_position')
+```
+
+\pagebreak
+
+# Figure captions
+
+**Figure 1** (A) Example of simulated branching networks.
+Node color is reflective of the disturbance value calculated in network generation.
+(B) Density distributions of disturbance intensity (i.e., proportional reduction of a population).
+Three networks of 50 patches were produced with varied branching probabilities, which were shown in different colors.
+The vertical line indicates the mean disturbance intensity at headwaters ($\mu_m = 0.8$).
+(C) Schematic diagram of trophic interactions in a three-species community.
+$B$ refers to a basal species whose population level is controlled by a density-dependent function.
+The consumer species (or intraguild prey), $C$, consumes $B$ as a resource.
+The predator (or intraguild predator), $P$, is able to consume both $B$ and $C$.
+The structure and strength of these trophic interactions determined by functional responses.
+On the far left, a simple linear food chain exists when the predator does not consume the basal species.
+Moving to the right, basal species' biomass is converted to predator biomass at higher efficiencies with greater attack rates.
+(D) Proportional contribution of basal species $B$ to predator $P$ under different omnivory scenarios.
+The contribution changes with the relative density of $B$ to $C$ ($C=50$ in this example), and the rate of change is determined by attack rate and handling time.
+
+**Figure 2** Relationships between food chain length and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
+Lines indicate loess curves, and shades are associated 95% confidence intervals.
+Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
+Different line colors indicate different disturbance probabilities.
+
+**Figure 3** Relationships between food chain length and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
+Lines indicate loess curves, and shades are associated 95% confidence intervals.
+Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
+Different line colors indicate different disturbance probabilities.
+
+**Figure 4** Relationships between patch state occupancy and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
+Lines indicate loess curves, and shades are associated 95% confidence intervals.
+Panels indicate different food web motifs.
+In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios.
+Different line colors indicate different disturbance probabilities.
+
+**Figure 5** Relationships between patch state occupancy and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
+Lines indicate loess curves, and shades are associated 95% confidence intervals.
+Panels indicate different food web motifs.
+In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios.
+Different line colors indicate different disturbance probabilities.

### 2023-05-16 rmd/reply2.Rmd
edit: text

@@
 mainfont: Times New Roman
+fontsize: 12pt
 ---
@@
 
-> This is a great idea! We included a new hierarchical tree figure to show how different factors detemine the FCL response to ecosystem size/complexity (Figure 2 in the revised MS).
+> This is a great idea! We included a new hierarchical tree figure to show how different factors determine the FCL response to ecosystem size/complexity (Figure 2 in the revised MS).
 Also, we revised the figure captions to better guide readers.
@@
 "*...is food chain length a product of the dynamic interactions that can be influenced by the magnitude of the trophic contributions from different trophic levels.*" This is how we defined FCL in our model, which was articulated in equations 12 & 13.
+The FCL can change due to (1) the P's degree of omnivory, and/or (2) species extinction.
+We have clarified these in the Method.
 
@@
 
-> We appreciate your feedback. The figure captions were editted so the readers can better understand what those figures mean.
-In particular, we have reworded the confusing term "occupancy" to "proportion of patch state" to clarify the y-axis.
+> We greatly appreciate your input. We have made improvements to the figure captions in order to enhance reader comprehension regarding the meaning of the figures.
+
+> Specifically, we have carefully edited the captions to provide a clearer understanding of the y-axis by replacing the confusing term "occupancy" with the more explicit phrase "proportion of patch state."
 
@@
 > Thank you, please let us clarify.
-First, we defined FCL as trophic position, not as trophic level. Thus, the FCL can take continuous values.
-With omnivory, the max is greater than 2 becuase of the coexistence of C and P.
-P's trophic level depends on the relative abundance of B.
-Thus, with omnivory, the FCL can be shorter than the chain motif even if the same set of species coexist.
+We defined FCL as trophic position, not trophic level.
+Thus, the FCL can take continuous values.
+With omnivory, the max is greater than 2. This occurs when P and C coexist.
+P's trophic level depends on the relative abundance of B and C (see equations 12 & 13).
+Thus, with omnivory, the FCL is unavoidably shorter than the chain motif even if the same set of species coexist.
 
@@
 "*Looking at the figure 1 c (and from experience) the expectation is that as omnivory increases the FCL decreases.*" -- The response of FCL to increasing omnivory can be more complex.
-Weak omnivory can cause similar or even lower FCL than strong omnivory because the intraguild predator has a greater predation pressure on intraguild prey.
+Weak omnivory can cause similar or even lower FCL than strong omnivory because the intraguild predator can have a greater predation pressure on intraguild prey (due to higher attack rate to C).
 For the definition of FCL in our study, please see equations 12 & 13 -- it accounts for relative availability of B and C in a given habitat patch.
+Please see also our response to your earlier comments.
 
@@
 > Thank you, we have revised figure captions.
+Specifically, we have carefully edited the captions to provide a clearer understanding of the y-axis by replacing the confusing term "occupancy" with the more explicit phrase "proportion of patch state."
 
@@
 
-> Thank you! We have now added a color scale that indicate the relative reliance on B and C.
+> Thank you. We have now added a color scale that indicates the relative reliance on B and C.
 Meanwhile, we avoided to include FCL in the figure; this is because the relative reliance per se may not necessarily determine the FCL.
@@
 
-> Agreed -- there are two levels of productivity in our model, and this set up was unavoidable.
-In the meantime, we feel that too much emphasis on this part may over-complicate the content, which is complex enough already (as you pointed out earlier in this round of review).
-In looking at the one-patch simulation, the FCL response to increased productivity can be well understood based on the single-patch analog.
-Thus, we decided not to delve into the specifics on this issue.
+> We appreciate your understanding regarding the two levels of productivity in our model
+
+> However, we feel that placing excessive emphasis on this particular aspect might further complicate an already complex content, as you astutely pointed out earlier during this round of review.
+
+> Upon analyzing the one-patch simulation, we found that the response of the FCL to increased productivity at the network level aligned well with that of a single-patch system.
+Consequently, we made a decision not to delve into the specifics of this issue.
+Thank you for your valuable insights though!
 
@@
 
-> We are not sure if we correctly understood this comment, but we interpreted this comment suggesting to move the patch state figures to the maintext.
-We have now moved those figures to the maintext.
-In the meantime, we kept the text to explain those figures in the result section, since it seems inappropriate to mention those findings in the discussion section without mentioning them in the result.
+> We are uncertain if we have fully grasped the intended meaning of this comment, but our understanding is that it suggests relocating the patch state figures to the main text.
+Consequently, we have moved these figures to the main body of the manuscript.
+
+> Simultaneously, we have retained the accompanying text that explains the figures in the results section.
+This decision was made because it would be inappropriate to discuss these findings in the discussion section without prior mention of them in the results section.
+
+> Thank you for bringing this matter to our attention, and we apologize for any confusion or misinterpretation that may have arisen.
 
@@
 
-> We have defined the term in the Intro.
+> Thank you.
+In the previous version of the manuscript, we have introduced "ecosystem geometry" as the term including ecosystem size and complexity in the Introduction.
+However, it was probably difficult to find.
+We have re-worded so the readers can find the information more easily.

### 2023-05-16 rmd/suppl_info.Rmd
edit: text

@@
 
-```{r fig5, fig.cap="Relationships between patch state occupancy and ecosystem size with high disturbance intensity $\\mu_m = 0.8$.
-Lines indicate loess curves, and shades are associated 95% confidence intervals.
-Panels indicate different food web motifs.
-In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios.
-Different line colors indicate different disturbance probabilities."}
-knitr::include_graphics(here::here("figure/figure_si_state_n_patch.pdf"))
-```
-
-\newpage
-
-```{r fig6, fig.cap="Relationships between patch state occupancy and ecosystem complexity with high disturbance intensity $\\mu_m = 0.8$.
-Lines indicate loess curves, and shades are associated 95% confidence intervals.
-Panels indicate different food web motifs.
-In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios.
-Different line colors indicate different disturbance probabilities."}
-knitr::include_graphics(here::here("figure/figure_si_state_p_branch.pdf"))
-```
-
-\newpage
-
 ```{r fig7, fig.cap="Effects of productivity (left column) and dispersal (right column) on food chain length grouped by disturbance frequency ($p_m$; x-axis).

### 2023-05-17 code/figure_patch_state.R
edit: figure

@@ -18,7 +18,13 @@ df_plot <- df_sim %>%
          (s == 0 & a_bp == 0) | s == s_set) %>% 
   pivot_longer(cols = paste0("s", 0:4),
                values_to = "p_state",
-               names_to = "state")
+               names_to = "state") %>% 
+  mutate(state = case_when(state == "s0" ~ "No species",
+                           state == "s1" ~ "B",
+                           state == "s2" ~ "B + C",
+                           state == "s3" ~ "B + P",
+                           state == "s4" ~ "B + C + P"),
+         state = fct_relevel(state, "No species", "B", "B + C", "B + P"))
 
 
 # plot: state transition --------------------------------------------------
@@ -28,11 +34,11 @@ theme_set(plt_theme)
 lab <- c(`8` = "Low~productivity",
          `16` = "High~productivity")
 
-state_lab <- c(`s0` = "No species",
-               `s1` = "B",
-               `s2` = "B + C",
-               `s3` = "B + P",
-               `s4` = "B + C + P")
+pm_lab <- c(`0` = "p[m]==0",
+            `0.05` = "p[m]==0.05",
+            `0.1` = "p[m]==0.10",
+            `0.15` = "p[m]==0.15",
+            `0.2` = "p[m]==0.20")
 
 ## ecosystem size effect
 list_g_np <- foreach(x = unique(df_plot$omn)) %do% {
@@ -41,21 +47,20 @@ list_g_np <- foreach(x = unique(df_plot$omn)) %do% {
     filter(omn == x) %>% 
     ggplot(aes(x = n_patch,
                y = p_state,
-               color = factor(p_disturb),
-               fill = factor(p_disturb))) +
-    #geom_point(alpha = 0.1) +
-    geom_smooth(method = "loess") +
-    facet_grid(rows = vars(state),
+               color = factor(state),
+               fill = factor(state))) +
+    geom_smooth(method = "loess",
+                linewidth = 0.75) +
+    facet_grid(rows = vars(p_disturb),
                cols = vars(disp, r_b),
-               #scales = "free",
                labeller = labeller(r_b = as_labeller(lab, label_parsed),
-                                   state = state_lab)) +
-    labs(y = "Proportion of patch state",
+                                   p_disturb = as_labeller(pm_lab, label_parsed))) +
+    labs(y = expression("Proportion of patch state"~bar(psi)),
          x = "Ecosystem size (number of habitat patches)",
-         color = "Disturbance prob.",
[…]

### 2023-05-17 rmd/manuscript.Rmd
edit: text

@@
 output: 
-  # word_document:
-  #   reference_docx: "word_format.docx"
-  pdf_document:
-   latex_engine: xelatex
+  word_document:
+    reference_docx: "word_format.docx"
+  # pdf_document:
+  #  latex_engine: xelatex
 knit: (function(inputFile, encoding) {
@@
 We simulate trophic community dynamics within branching habitats to investigate the effects of extrinsic controls of FCL.
-In our study, extrinsic controls include ecosystem geometry [size (number of patches) and complexity (branching probability] and disturbance regime.
+In our study, extrinsic controls include "ecosystem geometry" [size (number of patches) and complexity (branching probability] and disturbance regime.
 In addition, our simulation model incorporates established intrinsic drivers of FCL, including the degree of omnivory, productivity level, and species' dispersal ability, allowing us to assess the context dependency on drivers of FCL.
@@
 
-## State transitions
-
 State transitions largely depended on motif structure and shed light on how FCL changes occurred.
@@
 Ecosystem size increased FCL primarily through additive mechanisms, facilitating the state transition of "B" $\rightarrow$ "B + P" (low productivity) or "No species" $\rightarrow$ "B + P" (high productivity), particularly with frequent disturbance (**Figure 5**).
-
 In contrast, insertion (i.e., the insertion of intermediate trophic species) and additive mechanisms were responsible for the effect of ecosystem complexity.
@@
 
-<!-- ## Productivity effect -->
+## Direct effects of productivity and dispersal
 
-<!-- High productivity strengthened the relationship between FCL and ecosystem geometry.
-In the chain motif, productivity greatly increased the overall FCL (**Figure S7**), thereby shifting the upper bound of FCL (**Figures 2 and 3**).
-In contrast, increased productivity decreased FCL with weak and strong omnivory (**Figures 2, 3, and S7**).
-This negative effect of increased productivity was apparent in smaller and/or simpler ecosystems, steepening the slopes between FCL and ecosystem geometry (**Figures 2 and 3**).
--->
-
-<!-- State transitions associated with increased productivity depended on the motif structure.
+In the chain motif, productivity greatly increased the overall FCL (**Figure S5**), thereby shifting the upper bound of FCL (**Figures 3 and 4**).
+In contrast, increased productivity decreased FCL with weak and strong omnivory (**Figures 3, 4, and S5**).
+This negative effect of increased productivity was apparent in smaller and/or simpler ecosystems, steepening the slopes between FCL and ecosystem geometry (**Figures 3 and 4**).
+State transitions associated with increased productivity depended on the motif structure.
 In the chain motif, the transition from "B + C" to "B + C + P" was dominant regardless of disturbance frequency (**Figure S8**).
@@
 Instead, productivity significantly increased the proportion of the "No species" state (**Figure S8**) in disturbed scenarios, especially in small and/or simple ecosystems (**Figures S5 and S6**).
--->
 
-<!-- ## Dispersal effect -->
-
-<!-- Dispersal distance had a positive influence on FCL regardless of simulation scenarios (**Figure S7**). However, this direct effect was weak in comparison to other factors. -->
-
-<!-- State transitions occurred through an additive (chain) or insertion mechanism (omnivory).
+Dispersal distance had a positive influence on FCL regardless of simulation scenarios (**Figure S5**).
+State transitions occurred through an additive (chain) or insertion mechanism (omnivory).
 In the chain motif, the transitions were either "B" $\rightarrow$ "B + C" or "B + C" $\rightarrow$ "B + C+ P", depending on disturbance frequency (**Figure S9**).
 In the omnivory motifs, the predominant transition was from "B + P" to "B + C + P" (**Figure S9**).
--->
+However, this direct effect was weak in comparison to other factors with no qualitative alteration to the relationships between FCL and ecosystem geometry (**Figures 3 and 4**).
 
@@
 
-```{r}
-source(here::here("code/table_param.R"))
+<!-- ```{r} -->
+<!-- source(here::here("code/table_param.R")) -->
 
-knitr::kable(df_value %>%
-               dplyr::select(-"One-patch") %>% 
-               rename(Value = "Main"),
-             escape = FALSE,
-             booktabs = TRUE,
-             caption = "Parameter values used in the main simulation") %>% 
-  kableExtra::kable_styling(latex_options = 'HOLD_position')
-```
+<!-- knitr::kable(df_value %>% -->
+<!--                dplyr::select(-"One-patch") %>%  -->
+<!--                rename(Value = "Main"), -->
+<!--              escape = FALSE, -->
+<!--              booktabs = TRUE, -->
+<!--              caption = "Parameter values used in the main simulation") %>%  -->
+<!--   kableExtra::kable_styling(latex_options = 'HOLD_position') -->
+<!-- ``` -->
 
@@
 Moving to the right, basal species' biomass is converted to predator biomass at higher efficiencies with greater attack rates.
+Arrow colors schematically indicate the relative reliance of P on B and C, with black indicating trophic specialization.
 (D) Proportional contribution of basal species $B$ to predator $P$ under different omnivory scenarios.
@@
 
-**Figure 2** Relationships between food chain length and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
-Lines indicate loess curves, and shades are associated 95% confidence intervals.
+**Figure 2** Summary of our simulation results.
+Black text in each node indicates the potential factors that may influence the association between food chain length and ecosystem geometry, and black text on arrows indicates simulated scenarios for a given factor.
+Blue text denotes the resultant association between food chain length and ecosystem geometry.
+Panel (A) indicates the effect of ecosystem size.
+Panel (B) indicates the effect of ecosystem complexity.
+
+**Figure 3** Relationships between food chain length and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
+Lines in each panel indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.
 Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
 Different line colors indicate different disturbance probabilities.
+In this figure, ecosystem complexity values for individual simulation replicates were drawn randomly from a uniform distribution $P_b \sim \text{Unif}(0.01, 0.99)$.
 
-**Figure 3** Relationships between food chain length and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
-Lines indicate loess curves, and shades are associated 95% confidence intervals.
+**Figure 4** Relationships between food chain length and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
+Lines in each panel indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.
 Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
 Different line colors indicate different disturbance probabilities.
+In this figure, ecosystem size values for individual simulation replicates were drawn randomly from a uniform distribution $N_p \sim \text{Unif}(10, 150)$.
 
-**Figure 4** Relationships between patch state occupancy and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
-Lines indicate loess curves, and shades are associated 95% confidence intervals.
+**Figure 5** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch state) and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
+The proportion of each patch state $\bar{\psi}$ is averaged over time for each simulation replicate.
+Lines indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.
 Panels indicate different food web motifs.
-In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios.
+In each panel, rows distinguish disturbance probability $p_m$, and columns indicate different dispersal and productivity scenarios.
 Different line colors indicate different disturbance probabilities.
 
-**Figure 5** Relationships between patch state occupancy and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
-Lines indicate loess curves, and shades are associated 95% confidence intervals.
+**Figure 6** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch state) and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
+The proportion of each patch state $\bar{\psi}$ is averaged over time for each simulation replicate.
+Lines indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.
 Panels indicate different food web motifs.
-In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios.
+In each panel, rows distinguish disturbance probability $p_m$, and columns indicate different dispersal and productivity scenarios.
 Different line colors indicate different disturbance probabilities.

### 2023-05-17 rmd/reply2.Rmd
edit: text

@@
 > This is a great idea! We included a new hierarchical tree figure to show how different factors determine the FCL response to ecosystem size/complexity (Figure 2 in the revised MS).
-Also, we revised the figure captions to better guide readers.
+We revised the results section to reflect the addition of this new figure.
+Also, we editted figure captions as appropriate.
 
@@
 > We greatly appreciate your input. We have made improvements to the figure captions in order to enhance reader comprehension regarding the meaning of the figures.
-
-> Specifically, we have carefully edited the captions to provide a clearer understanding of the y-axis by replacing the confusing term "occupancy" with the more explicit phrase "proportion of patch state."
+Specifically, we have carefully edited the captions to provide a clearer understanding of the y-axis by replacing the confusing term "occupancy" with the more explicit phrase "proportion of patch state." We also changed the presentation so the readers can understand what transitions happened (color codes changed).
 
@@
 With omnivory, the max is greater than 2. This occurs when P and C coexist.
-P's trophic level depends on the relative abundance of B and C (see equations 12 & 13).
+P's trophic position depends on the relative abundance of B and C (see equations 12 & 13).
 Thus, with omnivory, the FCL is unavoidably shorter than the chain motif even if the same set of species coexist.
@@
 > Thank you for your input.
-"*Looking at the figure 1 c (and from experience) the expectation is that as omnivory increases the FCL decreases.*" -- The response of FCL to increasing omnivory can be more complex.
+"*Looking at the figure 1 c (and from experience) the expectation is that as omnivory increases the FCL decreases.*" -- After some thought, we realized that the response of FCL to increasing omnivory can be more complex.
 Weak omnivory can cause similar or even lower FCL than strong omnivory because the intraguild predator can have a greater predation pressure on intraguild prey (due to higher attack rate to C).
-For the definition of FCL in our study, please see equations 12 & 13 -- it accounts for relative availability of B and C in a given habitat patch.
+We have rephrased the corresponding paragraph in the Introduction to reflect this.
+
+> For the definition of FCL in our study, please see equations 12 & 13 -- it accounts for relative availability of B and C in a given habitat patch.
 Please see also our response to your earlier comments.
@@
 
-> Thank you, we have revised figure captions.
-Also, we clarified the values of ecosystem size or complexity in the main figures (both are randomly drawn from uniform distributions).
+> Thank you, we have revised figure captions throughout the MS.
+We clarified the values of ecosystem size or complexity in the main figures (both are randomly drawn from uniform distributions).
 
@@
 Specifically, we have carefully edited the captions to provide a clearer understanding of the y-axis by replacing the confusing term "occupancy" with the more explicit phrase "proportion of patch state."
+
+> Also, we thought those state transtion figures are better placed in the main text so the readers can see the associated text more easily.
+We have moved those figures to the main text.
 
@@
 
-> We appreciate your understanding regarding the two levels of productivity in our model
-
-> However, we feel that placing excessive emphasis on this particular aspect might further complicate an already complex content, as you astutely pointed out earlier during this round of review.
+> We appreciate your understanding regarding the two levels of productivity in our model.
+In the meantime, we feel that placing excessive emphasis on this particular aspect might further complicate an already complex content, as you astutely pointed out earlier during this round of review.
 
@@
 However, it was probably difficult to find.
-We have re-worded so the readers can find the information more easily.
+We have editted text so the readers can find the information more easily.

### 2023-05-17 rmd/suppl_info.Rmd
edit: text

@@
 ```{r fig3, fig.cap="Relationships between food chain length and ecosystem size with low disturbance intensity $\\mu_m = 0.2$.
-Lines indicate loess curves, and shades are associated 95% confidence intervals.
+Lines in each panel indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.
 Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
-Different line colors indicate different disturbance probabilities."}
+Different line colors indicate different disturbance probabilities.
+In this figure, ecosystem complexity values for individual simulation replicates were drawn randomly from a uniform distribution $P_b \\sim \\text{Unif}(0.01, 0.99)$."}
 knitr::include_graphics(here::here("figure/figure_si_n_patch.pdf"))
@@
 ```{r fig4, fig.cap="Relationships between food chain length and ecosystem complexity with low disturbance intensity $\\mu_m = 0.2$.
-Lines indicate loess curves, and shades are associated 95% confidence intervals.
+Lines in each panel indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.
 Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
-Different line colors indicate different disturbance probabilities."}
+Different line colors indicate different disturbance probabilities.
+In this figure, ecosystem complexity values for individual simulation replicates were drawn randomly from a uniform distribution $N_p \\sim \\text{Unif}(10, 150)$."}
 knitr::include_graphics(here::here("figure/figure_si_p_branch.pdf"))
@@
 
-```{r fig7, fig.cap="Effects of productivity (left column) and dispersal (right column) on food chain length grouped by disturbance frequency ($p_m$; x-axis).
+```{r fig5, fig.cap="Effects of productivity (left column) and dispersal (right column) on food chain length grouped by disturbance frequency ($p_m$; x-axis).
+Box plots show median (horizontal center line) and quartiles (box limits) of $500$ simulation replicates with whiskers extending up to the most extreme data points that are within ±1.5 interquartile range.
+Dots are outliers extending beyond ±1.5 interquartile range.
 Colors distinguish productivity or dispersal scenarios, and rows identify food web motifs.
@@
 
-```{r fig8, fig.cap="Effects of productivity on state transitions with high disturbance intensity $\\mu_m = 0.8$.
+```{r fig6, fig.cap="Effects of productivity on state transitions with high disturbance intensity $\\mu_m = 0.8$.
+The proportion of each patch state $\\bar{\\psi}$ is averaged over time for each simulation replicate.
+Box plots show median (horizontal center line) and quartiles (box limits) of $500$ simulation replicates with whiskers extending up to the most extreme data points that are within ±1.5 interquartile range.
+Dots are outliers extending beyond ±1.5 interquartile range.
 Colors distinguish productivity scenarios, columns represent different disturbance probability $p_m$, and rows identify food web motifs."}
@@
 
-```{r fig9, fig.cap="Effects of dispersal on state transitions with high disturbance intensity $\\mu_m = 0.8$.
+```{r fig7, fig.cap="Effects of dispersal on state transitions with high disturbance intensity $\\mu_m = 0.8$.
+The proportion of each patch state $\\bar{\\psi}$ is averaged over time for each simulation replicate.
+Box plots show median (horizontal center line) and quartiles (box limits) of $500$ simulation replicates with whiskers extending up to the most extreme data points that are within ±1.5 interquartile range.
+Dots are outliers extending beyond ±1.5 interquartile range.
 Colors distinguish dispersal scenarios, columns represent different disturbance probability $p_m$, and rows identify food web motifs."}
@@
 
-```{r fig10, fig.cap="Ecosystem size and complexity dictate total carrying capacity (= $\\sum_x K_x$) and mean distance between a given pair of habitat patches within a network."}
+```{r fig8, fig.cap="Ecosystem size and complexity dictate total carrying capacity (= $\\sum_x K_x$) and mean distance between a given pair of habitat patches within a network."}
 knitr::include_graphics(here::here("figure/figure_si_k_distance.pdf"))

### 2023-05-17 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 **Figure 1** (A) Example of simulated branching networks.
-Node color is reflective of the disturbance value calculated in network generation.
+Node color reflects the disturbance value calculated during network generation.
 (B) Density distributions of disturbance intensity (i.e., proportional reduction of a population).
-Three networks of 50 patches were produced with varied branching probabilities, which were shown in different colors.
-The vertical line indicates the mean disturbance intensity at headwaters ($\mu_m = 0.8$).
+Three networks of 50 patches were produced with varied branching probabilities, shown in different colors.
+The vertical line indicates the mean disturbance intensity at the headwaters ($\mu_m = 0.8$).
 (C) Schematic diagram of trophic interactions in a three-species community.
@@
 The predator (or intraguild predator), $P$, is able to consume both $B$ and $C$.
-The structure and strength of these trophic interactions determined by functional responses.
+The structure and strength of these trophic interactions are determined by functional responses.
 On the far left, a simple linear food chain exists when the predator does not consume the basal species.
-Moving to the right, basal species' biomass is converted to predator biomass at higher efficiencies with greater attack rates.
-Arrow colors schematically indicate the relative reliance of P on B and C, with black indicating trophic specialization.
+Moving to the right, the basal species' biomass is converted to predator biomass at higher efficiencies with greater attack rates.
+Arrow colors schematically indicate the relative reliance of $P$ on $B$ and $C$, with black indicating trophic specialization.
 (D) Proportional contribution of basal species $B$ to predator $P$ under different omnivory scenarios.
-The contribution changes with the relative density of $B$ to $C$ ($C=50$ in this example), and the rate of change is determined by attack rate and handling time.
+The contribution changes with the relative density of $B$ to $C$ ($C = 50$ in this example), and the rate of change is determined by attack rate and handling time.
 
@@
 
-**Figure 5** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch state) and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
+**Figure 5** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
 The proportion of each patch state $\bar{\psi}$ is averaged over time for each simulation replicate.
@@
 
-**Figure 6** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch state) and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
+**Figure 6** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
 The proportion of each patch state $\bar{\psi}$ is averaged over time for each simulation replicate.
-Lines indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.
+Lines indicate loess curves fitted to 500 simulation replicates, and shades are associated with 95% confidence intervals.
 Panels indicate different food web motifs.

### 2023-05-17 code/figure_patch_state.R
Update figure_patch_state.R

@@ -82,7 +82,8 @@ list_g_bp <- foreach(x = unique(df_plot$omn)) %do% {
                y = p_state,
                color = factor(state),
                fill = factor(state))) +
-    geom_smooth(method = "loess") +
+    geom_smooth(method = "loess",
+                linewidth = 0.75) +
     facet_grid(rows = vars(p_disturb),
                cols = vars(disp, r_b),
                labeller = labeller(r_b = as_labeller(lab, label_parsed),

### 2023-05-17 rmd/reply2.Rmd
Update reply2.Rmd

@@
 *This is a thorough revision of the previous manuscript (ms) and has taken into account all of my previous criticisms.
-In particular, I highly appreciate the full reformulation and analysis of the simulation model, which now seems theoretically sound.
-One thing that I noticed is that the new model assumes non-overlapping generations of consumer (C) and predator (P).
+In particular, I highly appreciate the full reformulation and analysis of the simulation model, which now seems theoretically sound.*
+
+> Thank you, your input was very helpful.
+
+*One thing that I noticed is that the new model assumes non-overlapping generations of consumer (C) and predator (P).
 Because all consumer and predator individuals die each time step, they go extinct locally (without immigration) in the next time step if their resource species are totally absent to preclude their reproduction.
@@
 
-> XXX
+> True.
+It would be worthwhile to conduct the additional simulations, but we believe the assumption of non-overlapping generations is justifiable.
+Many aquatic species, if not all, have semelparous life histories.
+We have mentioned this fact in the methods section.
 
@@
 
-> This is a great idea! We included a new hierarchical tree figure to show how different factors determine the FCL response to ecosystem size/complexity (Figure 2 in the revised MS).
-We revised the results section to reflect the addition of this new figure.
-Also, we editted figure captions as appropriate.
+> Great!
+We have incorporated a new hierarchical tree figure (Figure 2) into the revised manuscript.
+This figure visually demonstrates the influence of different factors on the response of food chain length (FCL) to ecosystem size/complexity, providing a comprehensive understanding of the relationships involved.
+
+> Furthermore, we have revised the results section to incorporate the insights and findings presented in the new figure.
+This ensures that the revised manuscript accurately reflects the added information and maintains a cohesive narrative.
+
+> In addition to incorporating the new figure, we have also made appropriate edits to the figure captions.
+This ensures that the captions are clear, informative, and provide the necessary details for readers to interpret the figures accurately.
+
+> These updates strengthen the manuscript by providing a clearer focus on the determinants of FCL in relation to ecosystem characteristics.
+The hierarchical tree figure, along with the revised results section and improved figure captions, enhances the overall clarity and impact of the manuscript.
 
@@
 
-> Our unclear explanation may have confused the reviewer.
-There are a number of cases where all species (B, C, P) coexist.
-"*...is food chain length a product of the dynamic interactions that can be influenced by the magnitude of the trophic contributions from different trophic levels.*" This is how we defined FCL in our model, which was articulated in equations 12 & 13.
-The FCL can change due to (1) the P's degree of omnivory, and/or (2) species extinction.
-We have clarified these in the Method.
+> Thank you for pointing out the need for clarification.
+
+> Regarding the effect of omnivory on food chain length (FCL) in the simulations, it is important to note that there are indeed scenarios in which all trophic species (B, C, P) coexist.
+In these cases, the dynamic interactions and trophic contributions from different trophic positions influence the FCL.
+Our model defines FCL based on these dynamic interactions, as articulated in equations 12 and 13, which take into account the trophic contributions from different trophic positions.
+
+> In the simulations, FCL can change through two main mechanisms: (1) the degree of omnivory exhibited by the predator (P), and (2) species extinction.
+Changes in the magnitude of trophic contributions and the structure of trophic interactions can influence the FCL.
+
+> We have clarified these in the Methods.
 
@@
 
-> We greatly appreciate your input. We have made improvements to the figure captions in order to enhance reader comprehension regarding the meaning of the figures.
-Specifically, we have carefully edited the captions to provide a clearer understanding of the y-axis by replacing the confusing term "occupancy" with the more explicit phrase "proportion of patch state." We also changed the presentation so the readers can understand what transitions happened (color codes changed).
+> Thank you for providing your valuable input. We sincerely appreciate it.
+
+> In response to your feedback, we have made significant improvements to the figure captions to ensure better reader comprehension and understanding of the figures.
+Specifically, we have taken care to edit the captions in a way that clarifies the meaning of the y-axis.
+To achieve this, we have replaced the potentially confusing term "occupancy" with the more explicit phrase "proportion of patch state." This modification aims to provide a clearer understanding of the data presented in the figures.
+
+> Additionally, we have modified the presentation of the figures to make it easier for readers to discern the transitions that occurred.
+This was achieved by changing the color codes to more effectively represent the different transitions and highlight the relevant information (see Figures 5 & 6 in the revised MS, for example).
+
+> By making these changes to the figure captions and presentation, we believe that readers will have an improved understanding of the figures and the associated data.
 
@@
 
-> Thank you, please let us clarify.
-We defined FCL as trophic position, not trophic level.
-Thus, the FCL can take continuous values.
-With omnivory, the max is greater than 2. This occurs when P and C coexist.
-P's trophic position depends on the relative abundance of B and C (see equations 12 & 13).
-Thus, with omnivory, the FCL is unavoidably shorter than the chain motif even if the same set of species coexist.
+> In our study, we defined FCL based on trophic position rather than trophic level.
+This distinction is important because trophic position allows for a continuous range of values, while trophic level typically represents discrete categories.
+
+> With the presence of omnivory, the maximum FCL can exceed 2.
+This occurs when both the consumer (C) and predator (P) coexist in the system.
+The trophic position of the predator (P) is influenced by the relative abundances of the basal species (B) and the consumer species (C), as described in equations 12 and 13.
+
+> It is worth noting that even when the same set of species coexist, the FCL in systems with omnivory is inevitably shorter than the chain motif.
+This is because omnivory introduces additional trophic interactions and modifies the flow of energy and resources within the food web.
 
@@
 
-> Thank you for your input.
-"*Looking at the figure 1 c (and from experience) the expectation is that as omnivory increases the FCL decreases.*" -- After some thought, we realized that the response of FCL to increasing omnivory can be more complex.
-Weak omnivory can cause similar or even lower FCL than strong omnivory because the intraguild predator can have a greater predation pressure on intraguild prey (due to higher attack rate to C).
-We have rephrased the corresponding paragraph in the Introduction to reflect this.
+> Thank you for your valuable input.
+Upon careful consideration, we have come to realize that the relationship between food chain length (FCL) and increasing omnivory can be more complex than initially anticipated.
 
-> For the definition of FCL in our study, please see equations 12 & 13 -- it accounts for relative availability of B and C in a given habitat patch.
-Please see also our response to your earlier comments.
+> In fact, weak omnivory can lead to similar or even lower FCL compared to strong omnivory.
+This is because the presence of an intraguild predator can exert higher predation pressure on the intraguild prey (C) due to its higher attack rate.
+As a result, the FCL may be reduced in cases of weak omnivory.
+
+> To reflect this insight, we have rephrased the corresponding paragraph in the Introduction to incorporate this understanding and acknowledge the potential complexities of the FCL response to increasing omnivory.
+
+> For a comprehensive understanding of how FCL is defined in our study, we kindly direct your attention to equations 12 and 13.
+These equations consider the relative availability of basal species (B) and consumer species (C) within a given habitat patch, which contributes to the determination of FCL.
+
+> We also encourage you to review our response to your previous comments, as it provides additional information related to the definition and dynamics of FCL in our study.
 
@@
 > Thank you, we have revised figure captions throughout the MS.
-We clarified the values of ecosystem size or complexity in the main figures (both are randomly drawn from uniform distributions).
+We clarified the values of ecosystem size or complexity in the main figures (both are randomly drawn from uniform distributions, as described in the Methods section).
 
@@
 
-> Thank you, we have revised figure captions.
-Specifically, we have carefully edited the captions to provide a clearer understanding of the y-axis by replacing the confusing term "occupancy" with the more explicit phrase "proportion of patch state."
+> Thank you for your feedback. We greatly appreciate it.
 
-> Also, we thought those state transtion figures are better placed in the main text so the readers can see the associated text more easily.
-We have moved those figures to the main text.
+> In response to your suggestion, we have made significant improvements to the figure captions to ensure a clearer understanding of the y-axis.
+To achieve this, we have replaced the potentially confusing term "occupancy" with the more explicit phrase "proportion of patch state." This modification aims to provide a clearer and more accessible description of the data presented in the figures.
+
+> Furthermore, we agree with your recommendation to enhance the visibility of the state transition figures.
+As a result, we have moved these figures from the supplementary material to the main text.
+By placing them in the main text, readers can now easily refer to the associated text while examining the figures, facilitating a more coherent and comprehensive understanding of the results.
 

### 2023-05-17 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 A discrete model is suitable for organisms with predictable reproduction cycles, a common life history for many lotic species, including macroinvertebrates and fish.
+We assume non-overlapping generations of species because many lotic species, if not all, have semelparous life histories.
+
 Our simulation model consists of three species; a basal species $B$ with positive logistic growth up to a carrying capacity, $K$; a primary consumer species or intraguild prey, $C$, which grows based on the consumption of $B$; and an intraguild predator $P$, which can consume both $B$ and $C$ (**Figure 1C**).
@@
 In the chain scenario (no omnivory) with low productivity, ecosystem size had a positive effect on FCL (**Figure 3**), while ecosystem complexity had a negative influence (**Figure 4**).
-These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increased the total carrying capacity, whereas ecosystem complexity decreased it (**Figure S10**).
+These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increased the total carrying capacity, whereas ecosystem complexity decreased it (**Figure S8**).
 However, when considering some levels of omnivory, these patterns characteristic of the chain scenario diminished.
@@
 
-<!-- ```{r} -->
-<!-- source(here::here("code/table_param.R")) -->
+```{r}
+source(here::here("code/table_param.R"))
 
-<!-- knitr::kable(df_value %>% -->
-<!--                dplyr::select(-"One-patch") %>%  -->
-<!--                rename(Value = "Main"), -->
-<!--              escape = FALSE, -->
-<!--              booktabs = TRUE, -->
-<!--              caption = "Parameter values used in the main simulation") %>%  -->
-<!--   kableExtra::kable_styling(latex_options = 'HOLD_position') -->
-<!-- ``` -->
+knitr::kable(df_value %>%
+               dplyr::select(-"One-patch") %>%
+               rename(Value = "Main"),
+             escape = FALSE,
+             booktabs = TRUE,
+             caption = "Parameter values used in the main simulation") %>%
+  kableExtra::kable_styling(latex_options = 'HOLD_position')
+```
 

### 2023-05-17 rmd/manuscript.Rmd
edit

@@
 State transitions associated with increased productivity depended on the motif structure.
-In the chain motif, the transition from "B + C" to "B + C + P" was dominant regardless of disturbance frequency (**Figure S8**).
+In the chain motif, the transition from "B + C" to "B + C + P" was dominant regardless of disturbance frequency (**Figure S6**).
 The introduction of omnivory allowed the transition between "B + P" and "B + C + P".
-With weak omnivory, the transition of "B + C + P" $\rightarrow$ "B + P" was prevalent and shortened FCL across disturbance regimes (**Figure S8**).
+With weak omnivory, the transition of "B + C + P" $\rightarrow$ "B + P" was prevalent and shortened FCL across disturbance regimes (**Figure S6**).
 However, this transition was marginal in the strong omnivory motif.
-Instead, productivity significantly increased the proportion of the "No species" state (**Figure S8**) in disturbed scenarios, especially in small and/or simple ecosystems (**Figures S5 and S6**).
+Instead, productivity significantly increased the proportion of the "No species" state (**Figure S6**) in disturbed scenarios, especially in small and/or simple ecosystems (**Figures 5 and 6**).
 
@@
 State transitions occurred through an additive (chain) or insertion mechanism (omnivory).
-In the chain motif, the transitions were either "B" $\rightarrow$ "B + C" or "B + C" $\rightarrow$ "B + C+ P", depending on disturbance frequency (**Figure S9**).
-In the omnivory motifs, the predominant transition was from "B + P" to "B + C + P" (**Figure S9**).
+In the chain motif, the transitions were either "B" $\rightarrow$ "B + C" or "B + C" $\rightarrow$ "B + C+ P", depending on disturbance frequency (**Figure S7**).
+In the omnivory motifs, the predominant transition was from "B + P" to "B + C + P" (**Figure S7**).
 However, this direct effect was weak in comparison to other factors with no qualitative alteration to the relationships between FCL and ecosystem geometry (**Figures 3 and 4**).
@@
              booktabs = TRUE,
-             caption = "Parameter values used in the main simulation") %>%
-  kableExtra::kable_styling(latex_options = 'HOLD_position')
+             caption = "Parameter values used in the main simulation")
 ```

### 2023-05-17 rmd/reply2.Rmd
edit

@@
 
-This is a re-submission of a manuscript: ECS22-0646 "*Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems.*" Thank you for allowing us to revise our manuscript.
-The comments were very helpful in improving our manuscript.
+This is a re-submission of a manuscript: ECS22-0646 "*Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems.*" We appreciate the time you have taken for our manuscript.
 When revising the manuscript, we have put a particular emphasis on the following points:
 
-XXX
+1. Reorganized the results section to better focus on ecosystem size/complexity effects
+2. Improved figure captions
 

### 2023-05-17 rmd/tables_figures.Rmd
edit rmd
edit

--- 2023-03-09 rmd/figures.Rmd
@@
 **Figure 1** (A) Example of simulated branching networks.
-Node color is reflective of the disturbance value calculated in network generation.
+Node color reflects the disturbance value calculated during network generation.
 (B) Density distributions of disturbance intensity (i.e., proportional reduction of a population).
-Three networks of 50 patches were produced with varied branching probabilities, which were shown in different colors.
-The vertical line indicates the mean disturbance intensity at headwaters ($\mu_m = 0.8$).
-(C) Trophic interactions in a three-species community.
+Three networks of 50 patches were produced with varied branching probabilities, shown in different colors.
+The vertical line indicates the mean disturbance intensity at the headwaters ($\mu_m = 0.8$).
+(C) Schematic diagram of trophic interactions in a three-species community.
 $B$ refers to a basal species whose population level is controlled by a density-dependent function.
@@
 The predator (or intraguild predator), $P$, is able to consume both $B$ and $C$.
-The structure and strength of these trophic interactions determined by functional responses.
+The structure and strength of these trophic interactions are determined by functional responses.
 On the far left, a simple linear food chain exists when the predator does not consume the basal species.
-Moving to the right, basal species' biomass is converted to predator biomass at higher efficiencies with greater attack rates.
+Moving to the right, the basal species' biomass is converted to predator biomass at higher efficiencies with greater attack rates.
+Arrow colors schematically indicate the relative reliance of $P$ on $B$ and $C$, with black indicating trophic specialization.
 (D) Proportional contribution of basal species $B$ to predator $P$ under different omnivory scenarios.
-The contribution changes with the relative density of $B$ to $C$ ($C=50$ in this example), and the rate of change is determined by attack rate and handling time.
+The contribution changes with the relative density of $B$ to $C$ ($C = 50$ in this example), and the rate of change is determined by attack rate and handling time.
 
@@
 ```{r figure2, fig.width=12, fig.height=5.5}
+knitr::include_graphics(here::here("figure/figure_tree.pdf"))
+```
+
+**Figure 2** Summary of our simulation results.
+Black text in each node indicates the potential factors that may influence the association between food chain length and ecosystem geometry, and black text on arrows indicates simulated scenarios for a given factor.
+Blue text denotes the resultant association between food chain length and ecosystem geometry.
+Panel (A) indicates the effect of ecosystem size.
+Panel (B) indicates the effect of ecosystem complexity.
+
+\pagebreak
+
+```{r figure3, fig.width=12, fig.height=5.5}
 knitr::include_graphics(here::here("figure/figure_main_n_patch.pdf"))
@@
 
-**Figure 2** Relationships between food chain length and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
-Lines indicate loess curves, and shades are associated 95% confidence intervals.
+**Figure 3** Relationships between food chain length and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
+Lines in each panel indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.
 Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
+Different line colors indicate different disturbance probabilities.
+In this figure, ecosystem complexity values for individual simulation replicates were drawn randomly from a uniform distribution $P_b \sim \text{Unif}(0.01, 0.99)$.
+
+\pagebreak
+
+```{r figure4, fig.width=12, fig.height=5.5}
+knitr::include_graphics(here::here("figure/figure_main_p_branch.pdf"))
+```
+
+**Figure 4** Relationships between food chain length and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
+Lines in each panel indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.
+Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
+Different line colors indicate different disturbance probabilities.
+In this figure, ecosystem size values for individual simulation replicates were drawn randomly from a uniform distribution $N_p \sim \text{Unif}(10, 150)$.
+
+\pagebreak
+
+```{r figure5, fig.width=12, fig.height=5.5}
+knitr::include_graphics(here::here("figure/figure_state_n_patch.pdf"))
+```
+
+**Figure 5** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
+The proportion of each patch state $\bar{\psi}$ is averaged over time for each simulation replicate.
+Lines indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.
+Panels indicate different food web motifs.
+In each panel, rows distinguish disturbance probability $p_m$, and columns indicate different dispersal and productivity scenarios.
 Different line colors indicate different disturbance probabilities.
@@
 
-```{r figure3, fig.width=12, fig.height=5.5}
-knitr::include_graphics(here::here("figure/figure_main_p_branch.pdf"))
+```{r figure6, fig.width=12, fig.height=5.5}
+knitr::include_graphics(here::here("figure/figure_state_p_branch.pdf"))
 ```
 
-**Figure 3** Relationships between food chain length and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
-Lines indicate loess curves, and shades are associated 95% confidence intervals.
-Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios.
+**Figure 6** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
+The proportion of each patch state $\bar{\psi}$ is averaged over time for each simulation replicate.
+Lines indicate loess curves fitted to 500 simulation replicates, and shades are associated with 95% confidence intervals.
+Panels indicate different food web motifs.
+In each panel, rows distinguish disturbance probability $p_m$, and columns indicate different dispersal and productivity scenarios.
 Different line colors indicate different disturbance probabilities.

### 2023-05-17 code/figure_patch_state.R

@@ -59,8 +59,10 @@ list_g_np <- foreach(x = unique(df_plot$omn)) %do% {
          x = "Ecosystem size (number of habitat patches)",
          color = "Patch state",
          fill = "Patch state") +
-    scale_color_met_d("Austria", direction = -1) +
-    scale_fill_met_d("Austria", direction = -1) +
+    scale_color_viridis_d() +
+    scale_fill_viridis_d() +
+    # scale_color_met_d("Austria", direction = -1) +
+    # scale_fill_met_d("Austria", direction = -1) +
     theme(legend.title = element_text(size = 20),
           legend.text =  element_text(size = 18)) +
     ggtitle(x)
@@ -93,8 +95,10 @@ list_g_bp <- foreach(x = unique(df_plot$omn)) %do% {
          color = "Patch state",
          fill = "Patch state") +
     scale_x_continuous(breaks = c(0.1, 0.4, 0.7, 1.0)) +
-    scale_color_met_d("Austria", direction = -1) +
-    scale_fill_met_d("Austria", direction = -1) +
+    scale_color_viridis_d() +
+    scale_fill_viridis_d() +
+    # scale_color_met_d("Austria", direction = -1) +
+    # scale_fill_met_d("Austria", direction = -1) +
     theme(legend.title = element_text(size = 20),
           legend.text =  element_text(size = 18)) +
     ggtitle(x)

### 2023-06-03 rmd/manuscript.Rmd
edit: minor

@@
 
-\*Corresponding Author: JPF Pomeranz, [jfpomeranz\@gmail.com](mailto:jfpomeranz@gmail.com){.email}; A Terui, [hanabi0111\@gmail.com](mailto:hanabi0111@gmail.com){.email}
+\*Corresponding Author: A Terui, [hanabi0111\@gmail.com](mailto:hanabi0111@gmail.com){.email}
 
@@
 Here, we develop a spatially explicit theoretical model which provides an integrated framework for understanding variation in food chain length in branching networks.
-We show independent, positive influences of ecosystem size and complexity on food chain length.
+We show independent, positive influences of ecosystem size and complexity (as indicated by branching properties) on food chain length.
 However, the effects of ecosystem size and complexity were contingent upon other factors, appearing more clearly in high-disturbance and high-productivity regimes.
@@
 
-# References{-}
+# References {.unnumbered}
 
-<div id="refs"></div>
+::: {#refs}
+:::
 

### 2023-06-03 rmd/reply2.Rmd
edit: minor

@@
 
-> We also encourage you to review our response to your previous comments, as it provides additional information related to the definition and dynamics of FCL in our study.
+> We also encourage you to review our response to your earlier comments in this letter, as it provides additional information related to the definition and dynamics of FCL in our study.
 

### 2023-06-03 code/figure_treeplot.R

@@ -31,7 +31,8 @@ V(adj)$label[c(8:9, 12)] <- "Positive"
 V(adj)$color[c(1:dim(m1)[1])] <- "node"
 V(adj)$color[c(6, 7, 8:12)] <- "pattern"
 
-E(adj)$label[1] <- "Absent"; E(adj)$label[2] <- "Present"
+E(adj)$label[1] <- "Absent \n (Disturb prob. = 0)"
+E(adj)$label[2] <- "Present \n (Disturb prob. > 0)"
 E(adj)$label[3:5] <- c("None (Chain)", "Weak", "Strong")
 E(adj)$label[6:7] <- c("Weak", "Strong")
 E(adj)$label[8:11] <- rep(c("Low", "High"), 2)
@@ -75,7 +76,7 @@ V(adj)$color[1:10] <- "node"
 V(adj)$color[c(5:10)] <- "pattern"
 
 E(adj)$label[1:(dim(m2)[1] - 1)] <- "NA"
-E(adj)$label[1:2] <- c("Absent", "Present")
+E(adj)$label[1:2] <- c("Absent \n (Disturb prob. = 0)", "Present \n (Disturb prob. > 0)")
 E(adj)$label[3:5] <- c("None (Chain)", "Weak", "Strong")
 E(adj)$label[6:7] <- c("Weak", "Strong")
 E(adj)$label[8:9] <- c("Low", "High")

### 2023-06-03 rmd/readme.Rmd
edit: readme

---
title: "README"
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here())
      })
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F)

pacman::p_load(tidyverse)
```

## Article Information

Title: Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks

## File descriptions

`code/` includes all R codes to reproduce findings from this research:

```{r}

list.files(here::here("code")) %>% 
  as_tibble() %>% 
  mutate(Description = case_when(str_detect(value, "^figure_")~"Code for figures: '_concept', Figure 1; '_delta', Figure 1D; '_food_web', Figure 1C; '_treeplot', Figure2; '_main_geom', Figure 3, 4; '_network', Figure 1A; '_patch_state', Figure 5, 6",
                                 str_detect(value, "^figure_si")~"Code for SI figures: '_geom', Figure S3, S4; '_one_patch', Figure S1, S2; '_r_disp', Figure S5; '_within_net', Figure S8",
                                 str_detect(value, "^format_")~"Code for formatting",
                                 str_detect(value, "^library")~"Code for figures",
                                 str_detect(value, "^run_")~"Code for simulations",
                                 str_detect(value, "^table_")~"Code for tables",
                                 str_detect(value, "^set_")~"Code for setting figure themes"
                                 )) %>% 
  rename(Name = value) %>% 
  knitr::kable(format = "markdown")
```

## Session Information

```{r}

source(here::here("code/library.R"))
sessionInfo()

```

### 2023-06-03 code/table_stvy.R

@@ -1,75 +1,0 @@
-
-# setup -------------------------------------------------------------------
-
-rm(list = ls())
-source(here::here("code/library.R"))
-source(here::here("code/analysis_stvy.R"))
-
-
-# table -------------------------------------------------------------------
-
-table_stvy <- foreach(i = 1:2) %do% {
-  fit_sense$fit[[i]] %>% 
-    tidy() %>%
-    mutate(Model = case_when(term %in% c("r_b",
-                                         "s0",
-                                         "e_bc",
-                                         "e_bp",
-                                         "e_cp",
-                                         "a_bc",
-                                         "a_bp",
-                                         "a_cp",
-                                         "h_bc",
-                                         "h_bp",
-                                         "h_cp",
-                                         "theta",
-                                         "p_dispersal",
-                                         "p_disturb") ~ "Food web",
-                             term %in% c("mean_disturb_source",
-                                         "sd_disturb_source",
-                                         "sd_disturb_lon") ~ "Network"),
-           Term = case_when(term == "r_b" ~ "$r_b$",
-                            term == "e_bc" ~"$e_{BC}$",
-                            term == "e_bp" ~"$e_{BP}$",
-                            term == "e_cp" ~"$e_{CP}$",
-                            term == "a_bc" ~"$a_{BC}$",
-                            term == "a_bp" ~"$a_{BP}$",
-                            term == "a_cp" ~"$a_{CP}$",
-                            term == "h_bc" ~"$h_{BC}$",
-                            term == "h_bp" ~"$h_{BP}$",
-                            term == "h_cp" ~"$h_{CP}$",
-                            term == "theta" ~"$\\theta$",
-                            term == "p_disturb" ~"$p_m$",
-                            term == "sd_disturb_source" ~ "$\\sigma_h$",
-                            term == "sd_disturb_lon" ~ "$\\sigma_l$",
-                            term == "s0" ~ "$s_0$",
-                            term == "p_dispersal" ~ "$p_d$"),
-           Interpretation = case_when(term == "r_b" ~ "Reproductive rate of basal species",
-                                      term == "e_bc" ~ "Conversion efficiency (B to C)",
-                                      term == "e_bp" ~ "Conversion efficiency (B to P)",
-                                      term == "e_cp" ~ "Conversion efficiency (C to P)",
-                                      term == "a_bc" ~ "Attack rate (C on B)",
-                                      term == "a_bp" ~ "Attack rate (P on B)",
[…]

### 2023-06-03 code/figure_patch_state.R
edit: figure theme

@@ -40,6 +40,12 @@ pm_lab <- c(`0` = "p[m]==0",
             `0.15` = "p[m]==0.15",
             `0.2` = "p[m]==0.20")
 
+cols <- c("No species" = "grey",
+          "B" = "steelblue",
+          "B + C" = "steelblue",
+          "B + P",
+          "B + C + P")
+
 ## ecosystem size effect
 list_g_np <- foreach(x = unique(df_plot$omn)) %do% {
 
@@ -59,10 +65,8 @@ list_g_np <- foreach(x = unique(df_plot$omn)) %do% {
          x = "Ecosystem size (number of habitat patches)",
          color = "Patch state",
          fill = "Patch state") +
-    scale_color_viridis_d() +
-    scale_fill_viridis_d() +
-    # scale_color_met_d("Austria", direction = -1) +
-    # scale_fill_met_d("Austria", direction = -1) +
+    scale_color_viridis_d(end = 0.9) +
+    scale_fill_viridis_d(end = 0.9) +
     theme(legend.title = element_text(size = 20),
           legend.text =  element_text(size = 18)) +
     ggtitle(x)
@@ -95,10 +99,8 @@ list_g_bp <- foreach(x = unique(df_plot$omn)) %do% {
          color = "Patch state",
          fill = "Patch state") +
     scale_x_continuous(breaks = c(0.1, 0.4, 0.7, 1.0)) +
-    scale_color_viridis_d() +
-    scale_fill_viridis_d() +
-    # scale_color_met_d("Austria", direction = -1) +
-    # scale_fill_met_d("Austria", direction = -1) +
+    scale_color_viridis_d(end = 0.9) +
+    scale_fill_viridis_d(end = 0.9) +
     theme(legend.title = element_text(size = 20),
           legend.text =  element_text(size = 18)) +
     ggtitle(x)

### 2023-06-03 code/figure_si_one_patch.R

@@ -56,7 +56,8 @@ foreach(i = 1:length(a1_set)) %do% {
                  cols = vars(h), labeller = labeller(h = as_labeller(lab[[1]], label_parsed),
                                                      r_b = as_labeller(lab[[2]], label_parsed))) +
       scale_fill_viridis_d(limits = factor(0:4),
-                           labels = state_lab) +
+                           labels = state_lab,
+                           end = 0.9) +
       scale_x_continuous(breaks = c(0.01, 0.05, 0.09)) +
       scale_y_continuous(breaks = c(0, 0.025, 0.05)) +
       theme_classic() +

### 2023-06-03 code/set_figure_theme.R

@@ -3,7 +3,7 @@ library(tidyverse)
 plt_theme <- theme_bw() + theme(
   plot.background = element_blank(),
   
-  panel.background = element_rect(grey(0.96)),
+  panel.background = element_rect(grey(0.98)),
   panel.border = element_rect(fill = NA),
   
   panel.grid = element_blank(),

### 2023-07-10 rmd/manuscript.Rmd
edit: proof correction
edit/correct in response to ecosphere's editorial office

@@
 As a result, the FCL in a single patch can change due to the P's degree of omnivory and/or species extinction.
-We defined the global FCL $\bar{\tau}$ across patches $N_p$ and time $T$ as $\bar{\tau} = \frac{\sum_x \sum_t \tau_{x,t}}{N_p T}$ (i.e., average over space and time).
+We defined the global FCL $\bar{\tau}$ across patches $N_p$ and time $T$ as $\bar{\tau} = \sum_x \sum_t \tau_{x,t}/(N_p T)$ (i.e., average over space and time).
 
 We also calculated the proportion of the patch state ("No species", "B", "B + C", "B + P", and "B + C + P") at each time step to infer mechanisms that drive changes in FCL.
-The time-averaged proportion of patch state $\bar{\psi}$ was obtained as $\bar{\psi} = \frac{q_t}{N_p T}$, where $q_t$ is the number of patches in a given patch state at time step $t$.
+The time-averaged proportion of patch state $\bar{\psi}$ was obtained as $\bar{\psi} = q_t/(N_p T)$, where $q_t$ is the number of patches in a given patch state at time step $t$.
 
@@
 
-Prior to the main simulation, we performed one-patch simulations to explore the predation parameter space ($a_{ij}$ and $h_{ij}$) that allows stable coexistence of the three species without disturbance and dispersal (see **Supporting Information**).
+Prior to the main simulation, we performed one-patch simulations to explore the predation parameter space ($a_{ij}$ and $h_{ij}$) that allows stable coexistence of the three species without disturbance and dispersal (see **Appendix S1: Supplementary Text**).
 In this simulation, we assumed $h_{ij} = h$ to reduce the dimension of parameter space.
@@
 Therefore, disturbance interacted with ecosystem geometry to produce the interdependence of these factors in determining FCL.
-Yet, the interactive effects were not observed with low disturbance intensity (**Figures S3 and S4**); this suggests that disturbance intensity must be strong enough to be influential.
+Yet, the interactive effects were not observed with low disturbance intensity (**Appendix S1: Figures S3 and S4**); this suggests that disturbance intensity must be strong enough to be influential.
 
@@
 In the chain scenario (no omnivory) with low productivity, ecosystem size had a positive effect on FCL (**Figure 3**), while ecosystem complexity had a negative influence (**Figure 4**).
-These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increased the total carrying capacity, whereas ecosystem complexity decreased it (**Figure S8**).
+These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increased the total carrying capacity, whereas ecosystem complexity decreased it (**Appendix S1: Supplementary Text8**).
 However, when considering some levels of omnivory, these patterns characteristic of the chain scenario diminished.
@@
 In the chain motif, an additive mechanism (i.e., the addition of higher trophic species) was always responsible for the changes in FCL.
-With disturbance, state transition occurs from "B" to "B + C" (low productivity) or "B + C" to "B + C + P" (high productivity) as ecosystem size or complexity increases (**Figures 5 and 6**).
+With disturbance, state transition occurs from "B" to "B + C" (low productivity) or "B + C" to "B + C + P" (high productivity) as ecosystem size or complexity increases (**Figures 5 and 8**).
 Exceptionally, the state transition of "B + C + P" $\rightarrow$ "B + C" occurred as ecosystem complexity increased with no disturbance and low productivity, a scenario in which ecosystem complexity decreased FCL.
@@
 
-In the chain motif, productivity greatly increased the overall FCL (**Figure S5**), thereby shifting the upper bound of FCL (**Figures 3 and 4**).
-In contrast, increased productivity decreased FCL with weak and strong omnivory (**Figures 3, 4, and S5**).
+In the chain motif, productivity greatly increased the overall FCL (**Appendix S1: Figure S5**), thereby shifting the upper bound of FCL (**Figures 3 and 4**).
+In contrast, increased productivity decreased FCL with weak and strong omnivory (**Figures 3, 4, and Appendix S1: Figure S5**).
 This negative effect of increased productivity was apparent in smaller and/or simpler ecosystems, steepening the slopes between FCL and ecosystem geometry (**Figures 3 and 4**).
 State transitions associated with increased productivity depended on the motif structure.
-In the chain motif, the transition from "B + C" to "B + C + P" was dominant regardless of disturbance frequency (**Figure S6**).
+In the chain motif, the transition from "B + C" to "B + C + P" was dominant regardless of disturbance frequency (**Appendix S1: Figure S6**).
 The introduction of omnivory allowed the transition between "B + P" and "B + C + P".
-With weak omnivory, the transition of "B + C + P" $\rightarrow$ "B + P" was prevalent and shortened FCL across disturbance regimes (**Figure S6**).
+With weak omnivory, the transition of "B + C + P" $\rightarrow$ "B + P" was prevalent and shortened FCL across disturbance regimes (**Appendix S1: Figure S6**).
 However, this transition was marginal in the strong omnivory motif.
-Instead, productivity significantly increased the proportion of the "No species" state (**Figure S6**) in disturbed scenarios, especially in small and/or simple ecosystems (**Figures 5 and 6**).
+Instead, productivity significantly increased the proportion of the "No species" state (**Appendix S1: Figure S6**) in disturbed scenarios, especially in small and/or simple ecosystems (**Figures 7 and 10**).
 
-Dispersal distance had a positive influence on FCL regardless of simulation scenarios (**Figure S5**).
+Dispersal distance had a positive influence on FCL regardless of simulation scenarios (**Appendix S1: Figure S5**).
 State transitions occurred through an additive (chain) or insertion mechanism (omnivory).
-In the chain motif, the transitions were either "B" $\rightarrow$ "B + C" or "B + C" $\rightarrow$ "B + C+ P", depending on disturbance frequency (**Figure S7**).
-In the omnivory motifs, the predominant transition was from "B + P" to "B + C + P" (**Figure S7**).
+In the chain motif, the transitions were either "B" $\rightarrow$ "B + C" or "B + C" $\rightarrow$ "B + C+ P", depending on disturbance frequency (**Appendix S1: Figure S7**).
+In the omnivory motifs, the predominant transition was from "B + P" to "B + C + P" (**Appendix S1: Figure S7**).
 However, this direct effect was weak in comparison to other factors with no qualitative alteration to the relationships between FCL and ecosystem geometry (**Figures 3 and 4**).
@@
 In the meantime, ecosystem complexity may increase the connectivity to disturbance refugia.
-Branching shortens the mean distance between patches and, therefore, may enhance immigration from less disturbed patches (**Figure S10**).
+Branching shortens the mean distance between patches and, therefore, may enhance immigration from less disturbed patches (**Appendix S1: Figure S8**).
 The proportion of disturbance refugia may also have increased in more branched networks.
@@
 In particular, with no disturbance and low productivity, the FCL of the chain motif had contrasting relationships with ecosystem size (positive) and complexity (negative).
-It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity (**Figures S1 and S2**); therefore, FCL followed the changes in the total carrying capacity along the gradients of ecosystem size and complexity.
-These relationships diminished in omnivory motifs as the dependence of FCL on local carrying capacity weakened or reversed (**Figures S1 and S2**).
+It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity (**Appendix S1: Figures S1 and S2**); therefore, FCL followed the changes in the total carrying capacity along the gradients of ecosystem size and complexity.
+These relationships diminished in omnivory motifs as the dependence of FCL on local carrying capacity weakened or reversed (**Appendix S1: Figures S1 and S2**).
 Importantly, however, disturbance overwhelmed the role of omnivory in modulating the FCL-geometry relationships: as the frequency of intense disturbance increases, the positive relationships between FCL and ecosystem geometry emerged regardless of omnivory levels.
@@
 However, this was not the case in the strong omnivory motif; the increased proportion of the "No species" state strongly impacted the minimum FCL, and this transition was mediated by disturbance.
-This result suggests that, although the strong omnivory motif is innately stable with high productivity (see one-patch simulations; **Figures S1 and S2**), external disturbance undermines the sensitive balance of predator-prey interactions with the abrupt increase of "No species" state, which must be countered by immigration from less disturbed habitats to maintain food chains.
+This result suggests that, although the strong omnivory motif is innately stable with high productivity (see one-patch simulations; **Appendix S1: Figures S1 and S2**), external disturbance undermines the sensitive balance of predator-prey interactions with the abrupt increase of "No species" state, which must be countered by immigration from less disturbed habitats to maintain food chains.
 
@@
 Black text in each node indicates the potential factors that may influence the association between food chain length and ecosystem geometry, and black text on arrows indicates simulated scenarios for a given factor.
-Blue text denotes the resultant association between food chain length and ecosystem geometry.
+Blue signs at the terminal nodes denote the resultant association between food chain length and ecosystem geometry (+: positive, -: negative, 0: no effect, (0): very weak effect).
 Panel (A) indicates the effect of ecosystem size.
@@
 
-**Figure 5** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem size with high disturbance intensity $\mu_m = 0.8$.
+**Figure 5** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem size under the chain motif scenario.
+Disturbance intensity is $\mu_m = 0.8$.
 The proportion of each patch state $\bar{\psi}$ is averaged over time for each simulation replicate.
-Lines indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.
-Panels indicate different food web motifs.
-In each panel, rows distinguish disturbance probability $p_m$, and columns indicate different dispersal and productivity scenarios.
+Lines indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.Rows distinguish disturbance probability $p_m$, and columns indicate different dispersal and productivity scenarios.
 Different line colors indicate different disturbance probabilities.
 
-**Figure 6** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$.
+**Figure 6** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem size under the weak omnivory scenario.
+Additional specifics can be found in Figure 5.
+
+**Figure 7** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem size under the strong omnivory scenario.
+Additional specifics can be found in Figure 5.
+
+**Figure 8** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem complexity under the chain motif scenario.
+Disturbance intensity is $\mu_m = 0.8$.
 The proportion of each patch state $\bar{\psi}$ is averaged over time for each simulation replicate.
 Lines indicate loess curves fitted to 500 simulation replicates, and shades are associated with 95% confidence intervals.
-Panels indicate different food web motifs.
-In each panel, rows distinguish disturbance probability $p_m$, and columns indicate different dispersal and productivity scenarios.
+Rows distinguish disturbance probability $p_m$, and columns indicate different dispersal and productivity scenarios.
 Different line colors indicate different disturbance probabilities.
+
+**Figure 9** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem complexity under the weak omnivory scenario.
+Disturbance intensity is $\mu_m = 0.8$.
+Additional specifics can be found in Figure 8.
+
+**Figure 10** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem complexity under the strong omnivory scenario.
+Disturbance intensity is $\mu_m = 0.8$.
+Additional specifics can be found in Figure 8.

### 2023-07-10 code/figure_main_geom.R

@@ -22,6 +22,8 @@ df_plot <- df_sim %>%
 
 # plot: geometry effect ---------------------------------------------------
 
+theme_set(plt_theme)
+
 lab <- c(`8` = "Low~productivity",
          `16` = "High~productivity")
 
@@ -37,10 +39,10 @@ g_np <-  df_plot %>%
              cols = vars(disp, r_b),
              scales = "free",
              labeller = labeller(r_b = as_labeller(lab, label_parsed))) +
-  labs(y = "Food chain length",
-       x = "Ecosystem size (number of habitat patches)",
-       color = "Disturbance prob.",
-       fill = "Disturbance prob.") +
+  labs(y = "Food chain length\n",
+       x = "\nEcosystem size (number of habitat patches)",
+       color = "Disturb prob.",
+       fill = "Disturb prob.") +
   scale_color_met_d("Hiroshige", direction = -1) +
   scale_fill_met_d("Hiroshige", direction = -1)
 
@@ -56,10 +58,10 @@ g_pb <-  df_plot %>%
              cols = vars(disp, r_b),
              scales = "free",
              labeller = labeller(r_b = as_labeller(lab, label_parsed))) +
-  labs(y = "Food chain length",
-       x = "Ecosystem complexity (branching probability)",
-       color = "Disturbance prob.",
-       fill = "Disturbance prob.") +
+  labs(y = "Food chain length\n",
+       x = "\nEcosystem complexity (branching probability)",
+       color = "Disturb prob.",
+       fill = "Disturb prob.") +
   scale_x_continuous(breaks = c(0.1, 0.4, 0.7, 1)) +
   scale_color_met_d("Hiroshige", direction = -1) +
   scale_fill_met_d("Hiroshige", direction = -1)

### 2023-07-10 code/figure_patch_state.R

@@ -48,7 +48,7 @@ cols <- c("No species" = "grey",
 
 ## ecosystem size effect
 list_g_np <- foreach(x = unique(df_plot$omn)) %do% {
-
+  # atop(): line break for expression at comma
   g_np <-  df_plot %>% 
     filter(omn == x) %>% 
     ggplot(aes(x = n_patch,
@@ -61,22 +61,21 @@ list_g_np <- foreach(x = unique(df_plot$omn)) %do% {
                cols = vars(disp, r_b),
                labeller = labeller(r_b = as_labeller(lab, label_parsed),
                                    p_disturb = as_labeller(pm_lab, label_parsed))) +
-    labs(y = expression("Proportion of patch state"~bar(psi)),
-         x = "Ecosystem size (number of habitat patches)",
+    labs(y = expression(atop("Proportion of patch state"~bar(psi),)),
+         x = "\nEcosystem size (number of habitat patches)",
          color = "Patch state",
          fill = "Patch state") +
     scale_color_viridis_d(end = 0.9) +
     scale_fill_viridis_d(end = 0.9) +
     theme(legend.title = element_text(size = 20),
-          legend.text =  element_text(size = 18)) +
-    ggtitle(x)
+          legend.text =  element_text(size = 18))
 }
 
-g_np_all <- list_g_np[[1]] + list_g_np[[2]] + 
-  list_g_np[[3]] + guide_area() +
-  plot_annotation(tag_levels = "A") +
-  plot_layout(guides = "collect", design = "AC
-                                            BD")
+# g_np_all <- list_g_np[[1]] + list_g_np[[2]] + 
+#   list_g_np[[3]] + guide_area() +
+#   plot_annotation(tag_levels = "A") +
+#   plot_layout(guides = "collect", design = "AC
+#                                             BD")
 
 
 ## ecosystem complexity effect
@@ -94,33 +93,42 @@ list_g_bp <- foreach(x = unique(df_plot$omn)) %do% {
                cols = vars(disp, r_b),
                labeller = labeller(r_b = as_labeller(lab, label_parsed),
                                    p_disturb = as_labeller(pm_lab, label_parsed))) +
-    labs(y = expression("Proportion of patch state"~bar(psi)),
-         x = "Ecosystem complexity (branching prob.)",
+    labs(y = expression(atop("Proportion of patch state"~bar(psi),)),
+         x = "\nEcosystem complexity (branching prob.)",
          color = "Patch state",
[…]

### 2023-07-10 code/figure_treeplot.R

@@ -24,9 +24,9 @@ V(adj)$label[c(2, 8)] <- "Omnivory"
 V(adj)$label[3] <- "Disturbance intensity"
 V(adj)$label[4:5] <- "Productivity"
 
-V(adj)$label[c(6, 10, 11)] <- "No effect"
-V(adj)$label[7] <- "Very weak effect \n (Fig. S3)"
-V(adj)$label[c(8:9, 12)] <- "Positive"
+V(adj)$label[c(6, 10, 11)] <- "0"
+V(adj)$label[7] <- "(0)"
+V(adj)$label[c(8:9, 12)] <- "+"
 
 V(adj)$color[c(1:dim(m1)[1])] <- "node"
 V(adj)$color[c(6, 7, 8:12)] <- "pattern"
@@ -37,14 +37,15 @@ E(adj)$label[3:5] <- c("None (Chain)", "Weak", "Strong")
 E(adj)$label[6:7] <- c("Weak", "Strong")
 E(adj)$label[8:11] <- rep(c("Low", "High"), 2)
 
-g_tree1 <- ggraph(adj, layout = 'tree') +
+g_tree1 <- ggraph(adj, layout = "tree") +
   geom_edge_link(aes(label = label),
                  color = grey(0.7),
                  linetype = "dashed",
                  arrow = arrow(length = unit(2,'mm'),
                                type = "closed"),
-                 start_cap = circle(8,'mm'),
-                 end_cap = circle(8,'mm')) +
+                 start_cap = circle(5,'mm'),
+                 end_cap = circle(5,'mm'),
+                 angle_calc = "along") +
   geom_node_label(aes(label = label,
                       color = color),
                   label.size = 0) +
@@ -67,10 +68,10 @@ V(adj)$label[2] <- "Omnivory"
 V(adj)$label[3] <- "Disturbance intensity"
 V(adj)$label[4] <- "Productivity"
 
-V(adj)$label[c(5, 8)] <- "Positive"
-V(adj)$label[c(6, 10)] <- "No effect"
-V(adj)$label[7] <- "Very weak effect \n (Fig. S4)"
-V(adj)$label[9] <- "Negative"
+V(adj)$label[c(5, 8)] <- "+"
+V(adj)$label[c(6, 10)] <- "0"
+V(adj)$label[7] <- "(0)"
+V(adj)$label[9] <- "-"
 
 V(adj)$color[1:10] <- "node"
[…]

### 2023-07-10 code/set_figure_theme.R

@@ -15,9 +15,14 @@ plt_theme <- theme_bw() + theme(
   panel.grid.minor.x = element_blank(),
   panel.grid.minor.y = element_blank(),
   
-  axis.text = element_text(size = 10),
+  legend.title = element_text(size = 18),
+  legend.text = element_text(size = 18),
+  
+  axis.text = element_text(size = 18),
     
   strip.background = element_blank(),
-  strip.text = element_text(size = 12),
-  axis.title = element_text(size = 14)
+  strip.text = element_text(size = 22),
+  axis.title = element_text(size = 22),
+  axis.title.x = element_text(margin = margin(t = 0, r = 0, b = 20, l = 0)),
+  axis.title.y = element_text(margin = margin(t = 0, r = 0, b = 0, l = 20))
 )

### 2023-07-10 rmd/manuscript.Rmd
edit: manuscript typo

@@
 
-$m_{1,up}$ and $m_{2,up}$ are the disturbance values of upstream contributing tributaries, $\varepsilon_1$ and $\varepsilon_2$ are independent local noises, and $\omega$ the relative size of one tributary over another defined as $\omega = \frac{Q_{1}}{Q_{1} + Q_{2}}$.
+$m_{1,up}$ and $m_{2,up}$ are the disturbance values of upstream contributing tributaries, $\varepsilon_1$ and $\varepsilon_2$ are independent local noises, and $\omega$ the relative size of one tributary over another defined as $\omega = Q_{1}/(Q_{1} + Q_{2})$.
 Thus, the larger tributary has a greater influence on the downstream environment, as observed in natural river networks [@bendaNetworkDynamicsHypothesis2004].
@@
 where the parameter $s$ determines the strength of switching ($0 \le s \le 1$).
-In this expression, the intraguild predator increases its attack rate to the basal species if $\phi > 0$, properly representing the switching behavior of the intraguild prey.
+In this expression, the intraguild predator increases its attack rate to the basal species if $\phi > 0$, properly representing the switching behavior.
 
@@
 We also calculated the proportion of the patch state ("No species", "B", "B + C", "B + P", and "B + C + P") at each time step to infer mechanisms that drive changes in FCL.
-The time-averaged proportion of patch state $\bar{\psi}$ was obtained as $\bar{\psi} = q_t/(N_p T)$, where $q_t$ is the number of patches in a given patch state at time step $t$.
+The time-averaged proportion of patch state $\bar{\psi}$ was obtained as $\bar{\psi} = \sum_t q_t/(N_p T)$, where $q_t$ is the number of patches in a given patch state at time step $t$.
 
@@
 This setup ($\sigma_h > \sigma_l$) produces high spatial variation among tributaries while allowing strong within-tributary downstream correlation [@teruiEmergentDualScaling2021], a typical pattern in real river networks [@nakamuraDisturbanceRegimesStream2000; @bendaNetworkDynamicsHypothesis2004].
-We used a combination of basal species' growth rate $r_b$ and base carrying capacity $K_{base}$ (low productivity: $r_b =$ `r min(df_sim$r_b)`, $K_{base} =$ `r min(df_sim$base_k)` ; high productivity = $r_b =$ `r max(df_sim$r_b)`, $K_{base} =$ `r max(df_sim$base_k)` ) as a proxy for resource availability.
+We used a combination of basal species' growth rate $r_b$ and base carrying capacity $K_{base}$ (low productivity: $r_b =$ `r min(df_sim$r_b)`, $K_{base} =$ `r min(df_sim$base_k)` ; high productivity: $r_b =$ `r max(df_sim$r_b)`, $K_{base} =$ `r max(df_sim$base_k)` ) as a proxy for resource availability.
 For dispersal, we examined short- and long-distance dispersal scenarios (short-distance dispersal $\theta=$ `r max(df_sim$theta)`; long-distance dispersal $\theta=$ `r min(df_sim$theta)`) with constant dispersal probability ($p_{d}=$ `r unique(df_sim$p_dispersal)`).
@@
 In the chain scenario (no omnivory) with low productivity, ecosystem size had a positive effect on FCL (**Figure 3**), while ecosystem complexity had a negative influence (**Figure 4**).
-These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increased the total carrying capacity, whereas ecosystem complexity decreased it (**Appendix S1: Supplementary Text8**).
+These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increased the total carrying capacity, whereas ecosystem complexity decreased it (**Appendix S1: Figure S8**).
 However, when considering some levels of omnivory, these patterns characteristic of the chain scenario diminished.
@@
 In the weak and strong omnivory motifs, different state transitions were associated with ecosystem size and complexity.
-Ecosystem size increased FCL primarily through additive mechanisms, facilitating the state transition of "B" $\rightarrow$ "B + P" (low productivity) or "No species" $\rightarrow$ "B + P" (high productivity), particularly with frequent disturbance (**Figure 5**).
+Ecosystem size increased FCL primarily through additive mechanisms, facilitating the state transition of "B" $\rightarrow$ "B + P" (low productivity) or "No species" $\rightarrow$ "B + P" (high productivity), particularly with frequent disturbance (**Figure 6--7**).
 In contrast, insertion (i.e., the insertion of intermediate trophic species) and additive mechanisms were responsible for the effect of ecosystem complexity.
-With weak omnivory, the transition from "B + P" to "B + C + P" (insertion) was predominant when productivity was low (**Figure 6B**).
-However, the additive mechanism ("No species" $\rightarrow$ "B + P") took primacy in the high productivity scenario (**Figure 6B**).
+With weak omnivory, the transition from "B + P" to "B + C + P" (insertion) was predominant when productivity was low (**Figure 9**).
+However, the additive mechanism ("No species" $\rightarrow$ "B + P") took primacy in the high productivity scenario (**Figure 9**).
 In the strong omnivory motif, transition patterns depended on disturbance frequency.
-When disturbance was rare ($p_m = 0.05$), the insertion mechanism ("B + P" $\rightarrow$ "B + C + P") drove the FCL change (**Figure 6C**).
-As disturbance frequency increased, the primary process switched to an additive mechanism ("B" $\rightarrow$ "B + P" or "No species" $\rightarrow$ "B + P"; **Figure 6C**).
+When disturbance was rare ($p_m = 0.05$), the insertion mechanism ("B + P" $\rightarrow$ "B + C + P") drove the FCL change (**Figure 10**).
+As disturbance frequency increased, the primary process switched to an additive mechanism ("B" $\rightarrow$ "B + P" or "No species" $\rightarrow$ "B + P"; **Figure 10**).
 

### 2023-07-10 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 The proportion of each patch state $\bar{\psi}$ is averaged over time for each simulation replicate.
-Lines indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.Rows distinguish disturbance probability $p_m$, and columns indicate different dispersal and productivity scenarios.
+Lines indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals.
+Rows distinguish disturbance probability $p_m$, and columns indicate different dispersal and productivity scenarios.
 Different line colors indicate different disturbance probabilities.

### 2023-07-10 rmd/readme.Rmd
edit: readme

@@
   as_tibble() %>% 
-  mutate(Description = case_when(str_detect(value, "^figure_")~"Code for figures: '_concept', Figure 1; '_delta', Figure 1D; '_food_web', Figure 1C; '_treeplot', Figure2; '_main_geom', Figure 3, 4; '_network', Figure 1A; '_patch_state', Figure 5, 6",
+  mutate(Description = case_when(str_detect(value, "^figure_")~"Code for figures: '_concept', Figure 1; '_delta', Figure 1D; '_food_web', Figure 1C; '_treeplot', Figure2; '_main_geom', Figure 3, 4; '_network', Figure 1A; '_patch_state', Figure 5-10",
                                  str_detect(value, "^figure_si")~"Code for SI figures: '_geom', Figure S3, S4; '_one_patch', Figure S1, S2; '_r_disp', Figure S5; '_within_net', Figure S8",

### 2023-07-10 rmd/readme.Rmd
edit: readme

@@
 
-Title: Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks
+[![DOI](https://zenodo.org/badge/469160066.svg)](https://zenodo.org/badge/latestdoi/469160066)
+
+Title: Ecosystem size and complexity as extrinsic drivers of food chain length in branching networks
+
+Author: Justin Pomeranz, Jacques Finlay, Akira Terui
+
+Journal: Ecosphere
 

### 2023-07-11 rmd/manuscript.Rmd
edit

@@
 
+Track: Freshwater Ecology
+
 # Abstract
@@
 No primary data were used or produced in this research.
-R functions for simulations are provided as the R package mcbrnet version 1.3.1 (available at <https://github.com/aterui/mcbrnet>).
-Codes for simulations, statistical analysis, figures, and tables are available at <https://github.com/aterui/public-proj_fcl-river-theory>.
+R functions for simulations are provided as the R package mcbrnet version 1.3.1 (available at <https://github.com/aterui/mcbrnet>) [@teruiMcbrnetMetacommunitySimulation2023].
+Codes for simulations, statistical analysis, figures, and tables are available at <https://github.com/aterui/public-proj_fcl-river-theory> [@pomeranzCodeEcosystemSize2023].
 
@@
              booktabs = TRUE,
-             caption = "Parameter values used in the main simulation")
+             caption = "Table 1: Parameter values used in the main simulation")
 ```

### 2023-07-17 code/figure_treeplot.R
edit: figure 2

@@ -19,9 +19,9 @@ m1[5, 11:12] <- 1 # freq:no x omn:weak x prod
 
 adj <- igraph::graph.adjacency(m1)
 
-V(adj)$label[1] <- "Disturbance frequency"
+V(adj)$label[1] <- "Disturb frequency"
 V(adj)$label[c(2, 8)] <- "Omnivory"
-V(adj)$label[3] <- "Disturbance intensity"
+V(adj)$label[3] <- "Disturb intensity"
 V(adj)$label[4:5] <- "Productivity"
 
 V(adj)$label[c(6, 10, 11)] <- "0"
@@ -63,9 +63,9 @@ m2[4, 9:10] <- 1
 
 adj <- igraph::graph.adjacency(m2)
 
-V(adj)$label[1] <- "Disturbance frequency"
+V(adj)$label[1] <- "Disturb frequency"
 V(adj)$label[2] <- "Omnivory"
-V(adj)$label[3] <- "Disturbance intensity"
+V(adj)$label[3] <- "Disturb intensity"
 V(adj)$label[4] <- "Productivity"
 
 V(adj)$label[c(5, 8)] <- "+"

### 2023-07-17 rmd/manuscript.Rmd
Update manuscript.Rmd

@@
 In the weak and strong omnivory motifs, different state transitions were associated with ecosystem size and complexity.
-Ecosystem size increased FCL primarily through additive mechanisms, facilitating the state transition of "B" $\rightarrow$ "B + P" (low productivity) or "No species" $\rightarrow$ "B + P" (high productivity), particularly with frequent disturbance (**Figure 6--7**).
+Ecosystem size increased FCL primarily through additive mechanisms, facilitating the state transition of "B" $\rightarrow$ "B + P" (low productivity) or "No species" $\rightarrow$ "B + P" (high productivity), particularly with frequent disturbance (**Figures 6 and 7**).
 In contrast, insertion (i.e., the insertion of intermediate trophic species) and additive mechanisms were responsible for the effect of ecosystem complexity.

### 2023-07-17 code/figure_treeplot.R
edit: figure 2

@@ -48,7 +48,8 @@ g_tree1 <- ggraph(adj, layout = "tree") +
                  angle_calc = "along") +
   geom_node_label(aes(label = label,
                       color = color),
-                  label.size = 0) +
+                  label.size = 0,
+                  size = 4) +
   scale_color_manual(values = c("black", "steelblue")) +
   theme_void() +
   guides(color = "none")
@@ -71,7 +72,7 @@ V(adj)$label[4] <- "Productivity"
 V(adj)$label[c(5, 8)] <- "+"
 V(adj)$label[c(6, 10)] <- "0"
 V(adj)$label[7] <- "(0)"
-V(adj)$label[9] <- "-"
+V(adj)$label[9] <- "\u2212"
 
 V(adj)$color[1:10] <- "node"
 V(adj)$color[c(5:10)] <- "pattern"
@@ -93,7 +94,8 @@ g_tree2 <- ggraph(adj, layout = 'tree') +
                  angle_calc = "along") +
   geom_node_label(aes(label = label,
                       color = color),
-                  label.size = 0) +
+                  label.size = 0,
+                  size = 4) +
   scale_color_manual(values = c("black", "steelblue")) +
   theme_void() +
   guides(color = "none")
